### 2016-02-07 Paper_v1.tex
Commit Paper_v4

\documentclass[12pt]{article}
% Load packages
\usepackage{url}  % Formatting web addresses
\usepackage{ifthen}  % Conditional
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
\usepackage{setspace}
%\usepackage{longtable}
\usepackage{colortbl}
%\usepackage{palatino,lettrine}
%\usepackage{times}
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage[wide]{sidecap}
%\usepackage[authoryear,round,comma,sort&compress]{natbib}
\usepackage[square,sort,comma,numbers,sort&compress]{natbib}
%\usepackage[authoryear,round]{natbib}
\usepackage{supertabular}
\usepackage{simplemargins}
\usepackage{fullpage}
\usepackage{comment}
\usepackage{lineno}
%\usepackage{chicago}
\usepackage{textcomp}

%\usepackage[space]{cite}
\urlstyle{rm}

%\textwidth = 6.50 in
%\textheight = 9.5 in
%\oddsidemargin =  0.0 in
%\evensidemargin = 0.0 in
%\topmargin = -0.50 in
%\headheight = 0.0 in
%\headsep = 0.25 in
%\parskip = 0.15in
%\linespread{1.75}
\doublespace

%\bibliographystyle{chicago}
\bibliographystyle{plos2009}

\makeatletter
\renewcommand\subsection{\@startsection
	{subsection}{2}{0mm}
	{-0.05in}
	{-0.5\baselineskip}
	{\normalfont\normalsize\bfseries}}
\renewcommand\subsubsection{\@startsection
	{subsubsection}{2}{0mm}
	{-0.05in}
	{-0.5\baselineskip}
	{\normalfont\normalsize\itshape}}
\renewcommand\section{\@startsection
	{subsection}{2}{0mm}
	{-0.2in}
	{0.05\baselineskip}
	{\normalfont\large\bfseries}}
\renewcommand\paragraph{\@startsection
	{paragraph}{2}{0mm}
	{-0.05in}
	{-0.5\baselineskip}
	{\normalfont\normalsize\itshape}}
\makeatother

%Review style settings
%\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings

% Single space'd bib -
\setlength\bibsep{0pt}

\renewcommand{\rmdefault}{phv}\renewcommand{\sfdefault}{phv}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}

% Change the number format in the ref list -
\renewcommand{\bibnumfmt}[1]{#1.}

% Change Figure to Fig.
\renewcommand{\figurename}{Fig.}

% Begin ...
\begin{document}
\begin{titlepage}
{\par\centering\textbf{\Large {Dynamically dimensioned particle swarm optimization (DdPSO): A meta-heuristic for parameter estimation in biochemical models}}}
\vspace{0.05in}
{\par \centering \large{Adithya Sagar, Christine Shoemaker$^{\dag}$ and Jeffrey D. Varner$^{*}$}}
\vspace{0.10in}
{\par \centering {School of Chemical and Biomolecular Engineering}}
{\par \centering {$^{\dag}$School of Civil and Environmental Engineering}}
{\par \centering {Cornell University, Ithaca NY 14853}}
\vspace{0.1in}
{\par \centering \textbf{Running Title:}~Parameter estimation in biochemical models}
\vspace{0.1in}
{\par \centering \textbf{To be submitted:}~\emph{PLoS ONE}}
\vspace{0.5in}
{\par \centering $^{*}$Corresponding author:}
{\par \centering Jeffrey D. Varner,}
{\par \centering Associate Professor, School of Chemical and Biomolecular Engineering,}
{\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853}
{\par \centering Email: jdv27@cornell.edu}
{\par \centering Phone: (607) 255 - 4258}
{\par \centering Fax: (607) 255 - 9166}
\end{titlepage}
\date{}
\thispagestyle{empty}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}
Mathematical modeling is as a powerful tool to analyze, and ultimately design biochemical networks.
However, estimation of biochemical model parameters can be a significant challenge.
Parameter estimation in biochemical models generally involves expensive function evaluations, making it difficult to quickly obtain optimal solutions.
Additionally, biochemical models often have many local minima or maxima which further complicates the parameter estimation problem.
In this study, we developed dynamically dimensioned particle swarm optimization (DdPSO), a novel meta-heurstic that combined the global search capability of multi-swarm particle swarm optimization with dynamic dimensioned search. DdPSO uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors which are then greedily updated using dynamic dimensioned search. We tested the performance of DdPSO on commonly used test functions for global optimization (Ackley and Rosenbrock), published biochemical parameter estimation benchmark problems and
lastly on a model of the human coagulation cascade. For each test we performed 25 trials with 4000 function evaluations per trial, and compared the performance of DdPSO with other commonly
used meta-heuristic approaches such as differential evolution (DE), particle swarm optimization (PSO), simulated annealing (SA) and dynamically dimensioned search (DDS).
For the wide range of problems that we considered, DdPSO outperformed the other meta-heuristic approaches despite a limited number of function evaluations.
We further tested the predictive power of the coagulation model parameters against data not used in training, and found good agreement between model simulations and experimental measurements.
Taken together, DdPSO is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.

{\noindent \textbf{Keywords:}~Parameter identification, Meta-heuristic optimization, Biochemical modeling}

% Extra abstract
% Mathematical modeling of biological systems with multiple feed back loops is one such area where parameter estimation is a difficult non-linear optimization problem. This difficulty is further compounded when dealing with parameter vectors of high dimensions.

%In this study, we present the dynamically dimensioned particle
%a novel meta-heuristic approach that combines a variant of particle swarm optimization (PSO) along with dynamically dimensioned search (DDS) to obtain near optimal solutions of high dimensional biochemical networks within a relatively few function evaluations.
%We use a particle swarm optimization technique that uses multi-swarms to generate candidate vectors which are then greedily updated using DDS by dynamically varying the perturbed parameter dimensions. We tested this algorithm (25 trials with 4000 function evaluations in each trial) on a biochemical network of coagulation (148 parameters and 92 species) and compared it's performance against other meta-heuristics like Differential Evolution (DE), Particle Swarm Optimization (PSO), Simulated Annealing (SA) and also against DDS alone. The new algorithm outperforms all the other meta-heuristics on the coagulation model. The parameter vectors obtained using this approach fit the experimental data well and also make accurate enough predictions on unseen experimental data. We also performed this comparison on commonly used test functions (Ackley and Rosenbrock) for global optimization and found the same behavior. Further we used two recently published benchmark problems, a genome wide kinetic model with 1759 parameters and a metabolic model of Chinese Hamster Ovary cells with 117 parameters to evaluate the performance of our approach. We  surprisingly performed well on these benchmarks and obtained the nominal parameter vector with just 4000 function evaluations in both cases.

\pagebreak

\setcounter{page}{1}

\linenumbers

\section*{Introduction}

Cells process nutrients and respond to changes in their environment using complex biochemical networks. These networks contain thousands of components
interconnected through nonlinear enzyme catalyzed reactions. Mathematical modeling has evolved as a powerful paradigm to analyze, and ultimately design these complex networks \cite{assmus2006dynamics, Riel:2006aa,Jaqaman:2006aa,kitano2002systems,hood2004systems}. Mathematical modeling of biochemical networks is often an iterative process.
First, models are formulated from biochemical knowledge, and then trained using using experimental data \cite{Aldridge:2006aa,banga2008optimization,ashyraliyev2009systems}.
Model training is typically framed as a non-linear optimization problem wherein the residual (or objective function) between experimental data and model simulations is minimized using an optimization strategy \cite{moles2003parameter}. The optimal parameters obtained from training are then used to validate the model on hitherto unseen experimental data. If validation fails, model construction and calibration are repeated iteratively until satisfactory results are obtained.

Parameter estimation is a major challenge in the development of large biochemical models.
Although parameter estimation has been a well studied problem in engineering for decades \cite{nieman1971review,beck1977parameter,young1981parameter,beck1998inverse},
the complex dynamics of large biological systems and noisy, often incomplete experimental data pose a unique estimation challenge.
Most of the optimization problems pertaining to biological systems are non-linear and multi-modal i.e. most of them have multiple local minima or maxima \cite{moles2003parameter,banga2008optimization}. The non-linearity of the problem coupled with multi-modality generally renders the local optimization techniques like pattern search method \cite{hooke1961direct}, Nelder-Mead simplex method \cite{nelder1965simplex} or Levenberg-Marquardt method\cite{more1978levenberg} incapable of obtaining an optimal solution. Though deterministic global optimization techniques can handle non-linearity and multi-modality \cite{esposito2000deterministic,horst2013global}, the absence of derivative information, discontinuity of the objective functions, non-smooth regions or the lack of any knowledge about the objective function severely hamper the use of these techniques.

Meta-heuristic approaches like Genetic Algorithms, Simulated Annealing \cite{kirkpatrick1983optimization}, Evolutionary Programming and population based searches like Differential Evolution\cite{storn1997differential} have shown promise in this regard \cite{sun2012parameter}. They do not make any assumptions about the structure of objective function or require any \textit{a priori} information about the objective function. Though they do not guarantee strong convergence, these approaches are effective in finding near optimal solutions.  Mendes et al. \cite{mendes1998non} used Simulated Annealing to estimate rate constants for the irreversible inhibition of HIV proteinase, Modchang et al. \cite {modchang2008mathematical} used Genetic Algorithms to estimate parameters for a mathematical model of signal transduction, Differential Evolution based approaches have been an effective approach on various systems \cite{tsai2005evolutionary,wang2001hybrid,noman2007inferring}. Tashkova et al. \cite{tashkova2011parameter} compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis and showed that DE was the most effective. Banga and co-workers have applied scatter-search based methods \cite{villaverde2012cooperative,rodriguez2006novel,egea2007scatter} to estimate parameters on non-linear biological processes. Hybrid approaches that combine a meta-heuristic with a local optimization search, wherein a near optimal solution that is obtained with a meta-heuristic is further refined using a local search have also become popular. Villaverde et al. \cite{villaverde2015biopredyn} combined scatter search with local search methods for parameter estimation in large scale systems biology models. Fan et al. recently showed that population based meta-heuristics along with decomposition based methods can be used to model gene circuits from mRNA data \cite{fan2015parameter}.Despite these successes, a  major drawback with most of these approaches is the vast number of objective function evaluations they take in arriving at good search regions. As the models grow in size and complexity and the number of dimensions of parameter vector increases,  parameter estimation becomes a computationally expensive task. In many of these high dimensional problems approaching an exact solution may not be necessary. Gutenkust et al. \cite{gutenkunst2007universally} showed that a number of systems biology models are 'sloppy'. Sloppy systems have specific parameter combinations that largely define the dynamics of the system. Large perturbations to the rest of the parameters does not greatly impact the system dynamics. Ensemble approaches \cite{song2010ensembles,luan2010ensembles} have exploited this aspect to describe the dynamics of biological systems including coagulation which can be described using only a set of key species or parameters \cite{sagar2015dynamic}. Tolson and Shoemaker \cite{tolson2007dynamically} showed through Dynamically Dimensioned Search (DDS) that high-dimensional watershed models can be calibrated quickly by perturbing only a subset of dimensions.

In this study, we developed dynamically dimensioned particle swarm optimization (DdPSO), a novel meta-heurstic that combined the global search capability of multi-swarm particle swarm optimization with dynamic dimensioned search. The objective of DdPSO was to obtain near optimal parameter estimates for large biochemical models within a relatively few function evaluations.
DdPSO uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors which are then greedily updated using dynamic dimensioned search.
We tested the performance of DdPSO on commonly used test functions for global optimization (Ackley and Rosenbrock), published biochemical parameter estimation benchmark problems and
lastly on a model of the human coagulation cascade. DdPSO outperformed common meta-heuristic approaches like Differential Evolution (DE), Particle Swarm Optimization (PSO), Simulated Annealing (SA) and dynamically dimensioned search (DDS) on the benchmark and coagulation problems.

% Extra text from the introduction -
%With the capacity to construct very large models using various mathematical formalisms \cite{chen2009input,tasseff2011modeling,luan2007computationally,mo2007genome,orth2011comprehensive,karr2012whole,buchel2013path2models,smallbone2010towards}, more often than not,

%this algorithm (25 trials with 4000 function evaluations in each trial) on a biochemical network of coagulation (148 parameters and 92 species) and compared it's performance against other meta-heuristics  The new algorithm outperforms all the other meta-heuristics on the coagulation model. The parameter vectors obtained using this approach fit the experimental data well and also make accurate enough predictions on unseen experimental data. We also performed this comparison on  and found the same behavior. Further we used two recently published benchmark problems, a genome wide kinetic model with 1759 parameters and a metabolic model of Chinese Hamster Ovary cells with 117 parameters to evaluate the performance of our approach. We  surprisingly performed well on these benchmarks and obtained the nominal parameter vector with just 4000 function evaluations in both cases.
%took into cognizance the fact that it may not be necessary to obtain the exact solution for high dimensional biological systems and that good enough solutions can be quickly obtained without expending a lot of objective function evaluations. Hence our current approach uses the power of population based heuristics along with DDS to obtain near optimal or good solutions. Though in theory we can combine any meta-heuristic with DDS, for the current purpose we used Particle Swarm Optimization (PSO). PSO unlike Genetic Algorithm (GA) or Differential Evolution (DE) does not have complex operations like cross over, mutation or recombination. It is simple to use and does not have a lot of parameters associated with other heuristics. However PSO is known to rapidly converge to a local optimum and thus several variants \cite{peer2003using,zhan2009adaptive,li2007fast} have been developed of which Multi swarm particle swarm optimization approaches (MLSPSO) \cite{zhao2008dynamic,liang2005dynamic} are one of the effective ones. We used MLSPSO to preclude bad regions of search and thereafter search within these regions using DDS.

\clearpage

\section*{Results}
\subsection*{Dynamically dimensioned particle swarm optimization (DdPSO).}
DdPSO is a novel meta-heuristic that combines multi-swarm based particle swarm methods with dynamically dimensioned search (Fig. \ref{fig-algorithm}).
The goal of this approach is to obtain optimal or near optimal parameters for high-dimensional complex biological systems within a pre-specified number of function evaluations. We randomly initialized a swarm of $\mathcal{K}$-dimensional particles (represented as ${x}_{i}$), wherein each of these particles corresponds to a $\mathcal{K}$-dimensional parameter vector. After initialization, the particles were grouped into different sub-swarms randomly. Thereafter within each sub-swarm ${S}_{k}$,  particles were updated according to the following rule.
\begin{eqnarray}
	\mathbf{x}_{i,j} &=&\theta_{1,j-1}\mathbf{x}_{i,j-1} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{x}_{i,j-1}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{GL}_{k} - \mathbf{x}_{i,j-1}\right)
\end{eqnarray}
where $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ are adjustable parameters, $\mathcal{L}_{i}$ denotes the best local solution found by particle $i$ till function evaluation $j$-$1$, and
$\mathcal{GL}_{k}$ denotes the best local solution found over the population of all particles within the swarm ${S}_{k}$. The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model
parameters ($\mathcal{K}\times{1}$). In our algorithm the parameter $\theta_{1,j-1}$ depends on the function evaluations and is controlled according to the following equation
\begin{eqnarray}
	\mathbf \theta_{1,j}&=&((\mathbf{N}-{j})*(\mathbf{w}_{max}-\mathbf{w}_{min}))/(\mathbf{N}-{1}) + \mathbf{w}_{min}
\end{eqnarray}
where $\mathbf{N}$ represents the total number of function evaluations, $\mathbf{w}_{max}$ and $\mathbf{w}_{min}$ are the maximum and minimum inertia weights respectively. While updating the particle, we made sure all dimensions of the solution represented by the particle were within bounds using a reflection boundary condition. After every $\mathbf{g}$  function evaluations, the particles within all sub-swarms were mixed and then randomly redistributed to a new sub-swarm. The particles were then again updated according Eq.1. This process continued till $\mathcal{FR}*\mathbf{N}$ number of functions evaluations, where $\mathcal{FR}$ represents the fraction of evaluations with the multi-swarms. At the end of these function evaluations, we froze all the solutions represented by various particles and chose the particle with best solution among $\mathcal{GL}_{1} \cdots \mathcal{GL}_{NS}$ as the initial candidate vector $\mathcal{G}$ for the remaining $({1}-\mathcal{FR})*\mathbf{N}$ number of function evaluations.

This particle was then updated according to the following rule
\begin{eqnarray}
	\mathcal{G}_{new}(\mathbf{J})& = &\mathcal{G}(\mathbf{J})+\mathbf{r}_{normal}(\mathbf{J})\sigma(\mathbf{J})
\end{eqnarray}
where $\mathbf{J}$ represents the set containing the specific dimensions being perturbed, ${r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$. $\sigma$ is the amplitude of perturbation given by following equation:
\begin{eqnarray}
	\sigma& = &\mathbf{R}(\mathcal{MAX} -\mathcal{MIN})
\end{eqnarray}
where $\mathbf{R}$ is the scalar perturbation size parameter, $\mathcal{MAX}$ and $\mathcal{MIN}$ are ($\mathcal{K}\times{1}$) vectors that represent the maximum and minimum bounds on each dimension. The set $\mathbf{J}$ was constructed using a probability function $\mathcal{P}_{j}$ that represented the probability whether a specific dimension $j$ was perturbed or not.  This function is a monotonically decreasing function that decreases with the number of function evaluations. $\mathcal{P}_{j}$ can be any monotonically decreasing function, in our approach we used the following function:
\begin{eqnarray}
	\mathcal{P}_{j}&=&{1}-\log(j/ (\beta({1}-\mathcal{FR})*\mathbf{N}))
\end{eqnarray}
where $\beta$ is the perturbation frequency probability modulator. Thus the number of dimensions of the candidate vector that are updated or perturbed decreases with the as the number of function evaluations increase. These updates are greedy in nature, so $\mathcal{G}_{new}$ becomes the new solution vector only if it is better than the old one $\mathcal{G}$.

\subsection*{Performance of DdPSO on benchmark problems.}
We compared the performance of DdPSO with algorithms on commonly used test functions for global optimization, as well as benchmark biochemical problems from literature (Table \ref{table:benchmark-problems}). We used a 300 dimensional Rastrigin function and 300 dimensional Ackley function. Both Ackley and Rastrigin have multiple local minima and maxima and attain a global minimum value of 0. We tested the performance of the 5 different heuristic approaches on these two functions. In each experiment we again fixed the number of function evaluations at 4000 and ran 25 experiments. In both cases (Figure S6 A and B) we see that the error convergence rate for DDSMLSPSO is much faster as compared to the other four s and it finds the global minimum of 0 in both cases.

Villaverde and co-workers recently published a set of benchmark problems to evaluate parameter estimation methods \cite{villaverde2015biopredyn}. From a computational cost perspective problems they categorized the problems as most expensive, intermediate and least expensive. We evaluated the performance of our algorithms on a problem from the most expensive and least expensive categories. The first problem is a genome wide kinetic model of \textit{Saccharomyces cerevisiae} with 1759 parameters (henceforth referred to as problem B1).  The second problem (henceforth referred to as problem B4) is a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters. In both cases pseudo time series data was generated. For problem B1, the time series data consisted of 44 observables and for problem B4 the data corresponded to 13 different metabolites. We fixed the number of function evaluations at 4000 for DDSMLSPSO and trained both the models against the pseudo experimental data. We recaptured the 'nominal' parameters (Figure S7 A and B)  in both cases within 4000 evaluations. The final objective function value (Table 1) is an order smaller for B1 possibly due to overfitting. In Table 1 we summarize the results obtained on the coagulation model and the benchmark problems.

\subsection*{Performance of DdPSO on a model of the human coagulation cascade.}
We compared the performance of DdPSO on a model of blood coagulation against commonly used meta-heuristics like simulated annealing (SA), differential evolution (DE), particle swarm optimization (PSO),and dynamically dimensioned search (DDS). Coagulation is an archetype biochemical network that is highly interconnected and tightly regulated with multiple positive and negative feedback loops (Fig. \ref{fig-coagulation-network}). The biochemistry underlying coagulation, though quite complex has been well studied \cite{mann2003dynamics,mann2003all,mann2003thrombin,vogler2009contact,diamond2013systems,fogelson2005coagulation,anand2003model}, and reliable experimental coagulation models have been developed \cite{hockin2002model,chatterjee2010systems,mann2006models,luan2007computationally}. This makes it an ideal system for mathematical modeling and parameter estimation. Coagulation is regulated by a set of serine proteases also known as coagulation factors and blood platelets. The coagulation factors are generally in an inactive state and are known as zymogens. These zymogens are activated through certain triggers. Trigger events like injury or trauma or sepsis expose factors like collagen, tissue factor and von Willebrand factor (vWF) to blood. The exposure of these factors to blood kick-starts a series of convergent cascades that lead to conversion of zymogen prothrombinase to thrombin.   Luan et al. modeled coagulation using coupled non-linear ordinary differential equations with 148 reactions and 92 species \cite{luan2007computationally} and validated model using 21 published datasets.

 To train the model parameters, we used data sets from TF-VIIa initiated coagulation with no anticoagulants. The objective function was a weighted linear combination of two different error functions that used data sets representing coagulation initiated with different concentrations of TF-VIIa (5pM, 5nM) \cite{hockin2002model}. We restricted the number of function evaluations to 4000 for each algorithm and performed 25 trials of this experiment. DdPSO exhibits a much faster rate of error convergence and has a much lower final error as compared to the other algorithms (Figure 3). Within the first 1000 function evaluations (swarm phase) of DDSMLSPSO there is a very rapid drop in error. Subsequently, again after 2000 function evaluations (dynamically dimensioned phase) the error drops quickly. Overall at the end of 4000 function evaluations DDSMLSPSO minimizes the error (final objective error is 90456.9) to a much greater extent than any of the other algorithms. Amongst the rest of the algorithms DDS has a comparable error drop for the first 2000 function evaluations. However the drop in error remains constant for DDS after 2000 function evaluations. Using the parameters obtained at the end of 4000 function evaluations we examined the 'fits' between models predictions and experimental data (Figure 2). The solid lines represent the mean value of prediction over 25 trials and the shaded region represents the 99\% confidence interval. Subsequently we compared the model predictions against completely 'unseen' or untrained experimental where coagulation was initiated with 500pM,50pM,10pM concentrations of TF-VIIa respectively (Figure 3).

\clearpage
\section*{Discussion}
Our study presents a novel approach for high-dimensional parameter estimation in complex biological systems with relatively few function evaluations. In this approach we combined a variant of a well known meta heuristic particle swarm optimization with Dynamically Dimensioned Search (DDS). We tested our approach on an ODE model of coagulation with 148 parameters and 92 species. Coagulation is an ideal system to test our approach since the biology is well known and complex, with multiple feed back loops that are tightly regulated. We used experimental data with under different conditions to obtain optimal parameters and used these parameters to make predictions against unseen experimental data. We obtained good fits and made accurate enough predictions using parameters obtained from 4000 function evaluations. Further, we also used high-dimensional forms of commonly used test functions of global optimization and showed that we were able to find the global minimum for 300 dimensional Ackley and Rastrigin functions faster than other meta-heuristics. We also considered two recently published benchmark problems to test parameter optimization approaches and showed that we were able to retrieve the nominal parameter vector within 4000 function evaluations.

meta-heuristic approaches are generally effective in finding close to optimum solutions of complex, multi-modal functions. In addition, they generally obviate the need for any \textit{a priori} knowledge (like function derivative). However they take an exorbitant number of objective function evaluations to come close to an optimum. When the objective function evaluations tend to become expensive it is infeasible to take up a large number of evaluations. As the dimensionality of parameter space increases, the search region gets widened and thus the problem becomes more challenging.  In addition, most of these approaches require optimization of  'algorithm parameters' before the actual optimization and also involve computationally expensive update operations.  Tolson and Shoemaker, through DDS, showed that randomly perturbing a subset of dimensions in high dimensional parameter space is an effective way to obtain near optimal solutions with few function evaluations. However their approach was based on a single solution. Single solutions based methods may prove ineffective in complex systems when they start from a 'bad' region of search or from a local minimum. In our approach, we tried to eliminate the probability of starting from a bad region of search using a variant of particle swarm optimization. Particle Swarm Optimization (PSO) is a population based meta heuristic that is known for rapid convergence without any complex operations like recombination, mutation or selection that are associated with other population based meta-heuristics like Differential Evolution (DE) or Genetic Algorithm (GA). Several particle swarm variants have been proposed to improve the search ability and rate of convergence, that involve different neighborhood structures, multi-swarms or adaptive parameters. Multi-swarm PSO with small particle neighborhoods have been shown to better in searching on complex multi-modal solutions \cite{zhao2008dynamic}. Multi swarm methods, in addition, avoid rapid convergence to a local optimum or stable point and are able to generate diverse solutions.  Generation of diverse solutions in the early stage gives a better exploratory capability and thus converge of upon multiple optima.

We utilized this capacity of multi swarms to generate diverse candidate solutions which can be used as initial solutions for a DDS like search. Thus the solution vectors obtained at the end of multi swarm search phase have a better propensity to not get stuck in bad regions in DDS phase as opposed to starting with a totally random initial solution. Given a fixed number of function evaluations, we have been able to show that we were able to obtain better solutions for the coagulation model and the test functions faster that other commonly known meta-heuristics and also DDS alone.  Choosing the number of function evaluations is largely a function of cost and complexity of the objective function. Traditionally the stopping condition for a parameter optimization problem can be the number of function evaluations, percentage of initial error achieved or an absolute error threshold. However in case of complex, expensive functions where we desire a value within a certain period of time, the number of evaluations are used as a stopping criterion. In our current study we used a value of 4000 which we based upon the time taken (approximately 8-10 seconds) for a single objective function evaluation in the coagulation case. We used the same value of 4000 for benchmarks published by Villaverde and co-workers \cite{villaverde2015biopredyn}. Quite surprisingly we took a couple of orders lesser number of function evaluations (1e3) to obtain the optimal parameter vector as compared to the enhanced Scatter Search (eSS) with a local optimizer, which took around the order of 1e5 number of evaluations.  The amount of CPU time taken (on an Intel Xeon processor 2.4 GHz) is lesser as compared to eSS on a similar architecture.

A surprisingly remarkable aspect about our algorithm has been that we did not 'pre optimize' any parameters of the algorithm to suit a specific problem. In the swarm phase this includes the number of particles, number of sub-swarms, acceleration constants or the number of generations after which the particles are redistributed and the neighborhood perturbation parameter in DDS phase. We used the same parameters for all the problems. The same rule was applied to the rest of the meta-heuristics barring Simulated Annealing. For SA, we optimized the cooling schedule for the coagulation model. Thus, in this approach any overhead that usually comes with additional function evaluations to optimize parameters was avoided. The only parameter that we adjusted was the fraction of evaluations within swarm search. For the coagulation model it was set to be 0.4 and for the benchmarks it is at 0.6. This parameter was set based on empirical observations rather than any definitive analysis. When the error convergence rates stagnated in the swarm phase we switched to DDS phase. A further improvement to our approach could  be to incorporate a meta intelligent way of switching in and switching out of both the phases.

The performance of our approach seems impressive given that it performed well on different complex systems with no specific conditioning being required. We comfortably outperformed common meta-heuristics and were also able to find minima of high dimensional global optimization test functions. Thus this approach may be well suited to large scale global optimization. In addition, surprisingly, we were able to obtain optimal parameter vectors for two different large scale systems biology models with a couple of orders fewer number of function evaluations as compared a hybrid approach. However it is quite possible that highly optimized versions of common meta heuristics may outperform us on these systems. This aspect is currently beyond the scope of this study. Our approach can also be combined with local derivative based searches to improve upon the accuracy of the solutions.

%\begin{figure}[H]
%\centering
%\includegraphics[width=1.00\textwidth,height=0.7\textheight]{./figs/Figure_2_CoagulationNetwork.pdf}
%\caption{Schematic of the extrinsic and intrinsic coagulation cascade. Inactive zymogens upstream (grey) are activated by exposure to tissue factor (TF) \cite{giesen1999blood} following vessel injury. Tissue factor and activated factor VIIa (FVIIa) form a complex that activates factor X (fX) and IX (fIX). FXa activates downstream factors including factor VIII (fVIII) and fIX. Factor V (fV) is primarily activated by thrombin (FIIa). In addition, we included a secondary fV activation route involving FXa. FXa and FVa form a complex (prothrombinase) on activated platelets that converts prothrombin (fII) to FIIa. FIXa and FVIIIa can also form a complex (tenase) on activated platelets which catalyzes FXa formation. Localized platelets are activated by external signals such as adenosine diphosphate (ADP) and thromboxane A2 (TXA2) or thrombin through protease-activated receptors (PARs) \protect\cite{coughlin2000thrombin} \protect \cite{coughlin1999protease} Thrombin also activates upstream coagulation factors, forming a strong positive feedback ensuring rapid activation. Tissue factor pathway inhibitor (TFPI) downregulates FXa formation and activity by sequestering free FXa and TF-FVIIa in a FXa-dependent manner. Antithrombin III (ATIII)  inhibits all proteases. Thrombin inhibits itself binding the surface protein thrombomodulin (TM) \protect\cite{esmon1989roles} The IIa-TM complex catalyzes the conversion of protein C (PC) to activated protein C (APC), which attenuates the coagulation response by the proteolytic cleavage of fV/FVa and fVIII/FVIIIa \protect\cite{esmon1989roles}}\label{fig-network}
%\end{figure}

\section*{Acknowledgements}
This study was supported by an award from the Army Research Office (ARO \#59155-LS).
\clearpage

%\bibliography{References_v1}
%\begin{thebibliography}{9999}
\bibliography{References_v2}

\clearpage

\begin{table}[h]
\centering
\caption{Table with optimization settings and results for coagulation and different benchmarks using DdPSO.}
\label{table:benchmark-problems}
\resizebox{\columnwidth}{!}{
\begin{tabular}{|cccccc|}
\hline
                        & Coagulation           & \multicolumn{4}{l}{Benchmarks}\\
 \hline
                        & Experimental Data Set & B1           & B4          & Ackley 300D & Rastrigin 300D \\
 \hline
Lower Bound             & 0.001.pnom            & 5.pnom       & 5.pnom      & 30          & 5.12           \\
Upper Bound             & 1000.pnom             & 0.2.pnom     & 0.2.pnom    & -15         & -5.12          \\
CPU Time                & 10.0835 hours         & 38.308 hours & 6.2 minutes & 2.8 seconds & 2.6 seconds    \\
Function Evaluations    & 4000                  & 4000         & 4000        & 4000        & 4000           \\
Initial Objective Value & 1.11837e+07           & 1.4275e+07   & 1.8536e+07  & 21.12       & 99.985         \\
Final Objective Value   & 90456.9               & 4.96e+05     & 38.9375     & 0           & 0              \\
Nominal Objective Value & 4.7785e+06            & 1.0986e+06   & 39.0676     & 0           & 0      \\
\hline
\end{tabular}
}
\end{table}

\clearpage

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth,height=0.5\textheight]{./figs/Figure_1_Algorithm.pdf}
\caption{Multi Swarm Particle Swarm Optimization with Dynamically Dimensioned Search. \textbf{A}: Each particle represents an N dimensional parameter vector. Particles are randomly initialized and grouped into different sub-swarms. Within each swarm the magnitude and direction of the movement a particle is influenced by the position of the best particle and also by its own experience. After every $\mathbf{g}$ number of function evaluations the particles are mixed and randomly assigned to different swarms. At the end of the evaluations ($\mathcal{FR}$) assigned to swarm search, the global best particle (orange color) amongst all sub-swarms is chosen as the candidate parameter vector for Dynamically Dimensioned Search \textbf{B}: The candidate vector does a greedy search in a dynamic neighborhood. The search neighborhood is dynamically adjusted by varying the number of dimensions that are perturbed (in black) in each evaluation step. The probability that a dimension is perturbed decreases as the number of function evaluations increase. Thus as the evaluations increase the optimality of the solution is preserved.
}\label{fig-algorithm}
\end{figure}

\clearpage

\begin{figure}[h]
\centering
\includegraphics[width=1.00\textwidth,height=0.7\textheight]{./figs/Figure_2_CoagulationNetwork.pdf}
\caption{Schematic of the extrinsic and intrinsic coagulation cascade. Inactive zymogens upstream (grey) are activated by exposure to tissue factor (TF)  following vessel injury. Tissue factor and activated factor VIIa (FVIIa) form a complex that activates factor X (fX) and IX (fIX). FXa activates downstream factors including factor VIII (fVIII) and fIX. Factor V (fV) is primarily activated by thrombin (FIIa). In addition, we included a secondary fV activation route involving FXa. FXa and FVa form a complex (prothrombinase) on activated platelets that converts prothrombin (fII) to FIIa. FIXa and FVIIIa can also form a complex (tenase) on activated platelets which catalyzes FXa formation.  Thrombin also activates upstream coagulation factors, forming a strong positive feedback ensuring rapid activation. Tissue factor pathway inhibitor (TFPI) downregulates FXa formation and activity by sequestering free FXa and TF-FVIIa in a FXa-dependent manner. Antithrombin III (ATIII)  inhibits all proteases. Thrombin inhibits itself binding the surface protein thrombomodulin (TM). The IIa-TM complex catalyzes the conversion of protein C (PC) to activated protein C (APC), which attenuates the coagulation response by the proteolytic cleavage of fV/FVa and fVIII/FVIIIa. }\label{fig-coagulation-network}
\end{figure}

\clearpage

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth,height=0.5\textheight]{./figs/Figure_3_Errors_convergence.pdf}
\caption{Error convergence rates of the 5 different algorithms on the coagulation model. The objective error is the mean over $N$= 25 trials. DDSMLSPSO and DDS have the steepest drop in error during first 2000 function evaluations. Thereafter the error drop in DDS remains nearly constant whereas DDSMLSPSO drops further from around 2000 to 4000 function evaluations. At the end of 4000 function evaluations DDSMLSPSO attains the lowest error.
}\label{fig-train}
\end{figure}

\clearpage

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth,height=0.5\textheight]{./figs/Figure_4_Sim_Train_E1_E5.pdf}
\caption{Model fits on experimental data using DDSMLSPSO. The model parameters were estimated using DDSMLSPSO. Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from $N$= 25 trials. The grey shaded region represents the 99\% confidence estimate of the mean simulated thrombin concentration. The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers. Thrombin generation is initiated by adding Factor VIIa-TF (5nM - Red and 5pM - Green) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
}\label{fig-train}
\end{figure}

\clearpage

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth,height=0.5\textheight]{./figs/Figure_5_Sim_Validate_E2_E4_E6.pdf}
\caption{Model predictions on unseen experimental data using parameters obtained from DDSMLSPSO. The parameter estimates that were obtained using DDSMLSPSO were tested against data that was not used in the model training. Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from $N$= 25 trials. The grey shaded region represents the 99\% confidence estimate of the mean simulated thrombin concentration. The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers. Thrombin generation is initiated by adding Factor VIIa-TF (500pM - Blue, 50pM - Pink and 10pM - Orange respectively) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
}\label{fig-validation}
\end{figure}

%\bibitem[Saltelli \em{et~al.}(2010)Saltelli, Annoni, Azzini, Campolongo, Ratto,
  %and Tarantola]{Saltelli:2010}
%Saltelli, A.; Annoni, P.; Azzini, I.; Campolongo, F.; Ratto, M.; Tarantola, S.
%\newblock Variance based sensitivity analysis of model output. Design and
  %estimator for the total sensitivity index.
%\newblock {\em Comput. Phys. Commun.} {\bf 2010}, {\em 181},~259--270.

%\end{thebibliography}

\clearpage

% Supplemental figures -
% Set the S-
\renewcommand\thefigure{S\arabic{figure}}
\renewcommand\thetable{T\arabic{table}}
\renewcommand\thepage{S-\arabic{page}}
\renewcommand\theequation{S\arabic{equation}}

% Reset the counters -
\setcounter{equation}{0}
\setcounter{table}{0}
\setcounter{figure}{0}
\setcounter{page}{1}

\begin{figure}[ht]
\centering
\includegraphics[width=1.02\textwidth,height=0.5\textheight]{./figs/Figure_6_Ackley_Rast}
\caption{Error convergence on 300 dimensional Ackley function and Rastrigin function. \textbf {(A)} Error convergence of 5 different meta-heuristics on 300 dimensional Ackley function \textbf {(B)} Error convergence of 5 different meta-heuristics on 300 dimensional Rastrigin function
}\label{fig-testfunctions}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.02\textwidth,height=0.5\textheight]{./figs/Figure_7_Benchmarks_Parameters}
\caption{Difference between optimal and nominal parameter vector values on benchmark problems. \textbf {(A)} Problem B1: Genome wide kinetic model of E.coli \textit{S.cerevisiae} with 1759 unknown parameters. \textbf {(B)} Problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
}\label{fig-benchmark}
\end{figure}

\end{document}

### 2016-02-07 Paper_v3.tex
Commit Paper_v4

--- 2016-02-07 Paper_v1.tex
@@
 \usepackage{textcomp}
+\usepackage{multirow}
+\usepackage{amsmath}
+\DeclareMathOperator*{\argmin}{\arg\!\min}
 
+\usepackage{algorithm2e}
 %\usepackage[space]{cite}
@@
 \begin{titlepage}
-{\par\centering\textbf{\Large {Dynamically dimensioned particle swarm optimization (DdPSO): A meta-heuristic for parameter estimation in biochemical models}}}
+{\par\centering\textbf{\Large {Dynamic Optimization with Particle Swarm (DOPS): A meta-heuristic for parameter estimation in biochemical models}}}
 \vspace{0.05in}
@@
 Additionally, biochemical models often have many local minima or maxima which further complicates the parameter estimation problem.
-In this study, we developed dynamically dimensioned particle swarm optimization (DdPSO), a novel meta-heurstic that combined the global search capability of multi-swarm particle swarm optimization with dynamic dimensioned search.
-DdPSO uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors which are then greedily updated using dynamic dimensioned search.
-We tested the performance of DdPSO on commonly used test functions for global optimization (Ackley and Rosenbrock), published biochemical parameter estimation benchmark problems and
+In this study, we developed dynamically dimensioned particle swarm optimization (DdPSO), a novel meta-heurstic that combined the global search capability of multi-swarm particle swarm optimization with dynamically dimensioned search.
+DdPSO uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors which are then greedily updated using dynamically dimensioned search.
+We tested the performance of DdPSO on commonly used test functions for global optimization (Ackley and Rosenbrock), on published biochemical parameter estimation benchmark problems and
 lastly on a model of the human coagulation cascade. For each test we performed 25 trials with 4000 function evaluations per trial, and compared the performance of DdPSO with other commonly
@@
 Mathematical modeling of biochemical networks is often an iterative process.
-First, models are formulated from biochemical knowledge, and then trained using using experimental data \cite{Aldridge:2006aa,banga2008optimization,ashyraliyev2009systems}.
-Model training is typically framed as a non-linear optimization problem wherein the residual (or objective function) between experimental data and model simulations is minimized using an optimization strategy \cite{moles2003parameter}.
+First, models are formulated from biochemical knowledge, and then model parameters are estimated using experimental data \cite{Aldridge:2006aa,banga2008optimization,ashyraliyev2009systems}.
+Parameter estimation is typically framed as a non-linear optimization problem wherein the residual (or objective function) between experimental data and model simulations is minimized using an optimization strategy \cite{moles2003parameter}.
 The optimal parameters obtained from training are then used to validate the model on hitherto unseen experimental data.
@@
 most of them have multiple local minima or maxima \cite{moles2003parameter,banga2008optimization}.
-The non-linearity of the problem coupled with multi-modality generally renders the local optimization techniques like pattern search method \cite{hooke1961direct}, Nelder-Mead simplex method \cite{nelder1965simplex} or Levenberg-Marquardt method\cite{more1978levenberg} incapable of obtaining an optimal solution.
-Though deterministic global optimization techniques can handle non-linearity and multi-modality \cite{esposito2000deterministic,horst2013global}, the absence of derivative information, discontinuity of the objective functions, non-smooth regions or the lack of any knowledge about the objective function severely hamper the use of these techniques.
+The non-linearity of the problem coupled with multi-modality generally renders the local optimization techniques like pattern search method \cite{hooke1961direct}, Nelder-Mead simplex method \cite{nelder1965simplex}, steepest descent method or Levenberg-Marquardt method \cite{more1978levenberg} incapable of reliably obtaining an optimal solution since they generally stop at the first local minimum.
+Though deterministic global optimization techniques (for example algorithms based on branch and bound framework) can handle non-linearity and multi-modality \cite{esposito2000deterministic,horst2013global}, the absence of derivative information, discontinuity of the objective functions, non-smooth regions or the lack of any knowledge about the objective function severely hamper the use of these techniques.
 
-Meta-heuristic approaches like Genetic Algorithms, Simulated Annealing \cite{kirkpatrick1983optimization}, Evolutionary Programming and population based searches like Differential Evolution\cite{storn1997differential} have shown promise in this regard \cite{sun2012parameter}.
+Meta-heuristic stochastic optimization approaches like Genetic Algorithms (GAs), Simulated Annealing (SA) \cite{kirkpatrick1983optimization}, Evolutionary Programming and population based searches like Differential Evolution (DE) \cite{storn1997differential} have shown promise in this regard \cite{sun2012parameter}.
 They do not make any assumptions about the structure of objective function or require any \textit{a priori} information about the objective function.
@@
 Banga and co-workers have applied scatter-search based methods \cite{villaverde2012cooperative,rodriguez2006novel,egea2007scatter} to estimate parameters on non-linear biological processes.
-Hybrid approaches that combine a meta-heuristic with a local optimization search, wherein a near optimal solution that is obtained with a meta-heuristic is further refined using a local search have also become popular.
+Hybrid approaches that combine a meta-heuristic with a local optimization search, wherein a near globally optimal solution that is obtained with a meta-heuristic is further refined using a local search have also become popular.
 Villaverde et al.
@@
 Fan et al.
-recently showed that population based meta-heuristics along with decomposition based methods can be used to model gene circuits from mRNA data \cite{fan2015parameter}.Despite these successes, a  major drawback with most of these approaches is the vast number of objective function evaluations they take in arriving at good search regions.
-As the models grow in size and complexity and the number of dimensions of parameter vector increases,  parameter estimation becomes a computationally expensive task.
+recently showed that population based meta-heuristics along with decomposition based methods can be used to model gene circuits from mRNA data \cite{fan2015parameter}.
+Despite these successes, a  major drawback with most of these metaheuristic approaches is the vast number of objective function evaluations they take in arriving at good search regions.
+As the models grow in size and complexity and the number of dimensions of parameter vector increases, evaluation of the objective function becomes computationally expensive.
+Thus performing a large number of evaluations is not computationally feasible.
+
 In many of these high dimensional problems approaching an exact solution may not be necessary.
@@
 
-\clearpage
+\section*{Problem Formulation}
+
+The problem of parameter estimation in dynamic biological systems consists of finding an optimal parameter vector that results in the best fit between the model predictions and observed experimental data.The cost function or objective that measures the goodness of fit is generally the Euclidean norm which is minimized under certain constraints.
+This can be mathematically expressed as follows.
+
+\begin{equation}\label{eq:static_opt}
+\begin{aligned}
+& \text{minimize}
+& & K(\mathbf{p})=(\sum_{i=1}^N g_i(t_i,\mathbf{x,p,u})-y_i)^2) \\
+& \text{subject to}
+&&\dot{x}=f(t,\mathbf{x(t,\mathbf{p}),p,u(\mathbf{t})})\\
+&&&x(t_0) = x_0\\
+&&&c(t,\mathbf{x,p,u}) \geqslant 0 \\
+&&& \mathbf{p}^L \leqslant \mathbf{p} \leqslant \mathbf{p}^U\\
+\end{aligned}
+\end{equation}
+
+where t is time, x is the state variable with an initial state x\_0 u is the input vector \, p is the parameter vector and f is the system of equations (like differential equations or set of algebraic constraints), c in the set of linear and non-linear constraints, g is the set of observables and p\textsuperscript{L} and p\textsuperscript{U} are the lower and upper bounds.
+Thus the problem eventually is to find {p}\textsuperscript{*} where
+
+\begin{equation}\label{eq:static_opt_2}
+\mathbf{p^{*}}= \arg\min_{p} K(\mathbf{p})
+\end{equation}
 
@@
 \subsection*{Dynamically dimensioned particle swarm optimization (DdPSO).}
-DdPSO is a novel meta-heuristic that combines multi-swarm based particle swarm methods with dynamically dimensioned search (Fig. \ref{fig-algorithm}).
+DdPSO is a novel meta-heuristic that combines multi-swarm based particle swarm methods with dynamically dimensioned search (Fig.\ref{fig-algorithm}).
 The goal of this approach is to obtain optimal or near optimal parameters for high-dimensional complex biological systems within a pre-specified number of function evaluations.
@@
 where $\mathbf{N}$ represents the total number of function evaluations, $\mathbf{w}_{max}$ and $\mathbf{w}_{min}$ are the maximum and minimum inertia weights respectively.
-While updating the particle, we made sure all dimensions of the solution represented by the particle were within bounds using a reflection boundary condition.
+While updating the particle, we made sure all dimensions of the solution represented by the particle were within bounds using a set of reflection boundary conditions.
+
+\begin{algorithm}
+\If{$x_{i,j}^{old} < x_{i}^{min}$}{  	
+  ${x}_{i,j}^{new}={x}_{i,j}^{old}+(x_{i}^{min} - {x}_{i,j}^{old})$
+  \If{${x}_{i,j}^{new}>x_{i}^{max}$}{
+    ${x}_{i,j}^{new}=x_{i}^{max}$
+  }
+}
+\If{$x_{i,j}^{old} > x_{i}^{max}$}{  	
+  ${x}_{i,j}^{new}={x}_{i,j}^{old}+({x}_{i,j}^{old}-x_{i}^{max})$
+  \If{${x}_{i,j}^{new}<x_{i}^{min}$}{
+    ${x}_{i,j}^{new}=x_{i}^{min}$
+  }
+}
+
+\end{algorithm}
+
 After every $\mathbf{g}$  function evaluations, the particles within all sub-swarms were mixed and then randomly redistributed to a new sub-swarm.
@@
 Coagulation is an ideal system to test our approach since the biology is well known and complex, with multiple feed back loops that are tightly regulated.
-We used experimental data with under different conditions to obtain optimal parameters and used these parameters to make predictions against unseen experimental data.
-We obtained good fits and made accurate enough predictions using parameters obtained from 4000 function evaluations.
+We used experimental data under different conditions to obtain optimal parameters and used these parameters to make predictions against unseen experimental data.
+We obtained good fits and made sufficiently accurate enough predictions using parameters obtained from 4000 function evaluations.
 Further, we also used high-dimensional forms of commonly used test functions of global optimization and showed that we were able to find the global minimum for 300 dimensional Ackley and Rastrigin functions faster than other meta-heuristics.
 We also considered two recently published benchmark problems to test parameter optimization approaches and showed that we were able to retrieve the nominal parameter vector within 4000 function evaluations.
-
-meta-heuristic approaches are generally effective in finding close to optimum solutions of complex, multi-modal functions.
+Meta-heuristic approaches are generally effective in finding close to optimum solutions of complex, multi-modal functions.
 In addition, they generally obviate the need for any \textit{a priori} knowledge (like function derivative).
@@
 \clearpage
+
+\begin{table}[h]
+\centering
+\caption{Error Analysis.}
+\label{table:Error-Analysis}
+\begin{center}
+\begin{tabular}{ |c|c|c|c| } 
+\hline
+Simulation&Concentration& Normalized Standard Error & R-Squared Value \\
+\hline
+\multirow{3}{4em}{FVIIa-TF level} & 5nM & 0.0367 & 0.7680\\ 
+& 500pM & 0.0636 & 0.8939\\
+& 50pM & 0.1204 & 0.8384\\ 
+& 10pM & 0.0813 & 0.9492\\
+& 5pM & 0.0574 & 0.9794\\ 
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 

### 2016-02-10 Paper_v5.tex
Started on Paper_v5 final version

--- 2016-02-07 Paper_v3.tex
@@
 \begin{titlepage}
-{\par\centering\textbf{\Large {Dynamic Optimization with Particle Swarm (DOPS): A meta-heuristic for parameter estimation in biochemical models}}}
+{\par\centering\textbf{\Large {Dynamic Optimization with Particle Swarms (DOPS): A meta-heuristic for parameter estimation in biochemical models}}}
 \vspace{0.05in}
@@
 \section*{Abstract}
-Mathematical modeling is as a powerful tool to analyze, and ultimately design biochemical networks.
-However, estimation of biochemical model parameters can be a significant challenge.
-Parameter estimation in biochemical models generally involves expensive function evaluations, making it difficult to quickly obtain optimal solutions.
-Additionally, biochemical models often have many local minima or maxima which further complicates the parameter estimation problem.
-In this study, we developed dynamically dimensioned particle swarm optimization (DdPSO), a novel meta-heurstic that combined the global search capability of multi-swarm particle swarm optimization with dynamically dimensioned search.
-DdPSO uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors which are then greedily updated using dynamically dimensioned search.
-We tested the performance of DdPSO on commonly used test functions for global optimization (Ackley and Rosenbrock), on published biochemical parameter estimation benchmark problems and
-lastly on a model of the human coagulation cascade. For each test we performed 25 trials with 4000 function evaluations per trial, and compared the performance of DdPSO with other commonly
-used meta-heuristic approaches such as differential evolution (DE), particle swarm optimization (PSO), simulated annealing (SA) and dynamically dimensioned search (DDS).
-For the wide range of problems that we considered, DdPSO outperformed the other meta-heuristic approaches despite a limited number of function evaluations.
-We further tested the predictive power of the coagulation model parameters against data not used in training, and found good agreement between model simulations and experimental measurements.
-Taken together, DdPSO is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.
+Mathematical modeling is a powerful tool to analyze, and ultimately design biochemical networks.
+However, the estimation of biochemical model parameters is a significant challenge.
+Parameter estimation in biochemical models typically involves expensive function evaluations and noisy data, making it difficult to quickly obtain optimal solutions.
+Biochemical models often also have many local extrema which further complicates parameter estimation.
+Toward these challenges, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel global meta-heuristic that combined features of multi-swarm particle swarm optimization with dynamically dimensioned search (DDS).
+DOPS uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors, the best of which is greedily updated using dynamically dimensioned search.
+We tested the performance of DOPS on a model of human coagulation cascade.
+We performed N = 25 trials with M = 4000 function evaluations per trial, and compared the performance of DOPS with other commonly
+used meta-heuristics such as differential evolution (DE), simulated annealing (SA) and dynamically dimensioned search (DDS).
+We further tested the predictive power of the coagulation model parameters against data not used in training, and found good agreement between simulations and experimental measurements.
+Lastly, we tested the performance of DOPS on commonly used test functions for global optimization and on published biochemical parameter estimation benchmark problems.
+For the wide range of problems that we considered, DOPS outperformed other meta-heuristic approaches despite a limited number of function evaluations.
+Taken together, DOPS is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.
 
+\vspace{0.1in}
 {\noindent \textbf{Keywords:}~Parameter identification, Meta-heuristic optimization, Biochemical modeling}
@@
 Parameter estimation is typically framed as a non-linear optimization problem wherein the residual (or objective function) between experimental data and model simulations is minimized using an optimization strategy \cite{moles2003parameter}.
-The optimal parameters obtained from training are then used to validate the model on hitherto unseen experimental data.
+Optimal parameters obtained from model training are then used to validate the model on unseen experimental data.
 If validation fails, model construction and calibration are repeated iteratively until satisfactory results are obtained.
 
-Parameter estimation is a major challenge in the development of large biochemical models.
+Parameter estimation is a major challenge in the iterative development of biochemical models.
 Although parameter estimation has been a well studied problem in engineering for decades \cite{nieman1971review,beck1977parameter,young1981parameter,beck1998inverse},
 the complex dynamics of large biological systems and noisy, often incomplete experimental data pose a unique estimation challenge.
-Most of the optimization problems pertaining to biological systems are non-linear and multi-modal i.e.
-most of them have multiple local minima or maxima \cite{moles2003parameter,banga2008optimization}.
-The non-linearity of the problem coupled with multi-modality generally renders the local optimization techniques like pattern search method \cite{hooke1961direct}, Nelder-Mead simplex method \cite{nelder1965simplex}, steepest descent method or Levenberg-Marquardt method \cite{more1978levenberg} incapable of reliably obtaining an optimal solution since they generally stop at the first local minimum.
-Though deterministic global optimization techniques (for example algorithms based on branch and bound framework) can handle non-linearity and multi-modality \cite{esposito2000deterministic,horst2013global}, the absence of derivative information, discontinuity of the objective functions, non-smooth regions or the lack of any knowledge about the objective function severely hamper the use of these techniques.
+Often optimization problems involving biological systems are non-linear and multi-modal i.e. typical models often have multiple local minima or maxima \cite{moles2003parameter,banga2008optimization}.
+Non-linearity coupled with multi-modality generally renders local optimization techniques such as pattern search \cite{hooke1961direct}, Nelder-Mead simplex methods \cite{nelder1965simplex},
+steepest descent methods or Levenberg-Marquardt \cite{more1978levenberg} incapable of reliably obtaining optimal solutions as they generally terminate at local minimum.
+Though deterministic global optimization techniques (for example algorithms based on branch and bound) can handle non-linearity and multi-modality \cite{esposito2000deterministic,horst2013global},
+the absence of derivative information, discontinuity of the objective functions, non-smooth regions or the lack of knowledge about the objective function hampers these techniques.
 
-Meta-heuristic stochastic optimization approaches like Genetic Algorithms (GAs), Simulated Annealing (SA) \cite{kirkpatrick1983optimization}, Evolutionary Programming and population based searches like Differential Evolution (DE) \cite{storn1997differential} have shown promise in this regard \cite{sun2012parameter}.
-They do not make any assumptions about the structure of objective function or require any \textit{a priori} information about the objective function.
+Meta-heuristic stochastic optimization approaches like Genetic Algorithms (GAs) [ADI-REF], Simulated Annealing (SA) \cite{kirkpatrick1983optimization},
+Evolutionary Programming [ADI-REF] and population based searches like Differential Evolution (DE) \cite{storn1997differential} have shown promise on nonlinear multi-modal problems \cite{sun2012parameter}.
+These techniques do not make any assumptions about the structure of objective function nor do they require \textit{a~priori} information about the objective function.
 Though they do not guarantee strong convergence, these approaches are effective in finding near optimal solutions.
-Mendes et al.
-\cite{mendes1998non} used Simulated Annealing to estimate rate constants for the irreversible inhibition of HIV proteinase, Modchang et al.
-\cite {modchang2008mathematical} used Genetic Algorithms to estimate parameters for a mathematical model of signal transduction, Differential Evolution based approaches have been an effective approach on various systems \cite{tsai2005evolutionary,wang2001hybrid,noman2007inferring}.
-Tashkova et al.
-\cite{tashkova2011parameter} compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis and showed that DE was the most effective.
-Banga and co-workers have applied scatter-search based methods \cite{villaverde2012cooperative,rodriguez2006novel,egea2007scatter} to estimate parameters on non-linear biological processes.
-Hybrid approaches that combine a meta-heuristic with a local optimization search, wherein a near globally optimal solution that is obtained with a meta-heuristic is further refined using a local search have also become popular.
-Villaverde et al.
-\cite{villaverde2015biopredyn} combined scatter search with local search methods for parameter estimation in large scale systems biology models.
-Fan et al.
-recently showed that population based meta-heuristics along with decomposition based methods can be used to model gene circuits from mRNA data \cite{fan2015parameter}.
-Despite these successes, a  major drawback with most of these metaheuristic approaches is the vast number of objective function evaluations they take in arriving at good search regions.
-As the models grow in size and complexity and the number of dimensions of parameter vector increases, evaluation of the objective function becomes computationally expensive.
-Thus performing a large number of evaluations is not computationally feasible.
+Mendes et al. \cite{mendes1998non} used simulated annealing to estimate rate constants for the irreversible inhibition of HIV proteinase,
+Modchang et al. \cite{modchang2008mathematical} used a genetic algorithm to estimate parameters for a model of signal transduction [ADI-EXPAND],
+differential evolution approaches have also been effective on various biological problems \cite{tsai2005evolutionary,wang2001hybrid,noman2007inferring}.
+Tashkova et al. compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis and showed that DE was the most effective \cite{tashkova2011parameter}.
+Banga and co-workers have also successfully applied scatter-search methods to estimate parameters on non-linear biological processes \cite{villaverde2012cooperative,rodriguez2006novel,egea2007scatter}.
+Hybrid approaches that combine a meta-heuristic with a local optimization search, wherein a near globally optimal solution obtained using a meta-heuristic is refined using a local search have also become popular.
+Villaverde et al. combined scatter search with local search methods for parameter estimation in a collection of systems biology models \cite{villaverde2015biopredyn}.
+Fan et al. recently showed that population based meta-heuristics along with decomposition based methods can be also used to model gene circuits from mRNA data \cite{fan2015parameter}.
+Despite these successes, a major drawback of most metaheuristic approaches is the large number of function evaluations required to explore the parameter space.
+Typically as models grow in size and complexity, evaluation of the objective function becomes computationally expensive.
+Thus performing a large number of function evaluations is not desirable (and perhaps not feasible).
 
-In many of these high dimensional problems approaching an exact solution may not be necessary.
+%In many of these high dimensional problems approaching an exact solution may not be necessary.
 Gutenkust et al.
@@
 
-In this study, we developed dynamically dimensioned particle swarm optimization (DdPSO), a novel meta-heurstic that combined the global search capability of multi-swarm particle swarm optimization with dynamic dimensioned search.
-The objective of DdPSO was to obtain near optimal parameter estimates for large biochemical models within a relatively few function evaluations.
-DdPSO uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors which are then greedily updated using dynamic dimensioned search.
-We tested the performance of DdPSO on commonly used test functions for global optimization (Ackley and Rosenbrock), published biochemical parameter estimation benchmark problems and
-lastly on a model of the human coagulation cascade.
-DdPSO outperformed common meta-heuristic approaches like Differential Evolution (DE), Particle Swarm Optimization (PSO), Simulated Annealing (SA) and dynamically dimensioned search (DDS) on the benchmark and coagulation problems.
+%Because the neighborhood in DDS is the entire parameter space, DDS is a global optimization technique although it is greedy.
+
+In this study, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel meta-heuristic that combines the global search capability of multi-swarm particle swarm optimization and dynamically dimensioned search (DDS).
+The objective of DOPS is to obtain near optimal parameter estimates for large biochemical models within a relatively few function evaluations.
+DOPS uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors which are then greedily updated using dynamically dimensioned search.
+We first considered a model of human coagulation cascade to test the performance of DOPS.
+Coagulation is a large, complex biochemical network involving strong positive feedback.
+We then tested the performance of DOPS on commonly used test functions for global optimization (Ackley and Rosenbrock) and published biochemical parameter estimation benchmark problems \cite{villaverde2015biopredyn}.
+DOPS outperformed common meta-heuristic approaches like Differential Evolution (DE), Simulated Annealing (SA) and dynamically dimensioned search (DDS) on both the test functions and the coagulation model.
+DOPS also performed well on the benchmark problems where it outperformed enhanced scatter search (eSS) and recovered the nominal parameters with only 4000 function evaluations ([ADI-compared with what?])
+across all the benchmark problems considered.
+Taken together, these studies suggest DOPS is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.
 
@@
 & \text{minimize}
-& & K(\mathbf{p})=(\sum_{i=1}^N g_i(t_i,\mathbf{x,p,u})-y_i)^2) \\
+& & K(\mathbf{p})=(\sum_{i=1}^N (g_i(t_i,\mathbf{x,p,u})-y_i)^2) \\
 & \text{subject to}
@@
 
-where t is time, x is the state variable with an initial state x\_0 u is the input vector \, p is the parameter vector and f is the system of equations (like differential equations or set of algebraic constraints), c in the set of linear and non-linear constraints, g is the set of observables and p\textsuperscript{L} and p\textsuperscript{U} are the lower and upper bounds.
+where t is time, x is the state variable with an initial state x\textsubscript{0}, u is the input vector \, p is the parameter vector and f is the system of equations (like differential equations or set of algebraic constraints), c in the set of linear and non-linear constraints, g is the set of observables and p\textsuperscript{L} and p\textsuperscript{U} are the lower and upper bounds.
 Thus the problem eventually is to find {p}\textsuperscript{*} where
@@
 \section*{Results}
-\subsection*{Dynamically dimensioned particle swarm optimization (DdPSO).}
-DdPSO is a novel meta-heuristic that combines multi-swarm based particle swarm methods with dynamically dimensioned search (Fig.\ref{fig-algorithm}).
+\subsection*{Dynamic Optimization with Particle Swarms (DOPS).}
+DOPS is a novel meta-heuristic that combines multi-swarm based particle swarm methods with dynamically dimensioned search (Fig. \ref{fig-algorithm}).
 The goal of this approach is to obtain optimal or near optimal parameters for high-dimensional complex biological systems within a pre-specified number of function evaluations.
@@
 \end{eqnarray}
-where $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ are adjustable parameters, $\mathcal{L}_{i}$ denotes the best local solution found by particle $i$ till function evaluation $j$-$1$, and
-$\mathcal{GL}_{k}$ denotes the best local solution found over the population of all particles within the swarm ${S}_{k}$.
+where $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ are adjustable parameters, $\mathcal{L}_{i}$ denotes the best solution found by particle $i$ within the sub-swarm till function evaluation $j$-$1$, and
+$\mathcal{GL}_{k}$ denotes the best solution found over the population of all particles within the sub-swarm ${S}_{k}$.
 The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model
-parameters ($\mathcal{K}\times{1}$). In our algorithm the parameter $\theta_{1,j-1}$ depends on the function evaluations and is controlled according to the following equation
+parameters ($\mathcal{K}\times{1}$).
+Equation (3) is suggested by the general form of particle swarm equations.
+However this equation does not contain the velocity terms that are generally associated with particle swarm optimization.
+In our algorithm the parameter $\theta_{1,j-1}$ depends on the function evaluations and is controlled according to the following equation
 \begin{eqnarray}
@@
 where $\mathbf{N}$ represents the total number of function evaluations, $\mathbf{w}_{max}$ and $\mathbf{w}_{min}$ are the maximum and minimum inertia weights respectively.
-While updating the particle, we made sure all dimensions of the solution represented by the particle were within bounds using a set of reflection boundary conditions.
+While updating the particle, we made sure all dimensions of the solution represented by the particle were within bounds using a set of reflection boundary conditions shown below.
+
+\begin{eqnarray}
+\end{eqnarray}
 
@@
 }
-
 \end{algorithm}
@@
 This process continued till $\mathcal{FR}*\mathbf{N}$ number of functions evaluations, where $\mathcal{FR}$ represents the fraction of evaluations with the multi-swarms.
+The fraction of evaluations within the swam phase is based on the drop in error.
+If the error does not drop by 1\% for a certain number of evaluations the search within swarm phase is terminated.
 At the end of these function evaluations, we froze all the solutions represented by various particles and chose the particle with best solution among $\mathcal{GL}_{1} \cdots \mathcal{GL}_{NS}$ as the initial candidate vector $\mathcal{G}$ for the remaining $({1}-\mathcal{FR})*\mathbf{N}$ number of function evaluations.
@@
 This particle was then updated according to the following rule
-\begin{eqnarray}
-	\mathcal{G}_{new}(\mathbf{J})& = &\mathcal{G}(\mathbf{J})+\mathbf{r}_{normal}(\mathbf{J})\sigma(\mathbf{J})
-\end{eqnarray}
+
+\begin{equation}
+  \mathcal{G}_{new}(\mathbf{J})=\begin{cases}
+    \mathcal{G}(\mathbf{J})+\mathbf{r}_{normal}(\mathbf{J})\sigma(\mathbf{J}), & \text{if $\mathcal{G}_{new}(\mathbf{J})<\mathcal{G}(\mathbf{J})$}.\\
+    \mathcal{G}(\mathbf{J}), & \text{otherwise}.
+  \end{cases}
+\end{equation}
+
 where $\mathbf{J}$ represents the set containing the specific dimensions being perturbed, ${r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$.
@@
 \begin{eqnarray}
-	\mathcal{P}_{j}&=&{1}-\log(j/ (\beta({1}-\mathcal{FR})*\mathbf{N}))
+	%\mathcal{P}_{j}&=&{1}-\log(j/ (\beta({1}-\mathcal{FR})*\mathbf{N}))
+	\mathcal{P}_{j}&=&{1}-\log(j/ (({1}-\mathcal{FR})*\mathbf{N}))
 \end{eqnarray}
-where $\beta$ is the perturbation frequency probability modulator.
+%where $\beta$ is the perturbation frequency probability modulator.
 Thus the number of dimensions of the candidate vector that are updated or perturbed decreases with the as the number of function evaluations increase.
 These updates are greedy in nature, so $\mathcal{G}_{new}$ becomes the new solution vector only if it is better than the old one $\mathcal{G}$.
+The decrease of weight function in equation 4, reflection boundary conditions in equation 5, the update function described in equations 6 and 7, the selection probability in equation 8 are the means by which DDS ideas are incorporated into DOPS.
+%Equation 5 describing the reflection boundary conditions, equations 6,7 and 8 are the means by which DDS ideas are incorporated into DOPS.
+The fraction of evaluations $\mathcal{FR}$  within the swarm phase is based on a switching strategy wherein the switch from swarm phase to DDS phase happens when the error due to the best solution does not drop more than 1\% of the original error, continuously for more than a prescribed number of function evaluations.
+This allows the solution to quickly jump out of local optima and avoid any convergence issues that are generally associated with swarm based searches.
 
-\subsection*{Performance of DdPSO on benchmark problems.}
-We compared the performance of DdPSO with algorithms on commonly used test functions for global optimization, as well as benchmark biochemical problems from literature (Table \ref{table:benchmark-problems}).
-We used a 300 dimensional Rastrigin function and 300 dimensional Ackley function.
-Both Ackley and Rastrigin have multiple local minima and maxima and attain a global minimum value of 0.
-We tested the performance of the 5 different heuristic approaches on these two functions.
-In each experiment we again fixed the number of function evaluations at 4000 and ran 25 experiments.
-In both cases (Figure S6 A and B) we see that the error convergence rate for DDSMLSPSO is much faster as compared to the other four s and it finds the global minimum of 0 in both cases.
-
-Villaverde and co-workers recently published a set of benchmark problems to evaluate parameter estimation methods \cite{villaverde2015biopredyn}.
-From a computational cost perspective problems they categorized the problems as most expensive, intermediate and least expensive.
-We evaluated the performance of our algorithms on a problem from the most expensive and least expensive categories.
-The first problem is a genome wide kinetic model of \textit{Saccharomyces cerevisiae} with 1759 parameters (henceforth referred to as problem B1).
-The second problem (henceforth referred to as problem B4) is a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters.
-In both cases pseudo time series data was generated.
-For problem B1, the time series data consisted of 44 observables and for problem B4 the data corresponded to 13 different metabolites.
-We fixed the number of function evaluations at 4000 for DDSMLSPSO and trained both the models against the pseudo experimental data.
-We recaptured the 'nominal' parameters (Figure S7 A and B)  in both cases within 4000 evaluations.
-The final objective function value (Table 1) is an order smaller for B1 possibly due to overfitting.
-In Table 1 we summarize the results obtained on the coagulation model and the benchmark problems.
-
-\subsection*{Performance of DdPSO on a model of the human coagulation cascade.}
-We compared the performance of DdPSO on a model of blood coagulation against commonly used meta-heuristics like simulated annealing (SA), differential evolution (DE), particle swarm optimization (PSO),and dynamically dimensioned search (DDS).
+\subsection*{Performance of DOPS on a model of the human coagulation cascade.}
+We compared the performance of DOPS on a model of blood coagulation against commonly used meta-heuristics like simulated annealing (SA), differential evolution (DE), and dynamically dimensioned search (DDS).
 Coagulation is an archetype biochemical network that is highly interconnected and tightly regulated with multiple positive and negative feedback loops (Fig.
@@
 Coagulation is regulated by a set of serine proteases also known as coagulation factors and blood platelets.
+
+%Added a little more description here
 The coagulation factors are generally in an inactive state and are known as zymogens.
@@
 The exposure of these factors to blood kick-starts a series of convergent cascades that lead to conversion of zymogen prothrombinase to thrombin.
+For example when coagulation is initiated through the tissue factor pathway, tissue factor and activated factor VIIa (FVIIa) form a complex that activates factors X (fX) and IX (fIX).
+Activated factor X (fXa) thereafter activates downstream factors VIII and IX.
+The initial activation leads to the production of picomolar amounts of thrombin (fIIa) which activates platelets and amplifies its own production through the formation of a prothrombinase complex (FXa-FVa) on the surface of the activated platelets.
+Thrombin also  downregulates its own production by forming a complex with thrombomodulin which then activates protein C (PC).
+PC inhibits the formation of prothrombinase complex.
+In addition, Tissue factor pathway inhibitor (TFPI) downregulates FXa formation and activity by sequestering free FXa and TF-FVIIa in a FXa-dependent manner.
+Antithrombin III (ATIII)  inhibits all proteases.
+surface protein thrombomodulin (TM).
 Luan et al.
@@
 The objective function was a weighted linear combination of two different error functions that used data sets representing coagulation initiated with different concentrations of TF-VIIa (5pM, 5nM) \cite{hockin2002model}.
+This choice of using a linear combination of two different error functions was motivated by poor validation results while using a single error function.
 We restricted the number of function evaluations to 4000 for each algorithm and performed 25 trials of this experiment.
-DdPSO exhibits a much faster rate of error convergence and has a much lower final error as compared to the other algorithms (Figure 3).
-Within the first 1000 function evaluations (swarm phase) of DDSMLSPSO there is a very rapid drop in error.
-Subsequently, again after 2000 function evaluations (dynamically dimensioned phase) the error drops quickly.
-Overall at the end of 4000 function evaluations DDSMLSPSO minimizes the error (final objective error is 90456.9) to a much greater extent than any of the other algorithms.
-Amongst the rest of the algorithms DDS has a comparable error drop for the first 2000 function evaluations.
-However the drop in error remains constant for DDS after 2000 function evaluations.
-Using the parameters obtained at the end of 4000 function evaluations we examined the 'fits' between models predictions and experimental data (Figure 2).
+DOPS exhibits a much faster rate of error convergence and has a much lower final error as compared to the other algorithms (Fig.\ref{fig-convergence}).
+
+ Within the first 1000 function evaluations  of DOPS there is a very rapid drop in error.
+Approximately between 500-1000 function evaluations a switch to dynamically dimensioned search phase happens (this switch varies from trial to trial since the switch is based on the error from the swarm phase).
+Overall at the end of 4000 function evaluations DOPS minimizes the error (final objective error is 5.5916e+04) to a much greater extent than any of the other algorithms.
+Using the parameters obtained at the end of 4000 function evaluations we examined the 'fits' between models predictions and experimental data (Fig.\ref{fig-train}).
 The solid lines represent the mean value of prediction over 25 trials and the shaded region represents the 99\% confidence interval.
-Subsequently we compared the model predictions against completely 'unseen' or untrained experimental where coagulation was initiated with 500pM,50pM,10pM concentrations of TF-VIIa respectively (Figure 3).
+Subsequently we used these optimal parameters to make model predictions that was compared against completely 'unseen' or untrained experimental data where coagulation was initiated with 500pM,50pM,10pM concentrations of TF-VIIa respectively (Fig.\ref{fig-validation}).
+
+\subsection*{Performance of DOPS on benchmark problems and test functions.}
+Villaverde and co-workers recently published a set of benchmark biochemical problems to evaluate parameter estimation methods \cite{villaverde2015biopredyn}.
+From a computational cost perspective problems they categorized the problems as most expensive, intermediate and least expensive.
+We evaluated the performance of our algorithms on a problem from the most expensive and least expensive categories.
+The first problem (most expensive) is a genome wide kinetic model of \textit{Saccharomyces cerevisiae} with 261 reactions, 262 variables and 1759 parameters (henceforth referred to as problem B1).
+The reactions were modeled using modular rate law and generalized form of Michaelis-Menten kinetics.
+The second problem (henceforth referred to as problem B4, least expensive) is a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters.
+In both cases pseudo time series data was generated by the authors.
+For problem B1, the time series data consisted of 44 observables and for problem B4 the data corresponded to 13 different metabolites.
+For a detailed description about the model architectures please refer to the work of Villaverde et al.
+\cite{villaverde2015biopredyn}
+
+We fixed the number of function evaluations at 4000 for DOPS and trained both the models against the pseudo experimental data.
+In both cases we found good fits  (Fig.
+\ref{fig-sims-b1} and Fig.
+\ref{fig-sims-b4}) to the problems within 4000 evaluations.
+We recaptured the 'nominal' parameters' in both cases within 4000 evaluations (Fig.
+\ref{fig-benchmark}).
+The final objective function value (Table \ref{table:benchmark-problems}) is an order smaller than nominal error for B1 possibly due to overfitting against the noise that was added to the synthetic data.
+
+Having obtained good fits on the coagulation problem and the benchmarks we proceeded to compare the performance of DOPS with other algorithms on commonly used test functions for global optimization.
+We used a 300 dimensional Rastrigin function and 300 dimensional Ackley function.
+Both Ackley and Rastrigin have multiple local minima and maxima and attain a global minimum value of 0.
+We tested the performance of the 4 different heuristic approaches on these two functions.
+In each experiment we again fixed the number of function evaluations at 4000 and ran 25 experiments.
+In both cases (Fig.
+\ref{fig-testfunctions}) we see that the error convergence rate for DOPS is much faster as compared to the other three heuristics and it finds the global minimum of 0 in both cases.
+In (Table \ref{table:benchmark-problems}) we summarize the results obtained on the coagulation model, the benchmark problems and the test functions.
 
@@
 Tolson and Shoemaker, through DDS, showed that randomly perturbing a subset of dimensions in high dimensional parameter space is an effective way to obtain near optimal solutions with few function evaluations.
-However their approach was based on a single solution.
-Single solutions based methods may prove ineffective in complex systems when they start from a 'bad' region of search or from a local minimum.
+Though their approach is based on a single solution, the decision vector carries forward from one iteration to the next.
+Hence it behaves similar to a population based search.
 In our approach, we tried to eliminate the probability of starting from a bad region of search using a variant of particle swarm optimization.
-Particle Swarm Optimization (PSO) is a population based meta heuristic that is known for rapid convergence without any complex operations like recombination, mutation or selection that are associated with other population based meta-heuristics like Differential Evolution (DE) or Genetic Algorithm (GA).
+Particle Swarm Optimization (PSO) is a population based meta heuristic which does not have any complex operations like recombination, mutation or selection that are associated with other population based meta-heuristics like Differential Evolution (DE) or Genetic Algorithm (GA).
 Several particle swarm variants have been proposed to improve the search ability and rate of convergence, that involve different neighborhood structures, multi-swarms or adaptive parameters.
@@
 Given a fixed number of function evaluations, we have been able to show that we were able to obtain better solutions for the coagulation model and the test functions faster that other commonly known meta-heuristics and also DDS alone.
+The computational cost was prohibitively high for parameter estimation on this problem using a standard PSO.
+Thus we used SA, DE and DDS for the purpose of comparison.
  Choosing the number of function evaluations is largely a function of cost and complexity of the objective function.
@@
 We used the same value of 4000 for benchmarks published by Villaverde and co-workers \cite{villaverde2015biopredyn}.
-Quite surprisingly we took a couple of orders lesser number of function evaluations (1e3) to obtain the optimal parameter vector as compared to the enhanced Scatter Search (eSS) with a local optimizer, which took around the order of 1e5 number of evaluations.
+In this work by Villaverde et al.
+they used a population based search enhanced Scatter Search (eSS) to estimate the biochemical model parameters.
+Quite surprisingly we took a couple of orders lesser number of function evaluations 10\textsuperscript{3} to obtain the optimal parameter vector as compared to the enhanced Scatter Search (eSS) with a local optimizer, which took around the order of 10\textsuperscript{5} number of evaluations.
 The amount of CPU time taken (on an Intel Xeon processor 2.4 GHz) is lesser as compared to eSS on a similar architecture.
@@
 For SA, we optimized the cooling schedule for the coagulation model.
-Thus, in this approach any overhead that usually comes with additional function evaluations to optimize parameters was avoided.
-The only parameter that we adjusted was the fraction of evaluations within swarm search.
-For the coagulation model it was set to be 0.4 and for the benchmarks it is at 0.6.
-This parameter was set based on empirical observations rather than any definitive analysis.
-When the error convergence rates stagnated in the swarm phase we switched to DDS phase.
-A further improvement to our approach could  be to incorporate a meta intelligent way of switching in and switching out of both the phases.
+Thus, in this approach any overhead that usually comes with additional function evaluations to pre optimize parameters was avoided.
 
-The performance of our approach seems impressive given that it performed well on different complex systems with no specific conditioning being required.
-We comfortably outperformed common meta-heuristics and were also able to find minima of high dimensional global optimization test functions.
+The performance of our approach seems impressive given that it performed well on different complex systems with no pre optimization of algorithm parameters being required.
+We comfortably outperformed existing, widely used meta-heuristics and were also able to find minima of high dimensional global optimization test functions.
 Thus this approach may be well suited to large scale global optimization.
-In addition, surprisingly, we were able to obtain optimal parameter vectors for two different large scale systems biology models with a couple of orders fewer number of function evaluations as compared a hybrid approach.
+In addition, surprisingly, we were able to obtain optimal parameter vectors for two different large scale systems biology models with a couple of orders fewer number of function evaluations as compared to enhanced Scatter Search (eSS).
 However it is quite possible that highly optimized versions of common meta heuristics may outperform us on these systems.
@@
 Our approach can also be combined with local derivative based searches to improve upon the accuracy of the solutions.
+In addition, the current implementation of the algorithm is designed to switch only once from the swarm phase to the DDS.
+In the DDS phase, the search uses only one candidate vector although there is a provision to start with a population of candidate vectors.
+Incorporating a more intelligent switching strategy that can do switch from swarm phase to DDS phase multiple times and having a population of candidate vectors are some aspects of the algorithm that can be studied further.
 
-%\begin{figure}[H]
-%\centering
-%\includegraphics[width=1.00\textwidth,height=0.7\textheight]{./figs/Figure_2_CoagulationNetwork.pdf}
-%\caption{Schematic of the extrinsic and intrinsic coagulation cascade.
-Inactive zymogens upstream (grey) are activated by exposure to tissue factor (TF) \cite{giesen1999blood} following vessel injury.
-Tissue factor and activated factor VIIa (FVIIa) form a complex that activates factor X (fX) and IX (fIX).
-FXa activates downstream factors including factor VIII (fVIII) and fIX.
-Factor V (fV) is primarily activated by thrombin (FIIa).
-In addition, we included a secondary fV activation route involving FXa.
-FXa and FVa form a complex (prothrombinase) on activated platelets that converts prothrombin (fII) to FIIa.
-FIXa and FVIIIa can also form a complex (tenase) on activated platelets which catalyzes FXa formation.
-Localized platelets are activated by external signals such as adenosine diphosphate (ADP) and thromboxane A2 (TXA2) or thrombin through protease-activated receptors (PARs) \protect\cite{coughlin2000thrombin} \protect \cite{coughlin1999protease} Thrombin also activates upstream coagulation factors, forming a strong positive feedback ensuring rapid activation.
-Tissue factor pathway inhibitor (TFPI) downregulates FXa formation and activity by sequestering free FXa and TF-FVIIa in a FXa-dependent manner.
-Antithrombin III (ATIII)  inhibits all proteases.
-Thrombin inhibits itself binding the surface protein thrombomodulin (TM) \protect\cite{esmon1989roles} The IIa-TM complex catalyzes the conversion of protein C (PC) to activated protein C (APC), which attenuates the coagulation response by the proteolytic cleavage of fV/FVa and fVIII/FVIIIa \protect\cite{esmon1989roles}}\label{fig-network}
-%\end{figure}
+%Need to talk more about biochemical benefits and importance for of biochemical problems
 
@@
 
-%\bibliography{References_v1}
-%\begin{thebibliography}{9999}
 \bibliography{References_v2}
@@
 \centering
-\caption{Error Analysis.}
-\label{table:Error-Analysis}
-\begin{center}
-\begin{tabular}{ |c|c|c|c| } 
-\hline
-Simulation&Concentration& Normalized Standard Error & R-Squared Value \\
-\hline
-\multirow{3}{4em}{FVIIa-TF level} & 5nM & 0.0367 & 0.7680\\ 
-& 500pM & 0.0636 & 0.8939\\
-& 50pM & 0.1204 & 0.8384\\ 
-& 10pM & 0.0813 & 0.9492\\
-& 5pM & 0.0574 & 0.9794\\ 
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[h]
-\centering
-\caption{Table with optimization settings and results for coagulation and different benchmarks using DdPSO.}
+\caption{Table with optimization settings and results for the coagulation problem, the benchmarks and test functions using DOPS.
+For each problem the bounds on the parameter vector, the total number of function evaluations, the best initial objective value and the best final objective value are specified.
+Nominal objective value represents the objective value using the true parameter vector or the nominal parameter vector.
+The CPU time is the time taken for the problem on a 2.4GHz Intel Xeon Architecture running Matlab 2014b.}
 \label{table:benchmark-problems}
@@
 \hline
-                        & Coagulation           & \multicolumn{4}{l}{Benchmarks}\\
- \hline
-                        & Experimental Data Set & B1           & B4          & Ackley 300D & Rastrigin 300D \\
+                        & Coagulation & B1           & B4          & Ackley 300D & Rastrigin 300D \\
  \hline
@@
 Initial Objective Value & 1.11837e+07           & 1.4275e+07   & 1.8536e+07  & 21.12       & 99.985         \\
-Final Objective Value   & 90456.9               & 4.96e+05     & 38.9375     & 0           & 0              \\
+Final Objective Value   & 5.5916e+04              & 3.5348e+04     & 38.1306     & 0           & 0              \\
 Nominal Objective Value & 4.7785e+06            & 1.0986e+06   & 39.0676     & 0           & 0      \\
@@
 \includegraphics[width=1.0\textwidth,height=0.5\textheight]{./figs/Figure_1_Algorithm.pdf}
-\caption{Multi Swarm Particle Swarm Optimization with Dynamically Dimensioned Search.
-\textbf{A}: Each particle represents an N dimensional parameter vector.
-Particles are randomly initialized and grouped into different sub-swarms.
+\caption{Dynamic Optimization with Particle Swarms.\textbf{A}: Each particle represents an N dimensional parameter vector.
+Particles are given randomly generated initial solutions and grouped into different sub-swarms.
 Within each swarm the magnitude and direction of the movement a particle is influenced by the position of the best particle and also by its own experience.
 After every $\mathbf{g}$ number of function evaluations the particles are mixed and randomly assigned to different swarms.
-At the end of the evaluations ($\mathcal{FR}$) assigned to swarm search, the global best particle (orange color) amongst all sub-swarms is chosen as the candidate parameter vector for Dynamically Dimensioned Search \textbf{B}: The candidate vector does a greedy search in a dynamic neighborhood.
+When the error due to the global best particle (best particle amongst all the sub-swarms - orange color) does not drop over a certain number of function evaluations, the swarm search is stopped and the search switches to a Dynamically Dimensioned Search with global best particle as the initial solution vector or candidate vector.
+\textbf{B}: The candidate vector performs a greedy global search for the remaining number of function evaluations.
 The search neighborhood is dynamically adjusted by varying the number of dimensions that are perturbed (in black) in each evaluation step.
@@
 \includegraphics[width=1.00\textwidth,height=0.7\textheight]{./figs/Figure_2_CoagulationNetwork.pdf}
-\caption{Schematic of the extrinsic and intrinsic coagulation cascade.
+\caption{Schematic of the extrinsic and intrinsic coagulation cascade\cite{luan2007computationally}.
 Inactive zymogens upstream (grey) are activated by exposure to tissue factor (TF)  following vessel injury.
@@
 The objective error is the mean over $N$= 25 trials.
-DDSMLSPSO and DDS have the steepest drop in error during first 2000 function evaluations.
-Thereafter the error drop in DDS remains nearly constant whereas DDSMLSPSO drops further from around 2000 to 4000 function evaluations.
-At the end of 4000 function evaluations DDSMLSPSO attains the lowest error.
-}\label{fig-train}
+DOPS, DDS and SA have the steepest drop in error during first 300 function evaluations.
+Thereafter the error drop in DDS and SA remains nearly constant whereas DOPS continues to drops further.
+At the end of 4000 function evaluations DOPS attains the lowest error.
+The next best estimate using DDS is nearly 3 times greater than the lowest error using DDS.
+}\label{fig-convergence}
 \end{figure}
@@
 \caption{Model fits on experimental data using DDSMLSPSO.
-The model parameters were estimated using DDSMLSPSO.
+The model parameters were estimated using DOPS.
 Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from $N$= 25 trials.
@@
 \includegraphics[width=1.0\textwidth,height=0.5\textheight]{./figs/Figure_5_Sim_Validate_E2_E4_E6.pdf}
-\caption{Model predictions on unseen experimental data using parameters obtained from DDSMLSPSO.
-The parameter estimates that were obtained using DDSMLSPSO were tested against data that was not used in the model training.
+\caption{Model predictions on unseen experimental data using parameters obtained from DOPS.
+The parameter estimates that were obtained using DOPS were tested against data that was not used in the model training.
 Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from $N$= 25 trials.
@@
 }\label{fig-validation}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth,height=0.5\textheight]{./figs/Figure_6_Ackley_Rast}
+\caption{Error convergence on 300 dimensional Ackley function and Rastrigin function.
+\textbf {(A)} Error convergence of 4 different meta-heuristics on 300 dimensional Ackley function \textbf {(B)} Error convergence of 4 different meta-heuristics on 300 dimensional Rastrigin function
+}\label{fig-testfunctions}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth,height=0.7\textheight]{./figs/Figure_B1_B4_params_measuredstates}
+\caption{ \textbf {(A)} Difference between nominal and optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
+\textbf {(B)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
+\textbf {(C)} Difference between nominal and optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
+\textbf {(D)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
+}\label{fig-benchmark}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth,height=0.7\textheight]{./figs/B1_sims.pdf}
+\caption{\textbf {(Data fits for Problem B1)} Pseudo-experimental data (red circles) vs.
+optimal solution obtained using DOPS (solid blue lines) for the 44 observed states.
+X axis: time [s]; Y axis: metabolite concentrations [mM].
+}\label{fig-sims-b1}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth,height=0.7\textheight]{./figs/FigB4_sims.pdf}
+\caption{\textbf {(Data fits for Problem B4)} Pseudo-experimental data (red x) vs.
+optimal solution obtained using DOPS (solid blue lines) for the 13 observed states.
+X axis: time [s]; Y axis: metabolite concentrations [mM].
+}\label{fig-sims-b4}
 \end{figure}
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.02\textwidth,height=0.5\textheight]{./figs/Figure_6_Ackley_Rast}
-\caption{Error convergence on 300 dimensional Ackley function and Rastrigin function.
-\textbf {(A)} Error convergence of 5 different meta-heuristics on 300 dimensional Ackley function \textbf {(B)} Error convergence of 5 different meta-heuristics on 300 dimensional Rastrigin function
-}\label{fig-testfunctions}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.02\textwidth,height=0.5\textheight]{./figs/Figure_7_Benchmarks_Parameters}
-\caption{Difference between optimal and nominal parameter vector values on benchmark problems.
-\textbf {(A)} Problem B1: Genome wide kinetic model of E.coli \textit{S.cerevisiae} with 1759 unknown parameters.
-\textbf {(B)} Problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
-}\label{fig-benchmark}
-\end{figure}
-
 \end{document}

### 2016-02-11 Paper_v5.tex
Toward a final draft

@@
 We tested the performance of DOPS on a model of human coagulation cascade.
-We performed N = 25 trials with M = 4000 function evaluations per trial, and compared the performance of DOPS with other commonly
+We performed $\mathcal{T}$ = 25 trials with $\mathcal{N}$ = 4000 function evaluations per trial, and compared the performance of DOPS with other commonly
 used meta-heuristics such as differential evolution (DE), simulated annealing (SA) and dynamically dimensioned search (DDS).
@@
 
-\section*{Problem Formulation}
+\clearpage
 
-The problem of parameter estimation in dynamic biological systems consists of finding an optimal parameter vector that results in the best fit between the model predictions and observed experimental data.The cost function or objective that measures the goodness of fit is generally the Euclidean norm which is minimized under certain constraints.
-This can be mathematically expressed as follows.
+\section*{Results}
+
+\subsection*{Optimization problem formulation.}
+The problem of parameter estimation in a dynamic biological model consists of finding an optimal parameter vector
+which minimizes the difference between model simulations and $\mathcal{E}$ experimental measurements. This difference is quantified by an objective function $K\left(\mathbf{p}\right)$
+which is typically the Euclidean norm of the simulation error subject to problem specific and parameter bounds constraints:
 
@@
 & \text{minimize}
-& & K(\mathbf{p})=(\sum_{i=1}^N (g_i(t_i,\mathbf{x,p,u})-y_i)^2) \\
+& & K(\mathbf{p})=\sum_{i=1}^\mathcal{E} \left(g_i(t_i,\mathbf{x,p,u})-y_i\right)^2 \\
 & \text{subject to}
-&&\dot{x}=f(t,\mathbf{x(t,\mathbf{p}),p,u(\mathbf{t})})\\
-&&&x(t_0) = x_0\\
-&&&c(t,\mathbf{x,p,u}) \geqslant 0 \\
+&&\dot{\mathbf{x}}=\mathbf{f}(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p})\\
+&&&\mathbf{x}(t_0) = \mathbf{x}_0\\
+&&&\mathbf{c}(t,\mathbf{x,p,u}) \geqslant \mathbf{0} \\
 &&& \mathbf{p}^L \leqslant \mathbf{p} \leqslant \mathbf{p}^U\\
@@
 \end{equation}
-
-where t is time, x is the state variable with an initial state x\textsubscript{0}, u is the input vector \, p is the parameter vector and f is the system of equations (like differential equations or set of algebraic constraints), c in the set of linear and non-linear constraints, g is the set of observables and p\textsuperscript{L} and p\textsuperscript{U} are the lower and upper bounds.
-Thus the problem eventually is to find {p}\textsuperscript{*} where
-
+where t is time, $\mathbf{x}\left(t,\mathbf{p}\right)$ is the state variable vector with an initial state $\mathbf{x}_{0}$, $\mathbf{u}(t)$ is a model input vector,
+$\mathbf{f}$ is the system of model equations (e.g., differential equations or algebraic constraints) and $\mathbf{p}$ is the model parameter vector.
+The parameter search (or model simulations) can be subject to $\mathbf{c}$ linear or non-linear constraints, and parameter bound constraints where
+p\textsuperscript{L} and p\textsuperscript{U} denote the lower and upper parameter bounds, respectively.
+The problem eventually is to find:
 \begin{equation}\label{eq:static_opt_2}
-\mathbf{p^{*}}= \arg\min_{p} K(\mathbf{p})
+\mathbf{p^{*}}= \arg\min_{\mathbf{p}} K\left(\mathbf{\mathbf{p}}\right)
 \end{equation}
 
-\section*{Results}
-\subsection*{Dynamic Optimization with Particle Swarms (DOPS).}
-DOPS is a novel meta-heuristic that combines multi-swarm based particle swarm methods with dynamically dimensioned search (Fig. \ref{fig-algorithm}).
-The goal of this approach is to obtain optimal or near optimal parameters for high-dimensional complex biological systems within a pre-specified number of function evaluations.
-We randomly initialized a swarm of $\mathcal{K}$-dimensional particles (represented as ${x}_{i}$), wherein each of these particles corresponds to a $\mathcal{K}$-dimensional parameter vector.
-After initialization, the particles were grouped into different sub-swarms randomly.
-Thereafter within each sub-swarm ${S}_{k}$,  particles were updated according to the following rule.
+\subsection*{Dynamic optimization with particle swarms (DOPS).}
+DOPS is a novel meta-heuristic which combines multi-swarm particle swarm methods with dynamically dimensioned search (Fig. \ref{fig-algorithm}).
+The goal of DOPS is to estimate optimal or near optimal parameter vectors for high-dimensional biological models within a specified number of function evaluations.
+Toward this objective, DOPS begins by using a particle swarm search and then dynamically switches, using an adaptive switching criteria, to the DDS search phase.
+We began the particle swarm phase by randomly initializing a swarm of $\mathcal{K}$-dimensional particles (represented as $z_{i}$), wherein each particle corresponded to a $\mathcal{K}$-dimensional parameter vector.
+After initialization, particles were randomly partitioned into $k$ equal sized sub-swarms $\mathcal{S}_{1},\hdots,\mathcal{S}_{k}$.
+Thereafter within each sub-swarm $\mathcal{S}_{k}$, particles were updated according to the rule:
+\begin{equation}\label{eqn:update-rule}
+	\mathbf{z}_{i,j} = \theta_{1,j-1}\mathbf{z}_{i,j-1} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{z}_{i,j-1}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G}_{k} - \mathbf{z}_{i,j-1}\right)
+\end{equation}
+where $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ were adjustable parameters, $\mathcal{L}_{i}$ denotes the best solution found by particle $i$ within sub-swarm $k$ for function evaluation $1\rightarrow j-1$, and
+$\mathcal{G}_{k}$ denotes the best solution found over all particles within sub-swarm $\mathcal{S}_{k}$.
+The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model parameters ($\mathcal{K}\times{1}$).
+Equation (3) is similar to the general particle swarm update rule, however, it does not contain velocity terms.
+In DOPS the parameter $\theta_{1,j-1}$ depends upon the function evaluations and is updated according to:
 \begin{eqnarray}
-	\mathbf{x}_{i,j} &=&\theta_{1,j-1}\mathbf{x}_{i,j-1} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{x}_{i,j-1}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{GL}_{k} - \mathbf{x}_{i,j-1}\right)
+	\mathbf \theta_{1,j}&=&\frac{(\mathcal{N}-{j})*(\mathbf{w}_{max}-\mathbf{w}_{min}))}{(\mathcal{N}-{1})} + \mathbf{w}_{min}
 \end{eqnarray}
-where $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ are adjustable parameters, $\mathcal{L}_{i}$ denotes the best solution found by particle $i$ within the sub-swarm till function evaluation $j$-$1$, and
-$\mathcal{GL}_{k}$ denotes the best solution found over the population of all particles within the sub-swarm ${S}_{k}$.
-The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model
-parameters ($\mathcal{K}\times{1}$).
-Equation (3) is suggested by the general form of particle swarm equations.
-However this equation does not contain the velocity terms that are generally associated with particle swarm optimization.
-In our algorithm the parameter $\theta_{1,j-1}$ depends on the function evaluations and is controlled according to the following equation
-\begin{eqnarray}
-	\mathbf \theta_{1,j}&=&((\mathbf{N}-{j})*(\mathbf{w}_{max}-\mathbf{w}_{min}))/(\mathbf{N}-{1}) + \mathbf{w}_{min}
-\end{eqnarray}
-where $\mathbf{N}$ represents the total number of function evaluations, $\mathbf{w}_{max}$ and $\mathbf{w}_{min}$ are the maximum and minimum inertia weights respectively.
-While updating the particle, we made sure all dimensions of the solution represented by the particle were within bounds using a set of reflection boundary conditions shown below.
+where $\mathcal{N}$ represents the total number of function evaluations, $\mathbf{w}_{max}$ and $\mathbf{w}_{min}$ are the maximum and minimum inertia weights, respectively.
+While updating the particles, we made sure all dimensions of the solution represented by the particle were within bounds using a set of reflection boundary conditions:
 
-\begin{eqnarray}
-\end{eqnarray}
+\begin{eqnarray*}
+\end{eqnarray*}
 
 \begin{algorithm}
-\If{$x_{i,j}^{old} < x_{i}^{min}$}{
-  ${x}_{i,j}^{new}={x}_{i,j}^{old}+(x_{i}^{min} - {x}_{i,j}^{old})$
-  \If{${x}_{i,j}^{new}>x_{i}^{max}$}{
-    ${x}_{i,j}^{new}=x_{i}^{max}$
+\If{$z_{i,j}^{old} < z_{i}^{min}$}{
+  ${z}_{i,j}^{new}={z}_{i,j}^{old}+(z_{i}^{min} - {z}_{i,j}^{old})$
+  \If{${z}_{i,j}^{new}>z_{i}^{max}$}{
+    ${z}_{i,j}^{new}=z_{i}^{max}$
   }
 }
-\If{$x_{i,j}^{old} > x_{i}^{max}$}{
-  ${x}_{i,j}^{new}={x}_{i,j}^{old}+({x}_{i,j}^{old}-x_{i}^{max})$
-  \If{${x}_{i,j}^{new}<x_{i}^{min}$}{
-    ${x}_{i,j}^{new}=x_{i}^{min}$
+\If{$z_{i,j}^{old} > z_{i}^{max}$}{
+  ${z}_{i,j}^{new}={z}_{i,j}^{old}+({z}_{i,j}^{old}-z_{i}^{max})$
+  \If{${z}_{i,j}^{new}<z_{i}^{min}$}{
+    ${z}_{i,j}^{new}=z_{i}^{min}$
   }
@@
 
-After every $\mathbf{g}$  function evaluations, the particles within all sub-swarms were mixed and then randomly redistributed to a new sub-swarm.
-The particles were then again updated according Eq.1.
-This process continued till $\mathcal{FR}*\mathbf{N}$ number of functions evaluations, where $\mathcal{FR}$ represents the fraction of evaluations with the multi-swarms.
+After every $\mathcal{M}$ function evaluations, particles were randomly redistributed to a new sub-swarm, and updated according to Eqn. \eqref{eqn:update-rule}.
+This process continued till $\mathcal{F}*\mathbf{N}$ number of functions evaluations, where $\mathcal{F}$ is the fraction of evaluations with the multi-swarms.
 The fraction of evaluations within the swam phase is based on the drop in error.
 If the error does not drop by 1\% for a certain number of evaluations the search within swarm phase is terminated.
-At the end of these function evaluations, we froze all the solutions represented by various particles and chose the particle with best solution among $\mathcal{GL}_{1} \cdots \mathcal{GL}_{NS}$ as the initial candidate vector $\mathcal{G}$ for the remaining $({1}-\mathcal{FR})*\mathbf{N}$ number of function evaluations.
-
-This particle was then updated according to the following rule
-
+At the end of these function evaluations, we froze all the solutions represented by various particles and chose the particle with best solution among $\mathcal{G}_{1} \cdots \mathcal{G}_{NS}$
+as the initial candidate vector $\mathcal{G}$ for the remaining $({1}-\mathcal{F})*\mathbf{N}$ number of function evaluations.
+This particle was then updated according to the rule:
 \begin{equation}
@@
 \end{equation}
-
-where $\mathbf{J}$ represents the set containing the specific dimensions being perturbed, ${r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$.
-$\sigma$ is the amplitude of perturbation given by following equation:
-\begin{eqnarray}
-	\sigma& = &\mathbf{R}(\mathcal{MAX} -\mathcal{MIN})
-\end{eqnarray}
+where $\mathbf{J}$ represents the set containing the specific dimensions being perturbed, ${r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$,
+and $\sigma$ is the amplitude of perturbation:
+\begin{equation}
+	\sigma = \mathbf{R}(\mathcal{MAX} -\mathcal{MIN})
+\end{equation}
 where $\mathbf{R}$ is the scalar perturbation size parameter, $\mathcal{MAX}$ and $\mathcal{MIN}$ are ($\mathcal{K}\times{1}$) vectors that represent the maximum and minimum bounds on each dimension.
@@
 
+\clearpage
+
 \subsection*{Performance of DOPS on a model of the human coagulation cascade.}
-We compared the performance of DOPS on a model of blood coagulation against commonly used meta-heuristics like simulated annealing (SA), differential evolution (DE), and dynamically dimensioned search (DDS).
-Coagulation is an archetype biochemical network that is highly interconnected and tightly regulated with multiple positive and negative feedback loops (Fig.
-\ref{fig-coagulation-network}).
-The biochemistry underlying coagulation, though quite complex has been well studied \cite{mann2003dynamics,mann2003all,mann2003thrombin,vogler2009contact,diamond2013systems,fogelson2005coagulation,anand2003model}, and reliable experimental coagulation models have been developed \cite{hockin2002model,chatterjee2010systems,mann2006models,luan2007computationally}.
-This makes it an ideal system for mathematical modeling and parameter estimation.
-Coagulation is regulated by a set of serine proteases also known as coagulation factors and blood platelets.
+We compared the performance of DOPS with simulated annealing (SA), differential evolution (DE), and dynamically dimensioned search (DDS) for a model of the human coagulation cascade.
+Coagulation is an archetype biochemical network that is highly interconnected, containing both positive and negative feedback (Fig. \ref{fig-coagulation-network}).
+In this study, we used the coagulation model of Luan et al \cite{luan2007computationally} which modeled coagulation as a system of coupled non-linear ordinary differential equations.
+The Luan model contained 148 parameters and 92 species and has been validated using 21 published datasets.
 
-%Added a little more description here
-The coagulation factors are generally in an inactive state and are known as zymogens.
-These zymogens are activated through certain triggers.
+The biochemistry underlying coagulation, though quite complex, has been well studied \cite{mann2003dynamics,mann2003all,mann2003thrombin,vogler2009contact,diamond2013systems,fogelson2005coagulation,anand2003model},
+and reliable experimental coagulation models have been developed \cite{hockin2002model,chatterjee2010systems,mann2006models,luan2007computationally}.
+Coagulation is mediated by a family proteases in the circulation, called factors and a key group of blood cells, called platelets.
+The central process in coagulation is the conversion of prothrombin (fII), an inactive coagulation factor, to the master protease thrombin (FIIa).
+Thrombin generation involves three phases, initiation, amplification and termination [13,14].
+Initiation requires a trigger event, for example vessel injury, which leads to the activation of factor VII (FVIIa).
+Two converging pathways, the extrinsic and intrinsic cascades, then process and amplify this initial coagulation signal.
+The extrinsic cascade is generally believed to be the main mechanism of thrombinogenesis in the blood [15–17].
+Initially, thrombin is produced upon cleavage of prothrombin by fluid phase activated factor X (FXa), which itself has been activated by Tissue Factor/activated factor VII (TF/FVIIa) [10].
+Picomolar amounts of thrombin then activate the cofactors factors V and VIII (fV and fVIII) and platelets, leading to the formation of the tenase and prothrombinase complexes on activated platelets.
+These complexes amplify the early coagulation signal by further activating FXa, and directly converting prothrombin to thrombin.
+There are several control points in the cascade that inhibit thrombin formation, and eventually terminate thrombin generation.
+Tissue Factor Pathway Inhibitor (TFPI) inhibits FXa formation catalyzed by TF/FVIIa, while antithrombin III (ATIII) neutralizes several of the proteases generated during coagulation, including thrombin.
+Thrombin itself also inadvertently plays a role in its own inhibition; thrombin, through interaction with thrombomodulin, protein C and endothelial cell protein C receptor (EPCR), converts protein C to activated protein C (APC) which attenuates the coagulation response by proteolytic cleavage of fV/FVa and fVIII/FVIIIa.
+Termination occurs after either prothrombin is consumed, or thrombin formation is neutralized by inhibitors such as APC or ATIII.
+
+%Coagulation is regulated by a set of serine proteases known as coagulation factors and a type of blood cell called a platelet.
+%In the circulation, coagulation factors are generally in an inactive state known as a zymogen.
+%These zymogens are activated through certain triggers.
 Trigger events like injury or trauma or sepsis expose factors like collagen, tissue factor and von Willebrand factor (vWF) to blood.
@@
 surface protein thrombomodulin (TM).
-Luan et al.
-modeled coagulation using coupled non-linear ordinary differential equations with 148 reactions and 92 species \cite{luan2007computationally} and validated model using 21 published datasets.
+%This choice of using a linear combination of two different error functions was motivated by poor validation results while using a single error function.
 
-To train the model parameters, we used data sets from TF-VIIa initiated coagulation with no anticoagulants.
-The objective function was a weighted linear combination of two different error functions that used data sets representing coagulation initiated with different concentrations of TF-VIIa (5pM, 5nM) \cite{hockin2002model}.
-This choice of using a linear combination of two different error functions was motivated by poor validation results while using a single error function.
-We restricted the number of function evaluations to 4000 for each algorithm and performed 25 trials of this experiment.
-DOPS exhibits a much faster rate of error convergence and has a much lower final error as compared to the other algorithms (Fig.\ref{fig-convergence}).
-
+To estimate coagulation model parameters, we used data sets from TF-VIIa initiated coagulation in the absence of anticoagulants.
+The objective function was a linear combination of two error functions that used data sets representing coagulation initiated with different concentrations of TF-VIIa (5pM, 5nM) \cite{hockin2002model}.
+We restricted the number of function evaluations to 4000 for each algorithm we tested, and performed 25 trials of each experiment.
+DOPS convergences faster and has a lower final error compared to the other algorithms (Fig. \ref{fig-convergence}).
 Within the first 1000 function evaluations of DOPS there is a very rapid drop in error.

### 2016-02-11 Paper_v5.tex
Toward a final draft

@@
 We restricted the number of function evaluations to 4000 for each algorithm we tested, and performed 25 trials of each experiment.
-DOPS convergences faster and has a lower final error compared to the other algorithms (Fig. \ref{fig-convergence}).
-Within the first 1000 function evaluations of DOPS there is a very rapid drop in error.
-Approximately between 500-1000 function evaluations a switch to dynamically dimensioned search phase happens (this switch varies from trial to trial since the switch is based on the error from the swarm phase).
-Overall at the end of 4000 function evaluations DOPS minimizes the error (final objective error is 5.5916e+04) to a much greater extent than any of the other algorithms.
+DOPS converged faster and had a lower final error compared to the other algorithms (Fig. \ref{fig-convergence}).
+Within the first 1000 function evaluations of DOPS, there was a rapid drop in error.
+Approximately between 500-1000 function evaluations a switch to dynamically dimensioned search phase happens (this switch varied from trial to trial since the switch was based on the error from the swarm phase).
+Overall at the end of 4000 function evaluations DOPS minimized the error (final objective error is 5.5916e+04) to a greater extent than the other algorithms.
 Using the parameters obtained at the end of 4000 function evaluations we examined the 'fits' between models predictions and experimental data (Fig.\ref{fig-train}).
 The solid lines represent the mean value of prediction over 25 trials and the shaded region represents the 99\% confidence interval.
-Subsequently we used these optimal parameters to make model predictions that was compared against completely 'unseen' or untrained experimental data where coagulation was initiated with 500pM,50pM,10pM concentrations of TF-VIIa respectively (Fig.\ref{fig-validation}).
+Subsequently we used these optimal parameters to make model predictions that was compared against completely 'unseen' or untrained experimental data where
+coagulation was initiated with 500pM,50pM,10pM concentrations of TF-VIIa respectively (Fig.\ref{fig-validation}).
 
@@
 The final objective function value (Table \ref{table:benchmark-problems}) is an order smaller than nominal error for B1 possibly due to overfitting against the noise that was added to the synthetic data.
-
 Having obtained good fits on the coagulation problem and the benchmarks we proceeded to compare the performance of DOPS with other algorithms on commonly used test functions for global optimization.

### 2016-02-13 Paper_v2.tex
Figs 4 and 5

--- 2016-02-07 Paper_v3.tex
@@
 \begin{titlepage}
-{\par\centering\textbf{\Large {Dynamic Optimization with Particle Swarm (DOPS): A meta-heuristic for parameter estimation in biochemical models}}}
+{\par\centering\textbf{\Large {Dynamically dimensioned particle swarm optimization (DdPSO): A meta-heuristic for parameter estimation in biochemical models}}}
 \vspace{0.05in}

### 2016-02-13 Paper_v4.tex
Figs 4 and 5

--- 2016-02-10 Paper_v5.tex
@@
 Mathematical modeling is a powerful tool to analyze, and ultimately design biochemical networks.
-However, the estimation of biochemical model parameters is a significant challenge.
-Parameter estimation in biochemical models typically involves expensive function evaluations and noisy data, making it difficult to quickly obtain optimal solutions.
-Biochemical models often also have many local extrema which further complicates parameter estimation.
-Toward these challenges, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel global meta-heuristic that combined features of multi-swarm particle swarm optimization with dynamically dimensioned search (DDS).
-DOPS uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors, the best of which is greedily updated using dynamically dimensioned search.
-We tested the performance of DOPS on a model of human coagulation cascade.
-We performed N = 25 trials with M = 4000 function evaluations per trial, and compared the performance of DOPS with other commonly
-used meta-heuristics such as differential evolution (DE), simulated annealing (SA) and dynamically dimensioned search (DDS).
-We further tested the predictive power of the coagulation model parameters against data not used in training, and found good agreement between simulations and experimental measurements.
-Lastly, we tested the performance of DOPS on commonly used test functions for global optimization and on published biochemical parameter estimation benchmark problems.
+However, estimation of biochemical model parameters is a significant challenge.
+Parameter estimation in biochemical models generally involves expensive function evaluations, making it difficult to quickly obtain optimal solutions.
+Additionally, biochemical models often have many local minima or maxima which further complicates the parameter estimation problem.
+In this study, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel meta-heuristic that combined features of multi-swarm particle swarm optimization with dynamically dimensioned search (DDS).
+DOPS uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors of which the best one is greedily updated using dynamically dimensioned search.
+We first tested the performance of DOPS on a model of human coagulation cascade.
+We performed 25 trials with 4000 function evaluations per trial, and compared the performance of DOPS with other commonly
+used meta-heuristic approaches such as differential evolution (DE), simulated annealing (SA) and dynamically dimensioned search (DDS).
+We further tested the predictive power of the coagulation model parameters against data not used in training, and found good agreement between model simulations and experimental measurements.
+Thereafter we tested the performance of DOPS on commonly used test functions for global optimization (Ackley and Rosenbrock) and on published biochemical parameter estimation benchmark problems.
 For the wide range of problems that we considered, DOPS outperformed other meta-heuristic approaches despite a limited number of function evaluations.
@@
 
-\vspace{0.1in}
 {\noindent \textbf{Keywords:}~Parameter identification, Meta-heuristic optimization, Biochemical modeling}
@@
 Parameter estimation is typically framed as a non-linear optimization problem wherein the residual (or objective function) between experimental data and model simulations is minimized using an optimization strategy \cite{moles2003parameter}.
-Optimal parameters obtained from model training are then used to validate the model on unseen experimental data.
+The optimal parameters obtained from training are then used to validate the model on hitherto unseen experimental data.
 If validation fails, model construction and calibration are repeated iteratively until satisfactory results are obtained.
 
-Parameter estimation is a major challenge in the iterative development of biochemical models.
+Parameter estimation is a major challenge in the development of large biochemical models.
 Although parameter estimation has been a well studied problem in engineering for decades \cite{nieman1971review,beck1977parameter,young1981parameter,beck1998inverse},
 the complex dynamics of large biological systems and noisy, often incomplete experimental data pose a unique estimation challenge.
-Often optimization problems involving biological systems are non-linear and multi-modal i.e. typical models often have multiple local minima or maxima \cite{moles2003parameter,banga2008optimization}.
-Non-linearity coupled with multi-modality generally renders local optimization techniques such as pattern search \cite{hooke1961direct}, Nelder-Mead simplex methods \cite{nelder1965simplex},
-steepest descent methods or Levenberg-Marquardt \cite{more1978levenberg} incapable of reliably obtaining optimal solutions as they generally terminate at local minimum.
-Though deterministic global optimization techniques (for example algorithms based on branch and bound) can handle non-linearity and multi-modality \cite{esposito2000deterministic,horst2013global},
-the absence of derivative information, discontinuity of the objective functions, non-smooth regions or the lack of knowledge about the objective function hampers these techniques.
+Most of the optimization problems pertaining to biological systems are non-linear and multi-modal i.e.
+most of them have multiple local minima or maxima \cite{moles2003parameter,banga2008optimization}.
+The non-linearity of the problem coupled with multi-modality generally renders the local optimization techniques like pattern search method \cite{hooke1961direct}, Nelder-Mead simplex method \cite{nelder1965simplex}, steepest descent method or Levenberg-Marquardt method \cite{more1978levenberg} incapable of reliably obtaining an optimal solution since they generally stop at the first local minimum.
+Though deterministic global optimization techniques (for example algorithms based on branch and bound framework) can handle non-linearity and multi-modality \cite{esposito2000deterministic,horst2013global}, the absence of derivative information, discontinuity of the objective functions, non-smooth regions or the lack of any knowledge about the objective function severely hamper the use of these techniques.
 
-Meta-heuristic stochastic optimization approaches like Genetic Algorithms (GAs) [ADI-REF], Simulated Annealing (SA) \cite{kirkpatrick1983optimization},
-Evolutionary Programming [ADI-REF] and population based searches like Differential Evolution (DE) \cite{storn1997differential} have shown promise on nonlinear multi-modal problems \cite{sun2012parameter}.
-These techniques do not make any assumptions about the structure of objective function nor do they require \textit{a~priori} information about the objective function.
+Meta-heuristic stochastic optimization approaches like Genetic Algorithms (GAs), Simulated Annealing (SA) \cite{kirkpatrick1983optimization}, Evolutionary Programming and population based searches like Differential Evolution (DE) \cite{storn1997differential} have shown promise in this regard \cite{sun2012parameter}.
+They do not make any assumptions about the structure of objective function or require any \textit{a priori} information about the objective function.
 Though they do not guarantee strong convergence, these approaches are effective in finding near optimal solutions.
-Mendes et al. \cite{mendes1998non} used simulated annealing to estimate rate constants for the irreversible inhibition of HIV proteinase,
-Modchang et al. \cite{modchang2008mathematical} used a genetic algorithm to estimate parameters for a model of signal transduction [ADI-EXPAND],
-differential evolution approaches have also been effective on various biological problems \cite{tsai2005evolutionary,wang2001hybrid,noman2007inferring}.
-Tashkova et al. compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis and showed that DE was the most effective \cite{tashkova2011parameter}.
-Banga and co-workers have also successfully applied scatter-search methods to estimate parameters on non-linear biological processes \cite{villaverde2012cooperative,rodriguez2006novel,egea2007scatter}.
-Hybrid approaches that combine a meta-heuristic with a local optimization search, wherein a near globally optimal solution obtained using a meta-heuristic is refined using a local search have also become popular.
-Villaverde et al. combined scatter search with local search methods for parameter estimation in a collection of systems biology models \cite{villaverde2015biopredyn}.
-Fan et al. recently showed that population based meta-heuristics along with decomposition based methods can be also used to model gene circuits from mRNA data \cite{fan2015parameter}.
-Despite these successes, a major drawback of most metaheuristic approaches is the large number of function evaluations required to explore the parameter space.
-Typically as models grow in size and complexity, evaluation of the objective function becomes computationally expensive.
-Thus performing a large number of function evaluations is not desirable (and perhaps not feasible).
+Mendes et al.
+\cite{mendes1998non} used Simulated Annealing to estimate rate constants for the irreversible inhibition of HIV proteinase, Modchang et al.
+\cite {modchang2008mathematical} used Genetic Algorithms to estimate parameters for a mathematical model of signal transduction, Differential Evolution based approaches have been an effective approach on various systems \cite{tsai2005evolutionary,wang2001hybrid,noman2007inferring}.
+Tashkova et al.
+\cite{tashkova2011parameter} compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis and showed that DE was the most effective.
+Banga and co-workers have applied scatter-search based methods \cite{villaverde2012cooperative,rodriguez2006novel,egea2007scatter} to estimate parameters on non-linear biological processes.
+Hybrid approaches that combine a meta-heuristic with a local optimization search, wherein a near globally optimal solution that is obtained with a meta-heuristic is further refined using a local search have also become popular.
+Villaverde et al.
+\cite{villaverde2015biopredyn} combined scatter search with local search methods for parameter estimation in large scale systems biology models.
+Fan et al.
+recently showed that population based meta-heuristics along with decomposition based methods can be used to model gene circuits from mRNA data \cite{fan2015parameter}.
+Despite these successes, a  major drawback with most of these metaheuristic approaches is the vast number of objective function evaluations they take in arriving at good search regions.
+As the models grow in size and complexity and the number of dimensions of parameter vector increases, evaluation of the objective function becomes computationally expensive.
+Thus performing a large number of evaluations is not computationally feasible.
 
-%In many of these high dimensional problems approaching an exact solution may not be necessary.
+In many of these high dimensional problems approaching an exact solution may not be necessary.
 Gutenkust et al.
@@
 
-%Because the neighborhood in DDS is the entire parameter space, DDS is a global optimization technique although it is greedy.
-
 In this study, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel meta-heuristic that combines the global search capability of multi-swarm particle swarm optimization and dynamically dimensioned search (DDS).
@@
 DOPS uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors which are then greedily updated using dynamically dimensioned search.
+Because the neighborhood in DDS is the entire space, DDS is a global optimization technique although it is greedy.
 We first considered a model of human  coagulation cascade to test the performance of DOPS.
-Coagulation is a large, complex biochemical network involving strong positive feedback.
+Coagulation is a large, complex biochemical network that is tightly regulated with several feedback loops.
 We then tested the performance of DOPS on commonly used test functions for global optimization (Ackley and Rosenbrock) and published biochemical parameter estimation benchmark problems \cite{villaverde2015biopredyn}.
-DOPS outperformed common meta-heuristic approaches like Differential Evolution (DE), Simulated Annealing (SA) and dynamically dimensioned search (DDS) on both the test functions and the coagulation model.
-DOPS also performed well on the benchmark problems where it outperformed enhanced scatter search (eSS) and recovered the nominal parameters with only 4000 function evaluations ([ADI-compared with what?])
-across all the benchmark problems considered.
-Taken together, these studies suggest DOPS is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.
+DOPS outperformed common meta-heuristic approaches like Differential Evolution (DE), Simulated Annealing (SA) and dynamically dimensioned search (DDS) on the test functions and the coagulation model.
+It also performed very well on the benchmark problems where it outperformed enhanced scatter search (eSS) and recovered the nominal parameters with only 4000 function evaluations across all the benchmark problems considered.
 

### 2016-02-13 Paper_v5.tex

@@
 \subsection*{Dynamic optimization with particle swarms (DOPS).}
-DOPS is a novel meta-heuristic which combines multi-swarm particle swarm methods with dynamically dimensioned search (Fig. \ref{fig-algorithm}).
+DOPS is a novel meta-heuristic which combines multi-swarm particle swarm methods with the dynamically dimensioned search approach of Shoemaker and colleagues (Fig. \ref{fig-algorithm}).
 The goal of DOPS is to estimate optimal or near optimal parameter vectors for high-dimensional biological models within a specified number of function evaluations.
 Toward this objective, DOPS begins by using a particle swarm search and then dynamically switches, using an adaptive switching criteria, to the DDS search phase.
-We began the particle swarm phase by randomly initializing a swarm of $\mathcal{K}$-dimensional particles (represented as $z_{i}$), wherein each particle corresponded to a $\mathcal{K}$-dimensional parameter vector.
+
+\subsubsection*{Phase 1: Swarm phase.}
+We began the particle swarm phase of DOPS by randomly initializing a swarm of $\mathcal{K}$-dimensional particles (represented as $z_{i}$),
+wherein each particle corresponded to a $\mathcal{K}$-dimensional parameter vector.
 After initialization, particles were randomly partitioned into $k$ equal sized sub-swarms $\mathcal{S}_{1},\hdots,\mathcal{S}_{k}$.
@@
 After every $\mathcal{M}$ function evaluations, particles were randomly redistributed to a new sub-swarm, and updated according to Eqn. \eqref{eqn:update-rule}.
-This process continued till $\mathcal{F}*\mathbf{N}$ number of functions evaluations, where $\mathcal{F}$ is the fraction of evaluations with the multi-swarms.
-The fraction of evaluations within the swam phase is based on the drop in error.
-If the error does not drop by 1\% for a certain number of evaluations the search within swarm phase is terminated.
-At the end of these function evaluations, we froze all the solutions represented by various particles and chose the particle with best solution among $\mathcal{G}_{1} \cdots \mathcal{G}_{NS}$
-as the initial candidate vector $\mathcal{G}$ for the remaining $({1}-\mathcal{F})*\mathbf{N}$ number of function evaluations.
-This particle was then updated according to the rule:
+This process continued for $\mathcal{F}*\mathcal{N}$ functions evaluations, where $\mathcal{F}$ is the fraction of evaluations in the particle swarm phase of DOPS.
+If the simulation error stagnates for example, does not change by more than 1\% for a specified number of evaluations, the swarm phase is terminated and DOPS switches to
+exploring parameter space using a DDS approach.
+
+\subsubsection*{Phase 2: DDS phase.}
+At the conclusion of the swarm phase, the overall best particle, $\mathcal{G}_{k}$, over the k sub-swarms was to initalize the DDS phase.
+DOPS takes $\left(1-\mathcal{F}\right)*\mathcal{N}$ function evaluations in the DDS phase and then terminates the search.
+For the DDS phase, the best parameter estimate was updated using the rule:
 \begin{equation}
@@
 where $\mathbf{J}$ represents the set containing the specific dimensions being perturbed, ${r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$,
-and $\sigma$ is the amplitude of perturbation:
+and $\sigma$ denotes the perturbation amplitude:
 \begin{equation}
@@
 The set $\mathbf{J}$ was constructed using a probability function $\mathcal{P}_{j}$ that represented the probability whether a specific dimension $j$ was perturbed or not.
-This function is a monotonically decreasing function that decreases with the number of function evaluations.
+T
+his function is a monotonically decreasing function that decreases with the number of function evaluations.
 $\mathcal{P}_{j}$ can be any monotonically decreasing function, in our approach we used the following function:
@@
 Coagulation is an archetype biochemical network that is highly interconnected, containing both positive and negative feedback (Fig. \ref{fig-coagulation-network}).
-In this study, we used the coagulation model of Luan et al \cite{luan2007computationally} which modeled coagulation as a system of coupled non-linear ordinary differential equations.
+The biochemistry of coagulation, though complex, has been well studied \cite{mann2003dynamics,mann2003all,mann2003thrombin,vogler2009contact,diamond2013systems,fogelson2005coagulation,anand2003model},
+and reliable experimental protocols have been developed to interrogate the system \cite{hockin2002model,chatterjee2010systems,mann2006models,luan2007computationally}.
+Thus, the human coagulation cascade is an ideal test case;
+coagulation is challenging because it contains both fast and slow dynamics, but also accessible because of the availability of comprehensive data sets for model identification and validation.
+In this study, we used the coagulation model of Luan et al \cite{luan2007computationally}. Luan and coworkers modeled coagulation as a system of coupled non-linear ordinary differential equations.
 The Luan model contained 148 parameters and 92 species and has been validated using 21 published datasets.
 
-The biochemistry underlying coagulation, though quite complex, has been well studied \cite{mann2003dynamics,mann2003all,mann2003thrombin,vogler2009contact,diamond2013systems,fogelson2005coagulation,anand2003model},
-and reliable experimental coagulation models have been developed \cite{hockin2002model,chatterjee2010systems,mann2006models,luan2007computationally}.
 Coagulation is mediated by a family proteases in the circulation, called factors and a key group of blood cells, called platelets.
@@
 Thrombin generation involves three phases, initiation, amplification and termination [13,14].
-Initiation requires a trigger event, for example vessel injury, which leads to the activation of factor VII (FVIIa).
+Initiation requires a trigger event, for example a vessel injury which exposes tissue factor (TF), which leads to the activation of factor VII (FVIIa) and the formation of the TF/FVIIa complex.
 Two converging pathways, the extrinsic and intrinsic cascades, then process and amplify this initial coagulation signal.
-The extrinsic cascade is generally believed to be the main mechanism of thrombinogenesis in the blood [15–17].
-Initially, thrombin is produced upon cleavage of prothrombin by fluid phase activated factor X (FXa), which itself has been activated by Tissue Factor/activated factor VII (TF/FVIIa) [10].
-Picomolar amounts of thrombin then activate the cofactors factors V and VIII (fV and fVIII) and platelets, leading to the formation of the tenase and prothrombinase complexes on activated platelets.
-These complexes amplify the early coagulation signal by further activating FXa, and directly converting prothrombin to thrombin.
 There are several control points in the cascade that inhibit thrombin formation, and eventually terminate thrombin generation.
-Tissue Factor Pathway Inhibitor (TFPI) inhibits FXa formation catalyzed by TF/FVIIa, while antithrombin III (ATIII) neutralizes several of the proteases generated during coagulation, including thrombin.
-Thrombin itself also inadvertently plays a role in its own inhibition; thrombin, through interaction with thrombomodulin, protein C and endothelial cell protein C receptor (EPCR), converts protein C to activated protein C (APC) which attenuates the coagulation response by proteolytic cleavage of fV/FVa and fVIII/FVIIIa.
+Tissue Factor Pathway Inhibitor (TFPI) inhibits upstream activation events, while antithrombin III (ATIII) neutralizes several of the proteases generated during coagulation, including thrombin.
+Thrombin itself also inadvertently plays a role in its own inhibition; thrombin, through interaction with thrombomodulin, protein C and endothelial cell protein C receptor (EPCR), converts protein C to activated protein C (APC) which attenuates the coagulation response by proteolytic cleavage of amplification complexes.
 Termination occurs after either prothrombin is consumed, or thrombin formation is neutralized by inhibitors such as APC or ATIII.
@@
 
-To estimate coagulation model parameters, we used data sets from TF-VIIa initiated coagulation in the absence of anticoagulants.
-The objective function was a linear combination of two error functions that used data sets representing coagulation initiated with different concentrations of TF-VIIa (5pM, 5nM) \cite{hockin2002model}.
-We restricted the number of function evaluations to 4000 for each algorithm we tested, and performed 25 trials of each experiment.
-DOPS converged faster and had a lower final error compared to the other algorithms (Fig. \ref{fig-convergence}).
+To estimate the coagulation model parameters, we used data sets from TF/VIIa initiated coagulation in the absence of anticoagulants.
+The objective function was a linear combination of two error functions, representing coagulation initiated with different concentrations of TF-VIIa (5pM, 5nM) \cite{hockin2002model}.
+We restricted the number of function evaluations to $\mathcal{N}$ = 4000 for each algorithm we tested, and performed 25 trials of each experiment to collect average performance data (Table ZZ).DOPS converged faster and had a lower final error compared to the other algorithms (Fig.
+\ref{fig-convergence}).
 Within the first 1000 function evaluations of DOPS, there was a rapid drop in error.
@@
 Overall at the end of 4000 function evaluations DOPS minimized the error (final objective error is 5.5916e+04) to a greater extent than the other algorithms.
+
 Using the parameters obtained at the end of 4000 function evaluations we examined the 'fits' between models predictions and experimental data (Fig.\ref{fig-train}).
@@
 
-\subsection*{Performance of DOPS on benchmark problems and test functions.}
+\subsection*{Performance of DOPS on benchmark problems.}
 Villaverde and co-workers recently published a set of benchmark biochemical problems to evaluate parameter estimation methods \cite{villaverde2015biopredyn}.
@@
 In (Table \ref{table:benchmark-problems}) we summarize the results obtained on the coagulation model, the benchmark problems and the test functions.
+
+\subsection*{Influence of the number of sub-swarms on DOPS performance.}
 
@@
 \resizebox{\columnwidth}{!}{
-\begin{tabular}{|cccccc|}
-\hline
-                        & Coagulation & B1           & B4          & Ackley 300D & Rastrigin 300D \\
- \hline
+\begin{tabular}{r|lllll}
+\hline\\
+                        & Coagulation 					& B1           	& B4          & Ackley 			& Rastrigin \\\\
+\hline\\
+Evaluations   					& 4000                  & 4000         	& 4000        & 4000        & 4000        \\
 Lower Bound             & 0.001.pnom            & 5.pnom       	& 5.pnom      & 30          & 5.12        \\
 Upper Bound             & 1000.pnom             & 0.2.pnom     	& 0.2.pnom    & -15         & -5.12       \\
-CPU Time                & 10.0835 hours         & 38.308 hours & 6.2 minutes & 2.8 seconds & 2.6 seconds    \\
-Function Evaluations    & 4000                  & 4000         & 4000        & 4000        & 4000           \\
-Initial Objective Value & 1.11837e+07           & 1.4275e+07   & 1.8536e+07  & 21.12       & 99.985         \\
-Final Objective Value   & 5.5916e+04              & 3.5348e+04     & 38.1306     & 0           & 0              \\
-Nominal Objective Value & 4.7785e+06            & 1.0986e+06   & 39.0676     & 0           & 0      \\
+CPU Time                & 10.1 hrs         			& 38.3 hrs 			& 6.2 min 		& 2.8 s 			& 2.6 s    		\\\\
+
+Scaled initial error 		& 1.0           				& 1.0   				& 1.0  				& 1.0       	& 1.0         \\
+Scaled final error  		& $<$ 0.01           		& $<$ 0.01   		& $<$ 0.01    & $<$ 0.01   	& $<$ 0.01    \\
+Scaled nominal error 		& 0.42            			& 0.1   				& $<$ 0.01    & 0           & 0      			\\
+
 \hline
@@
 \centering
-\includegraphics[width=1.0\textwidth,height=0.5\textheight]{./figs/Figure_3_Errors_convergence.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure_3_Errors_convergence_v2.pdf}
 \caption{Error convergence rates of the 5 different algorithms on the coagulation model.
@@
 \centering
-\includegraphics[width=1.0\textwidth,height=0.5\textheight]{./figs/Figure_4_Sim_Train_E1_E5.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure_4_Sim_Train_E1_E5.pdf}
 \caption{Model fits on experimental data using DDSMLSPSO.
@@
 \centering
-\includegraphics[width=1.00\textwidth,height=0.5\textheight]{./figs/Figure_6_Ackley_Rast}
+\includegraphics[width=1.00\textwidth]{./figs/Figure_6_Ackley_Rast}
 \caption{Error convergence on 300 dimensional Ackley function and Rastrigin function.
@@
 \centering
-\includegraphics[width=1.00\textwidth,height=0.7\textheight]{./figs/Figure_B1_B4_params_measuredstates}
+\includegraphics[width=1.00\textwidth]{./figs/Figure_B1_B4_params_measuredstates}
 \caption{ \textbf {(A)} Difference between nominal and optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
@@
 \centering
-\includegraphics[width=1.00\textwidth,height=0.7\textheight]{./figs/B1_sims.pdf}
+\includegraphics[width=1.00\textwidth]{./figs/B1_sims.pdf}
 \caption{\textbf {(Data fits for Problem B1)} Pseudo-experimental data (red circles) vs.
@@
 \centering
-\includegraphics[width=1.00\textwidth,height=0.7\textheight]{./figs/FigB4_sims.pdf}
+\includegraphics[width=1.00\textwidth]{./figs/FigB4_sims.pdf}
 \caption{\textbf {(Data fits for Problem B4)} Pseudo-experimental data (red x) vs.

### 2016-03-07 Paper_v5.tex
Standard Error Table updated
New figures

@@
 \centering
+\caption{Error Analysis.}
+\label{table:Error-Analysis}
+\begin{center}
+\begin{tabular}{ r|l }
+\hline\\
+TF/FVIIa concentration& Normalized S.E.  \\\\
+\hline\\
+5~nM & 0.0376 \\
+500~pM & 0.0564\\
+50~pM & 0.1125 \\
+10~pM & 0.0823 \\
+5~pM & 0.0338 \\\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}[h]
+\centering
 \caption{Table with optimization settings and results for the coagulation problem, the benchmarks and test functions using DOPS.
@@
 \centering
-\includegraphics[width=1.0\textwidth,height=0.5\textheight]{./figs/Figure_1_Algorithm.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure_1_Algorithm.pdf}
 \caption{Dynamic Optimization with Particle Swarms.\textbf{A}: Each particle represents an N dimensional parameter vector.
@@
 \centering
-\includegraphics[width=1.00\textwidth,height=0.7\textheight]{./figs/Figure_2_CoagulationNetwork.pdf}
+\includegraphics[width=1.00\textwidth]{./figs/Figure_2_Coagulation_v4.pdf}
 \caption{Schematic of the extrinsic and intrinsic coagulation cascade\cite{luan2007computationally}.
@@
 \centering
-\includegraphics[width=1.0\textwidth,height=0.5\textheight]{./figs/Figure_5_Sim_Validate_E2_E4_E6.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure_5_Sim_Validate_E2_E4_E6.pdf}
 \caption{Model predictions on unseen experimental data using parameters obtained from DOPS.
@@
 \centering
-\includegraphics[width=1.00\textwidth]{./figs/Figure_6_Ackley_Rast}
+\includegraphics[width=0.7\textwidth]{./figs/Figure_6_TestFunctions-crop.pdf}
 \caption{Error convergence on 300 dimensional Ackley function and Rastrigin function.

### 2016-03-22 Paper_v5.tex
Added discussion on sub-swarm influence

--- 2016-02-13 Paper_v5.tex
@@
 \subsection*{Influence of the number of sub-swarms on DOPS performance.}
+We attempted to quantify the influence of sub-swarms and DDS phase on the error convergence rates.
+To do so we considered only the multi swarm search without the DDS phase for 4000 function evaluations and 40 particles on the coagulation model.
+While doing the multi swarm search without DDS we varied the number of sub-swarms as one, two, four, five and eight.
+The number of sub-swarms were chosen so that each swarm contained the same number of particles.
+Hence we did not consider swarm numbers of three and seven.
+All the other parameters remained the same for each of these searches.
+We observed that the higher sub-swarm numbers had faster convergence rates with eight having the fastest rate.
+However this difference in convergences is not pronounced amongst four, five and eight.
+This suggests that there is an optimal number of particles per swarm beyond which there is no significant advantage.
+
+We also observe that when these convergence rates are compared with DOPS, DOPS shows a very rapid drop when the swarm searches begin to saturate.
+The green region indicates when swarm search switches to a DDS based search.
+As we see in (Fig.
+) the error convergence rates tend to remain mostly flat for the purely swarm searches with different swarm numbers.
+However in DOPS as soon as convergence rate begins to saturate it automatically switches to a DDS phase which leads to a more pronounced drop.
+Thus this automated switching strategy appears to be crucial in leading the algorithm out of local optima or saddle point regions.
 
@@
 \clearpage
+
+\begin{table}[h]
+\centering
+\caption{Error Analysis.}
+\label{table:Error-Analysis}
+\begin{center}
+\begin{tabular}{ |c|c|c|c| } 
+\hline
+Simulation&Concentration& Normalized Standard Error  \\
+\hline
+\multirow{3}{4em}{FVIIa-TF level} & 5nM & 0.0376\\ 
+& 500pM & 0.0564\\
+& 50pM & 0.1125\ 
+& 10pM & 0.0823 \\
+& 5pM & 0.0338 \\ 
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 

### 2016-03-22 Paper_v5.tex

@@
 \begin{center}
-\begin{tabular}{ |c|c|c|c| } 
-\hline
-Simulation&Concentration& Normalized Standard Error  \\
-\hline
-\multirow{3}{4em}{FVIIa-TF level} & 5nM & 0.0376\\ 
-& 500pM & 0.0564\\
-& 50pM & 0.1125\ 
-& 10pM & 0.0823 \\
-& 5pM & 0.0338 \\ 
+\begin{tabular}{ r|l }
+\hline\\
+TF/FVIIa concentration& Normalized S.E.  \\\\
+\hline\\
+5~nM & 0.0376 \\
+500~pM & 0.0564\\
+50~pM & 0.1125 \\
+10~pM & 0.0823 \\
+5~pM & 0.0338 \\\\
 \hline

### 2016-03-28 Paper_v5.tex
Algo and para switching

@@
 \usepackage{amsmath}
+\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
 \DeclareMathOperator*{\argmin}{\arg\!\min}
@@
 \usepackage{algorithm2e}
+\usepackage{algpseudocode}
 %\usepackage[space]{cite}
@@
 
+\newpage
+%\section*{Pseudo Code.} 
+%\begin{algorithm}[H]
+%
+% \While{$besterror\leq $}{
+%  read current\;
+%  \eIf{understand}{
+%   go to next section\;
+%   current section becomes this one\;
+%   }{
+%   go back to the beginning of current section\;
+%  }
+% }
+% \caption{Dynamic Optimization with Particle Swarms}
+%\end{algorithm}
+
+\IncMargin{1em}
+\begin{algorithm}
+\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
+\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
+\SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
+\Input{A randomized swarm of particles of size $NP\times K$ and fixed number of function evaluations N}
+\Output{Optimized parameter vector of size $1\times K$ }
+\BlankLine
+\emph{Initialize the particles randomly and assign particles randomly to various sub-swarms}\;
+\While{$j\leq N$}{
+
+\If{mod(j,G)=0} {
+ \emph{Reassign particles to different sub-swarms}\;
+}
+
+\For{$i\leftarrow 1$ \KwTo $NS$}{\label{forins}
+\emph{Update particles within sub-swarms according to equation 3}\;
+}
+
+\eIf{$besterror(j)\geq 0.99*besterror(j+1)$} {
+ $failurecounter\leftarrow failurecounter+1$\;
+ }
+ {
+ $failurecounter\leftarrow 0$\;
+ }
+
+ \If{$failurecounter\geq threshold$} {
+ 	\emph{Run Dynamically Dimensioned Search with N-j function evaluations}\;
+}
+
+$j\leftarrow j+1$\;
+}
+\caption{Dynamic Optimization with Particle Swarms}\label{algo_disjdecomp}
+\end{algorithm}\DecMargin{1em}
+
 \clearpage

### 2016-04-12 Paper_v5.tex
Algo schematic updated
Working on results and discussion

@@
 
+\subsection*{DOPS explores parameter space using a combination of global and greedy search.}
+DOPS is a novel meta-heuristic which combines multi-swarm particle swarm methods with the dynamically dimensioned search approach of Shoemaker and colleagues (Fig. \ref{fig-algorithm}).
+The goal of DOPS is to estimate optimal or near optimal parameter vectors for high-dimensional biological models within a specified number of function evaluations.
+Toward this objective, DOPS begins by using a particle swarm search and then dynamically switches, using an adaptive switching criteria, to the DDS search phase.
+[FINISH GENERIC DESCRIPTION OF APPROACH]
+
+\subsection*{DOPS minimized benchmark problems with a fixed number of function evaluations.}
+First, we tested DOPS and four other meta-heuristic approaches on the minimization of the Ackley and Rastrigin functions.
+DOPS performed similarly or outperformed the other meta-heuristic approaches on both of these test functions (Fig. \ref{fig-testfunctions}).
+The Ackley and Rastrigin functions both have multiple local extrema and attain a global minimum value of zero.
+In each case, we fixed the maximum number of function evaluations at 4000 and ran 25 independent experiments with different initial parameter vectors.
+DOPS found optimal or near optimal solutions for both the 10-dimensional Ackley (Fig.
+\ref{fig-testfunctions}A) and Rastrigin (Fig.
+\ref{fig-testfunctions}B) functions within the budget of function evaluations.
+In each of the 10-dimensional cases, other meta-heurtistics such as DDS and DE also performed well, however DOPS consistently outperformed all other approaches tested.
+This performance difference was more pronounced as the dimension of the search problem increased; for a 300-dimensional Rastrigin function DOPS was the only approach to find an optimal or near optimal solution
+within the function evaluation budget (Fig. \ref{fig-testfunctions}B).
+Taken together, these results suggested that DOPS could perform at least as well as other meta-heuristic approaches on small test problems, but was especially suited to large dimension search spaces.
+Next, we tested DOPS on benchmark biochemical models of varying complexity.
+
+Villaverde and co-workers published a set of benchmark biochemical problems to evaluate parameter estimation methods \cite{villaverde2015biopredyn}.
+They ranked ordered these example problems from most expensive to least expensive.
+We evaluated the performance of DOPS on problems from the least and most expensive categories.
+The least expensive problem (henceforth referred to as problem B4) was a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters.
+The biochemical reactions were modeled using modular rate laws and the generalized form of Michaelis-Menten kinetics.
+The expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters (henceforth referred to as problem B1).
+In both cases, synthetic time series data sets were generated using known parameter values.
+For problem B1, the time series data consisted of 44 observables, and for problem B4 the data corresponded to 13 different metabolite measurement sets.
+We fixed the number of function evaluations at 4000 for DOPS and trained both the models against the synthetic experimental data.
+In both cases we found good fits  (Fig. \ref{fig-sims-b1} and Fig. \ref{fig-sims-b4}) to the problems within 4000 evaluations.
+We recaptured the nominal parameter values within 4000 evaluations (Fig. \ref{fig-benchmark}).
+
+%The final objective function value (Table \ref{table:benchmark-problems}) is an order smaller than nominal error for B1 possibly due to overfitting against the noise that was added to the synthetic data.
+%Having obtained good fits on the coagulation problem and the benchmarks we proceeded to compare the performance of DOPS with other algorithms on commonly used test functions for global optimization.
+We used a 300 dimensional Rastrigin function and 300 dimensional Ackley function.
+In both cases we see that the error convergence rate for DOPS is much faster as compared to the other three heuristics and it finds the global minimum of 0 in both cases.
+In (Table \ref{table:benchmark-problems}) we summarize the results obtained on the coagulation model, the benchmark problems and the test functions.
+
+\subsection*{Performance of DOPS on a model of the human coagulation cascade.}
+We compared the performance of DOPS with simulated annealing (SA), differential evolution (DE), and dynamically dimensioned search (DDS) for a model of the human coagulation cascade.
+Coagulation is an archetype biochemical network that is highly interconnected, containing both positive and negative feedback (Fig. \ref{fig-coagulation-network}).
+The biochemistry of coagulation, though complex, has been well studied \cite{mann2003dynamics,mann2003all,mann2003thrombin,vogler2009contact,diamond2013systems,fogelson2005coagulation,anand2003model},
+and reliable experimental protocols have been developed to interrogate the system \cite{hockin2002model,chatterjee2010systems,mann2006models,luan2007computationally}.
+Thus, the human coagulation cascade is an ideal test case;
+coagulation is challenging because it contains both fast and slow dynamics, but also accessible because of the availability of comprehensive data sets for model identification and validation.
+In this study, we used the coagulation model of Luan et al \cite{luan2007computationally}. Luan and coworkers modeled coagulation as a system of coupled non-linear ordinary differential equations.
+The Luan model contained 148 parameters and 92 species and has been validated using 21 published datasets.
+
+Coagulation is mediated by a family proteases in the circulation, called factors and a key group of blood cells, called platelets.
+The central process in coagulation is the conversion of prothrombin (fII), an inactive coagulation factor, to the master protease thrombin (FIIa).
+Thrombin generation involves three phases, initiation, amplification and termination [13,14].
+Initiation requires a trigger event, for example a vessel injury which exposes tissue factor (TF), which leads to the activation of factor VII (FVIIa) and the formation of the TF/FVIIa complex.
+Two converging pathways, the extrinsic and intrinsic cascades, then process and amplify this initial coagulation signal.
+There are several control points in the cascade that inhibit thrombin formation, and eventually terminate thrombin generation.
+Tissue Factor Pathway Inhibitor (TFPI) inhibits upstream activation events, while antithrombin III (ATIII) neutralizes several of the proteases generated during coagulation, including thrombin.
+Thrombin itself also inadvertently plays a role in its own inhibition; thrombin, through interaction with thrombomodulin, protein C and endothelial cell protein C receptor (EPCR), converts protein C to activated protein C (APC) which attenuates the coagulation response by proteolytic cleavage of amplification complexes.
+Termination occurs after either prothrombin is consumed, or thrombin formation is neutralized by inhibitors such as APC or ATIII.
+
+%Coagulation is regulated by a set of serine proteases known as coagulation factors and a type of blood cell called a platelet.
+%In the circulation, coagulation factors are generally in an inactive state known as a zymogen.
+%These zymogens are activated through certain triggers.
+Trigger events like injury or trauma or sepsis expose factors like collagen, tissue factor and von Willebrand factor (vWF) to blood.
+The exposure of these factors to blood kick-starts a series of convergent cascades that lead to conversion of zymogen prothrombinase to thrombin.
+For example when coagulation is initiated through the tissue factor pathway, tissue factor and activated factor VIIa (FVIIa) form a complex that activates factors X (fX) and IX (fIX).
+Activated factor X (fXa) thereafter activates downstream factors VIII and IX.
+The initial activation leads to the production of picomolar amounts of thrombin (fIIa) which activates platelets and amplifies its own production through the formation of a prothrombinase complex (FXa-FVa) on the surface of the activated platelets.
+Thrombin also  downregulates its own production by forming a complex with thrombomodulin which then activates protein C (PC).
+PC inhibits the formation of prothrombinase complex.
+In addition, Tissue factor pathway inhibitor (TFPI) downregulates FXa formation and activity by sequestering free FXa and TF-FVIIa in a FXa-dependent manner.
+Antithrombin III (ATIII)  inhibits all proteases.
+surface protein thrombomodulin (TM).
+%This choice of using a linear combination of two different error functions was motivated by poor validation results while using a single error function.
+
+To estimate the coagulation model parameters, we used data sets from TF/VIIa initiated coagulation in the absence of anticoagulants.
+The objective function was a linear combination of two error functions, representing coagulation initiated with different concentrations of TF-VIIa (5pM, 5nM) \cite{hockin2002model}.
+We restricted the number of function evaluations to $\mathcal{N}$ = 4000 for each algorithm we tested, and performed 25 trials of each experiment to collect average performance data (Table ZZ).DOPS converged faster and had a lower final error compared to the other algorithms (Fig.
+\ref{fig-convergence}).
+Within the first 1000 function evaluations of DOPS, there was a rapid drop in error.
+Approximately between 500-1000 function evaluations a switch to dynamically dimensioned search phase happens (this switch varied from trial to trial since the switch was based on the error from the swarm phase).
+Overall at the end of 4000 function evaluations DOPS minimized the error (final objective error is 5.5916e+04) to a greater extent than the other algorithms.
+
+Using the parameters obtained at the end of 4000 function evaluations we examined the 'fits' between models predictions and experimental data (Fig.\ref{fig-train}).
+The solid lines represent the mean value of prediction over 25 trials and the shaded region represents the 99\% confidence interval.
+Subsequently we used these optimal parameters to make model predictions that was compared against completely 'unseen' or untrained experimental data where
+coagulation was initiated with 500pM,50pM,10pM concentrations of TF-VIIa respectively (Fig.\ref{fig-validation}).
+
+\subsection*{Influence of the number of sub-swarms on DOPS performance.}
+We attempted to quantify the influence of sub-swarms and DDS phase on the error convergence rates.
+To do so we considered only the multi swarm search without the DDS phase for 4000 function evaluations and 40 particles on the coagulation model.
+While doing the multi swarm search without DDS we varied the number of sub-swarms as one, two, four, five and eight.
+The number of sub-swarms were chosen so that each swarm contained the same number of particles.
+Hence we did not consider swarm numbers of three and seven.
+All the other parameters remained the same for each of these searches.
+We observed that the higher sub-swarm numbers had faster convergence rates with eight having the fastest rate.
+However this difference in convergences is not pronounced amongst four, five and eight.
+This suggests that there is an optimal number of particles per swarm beyond which there is no significant advantage.
+
+We also observe that when these convergence rates are compared with DOPS, DOPS shows a very rapid drop when the swarm searches begin to saturate.
+The green region indicates when swarm search switches to a DDS based search.
+As we see in (Fig.
+) the error convergence rates tend to remain mostly flat for the purely swarm searches with different swarm numbers.
+However in DOPS as soon as convergence rate begins to saturate it automatically switches to a DDS phase which leads to a more pronounced drop.
+Thus this automated switching strategy appears to be crucial in leading the algorithm out of local optima or saddle point regions.
+
+\clearpage
+\section*{Materials and Methods}
+
 \subsection*{Optimization problem formulation.}
@@
 The decrease of weight function in equation 4, reflection boundary conditions in equation 5, the update function described in equations 6 and 7, the selection probability in equation 8 are the means by which DDS ideas are incorporated into DOPS.
-%Equation 5 describing the reflection boundary conditions, equations 6,7 and 8 are the means by which DDS ideas are incorporated into DOPS.
 The fraction of evaluations $\mathcal{FR}$  within the swarm phase is based on a switching strategy wherein the switch from swarm phase to DDS phase happens when the error due to the best solution does not drop more than 1\% of the original error, continuously for more than a prescribed number of function evaluations.
 This allows the solution to quickly jump out of local optima and avoid any convergence issues that are generally associated with swarm based searches.
-
-\clearpage
-
-\subsection*{Performance of DOPS on a model of the human coagulation cascade.}
-We compared the performance of DOPS with simulated annealing (SA), differential evolution (DE), and dynamically dimensioned search (DDS) for a model of the human coagulation cascade.
-Coagulation is an archetype biochemical network that is highly interconnected, containing both positive and negative feedback (Fig. \ref{fig-coagulation-network}).
-The biochemistry of coagulation, though complex, has been well studied \cite{mann2003dynamics,mann2003all,mann2003thrombin,vogler2009contact,diamond2013systems,fogelson2005coagulation,anand2003model},
-and reliable experimental protocols have been developed to interrogate the system \cite{hockin2002model,chatterjee2010systems,mann2006models,luan2007computationally}.
-Thus, the human coagulation cascade is an ideal test case;
-coagulation is challenging because it contains both fast and slow dynamics, but also accessible because of the availability of comprehensive data sets for model identification and validation.
-In this study, we used the coagulation model of Luan et al \cite{luan2007computationally}. Luan and coworkers modeled coagulation as a system of coupled non-linear ordinary differential equations.
-The Luan model contained 148 parameters and 92 species and has been validated using 21 published datasets.
-
-Coagulation is mediated by a family proteases in the circulation, called factors and a key group of blood cells, called platelets.
-The central process in coagulation is the conversion of prothrombin (fII), an inactive coagulation factor, to the master protease thrombin (FIIa).
-Thrombin generation involves three phases, initiation, amplification and termination [13,14].
-Initiation requires a trigger event, for example a vessel injury which exposes tissue factor (TF), which leads to the activation of factor VII (FVIIa) and the formation of the TF/FVIIa complex.
-Two converging pathways, the extrinsic and intrinsic cascades, then process and amplify this initial coagulation signal.
-There are several control points in the cascade that inhibit thrombin formation, and eventually terminate thrombin generation.
-Tissue Factor Pathway Inhibitor (TFPI) inhibits upstream activation events, while antithrombin III (ATIII) neutralizes several of the proteases generated during coagulation, including thrombin.
-Thrombin itself also inadvertently plays a role in its own inhibition; thrombin, through interaction with thrombomodulin, protein C and endothelial cell protein C receptor (EPCR), converts protein C to activated protein C (APC) which attenuates the coagulation response by proteolytic cleavage of amplification complexes.
-Termination occurs after either prothrombin is consumed, or thrombin formation is neutralized by inhibitors such as APC or ATIII.
-
-%Coagulation is regulated by a set of serine proteases known as coagulation factors and a type of blood cell called a platelet.
-%In the circulation, coagulation factors are generally in an inactive state known as a zymogen.
-%These zymogens are activated through certain triggers.
-Trigger events like injury or trauma or sepsis expose factors like collagen, tissue factor and von Willebrand factor (vWF) to blood.
-The exposure of these factors to blood kick-starts a series of convergent cascades that lead to conversion of zymogen prothrombinase to thrombin.
-For example when coagulation is initiated through the tissue factor pathway, tissue factor and activated factor VIIa (FVIIa) form a complex that activates factors X (fX) and IX (fIX).
-Activated factor X (fXa) thereafter activates downstream factors VIII and IX.
-The initial activation leads to the production of picomolar amounts of thrombin (fIIa) which activates platelets and amplifies its own production through the formation of a prothrombinase complex (FXa-FVa) on the surface of the activated platelets.
-Thrombin also  downregulates its own production by forming a complex with thrombomodulin which then activates protein C (PC).
-PC inhibits the formation of prothrombinase complex.
-In addition, Tissue factor pathway inhibitor (TFPI) downregulates FXa formation and activity by sequestering free FXa and TF-FVIIa in a FXa-dependent manner.
-Antithrombin III (ATIII)  inhibits all proteases.
-surface protein thrombomodulin (TM).
-%This choice of using a linear combination of two different error functions was motivated by poor validation results while using a single error function.
-
-To estimate the coagulation model parameters, we used data sets from TF/VIIa initiated coagulation in the absence of anticoagulants.
-The objective function was a linear combination of two error functions, representing coagulation initiated with different concentrations of TF-VIIa (5pM, 5nM) \cite{hockin2002model}.
-We restricted the number of function evaluations to $\mathcal{N}$ = 4000 for each algorithm we tested, and performed 25 trials of each experiment to collect average performance data (Table ZZ).DOPS converged faster and had a lower final error compared to the other algorithms (Fig.
-\ref{fig-convergence}).
-Within the first 1000 function evaluations of DOPS, there was a rapid drop in error.
-Approximately between 500-1000 function evaluations a switch to dynamically dimensioned search phase happens (this switch varied from trial to trial since the switch was based on the error from the swarm phase).
-Overall at the end of 4000 function evaluations DOPS minimized the error (final objective error is 5.5916e+04) to a greater extent than the other algorithms.
-
-Using the parameters obtained at the end of 4000 function evaluations we examined the 'fits' between models predictions and experimental data (Fig.\ref{fig-train}).
-The solid lines represent the mean value of prediction over 25 trials and the shaded region represents the 99\% confidence interval.
-Subsequently we used these optimal parameters to make model predictions that was compared against completely 'unseen' or untrained experimental data where
-coagulation was initiated with 500pM,50pM,10pM concentrations of TF-VIIa respectively (Fig.\ref{fig-validation}).
-
-\subsection*{Performance of DOPS on benchmark problems.}
-Villaverde and co-workers recently published a set of benchmark biochemical problems to evaluate parameter estimation methods \cite{villaverde2015biopredyn}.
-From a computational cost perspective problems they categorized the problems as most expensive, intermediate and least expensive.
-We evaluated the performance of our algorithms on a problem from the most expensive and least expensive categories.
-The first problem (most expensive) is a genome wide kinetic model of \textit{Saccharomyces cerevisiae} with 261 reactions, 262 variables and 1759 parameters (henceforth referred to as problem B1).
-The reactions were modeled using modular rate law and generalized form of Michaelis-Menten kinetics.
-The second problem (henceforth referred to as problem B4, least expensive) is a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters.
-In both cases pseudo time series data was generated by the authors.
-For problem B1, the time series data consisted of 44 observables and for problem B4 the data corresponded to 13 different metabolites.
-For a detailed description about the model architectures please refer to the work of Villaverde et al.
-\cite{villaverde2015biopredyn}
-
-We fixed the number of function evaluations at 4000 for DOPS and trained both the models against the pseudo experimental data.
-In both cases we found good fits  (Fig.
-\ref{fig-sims-b1} and Fig.
-\ref{fig-sims-b4}) to the problems within 4000 evaluations.
-We recaptured the 'nominal' parameters' in both cases within 4000 evaluations (Fig.
-\ref{fig-benchmark}).
-The final objective function value (Table \ref{table:benchmark-problems}) is an order smaller than nominal error for B1 possibly due to overfitting against the noise that was added to the synthetic data.
-Having obtained good fits on the coagulation problem and the benchmarks we proceeded to compare the performance of DOPS with other algorithms on commonly used test functions for global optimization.
-We used a 300 dimensional Rastrigin function and 300 dimensional Ackley function.
-Both Ackley and Rastrigin have multiple local minima and maxima and attain a global minimum value of 0.
-We tested the performance of the 4 different heuristic approaches on these two functions.
-In each experiment we again fixed the number of function evaluations at 4000 and ran 25 experiments.
-In both cases (Fig.
-\ref{fig-testfunctions}) we see that the error convergence rate for DOPS is much faster as compared to the other three heuristics and it finds the global minimum of 0 in both cases.
-In (Table \ref{table:benchmark-problems}) we summarize the results obtained on the coagulation model, the benchmark problems and the test functions.
-
-\subsection*{Influence of the number of sub-swarms on DOPS performance.}
-We attempted to quantify the influence of sub-swarms and DDS phase on the error convergence rates.
-To do so we considered only the multi swarm search without the DDS phase for 4000 function evaluations and 40 particles on the coagulation model.
-While doing the multi swarm search without DDS we varied the number of sub-swarms as one, two, four, five and eight.
-The number of sub-swarms were chosen so that each swarm contained the same number of particles.
-Hence we did not consider swarm numbers of three and seven.
-All the other parameters remained the same for each of these searches.
-We observed that the higher sub-swarm numbers had faster convergence rates with eight having the fastest rate.
-However this difference in convergences is not pronounced amongst four, five and eight.
-This suggests that there is an optimal number of particles per swarm beyond which there is no significant advantage.
-
-We also observe that when these convergence rates are compared with DOPS, DOPS shows a very rapid drop when the swarm searches begin to saturate.
-The green region indicates when swarm search switches to a DDS based search.
-As we see in (Fig.
-) the error convergence rates tend to remain mostly flat for the purely swarm searches with different swarm numbers.
-However in DOPS as soon as convergence rate begins to saturate it automatically switches to a DDS phase which leads to a more pronounced drop.
-Thus this automated switching strategy appears to be crucial in leading the algorithm out of local optima or saddle point regions.
 
@@
 
+\emph{Find best particle $\mathcal{G}$ amongst all sub-swarms}\;
+
 \eIf{$besterror(j)\geq 0.99*besterror(j+1)$} {
@@
 
- \If{$failurecounter\geq threshold$} {
- 	\emph{Run Dynamically Dimensioned Search with N-j function evaluations}\;
+	 \eIf{$failurecounter\geq threshold$} {
+
+		$\mathcal{G} \leftarrow DDS(\mathcal{G}, N-j)$\;
+		\Return $\mathcal{G}$
+
 	}
-
+	{
 		$j\leftarrow j+1$\;
 	}
-\caption{Dynamic Optimization with Particle Swarms}\label{algo_disjdecomp}
+			\Return $\mathcal{G}$
+}
+\caption{Dynamic Optimization with Particle Swarms}\label{algo_ddspso}
 \end{algorithm}\DecMargin{1em}
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure_1_Algorithm.pdf}
-\caption{Dynamic Optimization with Particle Swarms.\textbf{A}: Each particle represents an N dimensional parameter vector.
+\caption{Dynamic Optimization with Particle Swarms.
+\textbf{A}: Each particle represents an N dimensional parameter vector.
 Particles are given randomly generated initial solutions and grouped into different sub-swarms.
@@
 After every $\mathbf{g}$ number of function evaluations the particles are mixed and randomly assigned to different swarms.
-When the error due to the global best particle (best particle amongst all the sub-swarms - orange color) does not drop over a certain number of function evaluations, the swarm search is stopped and the search switches to a Dynamically Dimensioned Search with global best particle as the initial solution vector or candidate vector.
+When the error due to the global best particle (best particle amongst all the sub-swarms) does not drop over a certain number of function evaluations, the swarm search is stopped and the search switches to a Dynamically Dimensioned Search with global best particle as the initial solution vector or candidate vector.
 \textbf{B}: The candidate vector performs a greedy global search for the remaining number of function evaluations.
@@
 }\label{fig-algorithm}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.7\textwidth]{./figs/Figure_6_TestFunctions-crop.pdf}
+\caption{Performance of DOPS and other meta-heuristics for the Ackley and Rastrigin functions.
+\textbf{A}: Mean scaled error versus the number of function evaluations for the 10-dimensional Ackley function.
+DOPS, DDS and DE find optimal or near optimal solutions within the specified number of function evaluations.
+\textbf{B}: Mean scaled error versus the number of function evaluations for the 10-dimensional Rastrigin function.
+DOPS and DDS find optimal or near optimal solutions within the specified number of function evaluations.
+\textbf{C}: Mean scaled error versus the number of function evaluations for the 300-dimensional Rastrigin function.
+DOPS is the only algorithm that finds an optimal or near optimal solution within the specified number of function evaluations.
+In all cases, the maximum number of function evaluations was 4000.
+Mean and standard deviation were calculated over 25 trials.
+}\label{fig-testfunctions}
 \end{figure}
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure_3_Errors_convergence_v2.pdf}
-\caption{Error convergence rates of the 5 different algorithms on the coagulation model.
+\caption{Error convergence rates of the five different algorithms on the coagulation model.
 The objective error is the mean over $N$= 25 trials.
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure_4_Sim_Train_E1_E5.pdf}
-\caption{Model fits on experimental data using DDSMLSPSO.
+\caption{Model fits on experimental data using DOPS.
 The model parameters were estimated using DOPS.
@@
 }\label{fig-validation}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.7\textwidth]{./figs/Figure_6_TestFunctions-crop.pdf}
-\caption{Error convergence on 300 dimensional Ackley function and Rastrigin function.
-\textbf {(A)} Error convergence of 4 different meta-heuristics on 300 dimensional Ackley function \textbf {(B)} Error convergence of 4 different meta-heuristics on 300 dimensional Rastrigin function
-}\label{fig-testfunctions}
 \end{figure}

### 2016-04-14 Paper_v5.tex
Results and disucssion updates

@@
 \subsection*{DOPS minimized benchmark problems with a fixed number of function evaluations.}
-First, we tested DOPS and four other meta-heuristic approaches on the minimization of the Ackley and Rastrigin functions.
+We tested DOPS and four other meta-heuristic approaches on the minimization of the Ackley and Rastrigin functions.
 DOPS performed similarly or outperformed the other meta-heuristic approaches on both of these test functions (Fig. \ref{fig-testfunctions}).
 The Ackley and Rastrigin functions both have multiple local extrema and attain a global minimum value of zero.
-In each case, we fixed the maximum number of function evaluations at 4000 and ran 25 independent experiments with different initial parameter vectors.
+In each case, we fixed the maximum number of function evaluations at $\mathcal{N}$ = 4000 and ran 25 independent experiments with different initial parameter vectors.
 DOPS found optimal or near optimal solutions for both the 10-dimensional Ackley (Fig.
@@
 within the function evaluation budget (Fig. \ref{fig-testfunctions}B).
-Taken together, these results suggested that DOPS could perform at least as well as other meta-heuristic approaches on small test problems, but was especially suited to large dimension search spaces.
+Taken together, DOPS performed at least as well as other meta-heuristic approaches on small test problems, but was especially suited to large dimension search spaces.
 Next, we tested DOPS on benchmark biochemical models of varying complexity.
@@
 Villaverde and co-workers published a set of benchmark biochemical problems to evaluate parameter estimation methods \cite{villaverde2015biopredyn}.
-They ranked ordered these example problems from most expensive to least expensive.
+They ranked ordered these example problems by computational cost from most to least expensive.
 We evaluated the performance of DOPS on problems from the least and most expensive categories.
@@
 The expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters (henceforth referred to as problem B1).
-In both cases, synthetic time series data sets were generated using known parameter values.
+In both cases, synthetic time series data sets were generated using known parameter values, these data were then used by the parameter estimation methods.
 For problem B1, the time series data consisted of 44 observables, and for problem B4 the data corresponded to 13 different metabolite measurement sets.
-We fixed the number of function evaluations at 4000 for DOPS and trained both the models against the synthetic experimental data.
-In both cases we found good fits  (Fig. \ref{fig-sims-b1} and Fig. \ref{fig-sims-b4}) to the problems within 4000 evaluations.
-We recaptured the nominal parameter values within 4000 evaluations (Fig. \ref{fig-benchmark}).
+We fixed the number of function evaluations at $\mathcal{N}$ = 4000, and trained both models against the synthetic experimental data.
+In both cases we found good fits  (Fig. \ref{fig-sims-b1} and Fig. \ref{fig-sims-b4}) to the problems.
+We also recaptured the nominal parameter values within the specified number of function evaluations (Fig. \ref{fig-benchmark}).
+Thus, DOPS estimated the parameters in benchmark biochemical models, and recovered the original parameters from synthetic data.
+Next, we compared the performance of DOPS with the four other meta-heuristic approaches for a model of the human coagulation cascade.
 
@@
 
-\subsection*{Performance of DOPS on a model of the human coagulation cascade.}
-We compared the performance of DOPS with simulated annealing (SA), differential evolution (DE), and dynamically dimensioned search (DDS) for a model of the human coagulation cascade.
-Coagulation is an archetype biochemical network that is highly interconnected, containing both positive and negative feedback (Fig. \ref{fig-coagulation-network}).
+\subsection*{DOPS estimated the parameters of a human coagulation model.}
+Coagulation is an archetype biochemical network that is highly interconnected, containing both negative and positive feedback (Fig. \ref{fig-coagulation-network}).
 The biochemistry of coagulation, though complex, has been well studied \cite{mann2003dynamics,mann2003all,mann2003thrombin,vogler2009contact,diamond2013systems,fogelson2005coagulation,anand2003model},
 and reliable experimental protocols have been developed to interrogate the system \cite{hockin2002model,chatterjee2010systems,mann2006models,luan2007computationally}.
-Thus, the human coagulation cascade is an ideal test case;
-coagulation is challenging because it contains both fast and slow dynamics, but also accessible because of the availability of comprehensive data sets for model identification and validation.
-In this study, we used the coagulation model of Luan et al \cite{luan2007computationally}. Luan and coworkers modeled coagulation as a system of coupled non-linear ordinary differential equations.
-The Luan model contained 148 parameters and 92 species and has been validated using 21 published datasets.
-
 Coagulation is mediated by a family proteases in the circulation, called factors and a key group of blood cells, called platelets.
@@
 There are several control points in the cascade that inhibit thrombin formation, and eventually terminate thrombin generation.
-Tissue Factor Pathway Inhibitor (TFPI) inhibits upstream activation events, while antithrombin III (ATIII) neutralizes several of the proteases generated during coagulation, including thrombin.
-Thrombin itself also inadvertently plays a role in its own inhibition; thrombin, through interaction with thrombomodulin, protein C and endothelial cell protein C receptor (EPCR), converts protein C to activated protein C (APC) which attenuates the coagulation response by proteolytic cleavage of amplification complexes.
+Tissue Factor Pathway Inhibitor (TFPI) inhibits upstream activation events, while antithrombin III (ATIII) neutralizes several of the proteases generated during coagulation, including thrombin.Thrombin itself also inadvertently plays a role in its own inhibition; thrombin, through interaction with thrombomodulin, protein C and endothelial cell protein C receptor (EPCR), converts protein C to activated protein C (APC) which attenuates the coagulation response by proteolytic cleavage of amplification complexes.
 Termination occurs after either prothrombin is consumed, or thrombin formation is neutralized by inhibitors such as APC or ATIII.
+Thus, the human coagulation cascade is an ideal test case;
+coagulation is challenging because it contains both fast and slow dynamics, but also accessible because of the availability of comprehensive data sets for model identification and validation.
+In this study, we used the coagulation model of Luan et al \cite{luan2007computationally}, which is a coupled system of non-linear ordinary differential equations where biochemical interactions are modeled using mass action kinetics.
+The Luan model contained 148 parameters and 92 species and has been validated using 21 published experimental datasets.
 
@@
 %This choice of using a linear combination of two different error functions was motivated by poor validation results while using a single error function.
+%Using the parameters obtained at the end of 4000 function evaluations we examined the 'fits' between models predictions and experimental data
+%The solid lines represent the mean value of prediction over 25 trials and the shaded region represents the 99\% confidence interval.
+%Subsequently we used these optimal parameters to make model predictions that was compared against completely 'unseen' or untrained experimental data where
+%coagulation was  respectively .
 
-To estimate the coagulation model parameters, we used data sets from TF/VIIa initiated coagulation in the absence of anticoagulants.
-The objective function was a linear combination of two error functions, representing coagulation initiated with different concentrations of TF-VIIa (5pM, 5nM) \cite{hockin2002model}.
-We restricted the number of function evaluations to $\mathcal{N}$ = 4000 for each algorithm we tested, and performed 25 trials of each experiment to collect average performance data (Table ZZ).DOPS converged faster and had a lower final error compared to the other algorithms (Fig.
+DOPS estimated the parameters of a human coagulation model for TF/VIIa initiated coagulation without anticoagulants (Fig.\ref{fig-train}).
+The objective function was an unweighted linear combination of two error functions,
+representing coagulation initiated with different concentrations of TF/FVIIa (5pM, 5nM) \cite{hockin2002model}.
+We restricted the number of function evaluations to $\mathcal{N}$ = 4000 for each algorithm we tested, and performed 25 trials of each experiment to collect average performance data (Table \ref{table:benchmark-problems}).
+DOPS converged faster and had a lower final error compared to the other algorithms (Fig.
 \ref{fig-convergence}).
-Within the first 1000 function evaluations of DOPS, there was a rapid drop in error.
-Approximately between 500-1000 function evaluations a switch to dynamically dimensioned search phase happens (this switch varied from trial to trial since the switch was based on the error from the swarm phase).
-Overall at the end of 4000 function evaluations DOPS minimized the error (final objective error is 5.5916e+04) to a greater extent than the other algorithms.
+Within the first 25\% of function evaluations, DOPS produced a rapid drop in error followed by a slower but steady decline.
+Approximately between 500-1000 function evaluations DOPS switched to the dynamically dimensioned search phase, however this switch iteration varied from trial to trial since the switch was based on the error from the swarm phase.
+Overall, DOPS minimized the error to a greater extent than the other meta-heuristic approaches.
+However, it was unclear if the parameters obtained by DOPS had predictive power on unseen data. To address this question, we used the final parameters estimated by
+DOPS to simulate data that was not used in training (coagulation initiated with 500pM,50pM,10pM TF/VIIa).
+The optimal or near optimal parameters obtained by DOPS predicted unseen coagulation datasets (Fig.\ref{fig-validation}).
+Taken together, DOPS estimated parameter sets with predictive power on unseen coagulation data using fewer function iterations than other meta-heuristics.
+Next, we explored how the number of sub-swarms and the switch to DDS during the initial phase of an optimization trial influenced the performance of the approach.
 
-Using the parameters obtained at the end of 4000 function evaluations we examined the 'fits' between models predictions and experimental data (Fig.\ref{fig-train}).
-The solid lines represent the mean value of prediction over 25 trials and the shaded region represents the 99\% confidence interval.
-Subsequently we used these optimal parameters to make model predictions that was compared against completely 'unseen' or untrained experimental data where
-coagulation was initiated with 500pM,50pM,10pM concentrations of TF-VIIa respectively (Fig.\ref{fig-validation}).
-
-\subsection*{Influence of the number of sub-swarms on DOPS performance.}
-We attempted to quantify the influence of sub-swarms and DDS phase on the error convergence rates.
-To do so we considered only the multi swarm search without the DDS phase for 4000 function evaluations and 40 particles on the coagulation model.
-While doing the multi swarm search without DDS we varied the number of sub-swarms as one, two, four, five and eight.
-The number of sub-swarms were chosen so that each swarm contained the same number of particles.
+\subsection*{The number of sub-swarms and phase switching influenced DOPS performance.}
+A differentiating feature of DOPS compared to other meta-heuristics was the rapid error drop during the first 25\% of an optimization trial.
+This drop occurred during the swarm phase and was punctuated by the switch to the dynamically dimensioned search phase of the approach.
+We quantified the influence of the number of sub-swarms and switch to the DDS phase on error convergence by comparing DOPS with and without DDS for different numbers of sub-swarms (Fig.
+\ref{fig-sub-swarms}).
+We considered only the multi swarm search without the DDS phase for $\mathcal{N}$ = 4000 function evaluations on the coagulation model.
+While doing the multi swarm search without DDS, we used one, two, four, five and eight sub-swarms, with a total of 40 particles dived evenly amongst the swarms.
 Hence we did not consider swarm numbers of three and seven.
 All the other parameters remained the same for each of these searches.
-We observed that the higher sub-swarm numbers had faster convergence rates with eight having the fastest rate.
-However this difference in convergences is not pronounced amongst four, five and eight.
-This suggests that there is an optimal number of particles per swarm beyond which there is no significant advantage.
-
-We also observe that when these convergence rates are compared with DOPS, DOPS shows a very rapid drop when the swarm searches begin to saturate.
+Generally, the higher sub-swarm numbers had faster convergence.
+However, the difference in convergences was not pronounced amongst four, five and eight, suggesting
+there was an optimal number of particles per swarm beyond which there was no significant advantage.
+We also observed that convergence rates are compared with DOPS, DOPS shows a very rapid drop when the swarm searches begin to saturate.
 The green region indicates when swarm search switches to a DDS based search.
-As we see in (Fig.
-) the error convergence rates tend to remain mostly flat for the purely swarm searches with different swarm numbers.
+The error convergence rates tend to remain mostly flat for the purely swarm searches with different swarm numbers.
 However in DOPS as soon as convergence rate begins to saturate it automatically switches to a DDS phase which leads to a more pronounced drop.
-Thus this automated switching strategy appears to be crucial in leading the algorithm out of local optima or saddle point regions.
+Thus, this automated switching strategy appears to be crucial in leading the algorithm out of local optima or saddle point regions.
 
 \clearpage
+
+\section*{Discussion}
+In this study, we developed dynamic optimization with particle swarms (DOPS), a novel meta-heuristic for parameter estimation in models of biological systems.
+DOPS combined multi-swarm particle swarm optimization, a global search approach, with the greedy strategy of dynamically dimensioned search to estimate optimal or nearly optimal solutions in a fixed number of function evaluations.
+We tested the performance of DOPS and four widely used meta-heuristics on the Ackley and Rastrigin test functions, a set of biochemical benchmark problems and a model of the human coagulation cascade.
+As the number of parameters increased, DOPS outperformed the other meta-heuristics, generating optimal or nearly optimal solutions using significantly fewer function evaluations compared with the other methods.
+We tested the solutions generated by DOPS by comparing the estimated and true parameters in the benchmark studies, and by using the coagulation model to predict unseen experimental data.
+For both benchmark problems, DOPS retrieved the true parameters in significantly fewer function evaluations than other meta-heuristics.
+For the coagulation model, we used experimental coagulation measurements under two different conditions to estimate optimal or nearly optimal parameters.
+These parameters were then used to predict unseen coagulation data;
+the coagulation model parameters estimated by DOPS predicted the correct thrombin dynamics following TF/FVIIa induced coagulation without anticoagulants.
+Lastly, we showed the average performance of DOPS improved when combined with dynamically dimensioned search phase, compared to an identical multi-swarm approach alone.
+Taken together, DOPS is a promising meta-heuristic for the estimation of parameters in large biochemical models.
+
+Meta-heuristics are effective techniques to estimate optimal or nearly optimal solutions for complex, multi-modal functions.
+However, they often take an exorbitant number of objective function evaluations to come close to an optimum.
+When the objective function evaluations tend to become expensive it is infeasible to take up a large number of evaluations.
+As the dimensionality of parameter space increases, the search region gets widened and thus the problem becomes more challenging.
+In addition, most of these approaches require optimization of  'algorithm parameters' before the actual optimization and also involve computationally expensive update operations.
+Tolson and Shoemaker, through DDS, showed that randomly perturbing a subset of dimensions in high dimensional parameter space is an effective way to obtain near optimal solutions with few function evaluations.
+Though their approach is based on a single solution, the decision vector carries forward from one iteration to the next.
+Hence it behaves similar to a population based search.
+In our approach, we tried to eliminate the probability of starting from a bad region of search using a variant of particle swarm optimization.
+Particle Swarm Optimization (PSO) is a population based meta heuristic which does not have any complex operations like recombination, mutation or selection that are associated with other population based meta-heuristics like Differential Evolution (DE) or Genetic Algorithm (GA).
+Several particle swarm variants have been proposed to improve the search ability and rate of convergence, that involve different neighborhood structures, multi-swarms or adaptive parameters.
+Multi-swarm PSO with small particle neighborhoods have been shown to better in searching on complex multi-modal solutions \cite{zhao2008dynamic}.
+Multi swarm methods, in addition, avoid rapid convergence to a local optimum or stable point and are able to generate diverse solutions.
+Generation of diverse solutions in the early stage gives a better exploratory capability and thus converge of upon multiple optima.
+
+We utilized this capacity of multi swarms to generate diverse candidate solutions which can be used as initial solutions for a DDS like search.
+Thus the solution vectors obtained at the end of multi swarm search phase have a better propensity to not get stuck in bad regions in DDS phase as opposed to starting with a totally random initial solution.
+Given a fixed number of function evaluations, we have been able to show that we were able to obtain better solutions for the coagulation model and the test functions faster that other commonly known meta-heuristics and also DDS alone.
+The computational cost was prohibitively high for parameter estimation on this problem using a standard PSO.
+Thus we used SA, DE and DDS for the purpose of comparison.
+Choosing the number of function evaluations is largely a function of cost and complexity of the objective function.
+Traditionally the stopping condition for a parameter optimization problem can be the number of function evaluations, percentage of initial error achieved or an absolute error threshold.
+However in case of complex, expensive functions where we desire a value within a certain period of time, the number of evaluations are used as a stopping criterion.
+In our current study we used a value of 4000 which we based upon the time taken (approximately 8-10 seconds) for a single objective function evaluation in the coagulation case.
+We used the same value of 4000 for benchmarks published by Villaverde and co-workers \cite{villaverde2015biopredyn}.
+In this work by Villaverde et al.
+they used a population based search enhanced Scatter Search (eSS) to estimate the biochemical model parameters.
+Quite surprisingly we took a couple of orders lesser number of function evaluations 10\textsuperscript{3} to obtain the optimal parameter vector as compared to the enhanced Scatter Search (eSS) with a local optimizer, which took around the order of 10\textsuperscript{5} number of evaluations.
+The amount of CPU time taken (on an Intel Xeon processor 2.4 GHz) is lesser as compared to eSS on a similar architecture.
+A surprisingly remarkable aspect about our algorithm has been that we did not 'pre optimize' any parameters of the algorithm to suit a specific problem.
+In the swarm phase this includes the number of particles, number of sub-swarms, acceleration constants or the number of generations after which the particles are redistributed and the neighborhood perturbation parameter in DDS phase.
+We used the same parameters for all the problems.
+The same rule was applied to the rest of the meta-heuristics barring Simulated Annealing.
+For SA, we optimized the cooling schedule for the coagulation model.
+Thus, in this approach any overhead that usually comes with additional function evaluations to pre optimize parameters was avoided.
+
+The performance of DOPS was impressive given that it performed well on different complex systems with no pre optimization of algorithm parameters being required.
+We comfortably outperformed existing, widely used meta-heuristics and were also able to find minima of high dimensional global optimization test functions.
+Thus this approach may be well suited to large scale global optimization.
+In addition, surprisingly, we were able to obtain optimal parameter vectors for two different large scale systems biology models with a couple of orders fewer number of function evaluations as compared to enhanced Scatter Search (eSS).
+However it is quite possible that highly optimized versions of common meta heuristics may outperform us on these systems.
+This aspect is currently beyond the scope of this study.
+Our approach can also be combined with local derivative based searches to improve upon the accuracy of the solutions.
+In addition, the current implementation of the algorithm is designed to switch only once from the swarm phase to the DDS.
+In the DDS phase, the search uses only one candidate vector although there is a provision to start with a population of candidate vectors.
+Incorporating a more intelligent switching strategy that can do switch from swarm phase to DDS phase multiple times and having a population of candidate vectors are some aspects of the algorithm that can be studied further.
+
+\clearpage
+
 \section*{Materials and Methods}
@@
 \clearpage
-\section*{Discussion}
-Our study presents a novel approach for high-dimensional parameter estimation in complex biological systems with relatively few function evaluations.
-In this approach we combined a variant of a well known meta heuristic particle swarm optimization with Dynamically Dimensioned Search (DDS).
-We tested our approach on an ODE model of coagulation with 148 parameters and 92 species.
-Coagulation is an ideal system to test our approach since the biology is well known and complex, with multiple feed back loops that are tightly regulated.
-We used experimental data under different conditions to obtain optimal parameters and used these parameters to make predictions against unseen experimental data.
-We obtained good fits and made sufficiently accurate enough predictions using parameters obtained from 4000 function evaluations.
-Further, we also used high-dimensional forms of commonly used test functions of global optimization and showed that we were able to find the global minimum for 300 dimensional Ackley and Rastrigin functions faster than other meta-heuristics.
-We also considered two recently published benchmark problems to test parameter optimization approaches and showed that we were able to retrieve the nominal parameter vector within 4000 function evaluations.
-Meta-heuristic approaches are generally effective in finding close to optimum solutions of complex, multi-modal functions.
-In addition, they generally obviate the need for any \textit{a priori} knowledge (like function derivative).
-However they take an exorbitant number of objective function evaluations to come close to an optimum.
-When the objective function evaluations tend to become expensive it is infeasible to take up a large number of evaluations.
-As the dimensionality of parameter space increases, the search region gets widened and thus the problem becomes more challenging.
-In addition, most of these approaches require optimization of  'algorithm parameters' before the actual optimization and also involve computationally expensive update operations.
-Tolson and Shoemaker, through DDS, showed that randomly perturbing a subset of dimensions in high dimensional parameter space is an effective way to obtain near optimal solutions with few function evaluations.
-Though their approach is based on a single solution, the decision vector carries forward from one iteration to the next.
-Hence it behaves similar to a population based search.
-In our approach, we tried to eliminate the probability of starting from a bad region of search using a variant of particle swarm optimization.
-Particle Swarm Optimization (PSO) is a population based meta heuristic which does not have any complex operations like recombination, mutation or selection that are associated with other population based meta-heuristics like Differential Evolution (DE) or Genetic Algorithm (GA).
-Several particle swarm variants have been proposed to improve the search ability and rate of convergence, that involve different neighborhood structures, multi-swarms or adaptive parameters.
-Multi-swarm PSO with small particle neighborhoods have been shown to better in searching on complex multi-modal solutions \cite{zhao2008dynamic}.
-Multi swarm methods, in addition, avoid rapid convergence to a local optimum or stable point and are able to generate diverse solutions.
-Generation of diverse solutions in the early stage gives a better exploratory capability and thus converge of upon multiple optima.
-
-We utilized this capacity of multi swarms to generate diverse candidate solutions which can be used as initial solutions for a DDS like search.
-Thus the solution vectors obtained at the end of multi swarm search phase have a better propensity to not get stuck in bad regions in DDS phase as opposed to starting with a totally random initial solution.
-Given a fixed number of function evaluations, we have been able to show that we were able to obtain better solutions for the coagulation model and the test functions faster that other commonly known meta-heuristics and also DDS alone.
-The computational cost was prohibitively high for parameter estimation on this problem using a standard PSO.
-Thus we used SA, DE and DDS for the purpose of comparison.
- Choosing the number of function evaluations is largely a function of cost and complexity of the objective function.
-Traditionally the stopping condition for a parameter optimization problem can be the number of function evaluations, percentage of initial error achieved or an absolute error threshold.
-However in case of complex, expensive functions where we desire a value within a certain period of time, the number of evaluations are used as a stopping criterion.
-In our current study we used a value of 4000 which we based upon the time taken (approximately 8-10 seconds) for a single objective function evaluation in the coagulation case.
-We used the same value of 4000 for benchmarks published by Villaverde and co-workers \cite{villaverde2015biopredyn}.
-In this work by Villaverde et al.
-they used a population based search enhanced Scatter Search (eSS) to estimate the biochemical model parameters.
-Quite surprisingly we took a couple of orders lesser number of function evaluations 10\textsuperscript{3} to obtain the optimal parameter vector as compared to the enhanced Scatter Search (eSS) with a local optimizer, which took around the order of 10\textsuperscript{5} number of evaluations.
-The amount of CPU time taken (on an Intel Xeon processor 2.4 GHz) is lesser as compared to eSS on a similar architecture.
-
-A surprisingly remarkable aspect about our algorithm has been that we did not 'pre optimize' any parameters of the algorithm to suit a specific problem.
-In the swarm phase this includes the number of particles, number of sub-swarms, acceleration constants or the number of generations after which the particles are redistributed and the neighborhood perturbation parameter in DDS phase.
-We used the same parameters for all the problems.
-The same rule was applied to the rest of the meta-heuristics barring Simulated Annealing.
-For SA, we optimized the cooling schedule for the coagulation model.
-Thus, in this approach any overhead that usually comes with additional function evaluations to pre optimize parameters was avoided.
-
-The performance of our approach seems impressive given that it performed well on different complex systems with no pre optimization of algorithm parameters being required.
-We comfortably outperformed existing, widely used meta-heuristics and were also able to find minima of high dimensional global optimization test functions.
-Thus this approach may be well suited to large scale global optimization.
-In addition, surprisingly, we were able to obtain optimal parameter vectors for two different large scale systems biology models with a couple of orders fewer number of function evaluations as compared to enhanced Scatter Search (eSS).
-However it is quite possible that highly optimized versions of common meta heuristics may outperform us on these systems.
-This aspect is currently beyond the scope of this study.
-Our approach can also be combined with local derivative based searches to improve upon the accuracy of the solutions.
-In addition, the current implementation of the algorithm is designed to switch only once from the swarm phase to the DDS.
-In the DDS phase, the search uses only one candidate vector although there is a provision to start with a population of candidate vectors.
-Incorporating a more intelligent switching strategy that can do switch from swarm phase to DDS phase multiple times and having a population of candidate vectors are some aspects of the algorithm that can be studied further.
 
@@
 \centering
-\includegraphics[width=0.7\textwidth]{./figs/Figure_6_TestFunctions-crop.pdf}
+\includegraphics[width=0.7\textwidth]{./figs/Figure_2_TestFunctions-crop.pdf}
 \caption{Performance of DOPS and other meta-heuristics for the Ackley and Rastrigin functions.
@@
 \centering
-\includegraphics[width=1.00\textwidth]{./figs/Figure_2_Coagulation_v4.pdf}
+\includegraphics[width=1.00\textwidth]{./figs/Figure_3_Coagulation_v4.pdf}
 \caption{Schematic of the extrinsic and intrinsic coagulation cascade\cite{luan2007computationally}.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure_3_Errors_convergence_v2.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure_4_Errors_convergence_v2.pdf}
 \caption{Error convergence rates of the five different algorithms on the coagulation model.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure_4_Sim_Train_E1_E5.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure_5_Sim_Train_E1_E5.pdf}
 \caption{Model fits on experimental data using DOPS.
 The model parameters were estimated using DOPS.
-Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from $N$= 25 trials.
+Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from 25 trials.
 The grey shaded region represents the 99\% confidence estimate of the mean simulated thrombin concentration.
 The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers.
-Thrombin generation is initiated by adding Factor VIIa-TF (5nM - Red and 5pM - Green) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
+Thrombin generation is initiated by adding Factor TF/VIIa (5nM and 5pM) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
 }\label{fig-train}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure_5_Sim_Validate_E2_E4_E6.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure_6_Sim_Validate_E2_E4_E6.pdf}
 \caption{Model predictions on unseen experimental data using parameters obtained from DOPS.
@@
 
-\begin{figure}[ht]
+\clearpage
+
+\begin{figure}[h]
 \centering
-\includegraphics[width=1.00\textwidth]{./figs/Figure_B1_B4_params_measuredstates}
-\caption{ \textbf {(A)} Difference between nominal and optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
-\textbf {(B)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
-\textbf {(C)} Difference between nominal and optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
-\textbf {(D)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
-}\label{fig-benchmark}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.00\textwidth]{./figs/B1_sims.pdf}
-\caption{\textbf {(Data fits for Problem B1)} Pseudo-experimental data (red circles) vs.
-optimal solution obtained using DOPS (solid blue lines) for the 44 observed states.
-X axis: time [s]; Y axis: metabolite concentrations [mM].
-}\label{fig-sims-b1}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.00\textwidth]{./figs/FigB4_sims.pdf}
-\caption{\textbf {(Data fits for Problem B4)} Pseudo-experimental data (red x) vs.
-optimal solution obtained using DOPS (solid blue lines) for the 13 observed states.
-X axis: time [s]; Y axis: metabolite concentrations [mM].
-}\label{fig-sims-b4}
+\includegraphics[width=1.0\textwidth]{./figs/Figure_7_ErrorSwarms_v2.pdf}
+\caption{Model predictions on unseen experimental data using parameters obtained from DOPS.
+The parameter estimates that were obtained using DOPS were tested against data that was not used in the model training.
+Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from $N$= 25 trials.
+The grey shaded region represents the 99\% confidence estimate of the mean simulated thrombin concentration.
+The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers.
+Thrombin generation is initiated by adding Factor VIIa-TF (500pM - Blue, 50pM - Pink and 10pM - Orange respectively) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
+}\label{fig-sub-swarms}
 \end{figure}
@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth]{./figs/Figure_S1_B1_sims.pdf}
+\caption{\textbf {(Data fits for Problem B1)} Pseudo-experimental data (red circles) vs.
+optimal solution obtained using DOPS (solid blue lines) for the 44 observed states.
+X axis: time [s]; Y axis: metabolite concentrations [mM].
+}\label{fig-sims-b1}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth]{./figs/Figure_S2_B4_sims.pdf}
+\caption{\textbf {(Data fits for Problem B4)} Pseudo-experimental data (red x) vs.
+optimal solution obtained using DOPS (solid blue lines) for the 13 observed states.
+X axis: time [s]; Y axis: metabolite concentrations [mM].
+}\label{fig-sims-b4}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth]{./figs/Figure_S3_B1_B4_params_measuredstates.pdf}
+\caption{ \textbf {(A)} Difference between nominal and optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
+\textbf {(B)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
+\textbf {(C)} Difference between nominal and optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
+\textbf {(D)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
+}\label{fig-benchmark}
+\end{figure}
+
 \end{document}

### 2016-04-15 Paper_v5.tex
Saftery commit

@@
 
+%As the dimensionality of  increases, the search region gets widened and thus the problem becomes more challenging.
+%
+
 \section*{Discussion}
@@
 
-Meta-heuristics are effective techniques to estimate optimal or nearly optimal solutions for complex, multi-modal functions.
-However, they often take an exorbitant number of objective function evaluations to come close to an optimum.
-When the objective function evaluations tend to become expensive it is infeasible to take up a large number of evaluations.
-As the dimensionality of parameter space increases, the search region gets widened and thus the problem becomes more challenging.
-In addition, most of these approaches require optimization of  'algorithm parameters' before the actual optimization and also involve computationally expensive update operations.
+Meta-heuristics can be effective techniques to estimate optimal or nearly optimal solutions for complex, multi-modal functions.
+However, they often require an exorbitant number of function evaluations to approach the optimum.
+Additionally, meta-heuristics require optimization of  algorithm parameters before the actual optimization, and also involve costly update operations to generate parameter diversity.
+Thus, for expensive objective functions in complex models with high dimension parameter spaces, meta-heuristics can be inefficient.
+Particle Swarm Optimization (PSO) is a population based meta heuristic which does not have any complex operations like recombination, mutation or selection that are associated with other population based meta-heuristics like differential evolution or genetic algorithms.
+Several particle swarm variants have been proposed to improve the search ability and rate of convergence.
+These variations involve different neighborhood structures, multi-swarms or adaptive parameters.
+Multi-swarm PSO with small particle neighborhoods have been shown to better in searching on complex multi-modal solutions \cite{zhao2008dynamic}.
+Multi swarm methods, in addition, avoid rapid convergence to a local optimum or stable point and are able to generate diverse solutions.
+Generation of diverse solutions in the early stage gives a better exploratory capability and thus converge of upon multiple optima.
 Tolson and Shoemaker, through DDS, showed that randomly perturbing a subset of dimensions in high dimensional parameter space is an effective way to obtain near optimal solutions with few function evaluations.
@@
 In our approach, we tried to eliminate the probability of starting from a bad region of search using a variant of particle swarm optimization.
-Particle Swarm Optimization (PSO) is a population based meta heuristic which does not have any complex operations like recombination, mutation or selection that are associated with other population based meta-heuristics like Differential Evolution (DE) or Genetic Algorithm (GA).
-Several particle swarm variants have been proposed to improve the search ability and rate of convergence, that involve different neighborhood structures, multi-swarms or adaptive parameters.
-Multi-swarm PSO with small particle neighborhoods have been shown to better in searching on complex multi-modal solutions \cite{zhao2008dynamic}.
-Multi swarm methods, in addition, avoid rapid convergence to a local optimum or stable point and are able to generate diverse solutions.
-Generation of diverse solutions in the early stage gives a better exploratory capability and thus converge of upon multiple optima.
-
 We utilized this capacity of multi swarms to generate diverse candidate solutions which can be used as initial solutions for a DDS like search.

### 2016-04-15 Paper_v5.tex
More discussion updates

@@
 
+%However, they often require a large number of function evaluations to approach the optimum,
+%the optimization of algorithm parameters before the actual optimization, and costly operations to generate parameter diversity.
+%Thus, for expensive objective functions in high dimensional parameter spaces, meta-heuristic approaches can be inefficient.
+
 Meta-heuristics can be effective techniques to estimate optimal or nearly optimal solutions for complex, multi-modal functions.
-However, they often require an exorbitant number of function evaluations to approach the optimum.
-Additionally, meta-heuristics require optimization of  algorithm parameters before the actual optimization, and also involve costly update operations to generate parameter diversity.
-Thus, for expensive objective functions in complex models with high dimension parameter spaces, meta-heuristics can be inefficient.
-Particle Swarm Optimization (PSO) is a population based meta heuristic which does not have any complex operations like recombination, mutation or selection that are associated with other population based meta-heuristics like differential evolution or genetic algorithms.
+DOPS is a combination of particle swarm optimization, which is a global search method, and dynamically dimensioned search, which is a greedy evolutionary technique.
+Particle swarm optimization is a population based meta-heuristic which uses collective information shared amongst swarms of computational particles to
+search for global extrema.
 Several particle swarm variants have been proposed to improve the search ability and rate of convergence.

### 2016-04-25 Paper_v5.tex
Updated

@@
 interconnected through nonlinear enzyme catalyzed reactions.
-Mathematical modeling has evolved as a powerful paradigm to analyze, and ultimately design these complex networks \cite{assmus2006dynamics, Riel:2006aa,Jaqaman:2006aa,kitano2002systems,hood2004systems}.
+Mathematical modeling has evolved as a powerful paradigm to analyze, and ultimately design these complex networks \cite{assmus2006dynamics, Riel2006aa,Jaqaman:2006aa,kitano2002systems,hood2004systems}.
 Mathematical modeling of biochemical networks is often an iterative process.
@@
 
-Meta-heuristic stochastic optimization approaches like Genetic Algorithms (GAs) [ADI-REF], Simulated Annealing (SA) \cite{kirkpatrick1983optimization},
-Evolutionary Programming [ADI-REF] and population based searches like Differential Evolution (DE) \cite{storn1997differential} have shown promise on nonlinear multi-modal problems \cite{sun2012parameter}.
+Meta-heuristic stochastic optimization approaches like Genetic Algorithms (GAs) \cite{goldberg2006genetic}, Simulated Annealing (SA) \cite{kirkpatrick1983optimization},
+Evolutionary Programming \cite{fogel2009artificial} and population based searches like Differential Evolution (DE) \cite{storn1997differential} have shown promise on nonlinear multi-modal problems \cite{sun2012parameter}.
 These techniques do not make any assumptions about the structure of objective function nor do they require \textit{a~priori} information about the objective function.
@@
 Mendes et al. \cite{mendes1998non} used simulated annealing to estimate rate constants for the irreversible inhibition of HIV proteinase,
-Modchang et al. \cite{modchang2008mathematical} used a genetic algorithm to estimate parameters for a model of signal transduction [ADI-EXPAND],
-differential evolution approaches have also been effective on various biological problems \cite{tsai2005evolutionary,wang2001hybrid,noman2007inferring}.
+Modchang et al.
+\cite{modchang2008mathematical} used a genetic algorithm (GA) to estimate parameters for a model of G-protein-coupled receptors(GPCRs) mediated signal transduction.
+The parameters obtained using GA helped qualitatively ascertain the efficacy and potency of two G-protein agonists - CPA and NECA.
+Differential evolution approaches have also been effective on various biological problems \cite{tsai2005evolutionary,wang2001hybrid,noman2007inferring}.
 Tashkova et al. compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis and showed that DE was the most effective \cite{tashkova2011parameter}.
@@
 DOPS outperformed common meta-heuristic approaches like Differential Evolution (DE), Simulated Annealing (SA) and dynamically dimensioned search (DDS) on both the test functions and the coagulation model.
-DOPS also performed well on the benchmark problems where it outperformed enhanced scatter search (eSS) and recovered the nominal parameters with only 4000 function evaluations ([ADI-compared with what?])
-across all the benchmark problems considered.
+We also compared the performance of DOPS against enhanced scatter search (eSS) on two benchmark problems published by Villaverde and co-workers\cite{villaverde2015biopredyn}.
+We fixed the number of function evaluations to 4000 for both the benchmark problems without pre-optimizing any algorithmic parameters.
+DOPS recovered the nominal parameters and surprisingly did so faster than eSS which took nearly an order of function evaluations more than DOPS in both the cases.
 Taken together, these studies suggest DOPS is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.
@@
 Toward this objective, DOPS begins by using a particle swarm search and then dynamically switches, using an adaptive switching criteria, to the DDS search phase.
-[FINISH GENERIC DESCRIPTION OF APPROACH]
+The particle swarm search uses multiple sub-swarms wherein each solution represented by a particle is updated using Equation \ref{eqn:update-rule}.
+The particle update is influenced by the best particle amongst all particles in the sub-swarm and the best solution found by the particle till the current iteration.
+This update rule however differs from a conventional PSO update wherein there are no velocity terms involved.
+The particle updates continue to happen within the sub-swarms for a certain number of iterations (or function evaluations) after which the sub-swarms are reorganized which is similar to the regrouping strategy described by Zhao et al.
+\cite{zhao2008dynamic}.
+DOPS switches out of this PSO phase based on an adaptive switching criteria that is a function of rate of error convergence.
+If the error represented by the best particle does not drop for a threshold number of function evaluations, DOPS switches automatically to the DDS search phase.
+The DDS search uses the best particle from the PSO search phase as its initial solution or candidate vector.
+Thereafter, the particle is greedily updated by perturbing a subset of dimensions for the remaining number of function evaluations.
+The number of dimensions perturbed is a monotonically decreasing function that generally depends on the number of function evaluations within the DDS phase.
 
@@
 We tested DOPS and four other meta-heuristic approaches on the minimization of the Ackley and Rastrigin functions.
-DOPS performed similarly or outperformed the other meta-heuristic approaches on both of these test functions (Fig. \ref{fig-testfunctions}).
+DOPS performed similarly or outperformed the other meta-heuristic approaches on both of these test functions (Fig.\ref{fig-testfunctions}).
 The Ackley and Rastrigin functions both have multiple local extrema and attain a global minimum value of zero.
@@
 Equation (3) is similar to the general particle swarm update rule, however, it does not contain velocity terms.
-In DOPS the parameter $\theta_{1,j-1}$ depends upon the function evaluations and is updated according to:
+In DOPS, the parameter $\theta_{1,j-1}$ is similar to the inertia weight parameter for the velocity term described by Shi and Eberhart \cite{shi1999empirical}.
+In this study Shi and Eberhart propose a linearly decreasing inertia weight to improve convergence properties of PSO.
+$\theta_{1,j-1}$ is an analogous equivalent to the inertia weight and uses the same rule described by Shi and Eberhart to adaptively change with the number of function evaluations and is updated according to:
 \begin{eqnarray}
-	\mathbf \theta_{1,j}&=&\frac{(\mathcal{N}-{j})*(\mathbf{w}_{max}-\mathbf{w}_{min}))}{(\mathcal{N}-{1})} + \mathbf{w}_{min}
+	\mathbf \theta_{1,j}&=&\frac{(\mathcal{N}-{j})*({w}_{max}-{w}_{min}))}{(\mathcal{N}-{1})} + {w}_{min}
 \end{eqnarray}
-where $\mathcal{N}$ represents the total number of function evaluations, $\mathbf{w}_{max}$ and $\mathbf{w}_{min}$ are the maximum and minimum inertia weights, respectively.
+where $\mathcal{N}$ represents the total number of function evaluations, ${w}_{max}$ and ${w}_{min}$ are the maximum and minimum inertia weights, respectively.
 While updating the particles, we made sure all dimensions of the solution represented by the particle were within bounds using a set of reflection boundary conditions:
@@
 \begin{equation}
-  \mathcal{G}_{new}(\mathbf{J})=\begin{cases}
+  \mathcal{G}_{new}({J})=\begin{cases}
     \mathcal{G}(\mathbf{J})+\mathbf{r}_{normal}(\mathbf{J})\sigma(\mathbf{J}), & \text{if $\mathcal{G}_{new}(\mathbf{J})<\mathcal{G}(\mathbf{J})$}.\\
@@
 \end{equation}
-where $\mathbf{J}$ represents the set containing the specific dimensions being perturbed, ${r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$,
+where $\mathbf{J}$ is a vector representing the subset of dimensions that are being perturbed, $\mathbf{r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$,
 and $\sigma$ denotes the perturbation amplitude:
 \begin{equation}
-	\sigma = \mathbf{R}(\mathcal{MAX} -\mathcal{MIN})
+	\sigma = {R}(\mathbf{p}^U - \mathbf{p}^L)
 \end{equation}
-where $\mathbf{R}$ is the scalar perturbation size parameter, $\mathcal{MAX}$ and $\mathcal{MIN}$ are ($\mathcal{K}\times{1}$) vectors that represent the maximum and minimum bounds on each dimension.
+where ${R}$ is the scalar perturbation size parameter, $\mathbf{p}^U$ and $\mathbf{p}^L$ are ($\mathcal{K}\times{1}$) vectors that represent the maximum and minimum bounds on each dimension.
 The set $\mathbf{J}$ was constructed using a probability function $\mathcal{P}_{j}$ that represented the probability whether a specific dimension $j$ was perturbed or not.
-T
-his function is a monotonically decreasing function that decreases with the number of function evaluations.
+This function is a monotonically decreasing function that decreases with the number of function evaluations.
 $\mathcal{P}_{j}$ can be any monotonically decreasing function, in our approach we used the following function:
@@
 	%\mathcal{P}_{j}&=&{1}-\log(j/ (\beta({1}-\mathcal{FR})*\mathbf{N}))
-	\mathcal{P}_{j}&=&{1}-\log(j/ (({1}-\mathcal{FR})*\mathbf{N}))
+	\mathcal{P}_{j}&=&{1}-\log(i/ (({1}-\mathcal{FR})*\mathbf{N}))
 \end{eqnarray}
 %where $\beta$ is the perturbation frequency probability modulator.
-Thus the number of dimensions of the candidate vector that are updated or perturbed decreases with the as the number of function evaluations increase.
+where $i$ is the current iteration. Thus the probability that a dimension $j$ is perturbed decreases as the number of iterations or function evaluations indicated by $i$ increase.
+%Thus the number of dimensions of the candidate vector that are updated or perturbed decreases with the as the number of function evaluations increase. T
 These updates are greedy in nature, so $\mathcal{G}_{new}$ becomes the new solution vector only if it is better than the old one $\mathcal{G}$.
-The decrease of weight function in equation 4, reflection boundary conditions in equation 5, the update function described in equations 6 and 7, the selection probability in equation 8 are the means by which DDS ideas are incorporated into DOPS.
+The reflection boundary conditions in equation 5, the update function described in equations 6 and 7, the selection probability in equation 8 are the means by which DDS ideas are incorporated into DOPS.
 The fraction of evaluations $\mathcal{FR}$  within the swarm phase is based on a switching strategy wherein the switch from swarm phase to DDS phase happens when the error due to the best solution does not drop more than 1\% of the original error, continuously for more than a prescribed number of function evaluations.
@@
 \begin{center}
-\begin{tabular}{ r|l }
+\begin{tabular}{ r|ll}
 \hline\\
-TF/FVIIa concentration& Normalized S.E.  \\\\
+TF/FVIIa concentration& Normalized S.E.& Data Set Category\\\\
 \hline\\
-5~nM & 0.0376 \\
-500~pM & 0.0564\\
-50~pM & 0.1125 \\
-10~pM & 0.0823 \\
-5~pM & 0.0338 \\\\
+5~nM & 0.0376 & Training \\
+500~pM & 0.0564 & Validation\\
+50~pM & 0.1125 & Validation\\
+10~pM & 0.0823 & Validation\\
+5~pM & 0.0338 & Training\\\\
 \hline
@@
 For each problem the bounds on the parameter vector, the total number of function evaluations, the best initial objective value and the best final objective value are specified.
+Here pnom indicates the nominal or true parameter vector of the model.
 Nominal objective value represents the objective value using the true parameter vector or the nominal parameter vector.

### 2016-04-25 Paper_v5.tex
Updated

@@
 Toward this objective, DOPS begins by using a particle swarm search and then dynamically switches, using an adaptive switching criteria, to the DDS search phase.
-The particle swarm search uses multiple sub-swarms wherein each solution represented by a particle is updated using Equation \ref{eqn:update-rule}.
+The particle swarm search uses multiple sub-swarms wherein each solution represented by a particle is updated using equation \ref{eqn:update-rule}.
 The particle update is influenced by the best particle amongst all particles in the sub-swarm and the best solution found by the particle till the current iteration.
@@
 \cite{zhao2008dynamic}.
-DOPS switches out of this PSO phase based on an adaptive switching criteria that is a function of rate of error convergence.
+DOPS switches out of this PSO phase based on an adaptive switching criteria that is a function of error convergence rate.
 If the error represented by the best particle does not drop for a threshold number of function evaluations, DOPS switches automatically to the DDS search phase.
@@
 In this study Shi and Eberhart propose a linearly decreasing inertia weight to improve convergence properties of PSO.
-$\theta_{1,j-1}$ is an analogous equivalent to the inertia weight and uses the same rule described by Shi and Eberhart to adaptively change with the number of function evaluations and is updated according to:
+$\theta_{1,j-1}$ is inspired from this idea and also from the idea of decreasing perturbation probability proposed by Tolson and Shoemaker \cite{tolson2007dynamically}.
+It is an analogous equivalent to inertia weight on velocity.
+However $\theta_{1,j-1}$ places inertia on the position rather than velocity and uses the same rule described by Shi and Eberhart to adaptively change with the number of function evaluations.
+It is updated according to:
 \begin{eqnarray}
@@
 \centering
-\caption{Error Analysis.}
+\caption[.]{Error Analysis.
+Normalized standard error showing the agreement between coagulation model dynamics and the coagulation experimental data.
+The model is trained on the 5~nM and the 5~pM cases to obtain the optimal parameters.
+Using these optimal parameters, dynamics of coagulation are predicted with varying initiator concentrations (500~pM, 50~pM and 10~pM).
+The normalized S.E.
+is defined as \( N.S.E.=(1/max(\mathbf{X}))*(\lVert (\mathbf{Y},\mathbf{X}) \rVert /sqrt$(N)$) \) where \textbf{X} is the experimental data, \textbf{Y} is the model simulation data interpolated on experimental time and $N$ is the total number of experimental time points.}
 \label{table:Error-Analysis}
@@
 5~nM & 0.0376 & Training \\
-500~pM & 0.0564 & Validation\\
-50~pM & 0.1125 & Validation\\
-10~pM & 0.0823 & Validation\\
+500~pM & 0.0564 & Prediction\\
+50~pM & 0.1125 & Prediction\\
+10~pM & 0.0823 & Prediction\\
 5~pM & 0.0338 & Training\\\\
@@
 For each problem the bounds on the parameter vector, the total number of function evaluations, the best initial objective value and the best final objective value are specified.
-Here pnom indicates the nominal or true parameter vector of the model.
+Here $pnom$ indicates the nominal or true parameter vector of the model.
 Nominal objective value represents the objective value using the true parameter vector or the nominal parameter vector.
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure_7_ErrorSwarms_v2.pdf}
-\caption{Model predictions on unseen experimental data using parameters obtained from DOPS.
-The parameter estimates that were obtained using DOPS were tested against data that was not used in the model training.
-Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from $N$= 25 trials.
-The grey shaded region represents the 99\% confidence estimate of the mean simulated thrombin concentration.
-The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers.
-Thrombin generation is initiated by adding Factor VIIa-TF (500pM - Blue, 50pM - Pink and 10pM - Orange respectively) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
+\caption{DOPS begins by using a particle swarm search (color:pale brown) and then dynamically switches (color:green), using an adaptive switching criteria, to the DDS search phase (color:white).
+The particle swarm search uses multiple sub-swarms wherein each solution represented by a particle is updated using equation \ref{eqn:update-rule}.
+The particle update is influenced by the best particle amongst all particles in the sub-swarm and the best solution found by the particle till the current iteration.
+This update rule however contains no velocity terms.
+The particle updates continue to happen within the sub-swarms for a certain number of iterations (or function evaluations) after which the sub-swarms are reorganized which is similar to the regrouping strategy described by Zhao et al.
+\cite{zhao2008dynamic}.
+DOPS switches out of this PSO phase based on an adaptive switching criteria (color:green) that is a function of error convergence rate.
+If the error represented by the best particle does not drop for a threshold number of function evaluations, DOPS switches automatically to the DDS search phase.
+Since the algorithm is stochastic, the switch out can happen at a different function evaluation in each trial.
+The DDS search then uses the best particle from the PSO search phase as its initial solution or candidate vector.
+This particle is greedily updated by perturbing a subset of dimensions for the remaining number of function evaluations.
+The number of dimensions perturbed is a monotonically decreasing function that generally depends on the number of function evaluations within the DDS phase.
 }\label{fig-sub-swarms}

### 2016-04-25 Paper_v5.tex
Equation 7 changed

@@
 We considered only the multi swarm search without the DDS phase for $\mathcal{N}$ = 4000 function evaluations on the coagulation model.
-While doing the multi swarm search without DDS, we used one, two, four, five and eight sub-swarms, with a total of 40 particles dived evenly amongst the swarms.
+While doing the multi swarm search without DDS, we used one, two, four, five and eight sub-swarms, with a total of 40 particles divided evenly amongst the swarms.
 Hence we did not consider swarm numbers of three and seven.
@@
 \subsubsection*{Phase 2: DDS phase.}
-At the conclusion of the swarm phase, the overall best particle, $\mathcal{G}_{k}$, over the k sub-swarms was to initalize the DDS phase.
+At the conclusion of the swarm phase, the overall best particle, $\mathcal{G}_{k}$, over the k sub-swarms was to initialize the DDS phase.
 DOPS takes $\left(1-\mathcal{F}\right)*\mathcal{N}$ function evaluations in the DDS phase and then terminates the search.
@@
 where ${R}$ is the scalar perturbation size parameter, $\mathbf{p}^U$ and $\mathbf{p}^L$ are ($\mathcal{K}\times{1}$) vectors that represent the maximum and minimum bounds on each dimension.
-The set $\mathbf{J}$ was constructed using a probability function $\mathcal{P}_{j}$ that represented the probability whether a specific dimension $j$ was perturbed or not.
+The set $\mathbf{J}$ was constructed using a probability function $\mathcal{P}_{i}$ that represents a threshold probability for determining whether a specific dimension $j$ was perturbed or not.
 This function is a monotonically decreasing function that decreases with the number of function evaluations.
-$\mathcal{P}_{j}$ can be any monotonically decreasing function, in our approach we used the following function:
+$\mathcal{P}_{i}$ can be any monotonically decreasing function, in our approach we used the following function:
 \begin{eqnarray}
 	%\mathcal{P}_{j}&=&{1}-\log(j/ (\beta({1}-\mathcal{FR})*\mathbf{N}))
-	\mathcal{P}_{j}&=&{1}-\log(i/ (({1}-\mathcal{FR})*\mathbf{N}))
+	\mathcal{P}_{i}&=&{1}-\log(i/ (({1}-\mathcal{F})*\mathbf{N}))
 \end{eqnarray}
 %where $\beta$ is the perturbation frequency probability modulator.
-where $i$ is the current iteration. Thus the probability that a dimension $j$ is perturbed decreases as the number of iterations or function evaluations indicated by $i$ increase.
+where $i$ is the current iteration.
+After $\mathcal{P}_{i}$ is determined, we assign a probability $\mathcal{P}_{j}$ from a uniform distribution for each dimension $j$.
+If $\mathcal{P}_{j}$ is less than $\mathcal{P}_{i}$ the dimension is included in \textbf{J}.
+Thus the probability that a dimension $j$ is perturbed decreases as the number of iterations or function evaluations indicated by $i$ increase.
 %Thus the number of dimensions of the candidate vector that are updated or perturbed decreases with the as the number of function evaluations increase. T
@@
 The reflection boundary conditions in equation 5, the update function described in equations 6 and 7, the selection probability in equation 8 are the means by which DDS ideas are incorporated into DOPS.
-The fraction of evaluations $\mathcal{FR}$  within the swarm phase is based on a switching strategy wherein the switch from swarm phase to DDS phase happens when the error due to the best solution does not drop more than 1\% of the original error, continuously for more than a prescribed number of function evaluations.
+The fraction of evaluations $\mathcal{F}$  within the swarm phase is based on a switching strategy wherein the switch from swarm phase to DDS phase happens when the error due to the best solution does not drop more than 1\% of the original error, continuously for more than a prescribed number of function evaluations.
 This allows the solution to quickly jump out of local optima and avoid any convergence issues that are generally associated with swarm based searches.
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure_7_ErrorSwarms_v2.pdf}
-\caption{DOPS begins by using a particle swarm search (color:pale brown) and then dynamically switches (color:green), using an adaptive switching criteria, to the DDS search phase (color:white).
+\caption{DOPS (color:black) begins by using a particle swarm search (color:pale brown) and then dynamically switches (color:green), using an adaptive switching criteria, to the DDS search phase (color:white).
 The particle swarm search uses multiple sub-swarms wherein each solution represented by a particle is updated using equation \ref{eqn:update-rule}.
@@
 The number of dimensions perturbed is a monotonically decreasing function that generally depends on the number of function evaluations within the DDS phase.
+We compared the performance of DOPS with multi-swarm searches without DDS to quantify the effect of number of sub-swarms.
+We used one, two, four, five and eight sub-swarms, with a total of 40 particles divided evenly amongst the swarms.
+The convergence rates with higher swarm numbers is typically higher but there is no pronounced difference amongst four, five and eight.
+When the multi-swarm only searches tend to saturate DOPS shows a rapid drop due to a switch to  the DDS phase.
 }\label{fig-sub-swarms}

### 2016-04-25 Paper_v5.tex
Updated Paper

@@
 Modchang et al.
-\cite{modchang2008mathematical} used a genetic algorithm (GA) to estimate parameters for a model of G-protein-coupled receptors(GPCRs) mediated signal transduction.
-The parameters obtained using GA helped qualitatively ascertain the efficacy and potency of two G-protein agonists - CPA and NECA.
+\cite{modchang2008mathematical} used a genetic algorithm (GA) to estimate parameters for a model of G-protein-coupled receptors (GPCRs) mediated signal transduction.
+The parameters obtained using GA helped qualitatively ascertain the efficacy and potency of two G-protein agonists - N6-cyclopentyladenosine (CPA) and 5'-N-ethylcarboxamidoadenosine (NECA).
 Differential evolution approaches have also been effective on various biological problems \cite{tsai2005evolutionary,wang2001hybrid,noman2007inferring}.
@@
 We evaluated the performance of DOPS on problems from the least and most expensive categories.
-The least expensive problem (henceforth referred to as problem B4) was a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters.
+The least expensive problem (henceforth referred to as problem B4) was a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters \cite{villaverde2014high}.
 The biochemical reactions were modeled using modular rate laws and the generalized form of Michaelis-Menten kinetics.
-The expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters (henceforth referred to as problem B1).
+The expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters \cite {smallbone2013large}(henceforth referred to as problem B1).
 In both cases, synthetic time series data sets were generated using known parameter values, these data were then used by the parameter estimation methods.

### 2016-04-26 Paper_v5.tex
DDS Portion added and references updated

@@
 
-%In many of these high dimensional problems approaching an exact solution may not be necessary.
+In many of these high dimensional problems approaching an exact solution may not be necessary.
 Gutenkust et al.
@@
 Tolson and Shoemaker \cite{tolson2007dynamically} showed through Dynamically Dimensioned Search (DDS) that high-dimensional watershed models can be calibrated quickly by perturbing only a subset of dimensions.
-
-%Because the neighborhood in DDS is the entire parameter space, DDS is a global optimization technique although it is greedy.
+DDS is a single solution based search where the solution is greedily updated.
+Because the neighborhood in DDS is the entire parameter space, DDS is a global optimization technique although it is greedy.
 
@@
 \begin{equation}\label{eqn:update-rule}
-	\mathbf{z}_{i,j} = \theta_{1,j-1}\mathbf{z}_{i,j-1} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{z}_{i,j-1}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G}_{k} - \mathbf{z}_{i,j-1}\right)
+	%\mathbf{z}_{i,j} = \theta_{1,j-1}\mathbf{z}_{i,j-1} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{z}_{i,j-1}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G}_{k} - \mathbf{z}_{i,j-1}\right)
+	{z}_{i,j} = \theta_{1,j-1}{z}_{i,j-1} + \theta_{2}{r}_{1}\left(\mathcal{L}_{i} - {z}_{i,j-1}\right) + \theta_{3}{r}_{2}\left(\mathcal{G}_{k} - {z}_{i,j-1}\right)
 \end{equation}

### 2016-04-26 Paper_v5.tex
Updates to the introduction

@@
 \vspace{0.1in}
-{\par \centering \textbf{To be submitted:}~\emph{PLoS ONE}}
+{\par \centering \textbf{To be submitted:}~\emph{Biotechnology~Journal}}
 \vspace{0.5in}
@@
 {\par \centering Jeffrey D. Varner,}
-{\par \centering Associate Professor, School of Chemical and Biomolecular Engineering,}
+{\par \centering Professor, School of Chemical and Biomolecular Engineering,}
 {\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853}
@@
 
-Cells process nutrients and respond to changes in their environment using complex biochemical networks. These networks contain thousands of components
-interconnected through nonlinear enzyme catalyzed reactions.
+Cells process nutrients and respond to changes in their environment using complex enzyme catalyzed biochemical networks.
 Mathematical modeling has evolved as a powerful paradigm to analyze, and ultimately design these complex networks \cite{assmus2006dynamics, Riel2006aa,Jaqaman:2006aa,kitano2002systems,hood2004systems}.
 Mathematical modeling of biochemical networks is often an iterative process.
-First, models are formulated from biochemical knowledge, and then model parameters are estimated using experimental data \cite{Aldridge:2006aa,banga2008optimization,ashyraliyev2009systems}.
-Parameter estimation is typically framed as a non-linear optimization problem wherein the residual (or objective function) between experimental data and model simulations is minimized using an optimization strategy \cite{moles2003parameter}.
-Optimal parameters obtained from model training are then used to validate the model on unseen experimental data.
-If validation fails, model construction and calibration are repeated iteratively until satisfactory results are obtained.
+First, models are formulated from existing biochemical knowledge, and then model parameters are estimated using experimental data \cite{Aldridge:2006aa,banga2008optimization,ashyraliyev2009systems}.
+Parameter estimation is typically framed as a non-linear optimization problem wherein the residual (or objective function) between experimental measurements and model simulations is minimized using an optimization strategy \cite{moles2003parameter}.
+Optimal parameter estimates obtained from model training are then used to validate the model on unseen experimental data.
+If the validation studies fail, model construction and calibration are repeated iteratively until satisfactory results are obtained.
 
@@
 Although parameter estimation has been a well studied problem in engineering for decades \cite{nieman1971review,beck1977parameter,young1981parameter,beck1998inverse},
-the complex dynamics of large biological systems and noisy, often incomplete experimental data pose a unique estimation challenge.
-Often optimization problems involving biological systems are non-linear and multi-modal i.e. typical models often have multiple local minima or maxima \cite{moles2003parameter,banga2008optimization}.
-Non-linearity coupled with multi-modality generally renders local optimization techniques such as pattern search \cite{hooke1961direct}, Nelder-Mead simplex methods \cite{nelder1965simplex},
-steepest descent methods or Levenberg-Marquardt \cite{more1978levenberg} incapable of reliably obtaining optimal solutions as they generally terminate at local minimum.
+the complex dynamics of large biological systems and noisy, often incomplete experimental data sets pose a unique challenge.
+Often optimization problems involving biological systems are non-linear and multi-modal i.e., typical models often have multiple local minima or maxima \cite{moles2003parameter,banga2008optimization}.
+Non-linearity coupled with multi-modality generally renders local optimization techniques such as pattern search \cite{hooke1961direct}, Nelder-Mead simplex methods \cite{nelder1965simplex}, steepest descent methods or Levenberg-Marquardt \cite{more1978levenberg} incapable of reliably obtaining globally optimal solutions as these methods often terminate at local minimum.
 Though deterministic global optimization techniques (for example algorithms based on branch and bound) can handle non-linearity and multi-modality \cite{esposito2000deterministic,horst2013global},
-the absence of derivative information, discontinuity of the objective functions, non-smooth regions or the lack of knowledge about the objective function hampers these techniques.
+the absence of derivative information, the discontinuity of the objective functions, non-smooth regions or the lack of knowledge about the objective function hampers these techniques.
+
+%Fan et al.
+recently showed that population based meta-heuristics along with decomposition based methods can also be used to identify gene circuit models from messenger RNA measurements \cite{fan2015parameter}.
 
@@
 Evolutionary Programming \cite{fogel2009artificial} and population based searches like Differential Evolution (DE) \cite{storn1997differential} have shown promise on nonlinear multi-modal problems \cite{sun2012parameter}.
-These techniques do not make any assumptions about the structure of objective function nor do they require \textit{a~priori} information about the objective function.
-Though they do not guarantee strong convergence, these approaches are effective in finding near optimal solutions.
-Mendes et al. \cite{mendes1998non} used simulated annealing to estimate rate constants for the irreversible inhibition of HIV proteinase,
-Modchang et al.
-\cite{modchang2008mathematical} used a genetic algorithm (GA) to estimate parameters for a model of G-protein-coupled receptors (GPCRs) mediated signal transduction.
-The parameters obtained using GA helped qualitatively ascertain the efficacy and potency of two G-protein agonists - N6-cyclopentyladenosine (CPA) and 5'-N-ethylcarboxamidoadenosine (NECA).
-Differential evolution approaches have also been effective on various biological problems \cite{tsai2005evolutionary,wang2001hybrid,noman2007inferring}.
-Tashkova et al. compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis and showed that DE was the most effective \cite{tashkova2011parameter}.
+These techniques do not make any assumptions about, nor do they require, \textit{a~priori} information about the structure of objective function.
+Though they do not guarantee strong convergence, meta-heuristic approaches are often effective in finding globally optimal or near optimal solutions.
+For example, Mendes et al. used simulated annealing to estimate rate constants for the irreversible inhibition of HIV proteinase \cite{mendes1998non},
+while Modchang et al.
+used a genetic algorithm (GA) to estimate parameters for a model of G-protein-coupled receptors (GPCRs) activity \cite{modchang2008mathematical}.
+Parameter estimates obtained using GA helped qualitatively determine the effectiveness of two G-protein agonists, N6-cyclopentyladenosine (CPA) and 5'-N-ethylcarboxamidoadenosine (NECA).
+Differential evolution (DE) have also been effective on various biological problems \cite{tsai2005evolutionary,wang2001hybrid,noman2007inferring}.
+For example, Tashkova et al. compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis
+and showed that DE was the most effective of the approaches tested \cite{tashkova2011parameter}.
 Banga and co-workers have also successfully applied scatter-search methods to estimate parameters on non-linear biological processes \cite{villaverde2012cooperative,rodriguez2006novel,egea2007scatter}.
-Hybrid approaches that combine a meta-heuristic with a local optimization search, wherein a near globally optimal solution obtained using a meta-heuristic is refined using a local search have also become popular.
-Villaverde et al. combined scatter search with local search methods for parameter estimation in a collection of systems biology models \cite{villaverde2015biopredyn}.
-Fan et al. recently showed that population based meta-heuristics along with decomposition based methods can be also used to model gene circuits from mRNA data \cite{fan2015parameter}.
-Despite these successes, a major drawback of most metaheuristic approaches is the large number of function evaluations required to explore the parameter space.
-Typically as models grow in size and complexity, evaluation of the objective function becomes computationally expensive.
-Thus performing a large number of function evaluations is not desirable (and perhaps not feasible).
+Hybrid approaches that combine a meta-heuristic with a local optimization search,
+where a near optimal solution estimated using a meta-heuristic is refined using a local search, have also become popular.
+For example, Villaverde et al. developed the enhanced scatter search (eSS) method, which combined scatter search with local search methods,
+for parameter estimation in a collection of systems biology models \cite{villaverde2015biopredyn}.
+However, despite these successes, a major drawback of most meta-heuristics is the large number of function evaluations required to explore parameter space.
+Performing a large number of potentially expensive function evaluations is not desirable (and perhaps not feasible) for many types of biochemical models.
 
-In many of these high dimensional problems approaching an exact solution may not be necessary.
+%Typically as models grow in complexity, evaluation of the objective function becomes more computationally expensive.
+%We fixed the number of function evaluations to 4000 for both the benchmark problems without pre-optimizing any algorithmic parameters.
+%Lastly, we also compared the performance of DOPS against eSS on published benchmark problems \cite{villaverde2015biopredyn}.
+
+In many high dimensional problems approaching an exact solution may not be necessary.
 Gutenkust et al.
-\cite{gutenkunst2007universally} showed that a number of systems biology models are 'sloppy'.
+showed that a number of systems biology models are 'sloppy' \cite{gutenkunst2007universally}.
 Sloppy systems have specific parameter combinations that largely define the dynamics of the system.
@@
 
-In this study, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel meta-heuristic that combines the global search capability of multi-swarm particle swarm optimization and dynamically dimensioned search (DDS).
+In this study, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel meta-heuristic that combines the global search capability of multi-swarm particle swarm optimization with the greedy refinement of dynamically dimensioned search.
 The objective of DOPS is to obtain near optimal parameter estimates for large biochemical models within a relatively few function evaluations.
-DOPS uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors which are then greedily updated using dynamically dimensioned search.
-We first considered a model of human coagulation cascade to test the performance of DOPS.
-Coagulation is a large, complex biochemical network involving strong positive feedback.
-We then tested the performance of DOPS on commonly used test functions for global optimization (Ackley and Rosenbrock) and published biochemical parameter estimation benchmark problems \cite{villaverde2015biopredyn}.
-DOPS outperformed common meta-heuristic approaches like Differential Evolution (DE), Simulated Annealing (SA) and dynamically dimensioned search (DDS) on both the test functions and the coagulation model.
-We also compared the performance of DOPS against enhanced scatter search (eSS) on two benchmark problems published by Villaverde and co-workers\cite{villaverde2015biopredyn}.
-We fixed the number of function evaluations to 4000 for both the benchmark problems without pre-optimizing any algorithmic parameters.
-DOPS recovered the nominal parameters and surprisingly did so faster than eSS which took nearly an order of function evaluations more than DOPS in both the cases.
+DOPS uses multi-swarm particle swarm optimization to generate candidate solutions, which are then greedily updated using dynamically dimensioned search.
+First, we tested the performance of DOPS on commonly used test functions for global optimization (Ackley and Rosenbrock) and published biochemical benchmark problems.
+Next, we used DOPS to identify a model of human coagulation cascade.
+DOPS outperformed common meta-heuristic approaches like differential evolution, simulated annealing, and dynamically dimensioned search on the test functions, benchmark problems and the coagulation model.
+For example, DOPS recovered the nominal parameters for the benchmark problems using an order of magnitude less function evaluations than eSS in all cases.
 Taken together, these studies suggest DOPS is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.

### 2016-04-26 Paper_v5.tex
Final draft = still needs review

@@
 Cells process nutrients and respond to changes in their environment using complex enzyme catalyzed biochemical networks.
-Mathematical modeling has evolved as a powerful paradigm to analyze, and ultimately design these complex networks \cite{assmus2006dynamics, Riel2006aa,Jaqaman:2006aa,kitano2002systems,hood2004systems}.
+Mathematical modeling has evolved as a powerful paradigm to analyze, and ultimately design these complex networks \cite{assmus2006dynamics, Riel:2006aa,Jaqaman:2006aa,kitano2002systems,hood2004systems}.
 Mathematical modeling of biochemical networks is often an iterative process.
@@
 recently showed that population based meta-heuristics along with decomposition based methods can also be used to identify gene circuit models from messenger RNA measurements \cite{fan2015parameter}.
+%Hybrid approaches use a global meta-heuristic to estimate a near optimal solution, which is then refined using a local search.
 
-Meta-heuristic stochastic optimization approaches like Genetic Algorithms (GAs) \cite{goldberg2006genetic}, Simulated Annealing (SA) \cite{kirkpatrick1983optimization},
-Evolutionary Programming \cite{fogel2009artificial} and population based searches like Differential Evolution (DE) \cite{storn1997differential} have shown promise on nonlinear multi-modal problems \cite{sun2012parameter}.
-These techniques do not make any assumptions about, nor do they require, \textit{a~priori} information about the structure of objective function.
-Though they do not guarantee strong convergence, meta-heuristic approaches are often effective in finding globally optimal or near optimal solutions.
+Meta-heuristic optimization approaches like Genetic Algorithms (GAs) \cite{goldberg2006genetic}, Simulated Annealing (SA) \cite{kirkpatrick1983optimization},
+Evolutionary Programming \cite{fogel2009artificial} and Differential Evolution (DE) \cite{storn1997differential,tsai2005evolutionary,wang2001hybrid,noman2007inferring} have shown promise on non-linear multi-modal problems \cite{sun2012parameter}.
+These techniques do not make any assumptions, nor do they require, \textit{a~priori} information about the structure of objective function.
+Meta-heuristics are often very effective at finding globally optimal or near optimal solutions.
 For example, Mendes et al. used simulated annealing to estimate rate constants for the irreversible inhibition of HIV proteinase \cite{mendes1998non},
@@
 Parameter estimates obtained using GA helped qualitatively determine the effectiveness of two G-protein agonists, N6-cyclopentyladenosine (CPA) and 5'-N-ethylcarboxamidoadenosine (NECA).
-Differential evolution (DE) have also been effective on various biological problems \cite{tsai2005evolutionary,wang2001hybrid,noman2007inferring}.
-For example, Tashkova et al. compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis
+Tashkova et al. compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis
 and showed that DE was the most effective of the approaches tested \cite{tashkova2011parameter}.
 Banga and co-workers have also successfully applied scatter-search methods to estimate parameters on non-linear biological processes \cite{villaverde2012cooperative,rodriguez2006novel,egea2007scatter}.
-Hybrid approaches that combine a meta-heuristic with a local optimization search,
-where a near optimal solution estimated using a meta-heuristic is refined using a local search, have also become popular.
+Hybrid approaches which combine meta-heuristics with local optimization techniques have also become popular.
 For example, Villaverde et al. developed the enhanced scatter search (eSS) method, which combined scatter search with local search methods,
@@
 Performing a large number of potentially expensive function evaluations is not desirable (and perhaps not feasible) for many types of biochemical models.
+Alternatively, Tolson and Shoemaker found, using high-dimensional watershed models, that perturbing only a subset of parameters was an effective strategy for estimating
+parameters in models with expensive function evaluations \cite{tolson2007dynamically}.
+This approach called Dynamically Dimensioned Search (DDS) is a simple stochastic single-solution based heuristic that estimates nearly optimal solutions within a specified maximum function (or model) evaluation limit.
+Taken together, meta-heuristics are often very effective at finding globally optimal or near optimal solutions, however the large number of function evaluations required to
+converge to a solution limits their application for expensive biochemical models.
 
@@
 %Lastly, we also compared the performance of DOPS against eSS on published benchmark problems \cite{villaverde2015biopredyn}.
-
-In many high dimensional problems approaching an exact solution may not be necessary.
+%In many high dimensional problems approaching an exact solution may not be necessary.
 Gutenkust et al.
 showed that a number of systems biology models are 'sloppy' \cite{gutenkunst2007universally}.
-Sloppy systems have specific parameter combinations that largely define the dynamics of the system.
-Large perturbations to the rest of the parameters does not greatly impact the system dynamics.
-Ensemble approaches \cite{song2010ensembles,luan2010ensembles} have exploited this aspect to describe the dynamics of biological systems including coagulation which can be described using only a set of key species or parameters \cite{sagar2015dynamic}.
+Sloppy systems have specific parameter combinations that largely define system dynamics, while large perturbations to the remaining parameters does not influence the system behavior.
+Ensemble approaches have exploited sloppiness to describe the dynamics of biological systems using families of parameters \cite{song2010ensembles,luan2010ensembles,sagar2015dynamic}.
 Tolson and Shoemaker \cite{tolson2007dynamically} showed through Dynamically Dimensioned Search (DDS) that high-dimensional watershed models can be calibrated quickly by perturbing only a subset of dimensions.
@@
 
-In this study, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel meta-heuristic that combines the global search capability of multi-swarm particle swarm optimization with the greedy refinement of dynamically dimensioned search.
+In this study, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel meta-heuristic that combines the global search capability of multi-swarm particle swarm optimization with the greedy refinement of dynamically dimensioned search (DDS).
 The objective of DOPS is to obtain near optimal parameter estimates for large biochemical models within a relatively few function evaluations.
@@
 \subsection*{Optimization problem formulation.}
-The problem of parameter estimation in a dynamic biological model consists of finding an optimal parameter vector
-which minimizes the difference between model simulations and $\mathcal{E}$ experimental measurements. This difference is quantified by an objective function $K\left(\mathbf{p}\right)$
-which is typically the Euclidean norm of the simulation error subject to problem specific and parameter bounds constraints:
+Model parameters were estimated by minimizing the difference between model simulations and $\mathcal{E}$ experimental measurements.
+Simulation error is quantified by an objective function $K\left(\mathbf{p}\right)$
+(typically the Euclidean norm of the difference between simulations and measurements) subject to problem and parameter constraints:
 
@@
 \begin{aligned}
-& \text{minimize}
-& & K(\mathbf{p})=\sum_{i=1}^\mathcal{E} \left(g_i(t_i,\mathbf{x,p,u})-y_i\right)^2 \\
+& \min_{\mathbf{p}} K(\mathbf{p}) & =& \sum_{i=1}^\mathcal{E} \left(g_i(t_i,\mathbf{x,p,u})-y_i\right)^2 \\
 & \text{subject to}
@@
 \end{equation}
-where t is time, $\mathbf{x}\left(t,\mathbf{p}\right)$ is the state variable vector with an initial state $\mathbf{x}_{0}$, $\mathbf{u}(t)$ is a model input vector,
-$\mathbf{f}$ is the system of model equations (e.g., differential equations or algebraic constraints) and $\mathbf{p}$ is the model parameter vector.
-The parameter search (or model simulations) can be subject to $\mathbf{c}$ linear or non-linear constraints, and parameter bound constraints where
-p\textsuperscript{L} and p\textsuperscript{U} denote the lower and upper parameter bounds, respectively.
-The problem eventually is to find:
+The term $K(\mathbf{p})$ denotes the objective function, t denotes time, $g_i(t_i,\mathbf{x,p,u})$ is the model output for experiment $i$,
+$\mathbf{x}\left(t,\mathbf{p}\right)$ is the state variable vector with an initial state $\mathbf{x}_{0}$, $\mathbf{u}(t)$ is a model input vector,
+$\mathbf{f}(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p})$ is the system of model equations (e.g., differential equations or algebraic constraints)
+and $\mathbf{p}$ denotes the model parameter vector.
+The parameter search (or model simulations) can be subject to $\mathbf{c}(t,\mathbf{x,p,u})$ linear or non-linear constraints, and parameter bound constraints where
+$\mathbf{p}^{L}$ and $\mathbf{p}^{U}$ denote the lower and upper parameter bounds, respectively.
+Optimal model parameters are then given by:
 \begin{equation}\label{eq:static_opt_2}
@@
 \subsection*{Dynamic optimization with particle swarms (DOPS).}
-DOPS is a novel meta-heuristic which combines multi-swarm particle swarm methods with the dynamically dimensioned search approach of Shoemaker and colleagues (Fig. \ref{fig-algorithm}).
+DOPS is a novel meta-heuristic which combines multi-swarm particle swarm optimization
+with the dynamically dimensioned search (Fig. \ref{fig-algorithm}) and (Algo. \ref{algo_ddspso}).
 The goal of DOPS is to estimate optimal or near optimal parameter vectors for high-dimensional biological models within a specified number of function evaluations.
-Toward this objective, DOPS begins by using a particle swarm search and then dynamically switches, using an adaptive switching criteria, to the DDS search phase.
-
-\subsubsection*{Phase 1: Swarm phase.}
-We began the particle swarm phase of DOPS by randomly initializing a swarm of $\mathcal{K}$-dimensional particles (represented as $z_{i}$),
-wherein each particle corresponded to a $\mathcal{K}$-dimensional parameter vector.
-After initialization, particles were randomly partitioned into $k$ equal sized sub-swarms $\mathcal{S}_{1},\hdots,\mathcal{S}_{k}$.
-Thereafter within each sub-swarm $\mathcal{S}_{k}$, particles were updated according to the rule:
-\begin{equation}\label{eqn:update-rule}
-	%\mathbf{z}_{i,j} = \theta_{1,j-1}\mathbf{z}_{i,j-1} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{z}_{i,j-1}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G}_{k} - \mathbf{z}_{i,j-1}\right)
-	{z}_{i,j} = \theta_{1,j-1}{z}_{i,j-1} + \theta_{2}{r}_{1}\left(\mathcal{L}_{i} - {z}_{i,j-1}\right) + \theta_{3}{r}_{2}\left(\mathcal{G}_{k} - {z}_{i,j-1}\right)
-\end{equation}
-where $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ were adjustable parameters, $\mathcal{L}_{i}$ denotes the best solution found by particle $i$ within sub-swarm $k$ for function evaluation $1\rightarrow j-1$, and
-$\mathcal{G}_{k}$ denotes the best solution found over all particles within sub-swarm $\mathcal{S}_{k}$.
-The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model parameters ($\mathcal{K}\times{1}$).
-Equation (3) is similar to the general particle swarm update rule, however, it does not contain velocity terms.
-In DOPS, the parameter $\theta_{1,j-1}$ is similar to the inertia weight parameter for the velocity term described by Shi and Eberhart \cite{shi1999empirical}.
-In this study Shi and Eberhart propose a linearly decreasing inertia weight to improve convergence properties of PSO.
-$\theta_{1,j-1}$ is inspired from this idea and also from the idea of decreasing perturbation probability proposed by Tolson and Shoemaker \cite{tolson2007dynamically}.
-It is an analogous equivalent to inertia weight on velocity.
-However $\theta_{1,j-1}$ places inertia on the position rather than velocity and uses the same rule described by Shi and Eberhart to adaptively change with the number of function evaluations.
-It is updated according to:
-\begin{eqnarray}
-	\mathbf \theta_{1,j}&=&\frac{(\mathcal{N}-{j})*({w}_{max}-{w}_{min}))}{(\mathcal{N}-{1})} + {w}_{min}
-\end{eqnarray}
-where $\mathcal{N}$ represents the total number of function evaluations, ${w}_{max}$ and ${w}_{min}$ are the maximum and minimum inertia weights, respectively.
-While updating the particles, we made sure all dimensions of the solution represented by the particle were within bounds using a set of reflection boundary conditions:
-
-\begin{eqnarray*}
-\end{eqnarray*}
-
-\begin{algorithm}
-\If{$z_{i,j}^{old} < z_{i}^{min}$}{
-  ${z}_{i,j}^{new}={z}_{i,j}^{old}+(z_{i}^{min} - {z}_{i,j}^{old})$
-  \If{${z}_{i,j}^{new}>z_{i}^{max}$}{
-    ${z}_{i,j}^{new}=z_{i}^{max}$
-  }
-}
-\If{$z_{i,j}^{old} > z_{i}^{max}$}{
-  ${z}_{i,j}^{new}={z}_{i,j}^{old}+({z}_{i,j}^{old}-z_{i}^{max})$
-  \If{${z}_{i,j}^{new}<z_{i}^{min}$}{
-    ${z}_{i,j}^{new}=z_{i}^{min}$
-  }
-}
-\end{algorithm}
-
-After every $\mathcal{M}$ function evaluations, particles were randomly redistributed to a new sub-swarm, and updated according to Eqn. \eqref{eqn:update-rule}.
-This process continued for $\mathcal{F}*\mathcal{N}$ functions evaluations, where $\mathcal{F}$ is the fraction of evaluations in the particle swarm phase of DOPS.
-If the simulation error stagnates for example, does not change by more than 1\% for a specified number of evaluations, the swarm phase is terminated and DOPS switches to
-exploring parameter space using a DDS approach.
-
-\subsubsection*{Phase 2: DDS phase.}
-At the conclusion of the swarm phase, the overall best particle, $\mathcal{G}_{k}$, over the k sub-swarms was to initialize the DDS phase.
-DOPS takes $\left(1-\mathcal{F}\right)*\mathcal{N}$ function evaluations in the DDS phase and then terminates the search.
-For the DDS phase, the best parameter estimate was updated using the rule:
-\begin{equation}
-  \mathcal{G}_{new}({J})=\begin{cases}
-    \mathcal{G}(\mathbf{J})+\mathbf{r}_{normal}(\mathbf{J})\sigma(\mathbf{J}), & \text{if $\mathcal{G}_{new}(\mathbf{J})<\mathcal{G}(\mathbf{J})$}.\\
-    \mathcal{G}(\mathbf{J}), & \text{otherwise}.
-  \end{cases}
-\end{equation}
-where $\mathbf{J}$ is a vector representing the subset of dimensions that are being perturbed, $\mathbf{r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$,
-and $\sigma$ denotes the perturbation amplitude:
-\begin{equation}
-	\sigma = {R}(\mathbf{p}^U - \mathbf{p}^L)
-\end{equation}
-where ${R}$ is the scalar perturbation size parameter, $\mathbf{p}^U$ and $\mathbf{p}^L$ are ($\mathcal{K}\times{1}$) vectors that represent the maximum and minimum bounds on each dimension.
-The set $\mathbf{J}$ was constructed using a probability function $\mathcal{P}_{i}$ that represents a threshold probability for determining whether a specific dimension $j$ was perturbed or not.
-This function is a monotonically decreasing function that decreases with the number of function evaluations.
-$\mathcal{P}_{i}$ can be any monotonically decreasing function, in our approach we used the following function:
-\begin{eqnarray}
-	%\mathcal{P}_{j}&=&{1}-\log(j/ (\beta({1}-\mathcal{FR})*\mathbf{N}))
-	\mathcal{P}_{i}&=&{1}-\log(i/ (({1}-\mathcal{F})*\mathbf{N}))
-\end{eqnarray}
-%where $\beta$ is the perturbation frequency probability modulator.
-where $i$ is the current iteration.
-After $\mathcal{P}_{i}$ is determined, we assign a probability $\mathcal{P}_{j}$ from a uniform distribution for each dimension $j$.
-If $\mathcal{P}_{j}$ is less than $\mathcal{P}_{i}$ the dimension is included in \textbf{J}.
-Thus the probability that a dimension $j$ is perturbed decreases as the number of iterations or function evaluations indicated by $i$ increase.
-%Thus the number of dimensions of the candidate vector that are updated or perturbed decreases with the as the number of function evaluations increase. T
-These updates are greedy in nature, so $\mathcal{G}_{new}$ becomes the new solution vector only if it is better than the old one $\mathcal{G}$.
-The reflection boundary conditions in equation 5, the update function described in equations 6 and 7, the selection probability in equation 8 are the means by which DDS ideas are incorporated into DOPS.
-The fraction of evaluations $\mathcal{F}$  within the swarm phase is based on a switching strategy wherein the switch from swarm phase to DDS phase happens when the error due to the best solution does not drop more than 1\% of the original error, continuously for more than a prescribed number of function evaluations.
-This allows the solution to quickly jump out of local optima and avoid any convergence issues that are generally associated with swarm based searches.
-
-\newpage
-%\section*{Pseudo Code.}
-%\begin{algorithm}[H]
-%
-% \While{$besterror\leq $}{
-%  read current\;
-%  \eIf{understand}{
-%   go to next section\;
-%   current section becomes this one\;
-%   }{
-%   go back to the beginning of current section\;
-%  }
-% }
-% \caption{Dynamic Optimization with Particle Swarms}
-%\end{algorithm}
+Toward this objective, DOPS begins by using a particle swarm search and then dynamically switches, using an adaptive switching criteria, to a DDS search phase.
 
@@
 }
-\caption{Dynamic Optimization with Particle Swarms}\label{algo_ddspso}
+\caption{Pseudo code for the dynamic optimization with particle swarms (DOPS) method.}\label{algo_ddspso}
 \end{algorithm}\DecMargin{1em}
+
+\subsubsection*{Phase 1: Particle swarm phase.}
+The particle swarm phase of DOPS begins by randomly initializing a swarm of $\mathcal{K}$-dimensional particles (represented as $z_{i}$),
+wherein each particle corresponded to a $\mathcal{K}$-dimensional parameter vector.
+After initialization, particles were randomly partitioned into $k$ equally sized sub-swarms $\mathcal{S}_{1},\hdots,\mathcal{S}_{k}$.
+Particles within each sub-swarm $\mathcal{S}_{k}$ were updated according to the rule:
+\begin{equation}\label{eqn:update-rule}
+	%\mathbf{z}_{i,j} = \theta_{1,j-1}\mathbf{z}_{i,j-1} + \theta_{2}\mathbf{r}_{1}\left(\mathcal{L}_{i} - \mathbf{z}_{i,j-1}\right) + \theta_{3}\mathbf{r}_{2}\left(\mathcal{G}_{k} - \mathbf{z}_{i,j-1}\right)
+	{z}_{i,j} = \theta_{1,j-1}{z}_{i,j-1} + \theta_{2}{r}_{1}\left(\mathcal{L}_{i} - {z}_{i,j-1}\right) + \theta_{3}{r}_{2}\left(\mathcal{G}_{k} - {z}_{i,j-1}\right)
+\end{equation}
+where $\left(\theta_{1},\theta_{2},\theta_{3}\right)$ were adjustable parameters, $\mathcal{L}_{i}$ denotes the best solution found by particle $i$ within sub-swarm
+$\mathcal{S}_{k}$ for function evaluation $1\rightarrow j-1$, and
+$\mathcal{G}_{k}$ denotes the best solution found over all particles within sub-swarm $\mathcal{S}_{k}$.
+The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model parameters ($\mathcal{K}\times{1}$).
+Equation (3) is similar to the general particle swarm update rule, however, it does not contain velocity terms.
+In DOPS, the parameter $\theta_{1,j-1}$ is similar to the inertia weight parameter for the velocity term described by Shi and Eberhart \cite{shi1999empirical}.
+In this study Shi and Eberhart propose a linearly decreasing inertia weight to improve convergence properties of PSO.
+$\theta_{1,j-1}$ is inspired from this idea and also from the idea of decreasing perturbation probability proposed by Tolson and Shoemaker \cite{tolson2007dynamically}.
+It is an analogous equivalent to inertia weight on velocity.
+However $\theta_{1,j-1}$ places inertia on the position rather than velocity and uses the same rule described by Shi and Eberhart to adaptively change with the number of function evaluations.
+It is updated according to:
+\begin{eqnarray}
+	\mathbf \theta_{1,j}&=&\frac{(\mathcal{N}-{j})*({w}_{max}-{w}_{min}))}{(\mathcal{N}-{1})} + {w}_{min}
+\end{eqnarray}
+where $\mathcal{N}$ represents the total number of function evaluations, ${w}_{max}$ and ${w}_{min}$ are the maximum and minimum inertia weights, respectively.
+While updating the particles, we made sure all dimensions of the solution represented by the particle were within bounds using reflection boundary conditions (Algo. \ref{algo_reflection}).
+
+\begin{algorithm}
+\If{$z_{i,j}^{old} < z_{i}^{min}$}{
+  ${z}_{i,j}^{new}={z}_{i,j}^{old}+(z_{i}^{min} - {z}_{i,j}^{old})$
+  \If{${z}_{i,j}^{new}>z_{i}^{max}$}{
+    ${z}_{i,j}^{new}=z_{i}^{max}$
+  }
+}
+\If{$z_{i,j}^{old} > z_{i}^{max}$}{
+  ${z}_{i,j}^{new}={z}_{i,j}^{old}+({z}_{i,j}^{old}-z_{i}^{max})$
+  \If{${z}_{i,j}^{new}<z_{i}^{min}$}{
+    ${z}_{i,j}^{new}=z_{i}^{min}$
+  }
+}
+\caption{Pseudo code for the reflective boundary conditions used by the dynamic optimization with particle swarms (DOPS) method.}\label{algo_reflection}
+\end{algorithm}
+
+After every $\mathcal{M}$ function evaluations, particles were randomly redistributed to a new sub-swarm, and updated according to Eqn. \eqref{eqn:update-rule}.
+This process continued for a maximum of $\mathcal{F}*\mathcal{N}$ functions evaluations, where $\mathcal{F}$ is the fraction of evaluations in the particle swarm phase of DOPS.
+However, if the simulation error stagnated e.g., did not change by more than 1\% for a specified number of evaluations, the swarm phase was terminated and DOPS switched to
+exploring parameter space using the DDS approach.
+
+\subsubsection*{Phase 2: DDS phase.}
+At the conclusion of the swarm phase, the overall best particle, $\mathcal{G}_{k}$, over the k sub-swarms was used to initialize the DDS phase.
+DOPS takes at least $\left(1-\mathcal{F}\right)*\mathcal{N}$ function evaluations during the DDS phase and then terminates the search.
+For the DDS phase, the best parameter estimate was updated using the rule:
+\begin{equation}
+  \mathcal{G}_{new}({J})=\begin{cases}
+    \mathcal{G}(\mathbf{J})+\mathbf{r}_{normal}(\mathbf{J})\sigma(\mathbf{J}), & \text{if $\mathcal{G}_{new}(\mathbf{J})<\mathcal{G}(\mathbf{J})$}.\\
+    \mathcal{G}(\mathbf{J}), & \text{otherwise}.
+  \end{cases}
+\end{equation}
+where $\mathbf{J}$ is a vector representing the subset of dimensions that are being perturbed, $\mathbf{r}_{normal}$ denotes a normal random vector of the same dimensions as $\mathcal{G}$,
+and $\sigma$ denotes the perturbation amplitude:
+\begin{equation}
+	\sigma = {R}(\mathbf{p}^U - \mathbf{p}^L)
+\end{equation}
+where ${R}$ is the scalar perturbation size parameter, $\mathbf{p}^U$ and $\mathbf{p}^L$ are ($\mathcal{K}\times{1}$) vectors that represent the maximum and minimum bounds on each dimension.
+The set $\mathbf{J}$ was constructed using a probability function $\mathcal{P}_{i}$ that represents a threshold for determining whether a specific dimension $j$ was perturbed or not; $\mathcal{P}_{i}$ is monotonically decreasing function of function evaluations:
+\begin{equation}
+	%\mathcal{P}_{j}&=&{1}-\log(j/ (\beta({1}-\mathcal{FR})*\mathbf{N}))
+	\mathcal{P}_{i}={1}-\log\left[\frac{i}{(1-\mathcal{F})*\mathcal{N}}\right]
+\end{equation}
+%where $\beta$ is the perturbation frequency probability modulator.
+where $i$ is the current iteration. After $\mathcal{P}_{i}$ is determined, we drew $\mathcal{P}_{j}$ from a uniform distribution for each dimension $j$.
+If $\mathcal{P}_{j}<\mathcal{P}_{i}$ was included in \textbf{J}.
+Thus, the probability that a dimension $j$ was perturbed was inversely proportional to the number of function evaluations.
+DDS updates are greedy; $\mathcal{G}_{new}$ becomes the new solution vector only if it is better than $\mathcal{G}$.
+
+%The reflection boundary conditions in equation 5, the update function described in equations 6 and 7, the selection probability in equation 8 are the means by which DDS ideas are incorporated into DOPS.
+The fraction of evaluations $\mathcal{F}$  within the swarm phase is based on a switching strategy wherein the switch from swarm phase to DDS phase happens when the error due to the best solution does not drop more than 1\% of the original error, continuously for more than a prescribed number of function evaluations.
+This allows the solution to quickly jump out of local optima and avoid any convergence issues that are generally associated with swarm based searches.
+%Thus the number of dimensions of the candidate vector that are updated or perturbed decreases with the as the number of function evaluations increase. T
+
+\newpage
+%\section*{Pseudo Code.}
+%\begin{algorithm}[H]
+%
+% \While{$besterror\leq $}{
+%  read current\;
+%  \eIf{understand}{
+%   go to next section\;
+%   current section becomes this one\;
+%   }{
+%   go back to the beginning of current section\;
+%  }
+% }
+% \caption{Dynamic Optimization with Particle Swarms}
+%\end{algorithm}
 

### 2016-04-27 Paper_v5.tex
Polish and refine

@@
 Parameter estimation is typically framed as a non-linear optimization problem wherein the residual (or objective function) between experimental measurements and model simulations is minimized using an optimization strategy \cite{moles2003parameter}.
-Optimal parameter estimates obtained from model training are then used to validate the model on unseen experimental data.
+Optimal parameter estimates are then used to predict unseen experimental data.
 If the validation studies fail, model construction and calibration are repeated iteratively until satisfactory results are obtained.
+As our biological knowledge increases, model formulation may not be as significant a challenge, but parameter estimation will likely remain difficult.
 
-Parameter estimation is a major challenge in the iterative development of biochemical models.
-Although parameter estimation has been a well studied problem in engineering for decades \cite{nieman1971review,beck1977parameter,young1981parameter,beck1998inverse},
-the complex dynamics of large biological systems and noisy, often incomplete experimental data sets pose a unique challenge.
-Often optimization problems involving biological systems are non-linear and multi-modal i.e., typical models often have multiple local minima or maxima \cite{moles2003parameter,banga2008optimization}.
-Non-linearity coupled with multi-modality generally renders local optimization techniques such as pattern search \cite{hooke1961direct}, Nelder-Mead simplex methods \cite{nelder1965simplex}, steepest descent methods or Levenberg-Marquardt \cite{more1978levenberg} incapable of reliably obtaining globally optimal solutions as these methods often terminate at local minimum.
+Parameter estimation is a major challenge to the development of biochemical models.
+Parameter estimation has been a well studied engineering problem for decades \cite{nieman1971review,beck1977parameter,young1981parameter,beck1998inverse}.
+However, the complex dynamics of large biological systems and noisy, often incomplete experimental data sets pose a unique estimation challenge.
+Often optimization problems involving biological systems are non-linear and multi-modal i.e., typical models have multiple local minima or maxima \cite{moles2003parameter,banga2008optimization}.
+Non-linearity coupled with multi-modality renders local optimization techniques such as pattern search \cite{hooke1961direct}, Nelder-Mead simplex methods \cite{nelder1965simplex}, steepest descent or Levenberg-Marquardt \cite{more1978levenberg} incapable of reliably obtaining globally optimal solutions as these methods often terminate at local minimum.
 Though deterministic global optimization techniques (for example algorithms based on branch and bound) can handle non-linearity and multi-modality \cite{esposito2000deterministic,horst2013global},
-the absence of derivative information, the discontinuity of the objective functions, non-smooth regions or the lack of knowledge about the objective function hampers these techniques.
+the absence of derivative information, discontinuous objective functions, non-smooth regions or the lack of knowledge about the objective function hampers these techniques.
 
@@
 
-Meta-heuristic optimization approaches like Genetic Algorithms (GAs) \cite{goldberg2006genetic}, Simulated Annealing (SA) \cite{kirkpatrick1983optimization},
-Evolutionary Programming \cite{fogel2009artificial} and Differential Evolution (DE) \cite{storn1997differential,tsai2005evolutionary,wang2001hybrid,noman2007inferring} have shown promise on non-linear multi-modal problems \cite{sun2012parameter}.
-These techniques do not make any assumptions, nor do they require, \textit{a~priori} information about the structure of objective function.
+Meta-heuristics like Genetic Algorithms (GAs) \cite{goldberg2006genetic}, Simulated Annealing (SA) \cite{kirkpatrick1983optimization},
+Evolutionary Programming \cite{fogel2009artificial} and Differential Evolution (DE) \cite{storn1997differential,tsai2005evolutionary,wang2001hybrid,noman2007inferring} have all shown promise on non-linear multi-modal problems \cite{sun2012parameter}.
+These techniques do not make any assumptions, nor do they require, \textit{a~priori} information about the structure of the objective function.
 Meta-heuristics are often very effective at finding globally optimal or near optimal solutions.
-For example, Mendes et al. used simulated annealing to estimate rate constants for the irreversible inhibition of HIV proteinase \cite{mendes1998non},
-while Modchang et al.
-used a genetic algorithm (GA) to estimate parameters for a model of G-protein-coupled receptors (GPCRs) activity \cite{modchang2008mathematical}.
-Parameter estimates obtained using GA helped qualitatively determine the effectiveness of two G-protein agonists, N6-cyclopentyladenosine (CPA) and 5'-N-ethylcarboxamidoadenosine (NECA).
-Tashkova et al. compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis
-and showed that DE was the most effective of the approaches tested \cite{tashkova2011parameter}.
-Banga and co-workers have also successfully applied scatter-search methods to estimate parameters on non-linear biological processes \cite{villaverde2012cooperative,rodriguez2006novel,egea2007scatter}.
-Hybrid approaches which combine meta-heuristics with local optimization techniques have also become popular.
-For example, Villaverde et al. developed the enhanced scatter search (eSS) method, which combined scatter search with local search methods,
-for parameter estimation in a collection of systems biology models \cite{villaverde2015biopredyn}.
-However, despite these successes, a major drawback of most meta-heuristics is the large number of function evaluations required to explore parameter space.
-Performing a large number of potentially expensive function evaluations is not desirable (and perhaps not feasible) for many types of biochemical models.
+For example, Mendes et al. used SA to estimate rate constants for the inhibition of HIV proteinase \cite{mendes1998non},
+while Modchang et al. used a GA to estimate parameters for a model of G-protein-coupled receptor (GPCR) activity \cite{modchang2008mathematical}.
+Parameter estimates obtained using the GA stratified the effectiveness of two G-protein agonists, N6-cyclopentyladenosine (CPA) and 5'-N-ethylcarboxamidoadenosine (NECA).
+Tashkova et al. compared different meta-heuristics for parameter estimation on a dynamic model of endocytosis;
+DE was the most effective of the approaches tested \cite{tashkova2011parameter}.
+Banga and co-workers have also successfully applied scatter-search to estimate model parameters \cite{villaverde2012cooperative,rodriguez2006novel,egea2007scatter}.
+Hybrid approaches, which combine meta-heuristics with local optimization techniques, have also become popular.
+For example, Villaverde et al. developed the enhanced scatter search (eSS) method, which combined scatter and local search methods,
+for parameter estimation in biological models \cite{villaverde2015biopredyn}.
+However, despite these successes, a major drawback of most meta-heuristics remains the large number of function evaluations required to explore parameter space.
+Performing numerous potentially expensive function evaluations is not desirable (and perhaps not feasible) for many types of biochemical models.
 Alternatively, Tolson and Shoemaker found, using high-dimensional watershed models, that perturbing only a subset of parameters was an effective strategy for estimating
-parameters in models with expensive function evaluations \cite{tolson2007dynamically}.
-This approach called Dynamically Dimensioned Search (DDS) is a simple stochastic single-solution based heuristic that estimates nearly optimal solutions within a specified maximum function (or model) evaluation limit.
-Taken together, meta-heuristics are often very effective at finding globally optimal or near optimal solutions, however the large number of function evaluations required to
-converge to a solution limits their application for expensive biochemical models.
+parameters in expensive models \cite{tolson2007dynamically}. Their approach, called Dynamically Dimensioned Search (DDS),
+is a simple stochastic single-solution heuristic that estimates nearly optimal solutions within a specified maximum number of function (or model) evaluations.
+Thus, while meta-heuristics are often effective at estimating globally optimal or nearly optimal solutions, they require a large number of function evaluations to
+converge to a solution.
 
@@
 
-In this study, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel meta-heuristic that combines the global search capability of multi-swarm particle swarm optimization with the greedy refinement of dynamically dimensioned search (DDS).
+In this study, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel hybrid meta-heuristic that combines the global search capability of multi-swarm particle swarm optimization with the greedy refinement of dynamically dimensioned search (DDS).
 The objective of DOPS is to obtain near optimal parameter estimates for large biochemical models within a relatively few function evaluations.
-DOPS uses multi-swarm particle swarm optimization to generate candidate solutions, which are then greedily updated using dynamically dimensioned search.
+DOPS uses multi-swarm particle swarm optimization to generate nearly optimal candidate solutions, which are then greedily updated using dynamically dimensioned search.
+We tested DOPS using a combination of classic optimization test functions, biochemical benchmark problems and real-world biochemical models.
 First, we tested the performance of DOPS on commonly used test functions for global optimization (Ackley and Rosenbrock) and published biochemical benchmark problems.
-Next, we used DOPS to identify a model of human coagulation cascade.
-DOPS outperformed common meta-heuristic approaches like differential evolution, simulated annealing, and dynamically dimensioned search on the test functions, benchmark problems and the coagulation model.
-For example, DOPS recovered the nominal parameters for the benchmark problems using an order of magnitude less function evaluations than eSS in all cases.
-Taken together, these studies suggest DOPS is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.
+Next, we used DOPS to estimate the parameters of a model of the human coagulation cascade.
+On average, DOPS outperformed other common meta-heuristics like differential evolution, simulated annealing, single-swarm particle swarm optimization, and dynamically dimensioned search on the optimization test functions, benchmark problems and the coagulation model.
+For example, DOPS recovered the nominal parameters for the benchmark problems using an order of magnitude fewer function evaluations than eSS in all cases.
+It also produced parameter estimates for the coagulation model that predicted unseen coagulation data sets.
+Thus, DOPS is a promising hybrid meta-heuristic for the estimation of biochemical model parameters in relatively few function evaluations.
 
@@
 
-\subsection*{DOPS explores parameter space using a combination of global and greedy search.}
-DOPS is a novel meta-heuristic which combines multi-swarm particle swarm methods with the dynamically dimensioned search approach of Shoemaker and colleagues (Fig. \ref{fig-algorithm}).
+\subsection*{DOPS explores parameter space using a combination of global methods.}
+DOPS is a novel hybrid meta-heuristic which combines a multi-swarm particle swarm method with the dynamically dimensioned search approach of Shoemaker and colleagues (Fig. \ref{fig-algorithm}).
 The goal of DOPS is to estimate optimal or near optimal parameter vectors for high-dimensional biological models within a specified number of function evaluations.
-Toward this objective, DOPS begins by using a particle swarm search and then dynamically switches, using an adaptive switching criteria, to the DDS search phase.
-The particle swarm search uses multiple sub-swarms wherein each solution represented by a particle is updated using equation \ref{eqn:update-rule}.
-The particle update is influenced by the best particle amongst all particles in the sub-swarm and the best solution found by the particle till the current iteration.
-This update rule however differs from a conventional PSO update wherein there are no velocity terms involved.
-The particle updates continue to happen within the sub-swarms for a certain number of iterations (or function evaluations) after which the sub-swarms are reorganized which is similar to the regrouping strategy described by Zhao et al.
+Toward this objective, DOPS begins by using a multi-swarm particle swarm search and then dynamically switches, using an adaptive switching criteria, to the DDS approach.
+The particle swarm search uses multiple sub-swarms wherein the update to each particle (corresponding to a parameter vector estimate) is influenced by the best particle amongst the sub-swarm, and the current globally best particle.
+Particle updates occur within sub-swarms for a certain number of function evaluations, after which the sub-swarms are reorganized.
+This sub-swarm mixing is similar to the regrouping strategy described by Zhao et al.
 \cite{zhao2008dynamic}.
-DOPS switches out of this PSO phase based on an adaptive switching criteria that is a function of error convergence rate.
-If the error represented by the best particle does not drop for a threshold number of function evaluations, DOPS switches automatically to the DDS search phase.
-The DDS search uses the best particle from the PSO search phase as its initial solution or candidate vector.
-Thereafter, the particle is greedily updated by perturbing a subset of dimensions for the remaining number of function evaluations.
-The number of dimensions perturbed is a monotonically decreasing function that generally depends on the number of function evaluations within the DDS phase.
+DOPS switches out of the particle swarm phase based upon an adaptive switching criteria that is a function of the rate of error convergence.
+If the error represented by the best particle does decrease for a threshold number of function evaluations, DOPS switches automatically to the DDS search phase.
+The DDS search is initialized with the globally best particle from the particle swarm phase, thereafter, the particle is greedily updated by perturbing a subset of dimensions for the remaining number of function evaluations.
+The identity of the parameters perturbed is chosen randomly, with fewer parameters perturbed the higher the number of function evaluations.
 
-\subsection*{DOPS minimized benchmark problems with a fixed number of function evaluations.}
-We tested DOPS and four other meta-heuristic approaches on the minimization of the Ackley and Rastrigin functions.
-DOPS performed similarly or outperformed the other meta-heuristic approaches on both of these test functions (Fig.\ref{fig-testfunctions}).
+%We tested DOPS and four other meta-heuristic approaches on the minimization of the Ackley and Rastrigin functions.
+
+\subsection*{DOPS minimized benchmark problems using fewer function evaluations.}
+On average, DOPS performed similarly or outperformed the four other meta-heuristics for the Ackley and Rastrigin test functions (Fig. \ref{fig-testfunctions}).
 The Ackley and Rastrigin functions both have multiple local extrema and attain a global minimum value of zero.
-In each case, we fixed the maximum number of function evaluations at $\mathcal{N}$ = 4000 and ran 25 independent experiments with different initial parameter vectors.
+In each case, we fixed the maximum number of function evaluations at $\mathcal{N}$ = 4000 and ran $\mathcal{T}$ = 25 independent experiments with different initial parameter vectors.
 DOPS found optimal or near optimal solutions for both the 10-dimensional Ackley (Fig.
@@
 \ref{fig-testfunctions}B) functions within the budget of function evaluations.
-In each of the 10-dimensional cases, other meta-heurtistics such as DDS and DE also performed well, however DOPS consistently outperformed all other approaches tested.
-This performance difference was more pronounced as the dimension of the search problem increased; for a 300-dimensional Rastrigin function DOPS was the only approach to find an optimal or near optimal solution
+In each of the 10-dimensional cases, other meta-heurtistics such as DDS and DE also performed well. However, DOPS consistently outperformed all other approaches tested.
+This performance difference was more pronounced as the dimension of the search problem increased; for a 300-dimensional Rastrigin function, DOPS was the only approach to find an optimal or near optimal solution
 within the function evaluation budget (Fig. \ref{fig-testfunctions}B).
-Taken together, DOPS performed at least as well as other meta-heuristic approaches on small test problems, but was especially suited to large dimension search spaces.
+Taken together, DOPS performed at least as well as other meta-heuristic approaches on small dimensional test problems, but seemed especially suited to large dimensional search spaces.
 Next, we tested DOPS on benchmark biochemical models of varying complexity.
@@
 Villaverde and co-workers published a set of benchmark biochemical problems to evaluate parameter estimation methods \cite{villaverde2015biopredyn}.
-They ranked ordered these example problems by computational cost from most to least expensive.
+They ranked the example problems by computational cost from most to least expensive.
 We evaluated the performance of DOPS on problems from the least and most expensive categories.
 The least expensive problem (henceforth referred to as problem B4) was a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters \cite{villaverde2014high}.
-The biochemical reactions were modeled using modular rate laws and the generalized form of Michaelis-Menten kinetics.
-The expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters \cite {smallbone2013large}(henceforth referred to as problem B1).
-In both cases, synthetic time series data sets were generated using known parameter values, these data were then used by the parameter estimation methods.
+The biochemical reactions were modeled using modular rate laws and generalized Michaelis-Menten kinetics.
+On the other hand, the expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters \cite{smallbone2013large} (henceforth referred to as problem B1).
+In both cases, synthetic time series data sets, generated using known parameter values, were used to estimate model parameters.
 For problem B1, the time series data consisted of 44 observables, and for problem B4 the data corresponded to 13 different metabolite measurement sets.
 We fixed the number of function evaluations at $\mathcal{N}$ = 4000, and trained both models against the synthetic experimental data.
-In both cases we found good fits  (Fig. \ref{fig-sims-b1} and Fig. \ref{fig-sims-b4}) to the problems.
-We also recaptured the nominal parameter values within the specified number of function evaluations (Fig. \ref{fig-benchmark}).
-Thus, DOPS estimated the parameters in benchmark biochemical models, and recovered the original parameters from synthetic data.
-Next, we compared the performance of DOPS with the four other meta-heuristic approaches for a model of the human coagulation cascade.
+In both cases DOPS, produced good fits to the synthetic data (Fig.
+\ref{fig-sims-b1} and Fig.
+\ref{fig-sims-b4}), and recapitulated the nominal parameter values using only $\mathcal{N}\leq~4000$ function evaluations (Fig.
+\ref{fig-benchmark}).
+On the other hand, enhanced scatter search (eSS) with a local optimizer took on order $10^{5}$ function evaluations for the same problems.
+Thus, DOPS estimated the parameters in benchmark biochemical models, and recovered the original parameters from synthetic data, using fewer function evaluations.
+Next, we compared the performance of DOPS with the four other meta-heuristics for a model of the human coagulation cascade.
 

### 2016-04-27 Paper_v5.tex
Polish and refine

@@
 On the other hand, the expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters \cite{smallbone2013large} (henceforth referred to as problem B1).
-In both cases, synthetic time series data sets, generated using known parameter values, were used to estimate model parameters.
+In both cases, synthetic time series data generated with known parameter values, was used to estimate model parameters.
 For problem B1, the time series data consisted of 44 observables, and for problem B4 the data corresponded to 13 different metabolite measurement sets.
 We fixed the number of function evaluations at $\mathcal{N}$ = 4000, and trained both models against the synthetic experimental data.
-In both cases DOPS, produced good fits to the synthetic data (Fig.
+In both cases, DOPS produced good fits to the synthetic data (Fig.
 \ref{fig-sims-b1} and Fig.

### 2016-04-27 Paper_v5.tex
Polish and refine

@@
 We fixed the number of function evaluations at $\mathcal{N}$ = 4000, and trained both models against the synthetic experimental data.
-In both cases, DOPS produced good fits to the synthetic data (Fig.
+DOPS produced good fits to the synthetic data (Fig.
 \ref{fig-sims-b1} and Fig.
@@
 The central process in coagulation is the conversion of prothrombin (fII), an inactive coagulation factor, to the master protease thrombin (FIIa).
-Thrombin generation involves three phases, initiation, amplification and termination [13,14].
+Thrombin generation involves three phases, initiation, amplification and termination.
 Initiation requires a trigger event, for example a vessel injury which exposes tissue factor (TF), which leads to the activation of factor VII (FVIIa) and the formation of the TF/FVIIa complex.
@@
 There are several control points in the cascade that inhibit thrombin formation, and eventually terminate thrombin generation.
-Tissue Factor Pathway Inhibitor (TFPI) inhibits upstream activation events, while antithrombin III (ATIII) neutralizes several of the proteases generated during coagulation, including thrombin.Thrombin itself also inadvertently plays a role in its own inhibition; thrombin, through interaction with thrombomodulin, protein C and endothelial cell protein C receptor (EPCR), converts protein C to activated protein C (APC) which attenuates the coagulation response by proteolytic cleavage of amplification complexes.
+Tissue Factor Pathway Inhibitor (TFPI) inhibits upstream activation events, while antithrombin III (ATIII) neutralizes several of the proteases generated during coagulation, including thrombin.
+Thrombin itself also inadvertently plays a role in its own inhibition; thrombin, through interaction with thrombomodulin, protein C and endothelial cell protein C receptor (EPCR), converts protein C to activated protein C (APC) which attenuates the coagulation response by proteolytic cleavage of amplification complexes.
 Termination occurs after either prothrombin is consumed, or thrombin formation is neutralized by inhibitors such as APC or ATIII.
@@
 coagulation is challenging because it contains both fast and slow dynamics, but also accessible because of the availability of comprehensive data sets for model identification and validation.
-In this study, we used the coagulation model of Luan et al \cite{luan2007computationally}, which is a coupled system of non-linear ordinary differential equations where biochemical interactions are modeled using mass action kinetics.
+In this study, we used the coagulation model of Luan et al.
+\cite{luan2007computationally}, which is a coupled system of non-linear ordinary differential equations where biochemical interactions were modeled using mass action kinetics.
 The Luan model contained 148 parameters and 92 species and has been validated using 21 published experimental datasets.
@@
 
-DOPS estimated the parameters of a human coagulation model for TF/VIIa initiated coagulation without anticoagulants (Fig.\ref{fig-train}).
+DOPS estimated the parameters of a human coagulation model for TF/VIIa initiated coagulation without anticoagulants (Fig. \ref{fig-train}).
 The objective function was an unweighted linear combination of two error functions,
 representing coagulation initiated with different concentrations of TF/FVIIa (5pM, 5nM) \cite{hockin2002model}.
-We restricted the number of function evaluations to $\mathcal{N}$ = 4000 for each algorithm we tested, and performed 25 trials of each experiment to collect average performance data (Table \ref{table:benchmark-problems}).
-DOPS converged faster and had a lower final error compared to the other algorithms (Fig.
-\ref{fig-convergence}).
+We restricted the number of function evaluations to $\mathcal{N}$ = 4000 for each algorithm we tested, and performed $\mathcal{T}$ = 25 trials of each experiment to collect average performance data (Table \ref{table:benchmark-problems}).
+DOPS converged faster and had a lower final error compared to the other algorithms (Fig. \ref{fig-convergence}).
 Within the first 25\% of function evaluations, DOPS produced a rapid drop in error followed by a slower but steady decline.
-Approximately between 500-1000 function evaluations DOPS switched to the dynamically dimensioned search phase, however this switch iteration varied from trial to trial since the switch was based on the error from the swarm phase.
-Overall, DOPS minimized the error to a greater extent than the other meta-heuristic approaches.
-However, it was unclear if the parameters obtained by DOPS had predictive power on unseen data. To address this question, we used the final parameters estimated by
-DOPS to simulate data that was not used in training (coagulation initiated with 500pM,50pM,10pM TF/VIIa).
-The optimal or near optimal parameters obtained by DOPS predicted unseen coagulation datasets (Fig.\ref{fig-validation}).
+Approximately between 500-1000 function evaluations DOPS switched to the dynamically dimensioned search phase, however this transition varied from trial to trial since the switch was based upon the local convergence rate.
+On average, DOPS minimized the coagulation model error to a greater extent than the other meta-heuristics.
+However, it was unclear if the parameters estimated by DOPS had predictive power on unseen data.
+To address this question, we used the final parameters estimated by
+DOPS to simulate data that was not used for training (coagulation initiated with 500pM, 50pM, and 10pM TF/VIIa).
+The optimal or near optimal parameters obtained by DOPS predicted unseen coagulation datasets (Fig. \ref{fig-validation}).
+The normalized standard error for the coagulation predictions was consistent with the training error, with the exception of the 50pM TF/VIIa case which was a factor 2.75 worse (Table \ref{table:Error-Analysis}).
+Coagulation initiation with 50pM TF/FVIIa was the farthest away from the training conditions.
+However, while the 50pM TF/FVIIa case was the worst of those tested, it still had an average error rate of only 11\% (which is near the margin of error for the experimental measurements).
 Taken together, DOPS estimated parameter sets with predictive power on unseen coagulation data using fewer function iterations than other meta-heuristics.
-Next, we explored how the number of sub-swarms and the switch to DDS during the initial phase of an optimization trial influenced the performance of the approach.
+Next, we explored how the number of sub-swarms and the switch to DDS influenced the performance of the approach.
 
@@
 \hline\\
-TF/FVIIa concentration& Normalized S.E.& Data Set Category\\\\
+TF/FVIIa concentration& Normalized S.E.& Category\\\\
 \hline\\

### 2016-04-28 Paper_v5.tex
Smooth and refine

@@
 
-\subsection*{The number of sub-swarms and phase switching influenced DOPS performance.}
-A differentiating feature of DOPS compared to other meta-heuristics was the rapid error drop during the first 25\% of an optimization trial.
-This drop occurred during the swarm phase and was punctuated by the switch to the dynamically dimensioned search phase of the approach.
+\subsection*{Phase switching was critical to DOPS performance.}
+A differentiating feature of DOPS compared to the other meta-heuristics is the switch to dynamically dimensioned search following stagnation of the initial particle swarm phase.
 We quantified the influence of the number of sub-swarms and switch to the DDS phase on error convergence by comparing DOPS with and without DDS for different numbers of sub-swarms (Fig.
 \ref{fig-sub-swarms}).
-We considered only the multi swarm search without the DDS phase for $\mathcal{N}$ = 4000 function evaluations on the coagulation model.
-While doing the multi swarm search without DDS, we used one, two, four, five and eight sub-swarms, with a total of 40 particles divided evenly amongst the swarms.
-Hence we did not consider swarm numbers of three and seven.
-All the other parameters remained the same for each of these searches.
-Generally, the higher sub-swarm numbers had faster convergence.
-However, the difference in convergences was not pronounced amongst four, five and eight, suggesting
-there was an optimal number of particles per swarm beyond which there was no significant advantage.
-We also observed that convergence rates are compared with DOPS, DOPS shows a very rapid drop when the swarm searches begin to saturate.
-The green region indicates when swarm search switches to a DDS based search.
-The error convergence rates tend to remain mostly flat for the purely swarm searches with different swarm numbers.
-However in DOPS as soon as convergence rate begins to saturate it automatically switches to a DDS phase which leads to a more pronounced drop.
-Thus, this automated switching strategy appears to be crucial in leading the algorithm out of local optima or saddle point regions.
+We considered multi swarm particle swarm optimization with and without the DDS phase for $\mathcal{N}$ = 4000 function evaluations and $\mathcal{T}$ = 25 trials on the coagulation model.
+We used one, two, four, five and eight sub-swarms, with a total of 40 particles divided evenly amongst the swarms.
+Hence, we did not consider swarm numbers of three and seven.
+All other algorithm parameters remained the same for each of these searches.
+Generally, the higher sub-swarm numbers converged in fewer function evaluations, where the optimum particle partitioning was in the neighborhood of five sub-swarms.
+However, the difference in convergence rate was qualitatively similar for four, five and eight sub-swarms, suggesting there was an optimal number of particles per swarm beyond
+which there was no significant advantage.
+The multi-swarm particle swarm optimization phase stagnated within the first 25\% of the available function evaluations irrespective of the number of sub-swarms.
+The DDS phase refined the globally best particle to produce significantly lower error on average when compared to multi-swarm particle swarm optimization alone.
+Thus, the automated switching strategy was critical to the overall performance of DOPS.
 
@@
 \centering
-\caption[.]{Error Analysis.
-Normalized standard error showing the agreement between coagulation model dynamics and the coagulation experimental data.
-The model is trained on the 5~nM and the 5~pM cases to obtain the optimal parameters.
-Using these optimal parameters, dynamics of coagulation are predicted with varying initiator concentrations (500~pM, 50~pM and 10~pM).
-The normalized S.E.
-is defined as \( N.S.E.=(1/max(\mathbf{X}))*(\lVert (\mathbf{Y},\mathbf{X}) \rVert /sqrt$(N)$) \) where \textbf{X} is the experimental data, \textbf{Y} is the model simulation data interpolated on experimental time and $N$ is the total number of experimental time points.}
-\label{table:Error-Analysis}
-\begin{center}
-\begin{tabular}{ r|ll}
-\hline\\
-TF/FVIIa concentration& Normalized S.E.& Category\\\\
-\hline\\
-5~nM & 0.0376 & Training \\
-500~pM & 0.0564 & Prediction\\
-50~pM & 0.1125 & Prediction\\
-10~pM & 0.0823 & Prediction\\
-5~pM & 0.0338 & Training\\\\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[h]
-\centering
 \caption{Table with optimization settings and results for the coagulation problem, the benchmarks and test functions using DOPS.
@@
 
+\begin{table}[h]
+\centering
+\caption[.]{Error Analysis.
+Normalized standard error showing the agreement between coagulation model dynamics and the coagulation experimental data.
+The model is trained on the 5~nM and the 5~pM cases to obtain the optimal parameters.
+Using these optimal parameters, dynamics of coagulation are predicted with varying initiator concentrations (500~pM, 50~pM and 10~pM).
+The normalized S.E.
+is defined as \( N.S.E.=(1/max(\mathbf{X}))*(\lVert (\mathbf{Y},\mathbf{X}) \rVert /sqrt$(N)$) \) where \textbf{X} is the experimental data, \textbf{Y} is the model simulation data interpolated on experimental time and $N$ is the total number of experimental time points.}
+\label{table:Error-Analysis}
+\begin{center}
+\begin{tabular}{ r|ll}
+\hline\\
+TF/FVIIa concentration& Normalized S.E.& Category\\\\
+\hline\\
+5~nM & 0.0376 & Training \\
+500~pM & 0.0564 & Prediction\\
+50~pM & 0.1125 & Prediction\\
+10~pM & 0.0823 & Prediction\\
+5~pM & 0.0338 & Training\\\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \clearpage
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure_1_Algorithm.pdf}
-\caption{Dynamic Optimization with Particle Swarms.
+\caption{Schematic of the dynamic optimization with particle swarms (DOPS) approach.
 \textbf{A}: Each particle represents an N dimensional parameter vector.
@@
 The probability that a dimension is perturbed decreases as the number of function evaluations increase.
-Thus as the evaluations increase the optimality of the solution is preserved.
 }\label{fig-algorithm}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{./figs/Figure_2_TestFunctions-crop.pdf}
+\includegraphics[width=0.60\textwidth]{./figs/Figure_2_TestFunctions-crop.pdf}
 \caption{Performance of DOPS and other meta-heuristics for the Ackley and Rastrigin functions.

### 2016-05-01 Paper_v5.tex
A draft second paragrpah

@@
 \subsection*{Phase switching was critical to DOPS performance.}
-A differentiating feature of DOPS compared to the other meta-heuristics is the switch to dynamically dimensioned search following stagnation of the initial particle swarm phase.
-We quantified the influence of the number of sub-swarms and switch to the DDS phase on error convergence by comparing DOPS with and without DDS for different numbers of sub-swarms (Fig.
+A differentiating feature of DOPS is the switch to dynamically dimensioned search following stagnation of the initial particle swarm phase.
+We quantified the influence of the number of sub-swarms and the switch to DDS on error convergence by comparing DOPS with and without DDS for different numbers of sub-swarms (Fig.
 \ref{fig-sub-swarms}).
@@
 Hence, we did not consider swarm numbers of three and seven.
-All other algorithm parameters remained the same for each of these searches.
+All other algorithm parameters remained the same for all cases.
 Generally, the higher sub-swarm numbers converged in fewer function evaluations, where the optimum particle partitioning was in the neighborhood of five sub-swarms.
@@
 which there was no significant advantage.
-The multi-swarm particle swarm optimization phase stagnated within the first 25\% of the available function evaluations irrespective of the number of sub-swarms.
+The multi-swarm particle swarm optimization stagnated after 25\% of the available function evaluations irrespective of the number of sub-swarms.
+However, DOPS (with five sub-swarms) switched to DDS after detecting the stagnation.
 The DDS phase refined the globally best particle to produce significantly lower error on average when compared to multi-swarm particle swarm optimization alone.
@@
 %Thus, for expensive objective functions in high dimensional parameter spaces, meta-heuristic approaches can be inefficient.
+%Generation of diverse solutions in the early stage gives a better exploratory capability and thus improves convergence.
+%Though dynamically dimensioned search is based on a single solution, the decision vector carries forward from one iteration to the next.
+%Hence, it behaves similar to a population based search.
 
 Meta-heuristics can be effective techniques to estimate optimal or nearly optimal solutions for complex, multi-modal functions.
+However, meta-heuristics typically require a large number of function evaluations to converge to a solution.
 DOPS is a combination of particle swarm optimization, which is a global search method, and dynamically dimensioned search, which is a greedy evolutionary technique.
-Particle swarm optimization is a population based meta-heuristic which uses collective information shared amongst swarms of computational particles to
-search for global extrema.
-Several particle swarm variants have been proposed to improve the search ability and rate of convergence.
+Particle swarm optimization uses collective information shared amongst swarms of computational particles to
+search for global extrema. Several particle swarm variants have been proposed to improve the search ability and rate of convergence.
 These variations involve different neighborhood structures, multi-swarms or adaptive parameters.
-Multi-swarm PSO with small particle neighborhoods have been shown to better in searching on complex multi-modal solutions \cite{zhao2008dynamic}.
-Multi swarm methods, in addition, avoid rapid convergence to a local optimum or stable point and are able to generate diverse solutions.
-Generation of diverse solutions in the early stage gives a better exploratory capability and thus converge of upon multiple optima.
-Tolson and Shoemaker, through DDS, showed that randomly perturbing a subset of dimensions in high dimensional parameter space is an effective way to obtain near optimal solutions with few function evaluations.
-Though their approach is based on a single solution, the decision vector carries forward from one iteration to the next.
-Hence it behaves similar to a population based search.
-In our approach, we tried to eliminate the probability of starting from a bad region of search using a variant of particle swarm optimization.
-We utilized this capacity of multi swarms to generate diverse candidate solutions which can be used as initial solutions for a DDS like search.
-Thus the solution vectors obtained at the end of multi swarm search phase have a better propensity to not get stuck in bad regions in DDS phase as opposed to starting with a totally random initial solution.
-Given a fixed number of function evaluations, we have been able to show that we were able to obtain better solutions for the coagulation model and the test functions faster that other commonly known meta-heuristics and also DDS alone.
+Multi-swarm particle swarm optimization with small particle neighborhoods has been shown to be better in searching on complex multi-modal solutions \cite{zhao2008dynamic}.
+Multi swarm methods generate diverse solutions, and avoid rapid convergence to local optima.
+However, at least for the coagulation problem used in this study, multi-swarm methods stagnated after approximately 25\% of the available function evaluations; only the introduction of dynamically dimensioned search improved the rate of error convergence.
+Dynamically dimensioned search, which greedily perturbs only a subset of parameter dimensions in high dimensional parameter spaces,
+refined the globally best particle and produced significantly lower error on average when compared to multi-swarm particle swarm optimization alone.
+However, dynamically dimensioned search, starting from a initial random parameter guess, was not as effective on average as DOPS.
+The initial solutions generated by the multi swarm search had a higher propensity to produce good parameter estimates when refined by dynamically dimensioned search.
+Thus, our hybrid combination of two meta-heuristics produced better results than either constituent approach, and better results than other meta-heuristic approaches on average.
+This was true of not only the convergence rate on the coagulation problem, but also the biochemical benchmark problems;
+DOPS required two-orders of magnitude fewer function evaluations compared with enhanced Scatter Search (eSS) to estimate the biochemical benchmark model parameters.
+Taken together, the combination of particle swarm optimization and dynamically dimensioned search produced better results than either of these constituent approaches alone,
+and required fewer function evaluations compared with other common meta-heuristics.
+
+%to not get stuck in bad regions in DDS phase as opposed to starting with a totally random initial solution.
+%In our approach, we tried to eliminate the probability of starting from a bad region of search using a variant of particle swarm optimization.
+%Tolson and Shoemaker, through DDS, showed that randomly perturbing a subset of dimensions  is an effective way to obtain near optimal solutions with few function evaluations.
+%We utilized this capacity of multi swarms to generate diverse candidate solutions which can be used as initial solutions for a DDS like search.
+Thus  Given a fixed number of function evaluations, we have been able to show that we were able to obtain better solutions for the coagulation model and the test functions faster that other commonly known meta-heuristics and also DDS alone.
 The computational cost was prohibitively high for parameter estimation on this problem using a standard PSO.
@@
 The amount of CPU time taken (on an Intel Xeon processor 2.4 GHz) is lesser as compared to eSS on a similar architecture.
-A surprisingly remarkable aspect about our algorithm has been that we did not 'pre optimize' any parameters of the algorithm to suit a specific problem.
+%A surprisingly remarkable aspect about our algorithm has been that we did not 'pre optimize' any parameters of the algorithm to suit a specific problem.
 In the swarm phase this includes the number of particles, number of sub-swarms, acceleration constants or the number of generations after which the particles are redistributed and the neighborhood perturbation parameter in DDS phase.

### 2016-05-01 Paper_v5.tex
Relase candidate

@@
 Multi-swarm particle swarm optimization with small particle neighborhoods has been shown to be better in searching on complex multi-modal solutions \cite{zhao2008dynamic}.
-Multi swarm methods generate diverse solutions, and avoid rapid convergence to local optima.
+Multi-swarm methods generate diverse solutions, and avoid rapid convergence to local optima.
 However, at least for the coagulation problem used in this study, multi-swarm methods stagnated after approximately 25\% of the available function evaluations; only the introduction of dynamically dimensioned search improved the rate of error convergence.
@@
 Thus, in this approach any overhead that usually comes with additional function evaluations to pre optimize parameters was avoided.
+%In addition, surprisingly, we were able to obtain optimal parameter vectors for two different large scale systems biology models with a couple of orders fewer number of function evaluations as compared to enhanced Scatter Search (eSS).
+%In addition, In the DDS phase, the search uses only one candidate vector although there is a provision to start with a population of candidate vectors.
+Incorporating a more intelligent switching strategy that can do switch from swarm phase to DDS phase multiple times and having a population of candidate vectors are some aspects of the algorithm that can be studied further.
 
-The performance of DOPS was impressive given that it performed well on different complex systems with no pre optimization of algorithm parameters being required.
-We comfortably outperformed existing, widely used meta-heuristics and were also able to find minima of high dimensional global optimization test functions.
-Thus this approach may be well suited to large scale global optimization.
-In addition, surprisingly, we were able to obtain optimal parameter vectors for two different large scale systems biology models with a couple of orders fewer number of function evaluations as compared to enhanced Scatter Search (eSS).
-However it is quite possible that highly optimized versions of common meta heuristics may outperform us on these systems.
-This aspect is currently beyond the scope of this study.
-Our approach can also be combined with local derivative based searches to improve upon the accuracy of the solutions.
-In addition, the current implementation of the algorithm is designed to switch only once from the swarm phase to the DDS.
-In the DDS phase, the search uses only one candidate vector although there is a provision to start with a population of candidate vectors.
-Incorporating a more intelligent switching strategy that can do switch from swarm phase to DDS phase multiple times and having a population of candidate vectors are some aspects of the algorithm that can be studied further.
+DOPS performed well on many different systems with no pre-optimization of algorithm parameters, however there are many research questions that should be pursued further.
+DOPS comfortably outperformed existing, widely used meta-heuristics for high dimensional global optimization functions, biochemical benchmark models and a model of the human coagulation system.
+However, it is possible that highly optimized versions of common meta-heuristics could surpass DOPS;
+we should compare the performance of DOPS with optimized versions of the common
+meta-heuristics on both test and real-world problems to determine if a performance advantage exists in practice.
+Next, DOPS has a hybrid architecture, thus the particle swarm phase could be combined with other search strategies such as local derivative based approaches to improve convergence rates.
+We could also consider multiple phases beyond particle swarm and dynamically dimensioned search, for example switching to a gradient based search following the dynamically
+dimensioned search phase.
+Next, the current implementation of the algorithm switches only once from the swarm phase to the dynamically dimensioned phase.
+An obvious modification would be to allow multiple transitions between the swarm and dynamically dimensioned phases.
+Lastly, we should update DOPS to treat multi-objective problems.
+The identification of large biochemical models sometimes requires training using qualitative, conflicting or even contradictory data sets.
+One strategy to address this challenge is to estimate experimentally constrained model ensembles using multiobjective optimization.
+Previously, we developed Pareto Optimal Ensemble Techniques (POETs) which integrates simulated annealing with Pareto optimality
+to identify models near the optimal tradeoff surface between competing training objectives \cite{song2010ensembles}.
+Since DOPS consistently outperformed simulated annealing on both test and real-world problems, we expect a multi-objective form of DOPS would more quickly estimate solutions
+which lie along high dimensional trade-off surfaces.
 

### 2016-05-02 Paper_v5.tex
Submitted. Keep your fingers crossed

@@
 \section*{Abstract}
-Mathematical modeling is a powerful tool to analyze, and ultimately design biochemical networks.
+Mathematical modeling is a powerful tool to analyze, and design biochemical networks.
 However, the estimation of biochemical model parameters is a significant challenge.
-Parameter estimation in biochemical models typically involves expensive function evaluations and noisy data, making it difficult to quickly obtain optimal solutions.
+Parameter estimation typically involves expensive function evaluations and noisy data, making it difficult to quickly obtain optimal solutions.
 Biochemical models often also have many local extrema which further complicates parameter estimation.
-Toward these challenges, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel global meta-heuristic that combined features of multi-swarm particle swarm optimization with dynamically dimensioned search (DDS).
+Toward these challenges, we developed Dynamic Optimization with Particle Swarms (DOPS), a novel hybrid meta-heuristic that combined multi-swarm particle swarm optimization with dynamically dimensioned search (DDS).
 DOPS uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors, the best of which is greedily updated using dynamically dimensioned search.
-We tested the performance of DOPS on a model of human coagulation cascade.
+We tested DOPS using a combination of classic optimization test functions, biochemical benchmark problems and real-world biochemical models.
 We performed $\mathcal{T}$ = 25 trials with $\mathcal{N}$ = 4000 function evaluations per trial, and compared the performance of DOPS with other commonly
 used meta-heuristics such as differential evolution (DE), simulated annealing (SA) and dynamically dimensioned search (DDS).
-We further tested the predictive power of the coagulation model parameters against data not used in training, and found good agreement between simulations and experimental measurements.
+On average, DOPS outperformed other common meta-heuristics on the optimization test functions,
+benchmark problems and a model of the human coagulation cascade.
+Taken together, DOPS is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.
+DOPS source code is available for download under a MIT license at http://www.varnerlab.org.
+
+%We further tested the predictive power of the coagulation model parameters against data not used in training, and found good agreement between simulations and experimental measurements.
 Lastly, we tested the performance of DOPS on commonly used test functions for global optimization and on published biochemical parameter estimation benchmark problems.
 For the wide range of problems that we considered, DOPS outperformed other meta-heuristic approaches despite a limited number of function evaluations.
-Taken together, DOPS is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.
 
 \vspace{0.1in}
-{\noindent \textbf{Keywords:}~Parameter identification, Meta-heuristic optimization, Biochemical modeling}
+{\noindent \textbf{Keywords:}~Biochemical engineering, systems biology, modeling, optimization}
 
@@
 We tested DOPS using a combination of classic optimization test functions, biochemical benchmark problems and real-world biochemical models.
-First, we tested the performance of DOPS on commonly used test functions for global optimization (Ackley and Rosenbrock) and published biochemical benchmark problems.
+First, we tested the performance of DOPS on the Ackley and Rosenbrock functions, and published biochemical benchmark problems.
 Next, we used DOPS to estimate the parameters of a model of the human coagulation cascade.
@@
 The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model parameters ($\mathcal{K}\times{1}$).
-Equation (3) is similar to the general particle swarm update rule, however, it does not contain velocity terms.
-In DOPS, the parameter $\theta_{1,j-1}$ is similar to the inertia weight parameter for the velocity term described by Shi and Eberhart \cite{shi1999empirical}.
-In this study Shi and Eberhart propose a linearly decreasing inertia weight to improve convergence properties of PSO.
-$\theta_{1,j-1}$ is inspired from this idea and also from the idea of decreasing perturbation probability proposed by Tolson and Shoemaker \cite{tolson2007dynamically}.
+Equation \eqref{qn:update-rule} is similar to the general particle swarm update rule, however, it does not contain velocity terms.
+In DOPS, the parameter $\theta_{1,j-1}$ is similar to the inertia weight parameter for the velocity term described by Shi and Eberhart \cite{shi1999empirical};
+Shi and Eberhart proposed a linearly decreasing inertia weight to improve convergence properties of particle swarm optimization.
+Our implementation of $\theta_{1,j-1}$ is inspired by this and the decreasing perturbation probability proposed by Tolson and Shoemaker \cite{tolson2007dynamically}.
 It is an analogous equivalent to inertia weight on velocity.
-However $\theta_{1,j-1}$ places inertia on the position rather than velocity and uses the same rule described by Shi and Eberhart to adaptively change with the number of function evaluations.
-It is updated according to:
+However $\theta_{1,j-1}$ places inertia on the position rather than velocity and uses the same rule described by Shi and Eberhart to adaptively change with the number of function evaluations:
 \begin{eqnarray}
@@
 where $\mathcal{N}$ represents the total number of function evaluations, ${w}_{max}$ and ${w}_{min}$ are the maximum and minimum inertia weights, respectively.
-While updating the particles, we made sure all dimensions of the solution represented by the particle were within bounds using reflection boundary conditions (Algo. \ref{algo_reflection}).
+While updating the particles, parameter bounds were enforced using reflection boundary conditions (Algo. \ref{algo_reflection}).
 
@@
 DDS updates are greedy; $\mathcal{G}_{new}$ becomes the new solution vector only if it is better than $\mathcal{G}$.
+DOPS source code is available for download under a MIT license at http://www.varnerlab.org.
 
@@
 \centering
-\caption[.]{Error Analysis.
-Normalized standard error showing the agreement between coagulation model dynamics and the coagulation experimental data.
-The model is trained on the 5~nM and the 5~pM cases to obtain the optimal parameters.
-Using these optimal parameters, dynamics of coagulation are predicted with varying initiator concentrations (500~pM, 50~pM and 10~pM).
-The normalized S.E.
-is defined as \( N.S.E.=(1/max(\mathbf{X}))*(\lVert (\mathbf{Y},\mathbf{X}) \rVert /sqrt$(N)$) \) where \textbf{X} is the experimental data, \textbf{Y} is the model simulation data interpolated on experimental time and $N$ is the total number of experimental time points.}
+\caption[.]{Error analysis for the human coagulation model. The coagulation model was trained on coagulation initiated with TF/FVIIa at 5~nM and the 5~pM to obtain the optimal parameters.
+Using these optimal parameters, coagulation dynamics were predicted for varying initiator concentrations (500~pM, 50~pM and 10~pM).
+Model agreement with measurements was quantified using normalized squared error.
+The normalized squared error is defined as \( N.S.E.=(1/max(\mathbf{X}))*(\lVert (\mathbf{Y},\mathbf{X}) \rVert /sqrt$(N)$) \) where \textbf{X} is the experimental data, \textbf{Y} is the model simulation data interpolated onto the experimental time scale and $N$ is the total number of experimental time points.}
 \label{table:Error-Analysis}
@@
 DOPS is the only algorithm that finds an optimal or near optimal solution within the specified number of function evaluations.
-In all cases, the maximum number of function evaluations was 4000.
-Mean and standard deviation were calculated over 25 trials.
+In all cases, the maximum number of function evaluations was $\mathcal{N}$ = 4000.
+Mean and standard deviation were calculated over $\mathcal{T}$ = 25 trials.
 }\label{fig-testfunctions}
@@
 \includegraphics[width=1.00\textwidth]{./figs/Figure_3_Coagulation_v4.pdf}
-\caption{Schematic of the extrinsic and intrinsic coagulation cascade\cite{luan2007computationally}.
+\caption{Schematic of the extrinsic and intrinsic coagulation cascade.
 Inactive zymogens upstream (grey) are activated by exposure to tissue factor (TF)  following vessel injury.
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure_4_Errors_convergence_v2.pdf}
-\caption{Error convergence rates of the five different algorithms on the coagulation model.
-The objective error is the mean over $N$= 25 trials.
-DOPS, DDS and SA have the steepest drop in error during first 300 function evaluations.
-Thereafter the error drop in DDS and SA remains nearly constant whereas DOPS continues to drops further.
-At the end of 4000 function evaluations DOPS attains the lowest error.
-The next best estimate using DDS is nearly 3 times greater than the lowest error using DDS.
+\caption{Error convergence rates of the five different algorithms on the coagulation model. The objective error is the mean over $\mathcal{T}$= 25 trials.
+DOPS, DDS and SA have the steepest drop in error during first 300 function evaluations. Thereafter the error drop in DDS and SA remains nearly constant whereas DOPS continues to drops further.
+At the end of 4000 function evaluations DOPS attains the lowest error. The next best estimate using DDS is nearly three times greater than the lowest error using DDS.
 }\label{fig-convergence}
@@
 The parameter estimates that were obtained using DOPS were tested against data that was not used in the model training.
-Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from $N$= 25 trials.
+Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from $\mathcal{T}$ = 25 trials.
 The grey shaded region represents the 99\% confidence estimate of the mean simulated thrombin concentration.
 The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers.
-Thrombin generation is initiated by adding Factor VIIa-TF (500pM - Blue, 50pM - Pink and 10pM - Orange respectively) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
+Thrombin generation is initiated by adding Factor VIIa-TF (500pM - Blue, 50pM - Pink and 10pM - orange, respectively) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
 }\label{fig-validation}
@@
 \includegraphics[width=1.0\textwidth]{./figs/Figure_7_ErrorSwarms_v2.pdf}
-\caption{DOPS (color:black) begins by using a particle swarm search (color:pale brown) and then dynamically switches (color:green), using an adaptive switching criteria, to the DDS search phase (color:white).
-The particle swarm search uses multiple sub-swarms wherein each solution represented by a particle is updated using equation \ref{eqn:update-rule}.
-The particle update is influenced by the best particle amongst all particles in the sub-swarm and the best solution found by the particle till the current iteration.
-This update rule however contains no velocity terms.
-The particle updates continue to happen within the sub-swarms for a certain number of iterations (or function evaluations) after which the sub-swarms are reorganized which is similar to the regrouping strategy described by Zhao et al.
-\cite{zhao2008dynamic}.
-DOPS switches out of this PSO phase based on an adaptive switching criteria (color:green) that is a function of error convergence rate.
-If the error represented by the best particle does not drop for a threshold number of function evaluations, DOPS switches automatically to the DDS search phase.
-Since the algorithm is stochastic, the switch out can happen at a different function evaluation in each trial.
-The DDS search then uses the best particle from the PSO search phase as its initial solution or candidate vector.
-This particle is greedily updated by perturbing a subset of dimensions for the remaining number of function evaluations.
-The number of dimensions perturbed is a monotonically decreasing function that generally depends on the number of function evaluations within the DDS phase.
-We compared the performance of DOPS with multi-swarm searches without DDS to quantify the effect of number of sub-swarms.
+\caption{Influence of the switching strategy and sub-swarms on DOPS performance for the coagulation model.
+DOPS begins by using a particle swarm search and then dynamically switches (switch region), using an adaptive switching criteria, to the DDS search phase.
+We compared the performance of DOPS with and without DDS for different sub-swarm searches to quantify the effect of number of sub-swarms and DDS.
 We used one, two, four, five and eight sub-swarms, with a total of 40 particles divided evenly amongst the swarms.
+The results presented are the average of $\mathcal{T}$ = 25 trials with $\mathcal{N}$ = 4000 function evaluations each.
 The convergence rates with higher swarm numbers is typically higher but there is no pronounced difference amongst four, five and eight.
-When the multi-swarm only searches tend to saturate DOPS shows a rapid drop due to a switch to  the DDS phase.
+The multi-swarm with DDS saturates while DOPS shows a rapid drop due to a switch to the DDS phase.
 }\label{fig-sub-swarms}

### 2016-05-02 DOPS/DDSPSO_strategy1.m
DOPS Code

@@ -1,0 +1,220 @@
+%Particle swarm optimization that implements dynamically dimensioned search
+
+function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=DDSPSO_strategy1(MAXJ,MINJ,NP,NI,NS,G,r)
+
+
+%se=load('randnseed.txt');
+%randn('seed',se);
+%rand('seed',se);
+
+init=0;
+Max_Inertia_weight=0.9;
+Min_Inertia_weight=0.4;
+C1=1.5;C2=1.5;
+beta=3.3;
+N=length(MINJ);
+SUB_SWARM_SIZE=round(NP/NS);
+failure_counter=0;
+success_counter=0;
+
+success_counter_threshold=4;
+failure_counter_threshold=4;
+rho(:,1)=ones(length(MINJ),1);
+dds_swarm_flag=0;
+%FR=0.4; %Fraction of iterations in DDSMLSPSO
+
+%Initializing the position of the particles within the swarms
+
+
+for i=1:NP
+    %Restricting the perturbation to be amongst the bounds specified
+    Z(:,i)=MINJ+(MAXJ-MINJ).*rand(N,1);
+    [Z(:,i)]=bind(Z(:,i),MINJ,MAXJ);
+    %Calculating the cost
+    particle(:,1,i)=Z(:,i);
+    fitness(i,1)=fit7(particle(:,1,i));
+
+end
+
+
+%Columns of S matrix have the indices of the particles in each swarm - NS*SUB_SWARM_SIZE matrix
+[S]=newswarms(NP,NS,SUB_SWARM_SIZE);
+velocity=0.729*particle;
+
+%Finding particle best and global best after initialization within each sub
+%swarm
+for j=1:NS
+    index_particle=S(:,j);
+
+    [ls_pbest_solution(j,:),ls_ITER(j,:)]=particlebest(fitness(index_particle,:));
+    [ls_gbest_solution(j),temp_particlenumber]=globalbest(ls_pbest_solution(j,:));
+    particlenumber(j)=index_particle(temp_particlenumber);
+end
+
+%Finding the best solution among all the swarms
+[g_best_solution(1),swarmnumber]=globalbest(ls_gbest_solution);
+bestparticle(:,1)=particle(:,ls_ITER(swarmnumber(1),S(:,swarmnumber(1))==particlenumber(swarmnumber(1))),particlenumber(swarmnumber(1)));
[…]

### 2016-05-02 DOPS/DOPS_DDS.m

@@ -1,0 +1,74 @@
+function[best,x_best,dds_swarm_flag]=DDS(IC,MAXJ,MINJ,r,NI)
+
+
+N=length(MINJ);
+x_best(:,1)=IC;
+x=IC;
+J=1:N;
+F_best=fit(x);
+F=fit(x);
+
+for i=1:NI
+
+	P_i=1-(log(i)/log(NI));
+	randomnumber=rand(N,1);
+	J=find(randomnumber<P_i);
+
+	if(isempty(J))
+
+		J=floor(1+length(IC)*rand);
+	end
+
+	SIGMA=r.*(MAXJ-MINJ);
+	x_new=x;
+	x_new(J)=(x(J))+SIGMA(J).*randn(length(J),1);
+	x_new=bind(x_new,MAXJ,MINJ);
+
+	F_new=fit(x_new);
+
+	%Greedy step
+	if(F_new<F)
+	    x=x_new;
+	    F=F_new;
+
+		if(F_new<=F_best)
+			F_best=F_new;
+			x_best(:,i)=x_new;
+        else
+            if i>1
+                x_best(:,i)=x_best(:,i-1);
+            end
+        end
+
+    else
+        if i>1
+            x_best(:,i)=x_best(:,i-1);
+        end
+	end
+
+
+best(i)=F_best;
+fprintf('Best value is %f and iteration is %d \n',best(i),i);          
+
+end
+
+dds_swarm_flag=1;
+end
+
+
+
+function [p]=bind(p,MAXJ,MINJ)
+
+	JMIN_NEW=find(p<MINJ);
+	p(JMIN_NEW)=MINJ(JMIN_NEW)+(MINJ(JMIN_NEW)-p(JMIN_NEW));
+
+	JTEMP1=find(p>MAXJ);
+	p(JTEMP1)=MINJ(JTEMP1);
+
+	JMAX_NEW=find(p>MAXJ);
+	p(JMAX_NEW)=MAXJ(JMAX_NEW)-(p(JMAX_NEW)-MAXJ(JMAX_NEW));
+
+	JTEMP2=find(p<MINJ);
+	p(JTEMP2)=MAXJ(JTEMP2);
+	
+end

### 2016-05-02 DOPS/DOPS_PSO.m

@@ -1,0 +1,212 @@
+%  ----------------------------------------------------------------------------------- 
+%  Copyright (c) 2016 Varnerlab
+%  School of Chemical and Biomolecular Engineering
+%  Cornell University, Ithaca NY 14853 USA
+% 
+%  Permission is hereby granted, free of charge, to any person obtaining a copy
+%  of this software and associated documentation files (the "Software"), to deal
+%  in the Software without restriction, including without limitation the rights
+%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+%  copies of the Software, and to permit persons to whom the Software is
+%  furnished to do so, subject to the following conditions:
+
+%  The above copyright notice and this permission notice shall be included
+%  in
+%  all copies or substantial portions of the Software.
+%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+%  THE SOFTWARE.
+%  ------------------------------------------------------------------------
+
+
+%% SWARM SEARCH WITH ADAPTIVE SWITCHING STRATEGY
+
+function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=DOPS_PSO(MAXJ,MINJ,NP,NI,NS,G,r)
+
+% Parameters for swarm search
+Max_Inertia_weight=0.9;
+Min_Inertia_weight=0.4;
+C1=1.5;C2=1.5;
+N=length(MINJ);
+SUB_SWARM_SIZE=round(NP/NS);
+failure_counter=0;
+failure_counter_threshold=4;
+dds_swarm_flag=0;
+
+%%
+%Initializing the position of the particles within the swarms
+
+for i=1:NP
[…]

### 2016-05-02 DOPS/fit.m

@@ -1,0 +1,28 @@
+%  ----------------------------------------------------------------------------------- 
+%  Copyright (c) 2016 Varnerlab
+%  School of Chemical and Biomolecular Engineering
+%  Cornell University, Ithaca NY 14853 USA
+% 
+%  Permission is hereby granted, free of charge, to any person obtaining a copy
+%  of this software and associated documentation files (the "Software"), to deal
+%  in the Software without restriction, including without limitation the rights
+%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+%  copies of the Software, and to permit persons to whom the Software is
+%  furnished to do so, subject to the following conditions:
+
+%  The above copyright notice and this permission notice shall be included
+%  in
+%  all copies or substantial portions of the Software.
+%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+%  THE SOFTWARE.
+%  ----------------------------------------------------------------------------------- 
+
+%%
+function fitness = fit(x)
+	    fitness=ackley(x');      %ackley function 
+end

### 2016-05-02 DOPS/run_DOPS.m

@@ -1,0 +1,80 @@
+
+%  ----------------------------------------------------------------------------------- 
+%  Copyright (c) 2016 Varnerlab
+%  School of Chemical and Biomolecular Engineering
+%  Cornell University, Ithaca NY 14853 USA
+% 
+%  Permission is hereby granted, free of charge, to any person obtaining a copy
+%  of this software and associated documentation files (the "Software"), to deal
+%  in the Software without restriction, including without limitation the rights
+%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+%  copies of the Software, and to permit persons to whom the Software is
+%  furnished to do so, subject to the following conditions:
+
+%  The above copyright notice and this permission notice shall be included
+%  in
+%  all copies or substantial portions of the Software.
+%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+%  THE SOFTWARE.
+%  ----------------------------------------------------------------------------------- 
+
+
+clear all;
+more off;
+warning('off','all');
+
+
+bounds=load('bounds.txt');
+MAXJ=bounds(:,2);
+MINJ=bounds(:,1);
+
+%% PARAMETERS FOR SWARM SEARCH AND DYNAMICALLY DIMENSIONED SEARCH
+
+NP=40;                     %Number of particles in the swarm
+NI=100;                    %Number of iterations
+NS=7;                      %Number of sub swarms
+G=10;                      %Number of iterations after which swarms are redistributed
+r=0.2;                     %Perturbation parameter for DDS
+
[…]

### 2016-05-02 DOPS/DDSPSO_strategy1.m
Updated Comments

@@ -1,220 +1,0 @@
-%Particle swarm optimization that implements dynamically dimensioned search
-
-function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=DDSPSO_strategy1(MAXJ,MINJ,NP,NI,NS,G,r)
-
-
-%se=load('randnseed.txt');
-%randn('seed',se);
-%rand('seed',se);
-
-init=0;
-Max_Inertia_weight=0.9;
-Min_Inertia_weight=0.4;
-C1=1.5;C2=1.5;
-beta=3.3;
-N=length(MINJ);
-SUB_SWARM_SIZE=round(NP/NS);
-failure_counter=0;
-success_counter=0;
-
-success_counter_threshold=4;
-failure_counter_threshold=4;
-rho(:,1)=ones(length(MINJ),1);
-dds_swarm_flag=0;
-%FR=0.4; %Fraction of iterations in DDSMLSPSO
-
-%Initializing the position of the particles within the swarms
-
-
-for i=1:NP
-    %Restricting the perturbation to be amongst the bounds specified
-    Z(:,i)=MINJ+(MAXJ-MINJ).*rand(N,1);
-    [Z(:,i)]=bind(Z(:,i),MINJ,MAXJ);
-    %Calculating the cost
-    particle(:,1,i)=Z(:,i);
-    fitness(i,1)=fit7(particle(:,1,i));
-
-end
-
-
-%Columns of S matrix have the indices of the particles in each swarm - NS*SUB_SWARM_SIZE matrix
-[S]=newswarms(NP,NS,SUB_SWARM_SIZE);
-velocity=0.729*particle;
-
-%Finding particle best and global best after initialization within each sub
-%swarm
-for j=1:NS
-    index_particle=S(:,j);
-
-    [ls_pbest_solution(j,:),ls_ITER(j,:)]=particlebest(fitness(index_particle,:));
-    [ls_gbest_solution(j),temp_particlenumber]=globalbest(ls_pbest_solution(j,:));
-    particlenumber(j)=index_particle(temp_particlenumber);
-end
-
-%Finding the best solution among all the swarms
-[g_best_solution(1),swarmnumber]=globalbest(ls_gbest_solution);
-bestparticle(:,1)=particle(:,ls_ITER(swarmnumber(1),S(:,swarmnumber(1))==particlenumber(swarmnumber(1))),particlenumber(swarmnumber(1)));
[…]

### 2016-05-02 DOPS/DOPS_DDS.m

@@ -1,32 +1,65 @@
-function[best,x_best,dds_swarm_flag]=DDS(IC,MAXJ,MINJ,r,NI)
+%  ----------------------------------------------------------------------------------- 
+%  Copyright (c) 2016 Varnerlab
+%  School of Chemical and Biomolecular Engineering
+%  Cornell University, Ithaca NY 14853 USA
+%
+%----------------*****  Code Author Information *****----------------------
+%   Code Author (Implementation Questions, Bug Reports, etc.): 
+%       Adithya Sagar: asg242@cornell.edu
+% 
+%  Permission is hereby granted, free of charge, to any person obtaining a copy
+%  of this software and associated documentation files (the "Software"), to deal
+%  in the Software without restriction, including without limitation the rights
+%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+%  copies of the Software, and to permit persons to whom the Software is
+%  furnished to do so, subject to the following conditions:
 
+%  The above copyright notice and this permission notice shall be included
+%  in
+%  all copies or substantial portions of the Software.
+%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+%  THE SOFTWARE.
+%  ------------------------------------------------------------------------
+%  ----------- 
 
[…]

### 2016-05-02 DOPS/DOPS_PSO.m

@@ -2,6 +2,10 @@
 %  Copyright (c) 2016 Varnerlab
 %  School of Chemical and Biomolecular Engineering
 %  Cornell University, Ithaca NY 14853 USA
+%
+%----------------*****  Code Author Information *****----------------------
+%   Code Author (Implementation Questions, Bug Reports, etc.): 
+%       Adithya Sagar: asg242@cornell.edu
 % 
 %  Permission is hereby granted, free of charge, to any person obtaining a copy
 %  of this software and associated documentation files (the "Software"), to deal
@@ -20,8 +24,7 @@
 %  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
 %  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
 %  THE SOFTWARE.
-%  ------------------------------------------------------------------------
-
+%  ----------------------------------------------------------------------------------- 
 
 %% SWARM SEARCH WITH ADAPTIVE SWITCHING STRATEGY
 

### 2016-05-02 DOPS/fit.m

@@ -2,6 +2,10 @@
 %  Copyright (c) 2016 Varnerlab
 %  School of Chemical and Biomolecular Engineering
 %  Cornell University, Ithaca NY 14853 USA
+%
+%----------------*****  Code Author Information *****----------------------
+%   Code Author (Implementation Questions, Bug Reports, etc.): 
+%       Adithya Sagar: asg242@cornell.edu
 % 
 %  Permission is hereby granted, free of charge, to any person obtaining a copy
 %  of this software and associated documentation files (the "Software"), to deal
@@ -22,7 +26,7 @@
 %  THE SOFTWARE.
 %  ----------------------------------------------------------------------------------- 
 
-%%
+%% Returns a numerical error value to DOPS 
 function fitness = fit(x)
-	    fitness=ackley(x');      %ackley function 
+	    fitness=ackley(x');      %An example here is ackley function. This function is user defined an basically determines the error between simulation and experiment. 
 end

### 2016-05-02 DOPS/run_DOPS.m

@@ -3,6 +3,10 @@
 %  Copyright (c) 2016 Varnerlab
 %  School of Chemical and Biomolecular Engineering
 %  Cornell University, Ithaca NY 14853 USA
+%
+%----------------*****  Code Author Information *****----------------------
+%   Code Author (Implementation Questions, Bug Reports, etc.): 
+%       Adithya Sagar: asg242@cornell.edu
 % 
 %  Permission is hereby granted, free of charge, to any person obtaining a copy
 %  of this software and associated documentation files (the "Software"), to deal
@@ -50,31 +54,29 @@ for i=1:NUMBER_TRIALS
     [g_best_solution(i,:),bestparticle(:,:,i),particle(:,:,:,i),fitness(:,:,i)]=DOPS(MAXJ,MINJ,NP,NI,NS,G,r);
     timeDOPS(i)=toc(tstart);
 
-   %-------This portion save results for every 5 trials. It is optional and the user can choose to comment it out--------% 
-   if(~mod(i,5))
-        cmd1 = ['save  ./DOPS_Results/DOPS_solution_iter',num2str(i),'.mat bestparticle'];                     %The best solution vector
+   %-------This portion save results and additional results for every trial. It is optional and the user can choose the results and number of trials --------% 
+   if(~mod(i,1))
+        cmd1 = ['save  ./DOPS_Results/DOPS_solution_iter',num2str(i),'.mat bestparticle'];                                                  % The best solution vector from swarm search
     	eval(cmd1)
 
-        cmd2 = ['save  ./DOPS_Results/DOPS_particle',num2str(i),'.mat particle'];
+        cmd2 = ['save  ./DOPS_Results/DOPS_particle',num2str(i),'.mat particle'];                                                           % The particle matrix - contains particle states for every iteration within the trial
     	eval(cmd2)
 
-        cmd3 = ['save  ./DOPS_Results/DOPS_fitness',num2str(i),'.mat fitness'];
+        cmd3 = ['save  ./DOPS_Results/DOPS_fitness',num2str(i),'.mat fitness'];                                                             % The fitness matrix - contains fitness states for the corresponding particle matrix 
     	eval(cmd3)
 
-
[…]

### 2016-09-25 Paper_v6.tex
Added minor changes

--- 2016-05-02 Paper_v5.tex
@@
 We evaluated the performance of DOPS on problems from the least and most expensive categories.
-The least expensive problem (henceforth referred to as problem B4) was a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters \cite{villaverde2014high}.
+The least expensive problem (henceforth referred to as CHO model) was a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters \cite{villaverde2014high}.
 The biochemical reactions were modeled using modular rate laws and generalized Michaelis-Menten kinetics.
-On the other hand, the expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters \cite{smallbone2013large} (henceforth referred to as problem B1).
+On the other hand, the expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters \cite{smallbone2013large} (henceforth referred to as yeast model).
 In both cases, synthetic time series data generated with known parameter values, was used to estimate model parameters.
-For problem B1, the time series data consisted of 44 observables, and for problem B4 the data corresponded to 13 different metabolite measurement sets.
+For yeast model, the time series data consisted of 44 observables, and for problem B4 the data corresponded to 13 different metabolite measurement sets.
 We fixed the number of function evaluations at $\mathcal{N}$ = 4000, and trained both models against the synthetic experimental data.
@@
 The quantities $r_{1}$ and $r_{2}$ denote uniform random vectors with the same dimension as the number of unknown model parameters ($\mathcal{K}\times{1}$).
-Equation \eqref{qn:update-rule} is similar to the general particle swarm update rule, however, it does not contain velocity terms.
+Equation \eqref{eqn:update-rule} is similar to the general particle swarm update rule, however, it does not contain velocity terms.
 In DOPS, the parameter $\theta_{1,j-1}$ is similar to the inertia weight parameter for the velocity term described by Shi and Eberhart \cite{shi1999empirical};
@@
 \hline\\
-                        & Coagulation 					& B1           	& B4          & Ackley 			& Rastrigin \\\\
+                        & Coagulation 					& Yeast model          	& CHO model          & Ackley 			& Rastrigin \\\\
 \hline\\
@@
 \includegraphics[width=1.00\textwidth]{./figs/Figure_S1_B1_sims.pdf}
-\caption{\textbf {(Data fits for Problem B1)} Pseudo-experimental data (red circles) vs.
+\caption{\textbf {(Data fits for Yeast Model (B1))} Pseudo-experimental data (red circles) vs.
 optimal solution obtained using DOPS (solid blue lines) for the 44 observed states.
@@
 \includegraphics[width=1.00\textwidth]{./figs/Figure_S2_B4_sims.pdf}
-\caption{\textbf {(Data fits for Problem B4)} Pseudo-experimental data (red x) vs.
+\caption{\textbf {(Data fits for CHO Model (B4))} Pseudo-experimental data (red x) vs.
 optimal solution obtained using DOPS (solid blue lines) for the 13 observed states.
@@
 \caption{ \textbf {(A)} Difference between nominal and optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
-\textbf {(B)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
+\textbf {(B)} Difference between experimental (measured) data and data simulated with optimal parameters for yeast model: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
 \textbf {(C)} Difference between nominal and optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
-\textbf {(D)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
+\textbf {(D)} Difference between experimental (measured) data and data simulated with optimal parameters for CHO model: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
 }\label{fig-benchmark}

### 2016-09-27 Paper_v6.tex
Updated references for particle swarm optimization to address reviewer 2's comments

@@
 DOPS is a novel meta-heuristic which combines multi-swarm particle swarm optimization
-with the dynamically dimensioned search (Fig. \ref{fig-algorithm}) and (Algo. \ref{algo_ddspso}). 
+with the dynamically dimensioned search (Fig.
+\ref{fig-algorithm}) and (Algo.
+\ref{algo_ddspso}).
+Particle swarm optimization is an evolutionary algorithm that uses a population of particles (solutions) to find an optimal solution.
+Each particle is updated based on its on experience (particle best) and the experience of all other particles within the swarm (global best) \cite{clerc2010particle}\cite{blondin2009particle}\cite{abraham2006swarm}.
+Multi-swarm particle swarm optimization is a variant of particle swarm optimization that uses multiple swarms instead of a single swarm.
+DOPS uses a novel update rule of particles within each swarm that is described in detail in the methods.
 The goal of DOPS is to estimate optimal or near optimal parameter vectors for high-dimensional biological models within a specified number of function evaluations.

### 2016-09-27 Paper_v6.tex
eSS reference added

@@
 Hybrid approaches, which combine meta-heuristics with local optimization techniques, have also become popular.
-For example, Villaverde et al. developed the enhanced scatter search (eSS) method, which combined scatter and local search methods,
+For example, Villaverde et al. developed the enhanced scatter search (eSS) method \cite{egea2010evolutionary}, which combined scatter and local search methods,
 for parameter estimation in biological models \cite{villaverde2015biopredyn}.

### 2016-10-08 Paper_v6.tex
Adding a more detailed description of DDS
Updated discussion about DDS

@@
 Multi-swarm particle swarm optimization is a variant of particle swarm optimization that uses multiple swarms instead of a single swarm.
-DOPS uses a novel update rule of particles within each swarm that is described in detail in the methods.
+DOPS uses a novel update rule of particles within each swarm.
+This rule uses only particle positions but not particle velocities in the update rule.
+This differs from existing PSO methods that rely both on particle position and velocity for update.
+The update rule further described in detail in the methods.
 The goal of DOPS is to estimate optimal or near optimal parameter vectors for high-dimensional biological models within a specified number of function evaluations.
 Toward this objective, DOPS begins by using a particle swarm search and then dynamically switches, using an adaptive switching criteria, to a DDS search phase.
+Dynamically Dimensioned Search (DDS) is a single solution based search algorithm.
+It was developed by Tolson and Shoemaker to obtain good solutions to high-dimensional search problems within a fixed number of function evaluations.
+DDS starts as a global search algorithm by perturbing all the dimensions. Later the number of dimensions that are perturbed is decreased with a certain probability. 
+The probability that a certain dimension is perturbed reduces (a minimum of one dimension is always perturbed) as the iterations increase. 
+This causes the algorithm to behave as a local search algorithm as the number of iterations increase.
+The perturbation magnitude of each dimension is from normal distribution with zero mean. The standard deviation that was used in the original DDS paper and the current study is 0.2.
+DDS performs a greedy search where the solution is updated only if it is better than the previous solution.
+The combination of perturbing a subset of dimensions along with greedy search indirectly relies on model sensitivity to a specific parameter combination. 
+For the sake of brevity we are not providing the entire pseudo code of DDS here.
+The reader is requested to refer to Figure 1 of the original paper by Tolson and Shoemaker \cite{tolson2007dynamically}.
 

### 2016-10-08 Paper_v6.tex
Updated discussion about algorithm includes a pseudo code section for DDS

@@
 \BlankLine
-\emph{Initialize the particles randomly and assign particles randomly to various sub-swarms}\;
+\emph{Initialize the particles randomly and assign particles randomly to $k$ sub-swarms}\;
 \While{$j\leq N$}{
@@
 \If{mod(j,G)=0} {
- \emph{Reassign particles to different sub-swarms}\;
+ \emph{Reassign particles to $k$ sub-swarms}\;
 }
@@
 \subsubsection*{Phase 2: DDS phase.}
+
+\IncMargin{1em}
+\begin{algorithm}
+\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
+\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
+\SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
+\Input{Candidate vector $\mathcal{G}$ from swarm search and $\left(1-\mathcal{F}\right)*\mathcal{N}$ evaluations}
+\Output{Optimized parameter vector of size $1\times K$ }
+\BlankLine
+\While{$j\leq \left(1-\mathcal{F}\right)*\mathcal{N}$}{
+
+\emph{Assign probability of perturbation to each dimension $\mathcal{P}_{i}$ according to equation 7}\;
+\emph{Select a subset of dimensions based on a threshold value for perturbation}\;
+\emph{Update candidate solution $\mathcal{G}({J})$  according to equation 5}\;
+\emph{Ensure updated solution $\mathcal{G}_{new}({J})$ is within bounds using Algorithm 2}\;
+
+}
+\caption{Pseudo code for the Dynamically Dimensioned Search (DDS) method.}\label{algo_dds}
+\end{algorithm}\DecMargin{1em}
+
 At the conclusion of the swarm phase, the overall best particle, $\mathcal{G}_{k}$, over the k sub-swarms was used to initialize the DDS phase.

### 2016-11-13 Paper_v6.tex
Added supplementary figures

@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth]{./figs/Figure_S3_B1_B4_params_measuredstates.pdf}
+\caption{ \textbf {(A)} Difference between nominal and optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
+\textbf {(B)} Difference between experimental (measured) data and data simulated with optimal parameters for yeast model: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
+\textbf {(C)} Difference between nominal and optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
+\textbf {(D)} Difference between experimental (measured) data and data simulated with optimal parameters for CHO model: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
+}\label{fig-benchmark}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[clip, trim=0.5cm 7cm 0.5cm 1.80cm, width=1.00\textwidth]{./figs/Figure_S4.pdf}
+\caption{ Convergence curves for different metaheuristics along with CMA-ES) Objective function value versus the number of function evaluations for \textbf {(A)} Ackley 300 dimensional \textbf {(B)} and Rastrigin 300 dimensional.
+}\label{fig-CMA-ES convergences}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth]{./figs/Figure_S5.pdf}
+\caption{ (Convergence curves for eSS and DOPS) Objective function value versus the number of function evaluations plotted for eSS and DOPS on \textbf {(A)} Coagulation \textbf {(B)} CHO Model \textbf {(C)} Ackley 300 dimensional and \textbf {(D)} Rastrigin 300 dimensional.
+}\label{fig-convergence eSS}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth]{./figs/Figure_S6.pdf}
+\caption{ (Dispersion curves for DOPS) Objective function value (25 trials) versus the number of function evaluations plotted for DOPS on \textbf {(A)} Ackley 300 dimensional \textbf {(B)} Rastrigin 300 dimensional \textbf {(C)}CHO Model - B4 \textbf {(D)} Coagulation.
+For CHO Model and Coagulation Y-axis is in log scale.
+}\label{fig-dispersion}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth]{./figs/Figure_S7.pdf}
+\caption{ Fig.
+S7: (Variability analysis for best objective value): Boxplots for different metaheuristics on \textbf {(A)} Ackley 300 dimensional \textbf {(B)} Rastrigin 300 dimensional \textbf {(C)} CHO Model - B4 \textbf {(D)} and Coagulation
+}\label{fig-boxplot}
+\end{figure}
+
 \end{document}

### 2017-09-07 Paper_v6.tex

@@
 within the function evaluation budget (Fig. \ref{fig-testfunctions}B).
+DOPS also had a lower variability in the best value obtained (Fig.
+\ref{fig-boxplot}), faster convergence (Fig.
+\ref{fig-CMA-ES convergences}, Fig.
+\ref{fig-convergence eSS}) and less stochasticity (Fig.
+\ref{fig-dispersion}) across multiple runs when compared to other metaheuristics.
+We also checked for statistical significance by doing a pairwise 2-sample t-test at 0.5\% significance level.
+In each case we found that the p-value was close to zero disproving the null hypothesis and confirming the statistical significance of our results.
 Taken together, DOPS performed at least as well as other meta-heuristic approaches on small dimensional test problems, but seemed especially suited to large dimensional search spaces.

### 2017-10-25 Paper_v6.tex
adding v6

--- 2016-10-08 Paper_v6.tex
@@
 \usepackage{algpseudocode}
+\usepackage{subfig}
 %\usepackage[space]{cite}
@@
 Hybrid approaches, which combine meta-heuristics with local optimization techniques, have also become popular.
-For example, Villaverde et al. developed the enhanced scatter search (eSS) method \cite{egea2010evolutionary}, which combined scatter and local search methods,
-for parameter estimation in biological models \cite{villaverde2015biopredyn}.
+For example, Villaverde et al. developed the enhanced scatter search (eSS) method \cite{villaverde2015biopredyn}, which combined scatter and local search methods,
+for parameter estimation in biological models  \cite{egea2007scatter}.
 However, despite these successes, a major drawback of most meta-heuristics remains the large number of function evaluations required to explore parameter space.
@@
 We evaluated the performance of DOPS on problems from the least and most expensive categories.
-The least expensive problem (henceforth referred to as CHO model) was a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters \cite{villaverde2014high}.
+The least expensive problem (henceforth referred to as problem B4) was a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters \cite{villaverde2014high}.
 The biochemical reactions were modeled using modular rate laws and generalized Michaelis-Menten kinetics.
-On the other hand, the expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters \cite{smallbone2013large} (henceforth referred to as yeast model).
+On the other hand, the expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters \cite{smallbone2013large} (henceforth referred to as problem B1).
 In both cases, synthetic time series data generated with known parameter values, was used to estimate model parameters.
-For yeast model, the time series data consisted of 44 observables, and for problem B4 the data corresponded to 13 different metabolite measurement sets.
+For problem B1, the time series data consisted of 44 observables, and for problem B4 the data corresponded to 13 different metabolite measurement sets.
 We fixed the number of function evaluations at $\mathcal{N}$ = 4000, and trained both models against the synthetic experimental data.
@@
 DOPS is a novel meta-heuristic which combines multi-swarm particle swarm optimization
-with the dynamically dimensioned search (Fig.
-\ref{fig-algorithm}) and (Algo.
-\ref{algo_ddspso}).
-Particle swarm optimization is an evolutionary algorithm that uses a population of particles (solutions) to find an optimal solution.
-Each particle is updated based on its on experience (particle best) and the experience of all other particles within the swarm (global best) \cite{clerc2010particle}\cite{blondin2009particle}\cite{abraham2006swarm}.
-Multi-swarm particle swarm optimization is a variant of particle swarm optimization that uses multiple swarms instead of a single swarm.
-DOPS uses a novel update rule of particles within each swarm.
-This rule uses only particle positions but not particle velocities in the update rule.
-This differs from existing PSO methods that rely both on particle position and velocity for update.
-The update rule further described in detail in the methods.
+with the dynamically dimensioned search (Fig. \ref{fig-algorithm}) and (Algo. \ref{algo_ddspso}).
 The goal of DOPS is to estimate optimal or near optimal parameter vectors for high-dimensional biological models within a specified number of function evaluations.
 Toward this objective, DOPS begins by using a particle swarm search and then dynamically switches, using an adaptive switching criteria, to a DDS search phase.
-Dynamically Dimensioned Search (DDS) is a single solution based search algorithm.
-It was developed by Tolson and Shoemaker to obtain good solutions to high-dimensional search problems within a fixed number of function evaluations.
-DDS starts as a global search algorithm by perturbing all the dimensions. Later the number of dimensions that are perturbed is decreased with a certain probability. 
-The probability that a certain dimension is perturbed reduces (a minimum of one dimension is always perturbed) as the iterations increase. 
-This causes the algorithm to behave as a local search algorithm as the number of iterations increase.
-The perturbation magnitude of each dimension is from normal distribution with zero mean. The standard deviation that was used in the original DDS paper and the current study is 0.2.
-DDS performs a greedy search where the solution is updated only if it is better than the previous solution.
-The combination of perturbing a subset of dimensions along with greedy search indirectly relies on model sensitivity to a specific parameter combination. 
-For the sake of brevity we are not providing the entire pseudo code of DDS here.
-The reader is requested to refer to Figure 1 of the original paper by Tolson and Shoemaker \cite{tolson2007dynamically}.
 
@@
 \BlankLine
-\emph{Initialize the particles randomly and assign particles randomly to $k$ sub-swarms}\;
+\emph{Initialize the particles randomly and assign particles randomly to various sub-swarms}\;
 \While{$j\leq N$}{
@@
 \If{mod(j,G)=0} {
- \emph{Reassign particles to $k$ sub-swarms}\;
+ \emph{Reassign particles to different sub-swarms}\;
 }
@@
 \subsubsection*{Phase 2: DDS phase.}
-
-\IncMargin{1em}
-\begin{algorithm}
-\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
-\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
-\SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-\Input{Candidate vector $\mathcal{G}$ from swarm search and $\left(1-\mathcal{F}\right)*\mathcal{N}$ evaluations}
-\Output{Optimized parameter vector of size $1\times K$ }
-\BlankLine
-\While{$j\leq \left(1-\mathcal{F}\right)*\mathcal{N}$}{
-
-\emph{Assign probability of perturbation to each dimension $\mathcal{P}_{i}$ according to equation 7}\;
-\emph{Select a subset of dimensions based on a threshold value for perturbation}\;
-\emph{Update candidate solution $\mathcal{G}({J})$  according to equation 5}\;
-\emph{Ensure updated solution $\mathcal{G}_{new}({J})$ is within bounds using Algorithm 2}\;
-
-}
-\caption{Pseudo code for the Dynamically Dimensioned Search (DDS) method.}\label{algo_dds}
-\end{algorithm}\DecMargin{1em}
+Dynamically Dimensioned Search (DDS) is a single solution based search algorithm.
+DDS is used to obtain good solutions to high-dimensional search problems within a fixed number of function evaluations.
+DDS starts as a global search algorithm by perturbing all the dimensions.
+Later the number of dimensions that are perturbed is decreased with a certain probability.
+The probability that a certain dimension is perturbed reduces (a minimum of one dimension is always perturbed) as the iterations increase.
+This causes the algorithm to behave as a local search algorithm as the number of iterations increase.
+The perturbation magnitude of each dimension is from normal distribution with zero mean.
+The standard deviation that was used in the original DDS paper and the current study is 0.2.
+DDS performs a greedy search where the solution is updated only if it is better than the previous solution.
+The combination of perturbing a subset of dimensions along with greedy search indirectly relies on model sensitivity to a specific parameter combination.
+The reader is requested to refer to the original paper by Tolson and Shoemaker for further detail \cite{tolson2007dynamically}.
 
@@
 \hline\\
-                        & Coagulation 					& Yeast model          	& CHO model          & Ackley 			& Rastrigin \\\\
+                        & Coagulation 					& B1           	& B4          & Ackley 			& Rastrigin \\\\
 \hline\\
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure_4_Errors_convergence_v2.pdf}
+\includegraphics[width=1.0\textwidth]{./figs/Figure_4_Errors_Convergence_v2.pdf}
 \caption{Error convergence rates of the five different algorithms on the coagulation model. The objective error is the mean over $\mathcal{T}$= 25 trials.
@@
 \includegraphics[width=1.00\textwidth]{./figs/Figure_S1_B1_sims.pdf}
-\caption{\textbf {(Data fits for Yeast Model (B1))} Pseudo-experimental data (red circles) vs.
+\caption{\textbf {(Data fits for Problem B1)} Pseudo-experimental data (red circles) vs.
 optimal solution obtained using DOPS (solid blue lines) for the 44 observed states.
@@
 \includegraphics[width=1.00\textwidth]{./figs/Figure_S2_B4_sims.pdf}
-\caption{\textbf {(Data fits for CHO Model (B4))} Pseudo-experimental data (red x) vs.
+\caption{\textbf {(Data fits for Problem B4)} Pseudo-experimental data (red x) vs.
 optimal solution obtained using DOPS (solid blue lines) for the 13 observed states.
@@
 \caption{ \textbf {(A)} Difference between nominal and optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
-\textbf {(B)} Difference between experimental (measured) data and data simulated with optimal parameters for yeast model: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
+\textbf {(B)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
 \textbf {(C)} Difference between nominal and optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
-\textbf {(D)} Difference between experimental (measured) data and data simulated with optimal parameters for CHO model: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
+\textbf {(D)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
 }\label{fig-benchmark}
@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[trim = {0 5cm 0 5cm }, clip, width=1.00\textwidth]{./rachelfigs/TimeComparsion.pdf}
+\caption{ Comparison of the runtime of the different optimization methods used for comparison with  $\mathcal{T}$ = 25 trials per method.
+All methods used take about the same amount of time to perform 4,000 function evaluations on the coagulation problem, as this problem is very stiff, so the majority of the time is spent solving the system of differential equations.
+}\label{fig-time-comparsion}
+\end{figure}
+
+\begin{figure}%
+    \centering
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/RecreateRebutalReviewer1_FigS4a} }}%
+    \qquad
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/RecreateRebutalReviewer1_FigS4b} }}%
+    \caption{Mean onvergence curves for different metaheuristics for (a) Ackley 300 dimensional and (b) Rastrigin 300 dimensional with $\mathcal{T}$ = 25 trials per method.
+DOPS not only finds a better solution than any other technique, it finds it with fewer function evaluations}%
+    \label{fig-more-methods}%
+\end{figure}
+
+\begin{figure}%
+    \centering
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/CompareExperimentalDOPSToESSOnProblemB4} }}%
+    \qquad
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/CompareExperimentalDOPSToESSOnCoag} }}%
+    \caption{Mean convergence curves for DOPS and ESS for (a) CHO model (problem B4) (b) coagulation with $\mathcal{T}$ = 25 trials per method.}%
+    \label{fig-compare-DOPS-ESS}%
+\end{figure}
+
+\begin{figure}%
+    \centering
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/DOPSDispersionCurvesOnb4} }}%
+    \qquad
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/DOPSDispersionCurvesOnCoag} }}%
+    \caption{Dispersion curves for DOPS on (a) CHO model (problem B4) (b) coagulation with $\mathcal{T}$ = 25 trials per problem .}%
+    \label{fig-DOPS-Dispersion-Curves}%
+\end{figure}
+
+\begin{figure}%
+    \centering
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/FigureS7a} }}%
+    \qquad
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/FigureS7b} }}%
+    \qquad
+        \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/FigureS7c} }}%
+    \qquad
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/FigureS7d} }}%
+    \caption{Variability  analysis in best objective value for $\mathcal{T}$ = 25 trials. (a) Ackley  300 dimensional (b) Rastrigin 300 dimensional (c) CHO model (problem b4) (d) coagulation. }%
+    \label{fig-DOPS-Dispersion-Curves}%
+\end{figure}
+
 \end{document}

### 2017-10-25 Paper_v6.tex
adding updated manuscript

@@
 On the other hand, enhanced scatter search (eSS) with a local optimizer took on order $10^{5}$ function evaluations for the same problems.
+DOPS also had lower variability in the best value obtained (Fig \ref{fig-DOPS-Variability}) and faster convergence (Fig \ref{fig-more-methods}, Fig \ref{fig-compare-DOPS-ESS}) across multiple runs when compared to other metaheuristics while requiring a comparable amount of time (Fig \ref{fig-time-comparsion}).
 Thus, DOPS estimated the parameters in benchmark biochemical models, and recovered the original parameters from synthetic data, using fewer function evaluations.
@@
 \subsubsection*{Phase 1: Particle swarm phase.}
+Particle warm optimization is an evolutionary algorithm that uses a population of particles (solutions) to find an optimal solution.
+Each particle is updated based on its experience (particle best) and the experience of all other particles within the swarm (global best).
 The particle swarm phase of DOPS begins by randomly initializing a swarm of $\mathcal{K}$-dimensional particles (represented as $z_{i}$),
@@
     \caption{Variability  analysis in best objective value for $\mathcal{T}$ = 25 trials. (a) Ackley  300 dimensional (b) Rastrigin 300 dimensional (c) CHO model (problem b4) (d) coagulation. }%
-    \label{fig-DOPS-Dispersion-Curves}%
+    \label{fig-DOPS-Variability}%
 \end{figure}

### 2017-10-26 Paper_v6.tex
updaing manuscript

@@
 DDS updates are greedy; $\mathcal{G}_{new}$ becomes the new solution vector only if it is better than $\mathcal{G}$.
-DOPS source code is available for download under a MIT license at http://www.varnerlab.org.
+
+\subsection*{Experimental DOPS}
+We sought to investigate if switching methods more than once would result in better performance.
+Experimental DOPS begins with the PSO phase and uses the same criteria as DOPS to switch to the DDS phase.
+The technique switches back to PSO phase when the DDS phase has reduced the functional value to 90\% of its initial value.
+Should the DDS phase fail to improve the functional value sufficiently, this version is identical to DOPS, but we found that for the CHO problem as well as for coagulation, the DDS phase did produce sufficient improvement in functional value, resulting in a switch back to the PSO phase from the DDS phase.
+When the switch from DDS to PSO occurs, we use the best solution from DDS to seed the particle swarm.
+
+DOPS and Experimental DOPS source code is available for download under a MIT license at http://www.varnerlab.org.
 
@@
 Evaluations   					& 4000                  & 4000         	& 4000        & 4000        & 4000        \\
-Lower Bound             & 0.001.pnom            & 5.pnom       	& 5.pnom      & 30          & 5.12        \\
-Upper Bound             & 1000.pnom             & 0.2.pnom     	& 0.2.pnom    & -15         & -5.12       \\
+Lower Bound             & 0.001.pnom            & 0.2.pnom       	& 0.2.pnom      & -15          & -5.12        \\
+Upper Bound             & 1000.pnom             & 5.pnom     	& 5.pnom    & 30         & 5.12       \\
 CPU Time                & 10.1 hrs         			& 38.3 hrs 			& 6.2 min 		& 2.8 s 			& 2.6 s    		\\\\
@@
 \hline\\
-5~nM & 0.0376 & Training \\
-500~pM & 0.0564 & Prediction\\
-50~pM & 0.1125 & Prediction\\
-10~pM & 0.0823 & Prediction\\
-5~pM & 0.0338 & Training\\\\
+5~nM & 0.1336 & Training \\
+500~pM & 0.2242 & Prediction\\
+50~pM & 0.3109 & Prediction\\
+10~pM & 0.2023 & Prediction\\
+5~pM & 0.1170 & Training\\\\
 \hline
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure_4_Errors_Convergence_v2.pdf}
+%\includegraphics[width=1.0\textwidth]{./figs/Figure_4_Errors_Convergence_v2.pdf}
+\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=1.0\textwidth]{./rachelfigs/RecreateFigure4.pdf}
 \caption{Error convergence rates of the five different algorithms on the coagulation model. The objective error is the mean over $\mathcal{T}$= 25 trials.
-DOPS, DDS and SA have the steepest drop in error during first 300 function evaluations. Thereafter the error drop in DDS and SA remains nearly constant whereas DOPS continues to drops further.
-At the end of 4000 function evaluations DOPS attains the lowest error. The next best estimate using DDS is nearly three times greater than the lowest error using DDS.
+DOPS and SA have the steepest drop in error during first 300 function evaluations. Thereafter the error drop in DDS and SA remains nearly constant whereas DOPS continues to drops further.
+At the end of 4000 function evaluations DOPS attains the lowest error. The next best estimate using DDS is nearly three times greater than the lowest error using DE.
 }\label{fig-convergence}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure_5_Sim_Train_E1_E5.pdf}
+%\includegraphics[width=1.0\textwidth]{./figs/Figure_5_Sim_Train_E1_E5.pdf}
+\includegraphics[trim = {0cm 6cm 7cm 6cm }, clip,width=1.0\textwidth]{./rachelfigs/ReplicateFigure5DivBySqrtN.pdf}
 \caption{Model fits on experimental data using DOPS.
@@
 The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers.
-Thrombin generation is initiated by adding Factor TF/VIIa (5nM and 5pM) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
+Thrombin generation is initiated by adding Factor TF/VIIa (5nM (blue) and 5pM (red)) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
 }\label{fig-train}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure_6_Sim_Validate_E2_E4_E6.pdf}
+%\includegraphics[width=1.0\textwidth]{./figs/Figure_6_Sim_Validate_E2_E4_E6.pdf}
+\includegraphics[trim = {3cm 6cm 3cm 6cm }, clip,width=1.0\textwidth]{./rachelfigs/ReplicateFigure6DivBySqrtN.pdf}
 \caption{Model predictions on unseen experimental data using parameters obtained from DOPS.
@@
 The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers.
-Thrombin generation is initiated by adding Factor VIIa-TF (500pM - Blue, 50pM - Pink and 10pM - orange, respectively) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
+Thrombin generation is initiated by adding Factor VIIa-TF (500pM - Blue, 50pM - Pink and 10pM - purple, respectively) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
 }\label{fig-validation}
@@
 \centering
-\includegraphics[width=1.00\textwidth]{./figs/Figure_S2_B4_sims.pdf}
+\includegraphics[width=1.00\textwidth]{./rachelfigs/RecreatedFigureS2}
 \caption{\textbf {(Data fits for Problem B4)} Pseudo-experimental data (red x) vs.

### 2017-10-27 Paper_v6.tex
updating manuscript and adding exp DOPS gamma results on coag

@@
 \subsubsection*{Phase 1: Particle swarm phase.}
-Particle warm optimization is an evolutionary algorithm that uses a population of particles (solutions) to find an optimal solution.
+Particle warm optimization is an evolutionary algorithm that uses a population of particles (solutions) to find an optimal solution \cite{clerc2010particle,abraham2006swarm}.
 Each particle is updated based on its experience (particle best) and the experience of all other particles within the swarm (global best).
@@
 \subsubsection*{Phase 2: DDS phase.}
+\IncMargin{1em}
+\begin{algorithm}
+\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
+\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
+\SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
+\Input{Candidate vector $\mathcal{G}$ from swarm search and $\left(1-\mathcal{F}\right)*\mathcal{N}$ evaluations}
+\Output{Optimized parameter vector of size $1\times K$ }
+\BlankLine
+\While{$j\leq \left(1-\mathcal{F}\right)*\mathcal{N}$}{
+
+\emph{Assign probability of perturbation to each dimension $\mathcal{P}_{i}$ according to equation 7}\;
+\emph{Select a subset of dimensions based on a threshold value for perturbation}\;
+\emph{Update candidate solution $\mathcal{G}({J})$  according to equation 5}\;
+\emph{Ensure updated solution $\mathcal{G}_{new}({J})$ is within bounds using Algorithm 2}\;
+
+}
+\caption{Pseudo code for the Dynamically Dimensioned Search (DDS) method.}\label{algo_dds}
+\end{algorithm}\DecMargin{1em}
+
 Dynamically Dimensioned Search (DDS) is a single solution based search algorithm.
@@
 \centering
-\includegraphics[width=0.60\textwidth]{./figs/Figure_2_TestFunctions-crop.pdf}
+%\includegraphics[width=0.60\textwidth]{./figs/Figure_2_TestFunctions-crop.pdf}
+    \subfloat[]{{\includegraphics[width=.6\textwidth]{./rachelfigs/ackley_10d} }}%
+    \newline
+    \subfloat[]{{\includegraphics[width=.6\textwidth]{./rachelfigs/rast_10d} }}%
+    \newline
+    \subfloat[]{{\includegraphics[width=.6\textwidth]{./rachelfigs/rast_300d} }}%
+
 \caption{Performance of DOPS and other meta-heuristics for the Ackley and Rastrigin functions.
 \textbf{A}: Mean scaled error versus the number of function evaluations for the 10-dimensional Ackley function.
-DOPS, DDS and DE find optimal or near optimal solutions within the specified number of function evaluations.
+DOPS, DDS and PSO find optimal or near optimal solutions within the specified number of function evaluations.
 \textbf{B}: Mean scaled error versus the number of function evaluations for the 10-dimensional Rastrigin function.

### 2017-11-02 Paper_v6.tex
updating manuscript

@@
 Thus, the automated switching strategy was critical to the overall performance of DOPS.
+
+\subsection*{Contiued Switching May Be Beneficial}
+We explored the performance of DOPS if it was permitted to switch between PSO and DDS modes.
+This mode ('Experimental DOPS') performed comparably to DOPS on 10-d Ackley and Rastrigin as well as on 300-dimensional Rastrigin.
+It performed better than DOPS on the CHO metabolism problem (b4) (Fig.
+\ref{fig-exp-DOPS-b4}a), with the average functional value for Experimental DOPS being nearly half of that of DOPS.
+To further distinguish DOPS from Experimental DOPS, we compared the performance of each algorithm on the Eggholder function, which is considered to be a difficult function to optimize, owing to its multiple minima.
+On this function, Experimental DOPS outperformed DOPS however, neither version of DOPS reached the true minimum at -959.6407 on any trial with a budget of 4,000 function evaluations.
 
@@
 
+\clearpage
+
+\begin{figure}[h]
+\centering
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnb4_obj.pdf} }}%
+    \qquad
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnEggholder}} }%
+\caption{Performance of DOPS and Experimental DOPS on the CHO metabolism problem (a) and the Eggholder function (b).
+Both methods have the same initial decrease in error, but as the number of function evaluations increases, experimental DOPS produces a larger decrease in error.
+The results presented are the average of $\mathcal{T}$ = 250 trials with for the CHO metabolism problem and $\mathcal{T}$ = 500 trials on the Eggholder function with $\mathcal{N}$ = 4000 function evaluations each.
+}\label{fig-exp-DOPS-b4}
+\end{figure}
+
 %\bibitem[Saltelli \em{et~al.}(2010)Saltelli, Annoni, Azzini, Campolongo, Ratto,

### 2017-11-03 Paper_v6.tex
updating manuscrip with exp DOPS results and figures

@@
 To further distinguish DOPS from Experimental DOPS, we compared the performance of each algorithm on the Eggholder function, which is considered to be a difficult function to optimize, owing to its multiple minima.
-On this function, Experimental DOPS outperformed DOPS however, neither version of DOPS reached the true minimum at -959.6407 on any trial with a budget of 4,000 function evaluations.
+On this function, Experimental DOPS outperformed DOPS however, neither version of DOPS reached the true minimum at -959.6407 on any trial with a budget of 4,000 function evaluations (Fig.
+\ref{fig-exp-DOPS-b4}b).
+We also explored the performance of both versions of DOPS on the 100 dimensional Styblinksi-Tang function (Fig.
+\ref{fig-exp-DOPS-b4}c).
+In this comparison, experimental DOPS far outperforms DOPS, succeeding in finding the true minimum before exhausting its budget of functional evaluations, whereas DOPS does not reach the minimum.
+Since the performance of Experimental DOPS was promising on these problems, we decided to measure its performance on the coagulation problem.
 
@@
     \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnEggholder}} }%
-\caption{Performance of DOPS and Experimental DOPS on the CHO metabolism problem (a) and the Eggholder function (b).
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnstybtang}} }%
+
+\caption{Performance of DOPS and Experimental DOPS on the CHO metabolism problem (a), the Eggholder function (b), and the 100 dimensional Styblinksi-Tang function (c).
 Both methods have the same initial decrease in error, but as the number of function evaluations increases, experimental DOPS produces a larger decrease in error.
-The results presented are the average of $\mathcal{T}$ = 250 trials with for the CHO metabolism problem and $\mathcal{T}$ = 500 trials on the Eggholder function with $\mathcal{N}$ = 4000 function evaluations each.
+The results presented are the average of $\mathcal{T}$ = 250 trials with for the CHO metabolism problem and $\mathcal{T}$ = 500 trials on the Eggholder and Styblinksi-Tang functions with $\mathcal{N}$ = 4000 function evaluations each.
 }\label{fig-exp-DOPS-b4}

### 2017-11-03 DOPS/DOPS_DDS.m
adding scripts

@@ -56,6 +56,22 @@ for i=1:NI
 	x_new=bind(x_new,MINJ,MAXJ);                     % Ensure the perturbed solution is within bounds
 
 	F_new=fit(x_new,optFunction);
+    
+    %make sure we haven't ended up with a NAN, if so repeat process until
+    %we get a a real number
+    
+    NANcount = 0;
+    while(isnan(F_new))
+        SIGMA=r.*(MAXJ-MINJ);
+        x_new=x;
+        x_new(J)=(x(J))+SIGMA(J).*randn(length(J),1);    % Perturb the current solution - this is user defined   
+        x_new=bind(x_new,MINJ,MAXJ);                     % Ensure the perturbed solution is within bounds
+
+        F_new=fit(x_new,optFunction);
+        NI = NI -1; %deduct the cost of finding a NA soltuion
+        NANcount = NANcount +1;
+        fprintf('In starting DDS found %d NAN parameter sets\n', NANcount);
+    end
 
 	%Greedy step - accept solution only if better than the previous
 	%solution

### 2017-11-03 DOPS/DOPS_PSO_ExperimentalV2gamma.m

@@ -53,7 +53,7 @@ bestval_dds_swarm=[];
 best_particle_dds_swarm=[];
 best_particles_ls=[];
 g_best_solution=[];bestparticle=[];particle=[];fitness = [];
-shrinkFactor = 10;
+shrinkFactor = 1E6;
 %%
 %Initializing the position of the particles within the swarms
 
@@ -65,6 +65,8 @@ if(size(initial_guess,1)==0) %if we're on the initial iteration and don't have a
         fitness(i,1)=fit(particle(:,1,i),optFunction);   %was fit7, now just fit                                  %Calculating fitness of each particle
     end
 else
+   % orders = getOrder(abs(initial_guess));
+   % shrinkFactor = repmat(10,N,1).^(orders-4);
     for i=1:NP
         if(i ==1)
             Z(:,i) = initial_guess; %we give the initial guess to one particle
@@ -74,7 +76,7 @@ else
                 fitness(i,1) = best_PSO_val;
             end
         else
-            Z(:,i) = initial_guess +(MAXJ-MINJ).*rand(N,1)/shrinkFactor; 
+            Z(:,i) = initial_guess +(MAXJ-MINJ).*rand(N,1).*shrinkFactor; 
         end
         [Z(:,i)]=bind(Z(:,i),MINJ,MAXJ);                                        %Restricting the perturbation to be amongst the bounds specified    
         particle(:,1,i)=Z(:,i);
@@ -188,7 +190,7 @@ w(j)=((NI - j)*(Max_Inertia_weight - Min_Inertia_weight))/(NI-1) + Min_Inertia_w
     else
       failure_counter=0;
     end
-   fprintf('In PSO. On iteration %d of %d failture counter = %d\n', i, NI, failure_counter);
+   %fprintf('In PSO. On iteration %d of %d failture counter = %d\n', i, NI, failure_counter);
     %Switch to DDS search if the solution has stagnated
     if(failure_counter > failure_counter_threshold)
        return;  
@@ -246,5 +248,10 @@ function [x]=bind(x,MINJ,MAXJ)
 
 	CHKMIN=find(x<MINJ);
 	x(CHKMIN)=MAXJ(CHKMIN);
+end
+
+function n=getOrder(val)
+    base = 10; %let's just assume we're in base ten
+    n = floor(log(abs(val))./log(base));
 end
 

### 2017-11-03 DOPS/compareAllTechniques.m

@@ -12,18 +12,22 @@ function[]= compareAllTechniques()
     plot(1:40:4000,mean(allDEResults,1), 'LineWidth', 2.0,'Color','r');
     set(gca, 'YScale', 'log');
     axis([0,4000,7E4,2E7])
-    legend('DOPS-N=25', 'Simulated Annealing-N=25', 'DDS-N=25', 'DE-N=25');
-    xlabel('Number of Function Evaluations')
-    ylabel('Average Functional Value');
+    legend('DOPS', 'Simulated Annealing', 'DDS', 'DE');
+    xlabel('Objective Function Evaluations')
+    ylabel({'Mean Objective FUnction Error', '(Number of Trials = 25)'});
     saveas(f,'../DOPS_Results/figures/RecreateFigure4.pdf', 'pdf')
-%     f=figure();
-%     boxplot([allSATimes, allDDSTimes, allDETimes,allDOPSTimes]/(60^2)) %convert to hours
-%     xticklabels({'Simulated Annealing', 'DDS', 'DE', 'DOPS'});
-%     xlabel('Metaheuristics')
-%     ylabel('Runtime per trial (number of hours)')
-%     saveas(f,'../DOPS_Results/figures/RecreateRebutalReviewer2.pdf', 'pdf')
-    
-    
+
+    %to create DOPS dispersion curve
+    f=figure();
+    hold('on')
+    for j=1:25
+       semilogy(adjustedAllDOPS(j,:), 'k') 
+       alpha(.5); %make lines somewhat transparent
+    end
+    xlabel('Number of Function Evaluations')
+    ylabel('Functional Value');
+    set(gca, 'YScale', 'log');
+     saveas(f,'../DOPS_Results/figures/DOPSDispersionCurvesOnCoag.pdf', 'pdf')
 end
 
 function [allSAResults,allTimes] =processSAResults()

### 2017-11-03 DOPS/compareDOPSVersionsOnB4.m

@@ -1,0 +1,123 @@
+function [allDOPS,adjustedAllDOPS,allExperimental,allDOPSTimes]=compareDOPSVersionsOnB4()
+    addpath('/home/rachel/Documents/DOPS/DOPS_Results/b4_obj');
+    close('all');
+    NUM_ITERS = 25;
+    NUM_EVALS = 4000;
+    NUM_PARTCILES = 40;
+    allDOPS = zeros(NUM_ITERS,NUM_EVALS);
+    adjustedAllDOPS = zeros(NUM_ITERS,NUM_EVALS);
+    allExperimental = zeros(NUM_ITERS,NUM_EVALS);
+    allExperimentalbeta = zeros(NUM_ITERS,NUM_EVALS);
+    allExperimentalgamma = zeros(NUM_ITERS,NUM_EVALS);
+    
+    allExperimentalgamma_shrinkfactor1E5 = zeros(NUM_ITERS,NUM_EVALS);
+    allExperimentalgamma_shrinkfactor1E6 = zeros(NUM_ITERS,NUM_EVALS);
+    allDOPSTimes = zeros(NUM_ITERS,1);
+    for j=1:NUM_ITERS
+       % if(j ==5) %need to skip 5 since rerunning at the moment
+       %    adjustedAllDOPS(j,:) = NaN;
+       %    continue;
+       % end
+        load(strcat('DOPS_error_iter',num2str(j),'.mat'));
+        PSO_error = g_best_solution{j};
+        corrected_PSO_error = fillInPSO(NUM_PARTCILES,PSO_error);
+        load(strcat('DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter', num2str(j),'.mat'));
+        DDS_error = bestval_dds_swarm{j};
+        alldata= cat(2,PSO_error,DDS_error);
+        correctedAllData = cat(2,corrected_PSO_error,DDS_error);
+        correctedAllData = correctedAllData(1:NUM_EVALS);
+        %allDOPS(j,:) =alldata;
+        adjustedAllDOPS(j,:) = correctedAllData;
+        
+        currTime = load(strcat('DOPS_time_iter', num2str(j), '.txt'));
+        currTime = currTime(j);
+        allDOPSTimes(j) = currTime;
+    end
+   
+    for k=1:NUM_ITERS
[…]

### 2017-11-03 DOPS/compareDOPSVersionsOnSelectedFunction.m

@@ -1,13 +1,14 @@
-function [] = compareESSToDOPSOnCoag()
+function [allDOPS,adjustedAllDOPS,allExperimental,allDOPSTimes]=compareDOPSVersionsOnSelectedFunction(functionName)
+    addpath(strcat('/home/rachel/Documents/DOPS/DOPS_Results/',functionName));
     close('all');
-    NUM_ITERS = 25;
+    NUM_ITERS = 250;
     NUM_EVALS = 4000;
     NUM_PARTCILES = 40;
     allDOPS = zeros(NUM_ITERS,NUM_EVALS);
     adjustedAllDOPS = zeros(NUM_ITERS,NUM_EVALS);
-    allESS = zeros(NUM_ITERS,NUM_EVALS+1);
-    addpath('/home/rachel/Documents/DOPS/DOPS_Results/fitCoag');
-    for j=1:25
+    allExperimental = zeros(NUM_ITERS,NUM_EVALS);
+    allDOPSTimes = zeros(NUM_ITERS,1);
+    for j=1:NUM_ITERS
         load(strcat('DOPS_error_iter',num2str(j),'.mat'));
         PSO_error = g_best_solution{j};
         corrected_PSO_error = fillInPSO(NUM_PARTCILES,PSO_error);
@@ -16,36 +17,31 @@ function [] = compareESSToDOPSOnCoag()
         alldata= cat(2,PSO_error,DDS_error);
         correctedAllData = cat(2,corrected_PSO_error,DDS_error);
         correctedAllData = correctedAllData(1:NUM_EVALS);
-       % allDOPS(j,:) =alldata;
+        %allDOPS(j,:) =alldata;
         adjustedAllDOPS(j,:) = correctedAllData;
+        
+        currTime = load(strcat('DOPS_time_iter', num2str(j), '.txt'));
+        currTime = currTime(j);
+        allDOPSTimes(j) = currTime;
     end
    
     for k=1:NUM_ITERS
-%         if(k ==18 || k ==19 || k ==20)
-%            allESS(k,:) = NaN;
-%            continue;
-%         end
-        load(strcat('/home/rachel/Documents/DOPS/ESS_Results/fitCoag/functionalInterpIter', num2str(k),'.mat'));
-        exp_err= f_interp(1:NUM_EVALS+1);
-        allESS(k,:) = exp_err;
[…]

### 2017-11-03 DOPS/compareExperimentalDOPS.m

@@ -8,6 +8,7 @@ function [allDOPS,adjustedAllDOPS,allExperimental,allDOPSTimes]=compareExperimen
     adjustedAllDOPS = zeros(NUM_ITERS,NUM_EVALS);
     allExperimental = zeros(NUM_ITERS,NUM_EVALS);
     allExperimentalbeta = zeros(NUM_ITERS,NUM_EVALS);
+    allExperimentalgamma = zeros(NUM_ITERS,NUM_EVALS);
     allDOPSTimes = zeros(NUM_ITERS,1);
     for j=1:NUM_ITERS
        % if(j ==5) %need to skip 5 since rerunning at the moment
@@ -51,18 +52,42 @@ function [allDOPS,adjustedAllDOPS,allExperimental,allDOPSTimes]=compareExperimen
         exp_err= g_best_solution{k};
         allExperimentalbeta(k,:) = exp_err;
     end
+    
+     for k=1:NUM_ITERS
+       if(k==7 || k==14 || k ==24)
+           allExperimental(k,:) = NaN;
+           continue;
+        end
+        
+        load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/ExperimentalV2gamma/fitCoagShrinkFactor1E6/DOPS_error_iter', num2str(k),'.mat'));
+        exp_err= g_best_solution{k};
+        allExperimentalgamma(k,:) = exp_err;
+    end
     f=figure();
     hold('on')
     semilogy(1:NUM_EVALS,nanmean(adjustedAllDOPS,1))
-    semilogy(1:NUM_EVALS,nanmean(allExperimental,1))
-    semilogy(1:NUM_EVALS,nanmean(allExperimentalbeta,1))
+    %semilogy(1:NUM_EVALS,nanmean(allExperimental,1))
+    %semilogy(1:NUM_EVALS,nanmean(allExperimentalbeta,1))
+    semilogy(1:NUM_EVALS,nanmean(allExperimentalgamma,1))
     %axis([0,4000,1E5,1E8])
     set(gca, 'YScale', 'log')
-    xlabel('Iteration Number')
+    xlabel('Number of Functional Evaluations')
     ylabel('Functional Value')
-    legend('DOPS-25 iters', 'Experimental DOPS-22 iters', 'Experimental DOPS beta-21 iters')
-    saveas(f, '../DOPS_Results/figures/DOPS_VS_BothExpDOPS.pdf', 'pdf');
[…]

### 2017-11-03 DOPS/compareExperimentalDOPSOnProblemB4.m

@@ -27,7 +27,7 @@ function [allDOPS,adjustedAllDOPS,allExperimental]=compareExperimentalDOPSOnProb
            %iter 9 didn't finish, so skip it for now
        %    continue; 
         %end
-        load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/ExperimentalV2/b4_obj/DOPS_error_iter', num2str(k),'.mat'));
+        load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/ExperimentalV2gamma/b4_obj/DOPS_error_iter', num2str(k),'.mat'));
         exp_err= g_best_solution{k};
         allExperimental(k,:) = exp_err;
     end
@@ -46,9 +46,9 @@ function [allDOPS,adjustedAllDOPS,allExperimental]=compareExperimentalDOPSOnProb
      hold('on')
      x = 1:NUM_EVALS;
      semilogy(1:NUM_EVALS,mean(adjustedAllDOPS,1), 'r', 'LineWidth', 2)
-     %semilogy(1:NUM_EVALS,mean(allExperimental,1), 'c')
+     semilogy(1:NUM_EVALS,mean(allExperimental,1), 'c')
      %semilogy(1:NUM_EVALS,mean(allExperimentalBeta,1), 'b')
-     semilogy(0:NUM_EVALS, mean(allESS,1), 'k', 'LineWidth', 2);
+     %semilogy(0:NUM_EVALS, mean(allESS,1), 'k', 'LineWidth', 2);
      
      %shade 1 stdev
 %      top1 = mean(adjustedAllDOPS,1)+1*std(adjustedAllDOPS,1);
@@ -60,23 +60,23 @@ function [allDOPS,adjustedAllDOPS,allExperimental]=compareExperimentalDOPSOnProb
      set(gca, 'YScale', 'log')
      xlabel('Number of Functional Evaluations')
      ylabel('Functional Value')
-     legend('DOPS', 'ESS')
-     saveas(f, '../DOPS_Results/figures/CompareExperimentalDOPSToESSOnProblemB4.pdf');
+     legend('DOPS', 'Exp DOPS ESS gamma')
+     saveas(f, '../DOPS_Results/figures/CompareExperimentalDOPSProblemB4.pdf');
      
-     f = figure();
-     hold('on')
-     for k = 1:NUM_ITERS
-         semilogy(1:NUM_EVALS,adjustedAllDOPS(k,:), 'g')
[…]

### 2017-11-03 DOPS/makeFig6Plot.m

@@ -1,6 +1,7 @@
 function makeFig6Plot(allcurve1, allcurve2,allcurve3,DFIN)
     close('all');
-   f=plot(mean(allcurve1,1), 'k');
+    f = figure();
+    plot(mean(allcurve1,1), 'k');
     hold('on')
     plot(mean(allcurve2,1), 'k');
     plot(mean(allcurve3,1), 'k');
@@ -33,5 +34,8 @@ function makeFig6Plot(allcurve1, allcurve2,allcurve3,DFIN)
     axis([0,200,0,1600])
     xlabel('Time (seconds)')
     ylabel('Thombin Concentration (nM)')
-    saveas(f,'../DOPS_Results/figures/ReplicateFigure6DivBySqrtN.pdf', 'pdf');
+    f.PaperPositionMode = 'auto';
+    f.PaperUnits = 'inches';
+    f.PaperPosition = [0 0 8 7];
+    print('../DOPS_Results/figures/ReplicateFigure5DivBySqrtN.pdf', '-dpdf','-r0');
 end

### 2017-11-03 DOPS/rast.m

@@ -1,10 +1,10 @@
 function y = rast(x,y)
-% 
-% Rastrigin function
-% Matlab Code by A. Hedar (Nov. 23, 2005).
-% The number of variables n should be adjusted below.
-% The default value of n = 2.
-% 
+
+%Rastrigin function
+%Matlab Code by A. Hedar (Nov. 23, 2005).
+%The number of variables n should be adjusted below.
+%The default value of n = 2.
+
 %for using with DE from http://www1.icsi.berkeley.edu/~storn/code.html#matl
 if(nargin<2)
     y = [];

### 2017-11-03 DOPS/recreateFigureS1.m

@@ -1,4 +1,5 @@
-function [x,bestparams,initialvals,functionalvals] = recreateFigureS1()
+tsfunction [x,bestparams,initialvals,functionalvals] = recreateFigureS1()
+    %also makes figures S3a and S3b
     close('all')
     NUM_REPEATS = 10;
     functionalvals = [];
@@ -15,6 +16,15 @@ function [x,bestparams,initialvals,functionalvals] = recreateFigureS1()
     [smallestf, minIdx] = nanmin(functionalvals); %ignore NAN
     bestparams = params(minIdx,:);
    [objective,constraints,residuals,x_sel,x,t,scaled_data] = b1_obj_forS1(bestparams);
+   containsNANS = max(isnan(x(:)));
+   repCounter = 1;
+   while(containsNANS ==1 && repCounter <100) %keep trying until we actually solve
+         fprintf('Attempting to solve problem b1. On attempt %d\n', repCounter);
+         [objective,constraints,residuals,x_sel,x,t,scaled_data] = b1_obj_forS1(bestparams);
+         containsNANS = max(isnan(x(:)));
+         repCounter = repCounter +1;
+   end
+   
     num_species = 44;
     scaled_data = scaled_data{1};
     f=figure();
@@ -32,7 +42,7 @@ function [x,bestparams,initialvals,functionalvals] = recreateFigureS1()
         [ydot,x_int,p] = b1([0]);
     f=figure();
     hold('on')
-    plot(log(p(1:size(bestparams,2))), log(bestparams), 'b.', 'MarkerSize', 20);
+    plot(log(p(1:size(bestparams,2))), log(bestparams), 'bo', 'MarkerSize', 5);
     plot([-15,10], [-15,10], 'k-');
     xlabel('Nominal Parameters (log)')
     ylabel('Optimal Parameters (log)')
@@ -43,11 +53,13 @@ function [x,bestparams,initialvals,functionalvals] = recreateFigureS1()
     scaled_data_o = scaled_data_o{1};
     f = figure();
     hold('on')
-    for j=1:num_states
-        plot(log(norm(x_o(:,j))), log(norm(x(:,j))), 'b.', 'MarkerSize',20);
+    %for j=1:num_states
+    for j = 1:size(x_sel_o,2)
+        %plot(log(norm(x_o(:,j))), log(norm(x(:,j))), 'b.', 'MarkerSize',20);
[…]

### 2017-11-03 DOPS/recreateFigureS2.m

@@ -1,4 +1,5 @@
 function [x,bestparams,functionalvals,initialvals] = recreateFigureS2()
+    %also makes s3b and s3d
     close('all')
     NUM_REPEATS = 25;
     functionalvals = [];
@@ -25,12 +26,12 @@ function [x,bestparams,functionalvals,initialvals] = recreateFigureS2()
     end
     f.PaperPositionMode = 'auto';
     f.PaperUnits = 'inches';
-    f.PaperPosition = [0 0 5 10];
+    f.PaperPosition = [0 0 6 11];
     print('../DOPS_Results/figures/RecreatedFigureS2.pdf', '-dpdf','-r0');
     [ydot,x_int,p] = b4([0]);
     f=figure();
     hold('on')
-    plot(log(p), log(bestparams), 'b.', 'MarkerSize', 20);
+    plot(log(p), log(bestparams), 'bo', 'MarkerSize', 5);
     plot([-6,8], [-6,8], 'k-');
     xlabel('Nominal Parameters (log)')
     ylabel('Optimal Parameters (log)')
@@ -40,11 +41,12 @@ function [x,bestparams,functionalvals,initialvals] = recreateFigureS2()
     [objective_o,constraints_o,residuals_o,x_all_o,x_o,t_o,timeseries_data_o,scaled_data_o] = b4_obj_forS2(p);
     f = figure();
     hold('on')
-    for j=1:size(x_all_o,2)
-        plot(log(norm(x_all_o(:,j))), log(norm(x_all(:,j))), 'b.', 'MarkerSize',20);
+    for j=1:size(x_o,2)
+        plot(log((x_o(:,j))), log((x(:,j))), 'bo', 'MarkerSize',5);
+        %plot(log((x_all_o(:,j))), log((x_all(:,j))), 'bo', 'MarkerSize',5);
     end
-    plot([0,14], [0,14], 'k-');
-    xlabel('Norm of Measured States With Nominal Paramters (log)')
-    ylabel('Norm of Measured States With Optimal Paramters (log)')
+    plot([-2,14], [-2,14], 'k-');
+    xlabel('Measured States With Nominal Paramters (log)')
+    ylabel('Measured States With Optimal Paramters (log)')
     print('../DOPS_Results/figures/RecreatedFigureS3d.pdf', '-dpdf','-r0');
 end

### 2017-11-03 DOPS/replicateFig5.m

@@ -40,7 +40,8 @@ end
 
 function makePlot(allcurve1, allcurve2,DFIN)
     close('all');
-   f=plot(mean(allcurve1,1), 'k')
+    f = figure();
+   plot(mean(allcurve1,1), 'k')
     hold('on')
     plot(mean(allcurve2,1), 'k')
     std(allcurve1,1)
@@ -61,7 +62,27 @@ function makePlot(allcurve1, allcurve2,DFIN)
     DATA2 = DFIN.EXPT_DATA_FIG2E5;
     plot(DATA1(:,1), DATA1(:,2), 'b.', 'MarkerSize', 12);
     plot(DATA2(:,1), DATA2(:,2), 'r.', 'MarkerSize',12);
-    legend(' 5pm Trigger', '5nm Trigger')
-    saveas(f,'../DOPS_Results/figures/ReplicateFigure5DivBySqrtN.pdf', 'pdf');
+   % legend('', '', '5pm Trigger', '5nm Trigger')
+    xlabel('Time (Seconds)');
+    ylabel('Thrombin Concentration (nM)');
+    %saveas(f,'../DOPS_Results/figures/ReplicateFigure5DivBySqrtN.pdf', 'pdf');
+    f.PaperPositionMode = 'auto';
+    f.PaperUnits = 'inches';
+    f.PaperPosition = [0 0 8 7];
+    print('../DOPS_Results/figures/ReplicateFigure5DivBySqrtN.pdf', '-dpdf','-r0');
+    NSE5pm = calculateNSE(DATA1,mean(allcurve1,1), x);
+    NSE5nm = calculateNSE(DATA2,mean(allcurve2,1),x);
+    fprintf('NSE for 5 pm trigger is %f\n', NSE5pm);
+    fprintf('NSE for 5 nm trigger is %f\n', NSE5nm);
+end
+
+function [NSE]= calculateNSE(exp_data, sim_data,sim_time)
+    time_pts = exp_data(:,1);
+    sim_data_interp = interp1(sim_time,sim_data, time_pts);
+    NSE = 0;
+    for j = 1:max(size(exp_data))
+       NSE = NSE +sqrt((exp_data(j,2)-sim_data_interp(j))^2); 
+    end
+    NSE = 1/max(exp_data(:,2))*1/sqrt(max(size(exp_data)))*NSE;
 end
 

### 2017-11-03 DOPS/replicateFig6.m

@@ -35,6 +35,17 @@ function [all_curve1,all_curve2,all_curve3,DFIN] = replicateFig6() %for DE, need
         all_curve3(j,:) = thrombin3;
     end
    makeFig6Plot(all_curve1,all_curve2,all_curve3,DFIN);
+    DATA1 = DFIN.EXPT_DATA_FIG2E2;
+    DATA2 = DFIN.EXPT_DATA_FIG2E3;
+    DATA3 = DFIN.EXPT_DATA_FIG2E4;
+    x = [1:1:250];
+    NSE10pm = calculateNSE(DATA1,mean(all_curve1,1), x);
+    NSE50pm = calculateNSE(DATA2,mean(all_curve2,1),x);
+    NSE500pm = calculateNSE(DATA3,mean(all_curve3,1),x);
+    
+    fprintf('NSE for 10 pm trigger is %f\n', NSE10pm);
+    fprintf('NSE for 50 pm trigger is %f\n', NSE50pm);
+    fprintf('NSE for 500 pm trigger is %f\n', NSE500pm);
 end
 function[all_params]= readInParams(num_iters)
     %num_iters = 25;
@@ -73,3 +84,12 @@ function [dynamics] = checkForDynamics(curve)
     end
 end
 
+function [NSE]= calculateNSE(exp_data, sim_data,sim_time)
+    time_pts = exp_data(:,1);
+    sim_data_interp = interp1(sim_time,sim_data, time_pts);
+    NSE = 0;
+    for j = 1:max(size(exp_data))
+       NSE = NSE +sqrt((exp_data(j,2)-sim_data_interp(j))^2); 
+    end
+    NSE = 1/max(exp_data(:,2))*1/sqrt(max(size(exp_data)))*NSE;
+end

### 2017-11-03 DOPS/runForTable1.m

@@ -1,8 +1,9 @@
 function[]=runForTable1(lowerIter, upperIter)
     addpath(genpath('CoagulationFiles')) %add coagulation functions to path
-    functions_to_test = {@rast, @ackley,@b4_obj, @b1_obj, @fitCoag};
-    bounds_files = {'rast_bounds.mat', 'ackley_bounds.mat', 'b4_bounds.mat','b1_bounds.mat', 'coagBounds.mat', 'coagBounds.mat'};
-    for j=4:size(functions_to_test,2)-1
+    functions_to_test = {@rast, @ackley,@b4_obj, @b1_obj, @fitCoag, @BukinF6, @rosenbrock, @Eggholder, @mccorm, @hart6, @stybtang};
+    bounds_files = {'rast_bounds.mat', 'ackley_bounds.mat', 'b4_bounds.mat','b1_bounds.mat', 'coagBounds.mat', 'BukinF6Bounds.mat','rosenbrock10d_bounds.mat', 'EggholderBounds.mat', 'mccorm_bounds.mat'...
+        'hart6_bounds.mat', 'stybtang100d_bounds.mat'};
+    for j=size(functions_to_test,2):size(functions_to_test,2)
         load(bounds_files{j});
         %deal with transposed files
         if(j>2 && j<=4)

### 2017-11-03 DOPS/runForTable1ExperimentalV2gamma.m

@@ -1,10 +1,11 @@
 function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls,DDS_iters]=runForTable1ExperimentalV2gamma(lowerIter,upperIter)
     addpath(genpath('CoagulationFiles')) %add coagulation functions to path
-    functions_to_test = {@rast, @ackley,@b4_obj, @b1_obj_v2, @fitCoag};
+    functions_to_test = {@rast, @ackley,@b4_obj, @b1_obj_v2, @fitCoag, @rast300, @BukinF6, @rosenbrock, @Eggholder, @mccorm, @hart6, @stybtang};
    % functions_to_test = {@rast, @ackley,@rosenbrock, @ButkinF6, @fitCoag};
-    bounds_files = {'rast_bounds.mat', 'ackley_bounds.mat', 'b4_bounds.mat','b1_bounds.mat', 'coagBounds.mat'};
+    bounds_files = {'rast_bounds.mat', 'ackley_bounds.mat', 'b4_bounds.mat','b1_bounds.mat', 'coagBounds.mat', 'rast300d_bounds.mat', 'BukinF6Bounds.mat','rosenbrock10d_bounds.mat', 'EggholderBounds.mat', 'mccorm_bounds.mat'...
+        'hart6_bounds' ,'stybtang100d_bounds'};
     %bounds_files = {'rast_bounds.mat', 'ackley_bounds.mat','rosenbrock10d_bounds.mat', 'BukinF6Bounds.mat','coagBounds.mat'}; 
-   for j=3:size(functions_to_test,2)-2
+   for j=5:5%size(functions_to_test,2):size(functions_to_test,2)%-4
         load(bounds_files{j});
         %deal with transposed files
         if(j ==3 || j ==4)

### 2017-11-03 DOPS/runSimulatedAnnealingOnCoag.m

@@ -7,11 +7,12 @@ function [allhistory]= runSimulatedAnnealingOnCoag()
         stop = false;
         if isequal(flag,'iter')
               history = [history; optimvalues.fval];
-		optimvalues.iteration %so I can see what's going on'
+            fprintf('Iter %d and functional value %f\n', optimvalues.iteration, optimvalues.fval);  
+            optimvalues.iteration; %so I can see what's going on'
         end
         optchanged = false;
     end
-	NUM_ITERATIONS = 500;
+	NUM_ITERATIONS = 2000;
 	options = optimoptions(@simulannealbnd,'MaxIterations',NUM_ITERATIONS, 'OutputFcn', @myoutput);
 	NUM_REPEATS = 25;
 	allhistory = zeros(NUM_REPEATS, NUM_ITERATIONS+1);

### 2017-11-03 DOPS/run_DOPS.m

@@ -54,7 +54,7 @@ function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_pa
         if(i ==4)
            rng(805); 
            
-        elseif(i ==5)
+        elseif(i ==5 || i ==20)
             rng(765); %let's try a different seed
         else
             rng(i); %for repeatability
@@ -68,26 +68,26 @@ function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_pa
 
        %-------This portion save results and additional results for every trial. It is optional and the user can choose the results and number of trials --------% 
        if(mod(i,1)==0)
-            cmd1 = ['save  ../DOPS_Results/',func2str(optFunction),'/DOPS_solution_iter',num2str(i),'.mat bestparticle'];                                                  % The best solution vector from swarm search
-            eval(cmd1)
+          %  cmd1 = ['save  ../DOPS_Results/',func2str(optFunction),'/DOPS_solution_iter',num2str(i),'.mat bestparticle'];                                                  % The best solution vector from swarm search
+           % eval(cmd1)
 
-            cmd2 = ['save  ../DOPS_Results/',func2str(optFunction),'/DOPS_particle',num2str(i),'.mat particle'];                                                           % The particle matrix - contains particle states for every iteration within the trial
-            eval(cmd2)
+            %cmd2 = ['save  ../DOPS_Results/',func2str(optFunction),'/DOPS_particle',num2str(i),'.mat particle'];                                                           % The particle matrix - contains particle states for every iteration within the trial
+            %eval(cmd2)
 
-            cmd3 = ['save  ../DOPS_Results/',func2str(optFunction),'/DOPS_fitness',num2str(i),'.mat fitness'];                                                             % The fitness matrix - contains fitness states for the corresponding particle matrix 
-            eval(cmd3)
+            %cmd3 = ['save  ../DOPS_Results/',func2str(optFunction),'/DOPS_fitness',num2str(i),'.mat fitness'];                                                             % The fitness matrix - contains fitness states for the corresponding particle matrix 
+            %eval(cmd3)
 
[…]

### 2017-11-03 DOPS/stybtang.m

@@ -1,0 +1,45 @@
+function [y] = stybtang(xx)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% STYBLINSKI-TANG FUNCTION
+%
+% Authors: Sonja Surjanovic, Simon Fraser University
+%          Derek Bingham, Simon Fraser University
+% Questions/Comments: Please email Derek Bingham at dbingham@stat.sfu.ca.
+%
+% Copyright 2013. Derek Bingham, Simon Fraser University.
+%
+% THERE IS NO WARRANTY, EXPRESS OR IMPLIED. WE DO NOT ASSUME ANY LIABILITY
+% FOR THE USE OF THIS SOFTWARE.  If software is modified to produce
+% derivative works, such modified software should be clearly marked.
+% Additionally, this program is free software; you can redistribute it 
+% and/or modify it under the terms of the GNU General Public License as 
+% published by the Free Software Foundation; version 2.0 of the License. 
+% Accordingly, this program is distributed in the hope that it will be 
+% useful, but WITHOUT ANY WARRANTY; without even the implied warranty 
+% of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU 
+% General Public License for more details.
+%
+% For function details and reference information, see:
+% http://www.sfu.ca/~ssurjano/
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% INPUT:
+%
+% xx = [x1, x2, ..., xd]
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+d = length(xx);
+sum = 0;
+for ii = 1:d
+	xi = xx(ii);
+	new = xi^4 - 16*xi^2 + 5*xi;
+	sum = sum + new;
+end
+
+y = sum/2;
+
+end

### 2017-11-03 Paper_v6.tex
adding updated manuscript

@@
 
-\subsection*{Contiued Switching May Be Beneficial}
+\subsection*{Continued Switching May Be Beneficial}
 We explored the performance of DOPS if it was permitted to switch between PSO and DDS modes.
@@
 \ref{fig-exp-DOPS-b4}b).
-We also explored the performance of both versions of DOPS on the 100 dimensional Styblinksi-Tang function (Fig.
+We also explored the performance of both versions of DOPS on the 100 dimensional Styblinksi-Tang function \cite{jamil2013literature} (Fig.
 \ref{fig-exp-DOPS-b4}c).
@@
 Since the performance of Experimental DOPS was promising on these problems, we decided to measure its performance on the coagulation problem.
+At 4,000 function evaluations, DOPS performed better than Experimental DOPS, however, the difference between the two method's final average functional values was small (Fig.
+\ref{fig-exp-DOPS-b4}d), with the final average objective value for DOPS reaching .9413\% of the initial functional value, compared to .9428\% for Experimental DOPS.
 
@@
 \centering
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnb4_obj.pdf} }}%
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnb4_obj.pdf} }}
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnEggholder}} }%
         \qquad
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnEggholder}} }%
     \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnstybtang}} }%    
+        \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnfitCoagShrinkFactor1E6}} }%
 
-\caption{Performance of DOPS and Experimental DOPS on the CHO metabolism problem (a), the Eggholder function (b), and the 100 dimensional Styblinksi-Tang function (c).
+\caption{Performance of DOPS and Experimental DOPS on the CHO metabolism problem (a), the Eggholder function (b), the 100 dimensional Styblinksi-Tang function (c) and the coagulation problem (d).
 Both methods have the same initial decrease in error, but as the number of function evaluations increases, experimental DOPS produces a larger decrease in error.
-The results presented are the average of $\mathcal{T}$ = 250 trials with for the CHO metabolism problem and $\mathcal{T}$ = 500 trials on the Eggholder and Styblinksi-Tang functions with $\mathcal{N}$ = 4000 function evaluations each.
+The results presented are the average of $\mathcal{T}$ = 250 trials with for the CHO metabolism problem and $\mathcal{T}$ = 500 trials on the Eggholder and Styblinksi-Tang functions with $\mathcal{N}$ = 4000 function evaluations each, and $\mathcal{T}$ = 25 trials for the coagulation problem.
 }\label{fig-exp-DOPS-b4}

### 2017-11-09 Paper_v6.tex
updating manuscript and figures

@@
 We evaluated the performance of DOPS on problems from the least and most expensive categories.
-The least expensive problem (henceforth referred to as problem B4) was a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters \cite{villaverde2014high}.
+The least expensive problem was a metabolic model of Chinese Hamster Ovary (CHO) with 35 metabolites, 32 reactions and 117 parameters \cite{villaverde2014high}.
 The biochemical reactions were modeled using modular rate laws and generalized Michaelis-Menten kinetics.
-On the other hand, the expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters \cite{smallbone2013large} (henceforth referred to as problem B1).
+On the other hand, the expensive problem was a genome scale kinetic model of \textit{Saccharomyces~cerevisiae} with 261 reactions, 262 variables and 1759 parameters \cite{smallbone2013large}.
 In both cases, synthetic time series data generated with known parameter values, was used to estimate model parameters.
-For problem B1, the time series data consisted of 44 observables, and for problem B4 the data corresponded to 13 different metabolite measurement sets.
+For the \textit{Saccharomyces~cerevisiae} model, the time series data consisted of 44 observables, and for the CHO metabolism problem the data corresponded to 13 different metabolite measurement sets.
 We fixed the number of function evaluations at $\mathcal{N}$ = 4000, and trained both models against the synthetic experimental data.
@@
 The optimal or near optimal parameters obtained by DOPS predicted unseen coagulation datasets (Fig. \ref{fig-validation}).
-The normalized standard error for the coagulation predictions was consistent with the training error, with the exception of the 50pM TF/VIIa case which was a factor 2.75 worse (Table \ref{table:Error-Analysis}).
+The normalized standard error for the coagulation predictions was consistent with the training error, with the exception of the 50pM TF/VIIa case which was a factor 2.65 worse (Table \ref{table:Error-Analysis}).
 Coagulation initiation with 50pM TF/FVIIa was the farthest away from the training conditions.
-However, while the 50pM TF/FVIIa case was the worst of those tested, it still had an average error rate of only 11\% (which is near the margin of error for the experimental measurements).
+%However, while the 50pM TF/FVIIa case was the worst of those tested, it still had an average error rate of only 11\% (which is near the margin of error for the experimental measurements).
 Taken together, DOPS estimated parameter sets with predictive power on unseen coagulation data using fewer function iterations than other meta-heuristics.
@@
 This mode ('Experimental DOPS') performed comparably to DOPS on 10-d Ackley and Rastrigin as well as on 300-dimensional Rastrigin.
-It performed better than DOPS on the CHO metabolism problem (b4) (Fig.
+It performed better than DOPS on the CHO metabolism problem (Fig.
 \ref{fig-exp-DOPS-b4}a), with the average functional value for Experimental DOPS being nearly half of that of DOPS.
@@
 dimensioned search phase.
-Next, the current implementation of the algorithm switches only once from the swarm phase to the dynamically dimensioned phase.
-An obvious modification would be to allow multiple transitions between the swarm and dynamically dimensioned phases.
+%Next, the current implementation of the algorithm switches only once from the swarm phase to the dynamically dimensioned phase.
+%An obvious modification would be to allow multiple transitions between the swarm and dynamically dimensioned phases.
 Lastly, we should update DOPS to treat multi-objective problems.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure_7_ErrorSwarms_v2.pdf}
+%\includegraphics[width=1.0\textwidth]{./figs/Figure_7_ErrorSwarms_v2.pdf}
+\includegraphics[trim = {3cm 6cm 3cm 6cm }, clip,width = 1.0\textwidth]{./rachelfigs/RecreateFigure7}
 \caption{Influence of the switching strategy and sub-swarms on DOPS performance for the coagulation model.
@@
 The convergence rates with higher swarm numbers is typically higher but there is no pronounced difference amongst four, five and eight.
-The multi-swarm with DDS saturates while DOPS shows a rapid drop due to a switch to the DDS phase.
+The multi-swarm with without DDS saturates while DOPS shows a rapid drop due to a switch to the DDS phase.
 }\label{fig-sub-swarms}

### 2017-11-28 Paper_v6.tex
updating manuscript and figures

@@
 \centering
-\includegraphics[width=1.00\textwidth]{./figs/Figure_S1_B1_sims.pdf}
-\caption{\textbf {(Data fits for Problem B1)} Pseudo-experimental data (red circles) vs.
+%\includegraphics[width=1.00\textwidth]{./figs/Figure_S1_B1_sims.pdf}
+\includegraphics[width=1.00\textwidth]{./rachelfigs/RecreatedFigureS1usingMS2_save.pdf}
+\caption{\textbf {(Data fits for Problem B1)} Pseudo-experimental data (red x) vs.
 optimal solution obtained using DOPS (solid blue lines) for the 44 observed states.
@@
 \centering
-\includegraphics[width=1.00\textwidth]{./figs/Figure_S3_B1_B4_params_measuredstates.pdf}
+%\includegraphics[width=1.00\textwidth]{./figs/Figure_S3_B1_B4_params_measuredstates.pdf}
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/RecreatedFigureS3a} }}%
+    \qquad
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/RecreatedFigureS3b} }}%
+    \qquad
+        \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/RecreatedFigureS3c} }}%
+    \qquad
+    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/RecreatedFigureS3d} }}%
 \caption{ \textbf {(A)} Difference between nominal and optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
@@
     \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/CompareExperimentalDOPSToESSOnCoag} }}%
-    \caption{Mean convergence curves for DOPS and ESS for (a) CHO model (problem B4) (b) coagulation with $\mathcal{T}$ = 25 trials per method.}%
+    \caption{Mean convergence curves for DOPS and ESS for (a) CHO model  (b) coagulation with $\mathcal{T}$ = 25 trials per method.}%
     \label{fig-compare-DOPS-ESS}%
@@
     \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/DOPSDispersionCurvesOnCoag} }}%
-    \caption{Dispersion curves for DOPS on (a) CHO model (problem B4) (b) coagulation with $\mathcal{T}$ = 25 trials per problem .}%
+    \caption{Dispersion curves for DOPS on (a) CHO model (b) coagulation with $\mathcal{T}$ = 25 trials per problem .}%
     \label{fig-DOPS-Dispersion-Curves}%
@@
     \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/FigureS7d} }}%
-    \caption{Variability  analysis in best objective value for $\mathcal{T}$ = 25 trials. (a) Ackley  300 dimensional (b) Rastrigin 300 dimensional (c) CHO model (problem b4) (d) coagulation. }%
+    \caption{Variability  analysis in best objective value for $\mathcal{T}$ = 25 trials. (a) Ackley  300 dimensional (b) Rastrigin 300 dimensional (c) CHO model (d) coagulation. }%
     \label{fig-DOPS-Variability}%

### 2017-12-04 Paper_v7.tex
adding v7 in BMC format

--- 2017-11-09 Paper_v6.tex
@@
-\documentclass[12pt]{article}
+%% BioMed_Central_Tex_Template_v1.06
+%%                                      %
+%  bmc_article.tex            ver: 1.06 %
+%                                       %
+
+%%IMPORTANT: do not delete the first line of this template
+%%It must be present to enable the BMC Submission system to
+%%recognise this template!!
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                     %%
+%%  LaTeX template for BioMed Central  %%
+%%     journal article submissions     %%
+%%                                     %%
+%%          <8 June 2012>              %%
+%%                                     %%
+%%                                     %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                                 %%
+%% For instructions on how to fill out this Tex template           %%
+%% document please refer to Readme.html and the instructions for   %%
+%% authors page on the biomed central website                      %%
+%% http://www.biomedcentral.com/info/authors/                      %%
+%%                                                                 %%
+%% Please do not use \input{...} to include other tex files.       %%
+%% Submit your LaTeX manuscript as one .tex document.              %%
+%%                                                                 %%
+%% All additional figures and files should be attached             %%
+%% separately and not embedded in the \TeX\ document itself.       %%
+%%                                                                 %%
+%% BioMed Central currently use the MikTex distribution of         %%
+%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
+%% http://www.miktex.org                                           %%
+%%                                                                 %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%% additional documentclass options:
+%  [doublespacing]
+%  [linenumbers]   - put the line numbers on margins
+
+%%% loading packages, author definitions
+
+%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
+\documentclass{bmcart}
+
+%%% Load packages
+%\usepackage{amsthm,amsmath}
+%\RequirePackage{natbib}
+%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
+%\RequirePackage{hyperref}
+\usepackage[utf8]{inputenc} %unicode support
+%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
+%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
 % Load packages
@@
 
-%\textwidth = 6.50 in
-%\textheight = 9.5 in
-%\oddsidemargin =  0.0 in
-%\evensidemargin = 0.0 in
-%\topmargin = -0.50 in
-%\headheight = 0.0 in
-%\headsep = 0.25 in
-%\parskip = 0.15in
-%\linespread{1.75}
-\doublespace
-
-%\bibliographystyle{chicago}
-\bibliographystyle{plos2009}
-
-\makeatletter
-\renewcommand\subsection{\@startsection
-	{subsection}{2}{0mm}
-	{-0.05in}
-	{-0.5\baselineskip}
-	{\normalfont\normalsize\bfseries}}
-\renewcommand\subsubsection{\@startsection
-	{subsubsection}{2}{0mm}
-	{-0.05in}
-	{-0.5\baselineskip}
-	{\normalfont\normalsize\itshape}}
-\renewcommand\section{\@startsection
-	{subsection}{2}{0mm}
-	{-0.2in}
-	{0.05\baselineskip}
-	{\normalfont\large\bfseries}}
-\renewcommand\paragraph{\@startsection
-	{paragraph}{2}{0mm}
-	{-0.05in}
-	{-0.5\baselineskip}
-	{\normalfont\normalsize\itshape}}
-\makeatother
-
-%Review style settings
-%\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}
-
-%Publication style settings
-
-% Single space'd bib -
-\setlength\bibsep{0pt}
-
-\renewcommand{\rmdefault}{phv}\renewcommand{\sfdefault}{phv}
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-
-% Change the number format in the ref list -
-\renewcommand{\bibnumfmt}[1]{#1.}
-
-% Change Figure to Fig.
-\renewcommand{\figurename}{Fig.}
-
-% Begin ...
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                             %%
+%%  If you wish to display your graphics for   %%
+%%  your own use using includegraphic or       %%
+%%  includegraphics, then comment out the      %%
+%%  following two lines of code.               %%
+%%  NB: These line *must* be included when     %%
+%%  submitting to BMC.                         %%
+%%  All figure files must be submitted as      %%
+%%  separate graphics through the BMC          %%
+%%  submission process, not included in the    %%
+%%  submitted article.                         %%
+%%                                             %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\def\includegraphic{}
+%\def\includegraphics{}
+
+%%% Put your definitions there:
+\startlocaldefs
+\endlocaldefs
+
+%%% Begin ...
 \begin{document}
-\begin{titlepage}
-{\par\centering\textbf{\Large {Dynamic Optimization with Particle Swarms (DOPS): A meta-heuristic for parameter estimation in biochemical models}}}
-\vspace{0.05in}
-{\par \centering \large{Adithya Sagar, Christine Shoemaker$^{\dag}$ and Jeffrey D. Varner$^{*}$}}
-\vspace{0.10in}
-{\par \centering {School of Chemical and Biomolecular Engineering}}
-{\par \centering {$^{\dag}$School of Civil and Environmental Engineering}}
-{\par \centering {Cornell University, Ithaca NY 14853}}
-\vspace{0.1in}
-{\par \centering \textbf{Running Title:}~Parameter estimation in biochemical models}
-\vspace{0.1in}
-{\par \centering \textbf{To be submitted:}~\emph{Biotechnology~Journal}}
-\vspace{0.5in}
-{\par \centering $^{*}$Corresponding author:}
-{\par \centering Jeffrey D. Varner,}
-{\par \centering Professor, School of Chemical and Biomolecular Engineering,}
-{\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853}
-{\par \centering Email: jdv27@cornell.edu}
-{\par \centering Phone: (607) 255 - 4258}
-{\par \centering Fax: (607) 255 - 9166}
-\end{titlepage}
-\date{}
-\thispagestyle{empty}
-\pagebreak
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Abstract}
+
+%%% Start of article front matter
+\begin{frontmatter}
+
+\begin{fmbox}
+\dochead{Research}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Enter the title of your article here     %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\title{Parameter estimation in biochemical models}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Enter the authors here                   %%
+%%                                          %%
+%% Specify information, if available,       %%
+%% in the form:                             %%
+%%   <key>={<id1>,<id2>}                    %%
+%%   <key>=                                 %%
+%% Comment or delete the keys which are     %%
+%% not used. Repeat \author command as much %%
+%% as required.                             %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\author[
+   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
+   corref={aff1},                       % id of corresponding address, if any
+   noteref={n1},                        % id's of article notes, if any
+   email={asg242@cornell.edu}   % email address
+]{\fnm{Adithya} \snm{Sagar}}
+\author[
+   addressref={aff1},
+   email={ril34@cornell.edu}
+]{\fnm{Rachel} \snm{LeCover}}
+\author[
+   addressref={aff1},
+   email={jdv27@cornell.edu}
+]{\inits{D}\fnm{Jeffrey} \snm{Varner}}
+\author[
+   addressref={aff2},
+   email={jdv27@cornell.edu}
+]{\fnm{Christine} \snm{Shoemaker}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Enter the authors' addresses here        %%
+%%                                          %%
+%% Repeat \address commands as much as      %%
+%% required.                                %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\address[id=aff1]{%                           % unique id
+  \orgname{Robert Fredrick Smith School of Chemical and Biomolecular Engineering, Cornell University}, % university, etc
+  \street{244 Olin Hall},                     %
+  %\postcode{}                                % post or zip code
+  \city{Ithaca},                              % city
+  \cny{NY}                                    % country
+}
+\address[id=aff2]{%
+  \orgname{School of Civil and Environmental Engineering},
+  \street{Cornell University},
+  %\postcode{24105}
+  \city{Ithaca},
+  \cny{NY}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Enter short notes here                   %%
+%%                                          %%
+%% Short notes will be after addresses      %%
+%% on first page.                           %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{artnotes}
+%\note{Sample of title note}     % note to the article
+\note[id=n1]{Equal contributor} % note, connected to author
+\end{artnotes}
+
+\end{fmbox}% comment this for two column layout
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% The Abstract begins here                 %%
+%%                                          %%
+%% Please refer to the Instructions for     %%
+%% authors on http://www.biomedcentral.com  %%
+%% and include the section headings         %%
+%% accordingly for your article type.       %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{abstractbox}
+
+\begin{abstract} % abstract
+\parttitle{Background}
 Mathematical modeling is a powerful tool to analyze, and design biochemical networks.
@@
 DOPS uses a multi-swarm particle swarm optimization technique to generate candidate solution vectors, the best of which is greedily updated using dynamically dimensioned search.
+\parttitle{Results}
 We tested DOPS using a combination of classic optimization test functions, biochemical benchmark problems and real-world biochemical models.
@@
 benchmark problems and a model of the human coagulation cascade.
-Taken together, DOPS is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.
+\parttitle{Conclusions}
+DOPS is a promising meta-heuristic approach for the estimation of biochemical model parameters in relatively few function evaluations.
 DOPS source code is available for download under a MIT license at http://www.varnerlab.org.
-
-%We further tested the predictive power of the coagulation model parameters against data not used in training, and found good agreement between simulations and experimental measurements.
-Lastly, we tested the performance of DOPS on commonly used test functions for global optimization and on published biochemical parameter estimation benchmark problems.
-For the wide range of problems that we considered, DOPS outperformed other meta-heuristic approaches despite a limited number of function evaluations.
-
-\vspace{0.1in}
-{\noindent \textbf{Keywords:}~Biochemical engineering, systems biology, modeling, optimization}
-
-% Extra abstract
-% Mathematical modeling of biological systems with multiple feed back loops is one such area where parameter estimation is a difficult non-linear optimization problem.
-This difficulty is further compounded when dealing with parameter vectors of high dimensions.
-
-%In this study, we present the dynamically dimensioned particle
-%a novel meta-heuristic approach that combines a variant of particle swarm optimization (PSO) along with dynamically dimensioned search (DDS) to obtain near optimal solutions of high dimensional biochemical networks within a relatively few function evaluations.
-%We use a particle swarm optimization technique that uses multi-swarms to generate candidate vectors which are then greedily updated using DDS by dynamically varying the perturbed parameter dimensions.
-We tested this algorithm (25 trials with 4000 function evaluations in each trial) on a biochemical network of coagulation (148 parameters and 92 species) and compared it's performance against other meta-heuristics like Differential Evolution (DE), Particle Swarm Optimization (PSO), Simulated Annealing (SA) and also against DDS alone.
-The new algorithm outperforms all the other meta-heuristics on the coagulation model.
-The parameter vectors obtained using this approach fit the experimental data well and also make accurate enough predictions on unseen experimental data.
-We also performed this comparison on commonly used test functions (Ackley and Rosenbrock) for global optimization and found the same behavior.
-Further we used two recently published benchmark problems, a genome wide kinetic model with 1759 parameters and a metabolic model of Chinese Hamster Ovary cells with 117 parameters to evaluate the performance of our approach.
-We  surprisingly performed well on these benchmarks and obtained the nominal parameter vector with just 4000 function evaluations in both cases.
-
-\pagebreak
-
-\setcounter{page}{1}
-
-\linenumbers
-
-\section*{Introduction}
-
+\end{abstract}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% The keywords begin here                  %%
+%%                                          %%
+%% Put each keyword in separate \kwd{}.     %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{keyword}
+\kwd{Biochemical engineering}
+\kwd{systems biology}
+\kwd{modeling}
+\kwd{optimization}
+\end{keyword}
+
+% MSC classifications codes, if any
+%\begin{keyword}[class=AMS]
+%\kwd[Primary ]{}
+%\kwd{}
+%\kwd[; secondary ]{}
+%\end{keyword}
+
+\end{abstractbox}
+%
+%\end{fmbox}% uncomment this for twcolumn layout
+
+\end{frontmatter}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% The Main Body begins here                %%
+%%                                          %%
+%% Please refer to the instructions for     %%
+%% authors on:                              %%
+%% http://www.biomedcentral.com/info/authors%%
+%% and include the section headings         %%
+%% accordingly for your article type.       %%
+%%                                          %%
+%% See the Results and Discussion section   %%
+%% for details on how to create sub-sections%%
+%%                                          %%
+%% use \cite{...} to cite references        %%
+%%  \cite{koon} and                         %%
+%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
+%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
+% <put your article body there>
+
+%%%%%%%%%%%%%%%%
+%% Background %%
+%%
+\section*{Background}
 Cells process nutrients and respond to changes in their environment using complex enzyme catalyzed biochemical networks.
@@
 
-\clearpage
-
 \section*{Results}
@@
 \ref{fig-exp-DOPS-b4}d), with the final average objective value for DOPS reaching .9413\% of the initial functional value, compared to .9428\% for Experimental DOPS.
-
-\clearpage
-
-%As the dimensionality of  increases, the search region gets widened and thus the problem becomes more challenging.
-%
 
@@
 
-\clearpage
-
-\section*{Materials and Methods}
+\section*{Methods}
 
@@
 
-\newpage
-%\section*{Pseudo Code.}
-%\begin{algorithm}[H]
-%
-% \While{$besterror\leq $}{
-%  read current\;
-%  \eIf{understand}{
-%   go to next section\;
-%   current section becomes this one\;
-%   }{
-%   go back to the beginning of current section\;
-%  }
-% }
-% \caption{Dynamic Optimization with Particle Swarms}
-%\end{algorithm}
-
-\clearpage
-
-%Need to talk more about biochemical benefits and importance for of biochemical problems
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Backmatter begins here                   %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{backmatter}
+
+\section*{Competing interests}
+  The authors declare that they have no competing interests.
+
+\section*{Author's contributions}
+    A.S. originally developed algorithm. R.L developed experimental DOPS and generated all of the figures. JDV assisted in manuscript preparation.
 
@@
  This study was supported by an award from the Army Research Office (ARO \#59155-LS).
-\clearpage
-
-\bibliography{References_v2}
-
-\clearpage
-
-\begin{table}[h]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                  The Bibliography                       %%
+%%                                                         %%
+%%  Bmc_mathpys.bst  will be used to                       %%
+%%  create a .BBL file for submission.                     %%
+%%  After submission of the .TEX file,                     %%
+%%  you will be prompted to submit your .BBL file.         %%
+%%                                                         %%
+%%                                                         %%
+%%  Note that the displayed Bibliography will not          %%
+%%  necessarily be rendered by Latex exactly as specified  %%
+%%  in the online Instructions for Authors.                %%
+%%                                                         %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% if your bibliography is in bibtex format, use those commands:
+\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
+\bibliography{References_v2}      % Bibliography file (usually '*.bib' )
+% for author-year bibliography (bmc-mathphys or spbasic)
+% a) write to bib file (bmc-mathphys only)
+% @settings{label, options="nameyear"}
+% b) uncomment next line
+%\nocite{label}
+
+% or include bibliography directly:
+% \begin{thebibliography}
+% \bibitem{b1}
+% \end{thebibliography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                               %%
+%% Figures                       %%
+%%                               %%
+%% NB: this is for captions and  %%
+%% Titles. All graphics must be  %%
+%% submitted separately and NOT  %%
+%% included in the Tex document  %%
+%%                               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% Do not use \listoffigures as most will included as separate files
+
+\section*{Figures}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.85\textwidth]{./figs/Figure_1_Algorithm.pdf}
+\caption{\csentence{Schematic of the dynamic optimization with particle swarms (DOPS) approach.} \textbf{A}: Each particle represents an N dimensional parameter vector.
+Particles are given randomly generated initial solutions and grouped into different sub-swarms.
+Within each swarm the magnitude and direction of the movement a particle is influenced by the position of the best particle and also by its own experience.
+After every $\mathbf{g}$ number of function evaluations the particles are mixed and randomly assigned to different swarms.
+When the error due to the global best particle (best particle amongst all the sub-swarms) does not drop over a certain number of function evaluations, the swarm search is stopped and the search switches to a Dynamically Dimensioned Search with global best particle as the initial solution vector or candidate vector.
+\textbf{B}: The candidate vector performs a greedy global search for the remaining number of function evaluations.
+The search neighborhood is dynamically adjusted by varying the number of dimensions that are perturbed (in black) in each evaluation step.
+The probability that a dimension is perturbed decreases as the number of function evaluations increase.
+}\label{fig-algorithm}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.60\textwidth]{./rachelfigs/fig2cropped.pdf}
+\caption{\csentence{Performance of DOPS and other meta-heuristics for the Ackley and Rastrigin functions.}
+\textbf{A}: Mean scaled error versus the number of function evaluations for the 10-dimensional Ackley function.
+DOPS, DDS and PSO find optimal or near optimal solutions within the specified number of function evaluations.
+\textbf{B}: Mean scaled error versus the number of function evaluations for the 10-dimensional Rastrigin function.
+DOPS and DDS find optimal or near optimal solutions within the specified number of function evaluations.
+\textbf{C}: Mean scaled error versus the number of function evaluations for the 300-dimensional Rastrigin function.
+DOPS is the only algorithm that finds an optimal or near optimal solution within the specified number of function evaluations.
+In all cases, the maximum number of function evaluations was $\mathcal{N}$ = 4000.
+Mean and standard deviation were calculated over $\mathcal{T}$ = 25 trials.
+}\label{fig-testfunctions}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=1.00\textwidth]{./figs/Figure_3_Coagulation_v4.pdf}
+\caption{\csentence{Schematic of the extrinsic and intrinsic coagulation cascade.} Inactive zymogens upstream (grey) are activated by exposure to tissue factor (TF)  following vessel injury.
+Tissue factor and activated factor VIIa (FVIIa) form a complex that activates factor X (fX) and IX (fIX).
+FXa activates downstream factors including factor VIII (fVIII) and fIX.
+Factor V (fV) is primarily activated by thrombin (FIIa).
+In addition, we included a secondary fV activation route involving FXa.
+FXa and FVa form a complex (prothrombinase) on activated platelets that converts prothrombin (fII) to FIIa.
+FIXa and FVIIIa can also form a complex (tenase) on activated platelets which catalyzes FXa formation.
+Thrombin also activates upstream coagulation factors, forming a strong positive feedback ensuring rapid activation.
+Tissue factor pathway inhibitor (TFPI) downregulates FXa formation and activity by sequestering free FXa and TF-FVIIa in a FXa-dependent manner.
+Antithrombin III (ATIII)  inhibits all proteases.
+Thrombin inhibits itself binding the surface protein thrombomodulin (TM).
+The IIa-TM complex catalyzes the conversion of protein C (PC) to activated protein C (APC), which attenuates the coagulation response by the proteolytic cleavage of fV/FVa and fVIII/FVIIIa.
+}\label{fig-coagulation-network}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+%\includegraphics[width=1.0\textwidth]{./figs/Figure_4_Errors_Convergence_v2.pdf}
+\includegraphics[trim = {3cm 8cm 3cm 7cm }, clip,width=1.0\textwidth]{./rachelfigs/RecreateFigure4.pdf}
+\caption{\csentence{Error convergence rates of the four different algorithms on the coagulation model.} The objective error is the mean over $\mathcal{T}$= 25 trials.
+DOPS and SA have the steepest drop in error during first 300 function evaluations. Thereafter the error drop in DDS and SA remains nearly constant whereas DOPS continues to drops further.
+At the end of 4000 function evaluations DOPS attains the lowest error. The next best estimate using DDS is nearly three times greater than the lowest error using DE.
+}\label{fig-convergence}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+%\includegraphics[width=1.0\textwidth]{./figs/Figure_5_Sim_Train_E1_E5.pdf}
+\includegraphics[trim = {0cm 6cm 7cm 6cm }, clip,width=1.0\textwidth]{./rachelfigs/ReplicateFigure5DivBySqrtN.pdf}
+\caption{\csentence{Model fits on experimental data using DOPS.} The model parameters were estimated using DOPS.
+Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from 25 trials.
+The grey shaded region represents the 99\% confidence estimate of the mean simulated thrombin concentration.
+The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers.
+Thrombin generation is initiated by adding Factor TF/VIIa (5nM (blue) and 5pM (red)) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
+}\label{fig-train}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+%\includegraphics[width=1.0\textwidth]{./figs/Figure_6_Sim_Validate_E2_E4_E6.pdf}
+\includegraphics[trim = {3cm 6cm 3cm 6cm }, clip,width=1.0\textwidth]{./rachelfigs/ReplicateFigure6DivBySqrtN.pdf}
+\caption{\csentence{Model predictions on unseen experimental data using parameters obtained from DOPS.} The parameter estimates that were obtained using DOPS were tested against data that was not used in the model training.
+Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from $\mathcal{T}$ = 25 trials.
+The grey shaded region represents the 99\% confidence estimate of the mean simulated thrombin concentration.
+The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers.
+Thrombin generation is initiated by adding Factor VIIa-TF (500pM - Blue, 50pM - Pink and 10pM - purple, respectively) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
+}\label{fig-validation}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+%\includegraphics[width=1.0\textwidth]{./figs/Figure_7_ErrorSwarms_v2.pdf}
+\includegraphics[width = 1.0\textwidth]{./rachelfigs/RecreateFigure7}
+\caption{\csentence{Influence of the switching strategy and sub-swarms on DOPS performance for the coagulation model.}
+DOPS begins by using a particle swarm search and then dynamically switches (switch region), using an adaptive switching criteria, to the DDS search phase.
+We compared the performance of DOPS with and without DDS for different sub-swarm searches to quantify the effect of number of sub-swarms and DDS.
+We used one, two, four, five and eight sub-swarms, with a total of 40 particles divided evenly amongst the swarms.
+The results presented are the average of $\mathcal{T}$ = 25 trials with $\mathcal{N}$ = 4000 function evaluations each.
+The convergence rates with higher swarm numbers is typically higher but there is no pronounced difference amongst four, five and eight.
+The multi-swarm with without DDS saturates while DOPS shows a rapid drop due to a switch to the DDS phase.
+}\label{fig-sub-swarms}
+\end{figure}
+
+  \begin{figure}[h!]
+  \caption{\csentence{Sample figure title.}
+      A short description of the figure content
+      should go here.}
+      \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                               %%
+%% Tables                        %%
+%%                               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Use of \listoftables is discouraged.
+%%
+\section*{Tables}
+\begin{table}[h!]
 \centering
@@
 
-\begin{table}[h]
+\begin{table}[h!]
 \centering
@@
 \end{table}
-
-\clearpage
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=1.0\textwidth]{./figs/Figure_1_Algorithm.pdf}
-\caption{Schematic of the dynamic optimization with particle swarms (DOPS) approach.
-\textbf{A}: Each particle represents an N dimensional parameter vector.
-Particles are given randomly generated initial solutions and grouped into different sub-swarms.
-Within each swarm the magnitude and direction of the movement a particle is influenced by the position of the best particle and also by its own experience.
-After every $\mathbf{g}$ number of function evaluations the particles are mixed and randomly assigned to different swarms.
-When the error due to the global best particle (best particle amongst all the sub-swarms) does not drop over a certain number of function evaluations, the swarm search is stopped and the search switches to a Dynamically Dimensioned Search with global best particle as the initial solution vector or candidate vector.
-\textbf{B}: The candidate vector performs a greedy global search for the remaining number of function evaluations.
-The search neighborhood is dynamically adjusted by varying the number of dimensions that are perturbed (in black) in each evaluation step.
-The probability that a dimension is perturbed decreases as the number of function evaluations increase.
-}\label{fig-algorithm}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[ht]
-\centering
-%\includegraphics[width=0.60\textwidth]{./figs/Figure_2_TestFunctions-crop.pdf}
-    \subfloat[]{{\includegraphics[width=.6\textwidth]{./rachelfigs/ackley_10d} }}%
-    \newline
-    \subfloat[]{{\includegraphics[width=.6\textwidth]{./rachelfigs/rast_10d} }}%
-    \newline
-    \subfloat[]{{\includegraphics[width=.6\textwidth]{./rachelfigs/rast_300d} }}%
-
-\caption{Performance of DOPS and other meta-heuristics for the Ackley and Rastrigin functions.
-\textbf{A}: Mean scaled error versus the number of function evaluations for the 10-dimensional Ackley function.
-DOPS, DDS and PSO find optimal or near optimal solutions within the specified number of function evaluations.
-\textbf{B}: Mean scaled error versus the number of function evaluations for the 10-dimensional Rastrigin function.
-DOPS and DDS find optimal or near optimal solutions within the specified number of function evaluations.
-\textbf{C}: Mean scaled error versus the number of function evaluations for the 300-dimensional Rastrigin function.
-DOPS is the only algorithm that finds an optimal or near optimal solution within the specified number of function evaluations.
-In all cases, the maximum number of function evaluations was $\mathcal{N}$ = 4000.
-Mean and standard deviation were calculated over $\mathcal{T}$ = 25 trials.
-}\label{fig-testfunctions}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=1.00\textwidth]{./figs/Figure_3_Coagulation_v4.pdf}
-\caption{Schematic of the extrinsic and intrinsic coagulation cascade.
-Inactive zymogens upstream (grey) are activated by exposure to tissue factor (TF)  following vessel injury.
-Tissue factor and activated factor VIIa (FVIIa) form a complex that activates factor X (fX) and IX (fIX).
-FXa activates downstream factors including factor VIII (fVIII) and fIX.
-Factor V (fV) is primarily activated by thrombin (FIIa).
-In addition, we included a secondary fV activation route involving FXa.
-FXa and FVa form a complex (prothrombinase) on activated platelets that converts prothrombin (fII) to FIIa.
-FIXa and FVIIIa can also form a complex (tenase) on activated platelets which catalyzes FXa formation.
-Thrombin also activates upstream coagulation factors, forming a strong positive feedback ensuring rapid activation.
-Tissue factor pathway inhibitor (TFPI) downregulates FXa formation and activity by sequestering free FXa and TF-FVIIa in a FXa-dependent manner.
-Antithrombin III (ATIII)  inhibits all proteases.
-Thrombin inhibits itself binding the surface protein thrombomodulin (TM).
-The IIa-TM complex catalyzes the conversion of protein C (PC) to activated protein C (APC), which attenuates the coagulation response by the proteolytic cleavage of fV/FVa and fVIII/FVIIIa.
-}\label{fig-coagulation-network}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[h]
-\centering
-%\includegraphics[width=1.0\textwidth]{./figs/Figure_4_Errors_Convergence_v2.pdf}
-\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=1.0\textwidth]{./rachelfigs/RecreateFigure4.pdf}
-\caption{Error convergence rates of the five different algorithms on the coagulation model. The objective error is the mean over $\mathcal{T}$= 25 trials.
-DOPS and SA have the steepest drop in error during first 300 function evaluations. Thereafter the error drop in DDS and SA remains nearly constant whereas DOPS continues to drops further.
-At the end of 4000 function evaluations DOPS attains the lowest error. The next best estimate using DDS is nearly three times greater than the lowest error using DE.
-}\label{fig-convergence}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[h]
-\centering
-%\includegraphics[width=1.0\textwidth]{./figs/Figure_5_Sim_Train_E1_E5.pdf}
-\includegraphics[trim = {0cm 6cm 7cm 6cm }, clip,width=1.0\textwidth]{./rachelfigs/ReplicateFigure5DivBySqrtN.pdf}
-\caption{Model fits on experimental data using DOPS.
-The model parameters were estimated using DOPS.
-Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from 25 trials.
-The grey shaded region represents the 99\% confidence estimate of the mean simulated thrombin concentration.
-The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers.
-Thrombin generation is initiated by adding Factor TF/VIIa (5nM (blue) and 5pM (red)) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
-}\label{fig-train}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[h]
-\centering
-%\includegraphics[width=1.0\textwidth]{./figs/Figure_6_Sim_Validate_E2_E4_E6.pdf}
-\includegraphics[trim = {3cm 6cm 3cm 6cm }, clip,width=1.0\textwidth]{./rachelfigs/ReplicateFigure6DivBySqrtN.pdf}
-\caption{Model predictions on unseen experimental data using parameters obtained from DOPS.
-The parameter estimates that were obtained using DOPS were tested against data that was not used in the model training.
-Solid black lines indicate the simulated mean thrombin concentration using parameter vectors from $\mathcal{T}$ = 25 trials.
-The grey shaded region represents the 99\% confidence estimate of the mean simulated thrombin concentration.
-The experimental data is reproduced from the synthetic plasma assays of Mann and co-workers.
-Thrombin generation is initiated by adding Factor VIIa-TF (500pM - Blue, 50pM - Pink and 10pM - purple, respectively) to synthetic plasma containing 200 $\mu$mol/L of phospholipid vesicles (PCPS) and a mixture of coagulation factors (II,V,VII,VIII,IX,X and XI) at their mean plasma concentrations.
-}\label{fig-validation}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[h]
-\centering
-%\includegraphics[width=1.0\textwidth]{./figs/Figure_7_ErrorSwarms_v2.pdf}
-\includegraphics[trim = {3cm 6cm 3cm 6cm }, clip,width = 1.0\textwidth]{./rachelfigs/RecreateFigure7}
-\caption{Influence of the switching strategy and sub-swarms on DOPS performance for the coagulation model.
-DOPS begins by using a particle swarm search and then dynamically switches (switch region), using an adaptive switching criteria, to the DDS search phase.
-We compared the performance of DOPS with and without DDS for different sub-swarm searches to quantify the effect of number of sub-swarms and DDS.
-We used one, two, four, five and eight sub-swarms, with a total of 40 particles divided evenly amongst the swarms.
-The results presented are the average of $\mathcal{T}$ = 25 trials with $\mathcal{N}$ = 4000 function evaluations each.
-The convergence rates with higher swarm numbers is typically higher but there is no pronounced difference amongst four, five and eight.
-The multi-swarm with without DDS saturates while DOPS shows a rapid drop due to a switch to the DDS phase.
-}\label{fig-sub-swarms}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[h]
-\centering
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnb4_obj.pdf} }}
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnEggholder}} }%
-        \qquad
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnstybtang}} }%    
-        \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/2versionsofDOPSOnfitCoagShrinkFactor1E6}} }%
-
-\caption{Performance of DOPS and Experimental DOPS on the CHO metabolism problem (a), the Eggholder function (b), the 100 dimensional Styblinksi-Tang function (c) and the coagulation problem (d).
-Both methods have the same initial decrease in error, but as the number of function evaluations increases, experimental DOPS produces a larger decrease in error.
-The results presented are the average of $\mathcal{T}$ = 250 trials with for the CHO metabolism problem and $\mathcal{T}$ = 500 trials on the Eggholder and Styblinksi-Tang functions with $\mathcal{N}$ = 4000 function evaluations each, and $\mathcal{T}$ = 25 trials for the coagulation problem.
-}\label{fig-exp-DOPS-b4}
-\end{figure}
-
-%\bibitem[Saltelli \em{et~al.}(2010)Saltelli, Annoni, Azzini, Campolongo, Ratto,
-  %and Tarantola]{Saltelli:2010}
-%Saltelli, A.; Annoni, P.; Azzini, I.; Campolongo, F.; Ratto, M.; Tarantola, S.
-%\newblock Variance based sensitivity analysis of model output. Design and
-  %estimator for the total sensitivity index.
-%\newblock {\em Comput. Phys. Commun.} {\bf 2010}, {\em 181},~259--270.
-
-%\end{thebibliography}
-
-\clearpage
-
-% Supplemental figures -
-% Set the S-
-\renewcommand\thefigure{S\arabic{figure}}
-\renewcommand\thetable{T\arabic{table}}
-\renewcommand\thepage{S-\arabic{page}}
-\renewcommand\theequation{S\arabic{equation}}
-
-% Reset the counters -
-\setcounter{equation}{0}
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\setcounter{page}{1}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.00\textwidth]{./figs/Figure_S1_B1_sims.pdf}
-\caption{\textbf {(Data fits for Problem B1)} Pseudo-experimental data (red circles) vs.
-optimal solution obtained using DOPS (solid blue lines) for the 44 observed states.
-X axis: time [s]; Y axis: metabolite concentrations [mM].
-}\label{fig-sims-b1}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.00\textwidth]{./rachelfigs/RecreatedFigureS2}
-\caption{\textbf {(Data fits for Problem B4)} Pseudo-experimental data (red x) vs.
-optimal solution obtained using DOPS (solid blue lines) for the 13 observed states.
-X axis: time [s]; Y axis: metabolite concentrations [mM].
-}\label{fig-sims-b4}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.00\textwidth]{./figs/Figure_S3_B1_B4_params_measuredstates.pdf}
-\caption{ \textbf {(A)} Difference between nominal and optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
-\textbf {(B)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B1: Genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
-\textbf {(C)} Difference between nominal and optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
-\textbf {(D)} Difference between experimental (measured) data and data simulated with optimal parameters for problem B4: Metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
-}\label{fig-benchmark}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[trim = {0 5cm 0 5cm }, clip, width=1.00\textwidth]{./rachelfigs/TimeComparsion.pdf}
-\caption{ Comparison of the runtime of the different optimization methods used for comparison with  $\mathcal{T}$ = 25 trials per method.
-All methods used take about the same amount of time to perform 4,000 function evaluations on the coagulation problem, as this problem is very stiff, so the majority of the time is spent solving the system of differential equations.
-}\label{fig-time-comparsion}
-\end{figure}
-
-\begin{figure}%
-    \centering
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/RecreateRebutalReviewer1_FigS4a} }}%
-    \qquad
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/RecreateRebutalReviewer1_FigS4b} }}%
-    \caption{Mean onvergence curves for different metaheuristics for (a) Ackley 300 dimensional and (b) Rastrigin 300 dimensional with $\mathcal{T}$ = 25 trials per method.
-DOPS not only finds a better solution than any other technique, it finds it with fewer function evaluations}%
-    \label{fig-more-methods}%
-\end{figure}
-
-\begin{figure}%
-    \centering
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/CompareExperimentalDOPSToESSOnProblemB4} }}%
-    \qquad
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/CompareExperimentalDOPSToESSOnCoag} }}%
-    \caption{Mean convergence curves for DOPS and ESS for (a) CHO model (problem B4) (b) coagulation with $\mathcal{T}$ = 25 trials per method.}%
-    \label{fig-compare-DOPS-ESS}%
-\end{figure}
-
-\begin{figure}%
-    \centering
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/DOPSDispersionCurvesOnb4} }}%
-    \qquad
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/DOPSDispersionCurvesOnCoag} }}%
-    \caption{Dispersion curves for DOPS on (a) CHO model (problem B4) (b) coagulation with $\mathcal{T}$ = 25 trials per problem .}%
-    \label{fig-DOPS-Dispersion-Curves}%
-\end{figure}
-
-\begin{figure}%
-    \centering
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/FigureS7a} }}%
-    \qquad
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/FigureS7b} }}%
-    \qquad
-        \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/FigureS7c} }}%
-    \qquad
-    \subfloat[]{{\includegraphics[trim = {3cm 7cm 3cm 7cm }, clip,width=.45\textwidth]{./rachelfigs/FigureS7d} }}%
-    \caption{Variability  analysis in best objective value for $\mathcal{T}$ = 25 trials. (a) Ackley  300 dimensional (b) Rastrigin 300 dimensional (c) CHO model (problem b4) (d) coagulation. }%
-    \label{fig-DOPS-Variability}%
-\end{figure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                               %%
+%% Additional Files              %%
+%%                               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\section*{Additional Files}
+%  \subsection*{Additional file 1 --- Sample additional file title}
+%    Additional file descriptions text (including details of how to
+%    view the file, if it is in a non-standard format or the file extension).  This might
+%    refer to a multi-page table or a figure.
+%
+%  \subsection*{Additional file 2 --- Sample additional file title}
+%    Additional file descriptions text.
+
+\end{backmatter}
 \end{document}

### 2017-12-05 Paper_v7.tex
updating manuscript

@@
 \begin{figure}[h!]
+\includegraphics[width = 1.0\textwidth]{./rachelfigs/expDOPSvsDOPS-crop.pdf} 
+\caption{\csentence{Comparsion of DOPS and Experimental DOPS} Performance of DOPS and Experimental DOPS on the CHO metabolism problem (a), the Eggholder function (b), the 100 dimensional Styblinksi-Tang function (c) and the coagulation problem (d).
+Both methods have the same initial decrease in error, but as the number of function evaluations increases, experimental DOPS produces a larger decrease in error.
+The results presented are the average of $\mathcal{T}$ = 250 trials with for the CHO metabolism problem and $\mathcal{T}$ = 250 trials on the Eggholder and Styblinksi-Tang functions with $\mathcal{N}$ = 250 function evaluations each, and $\mathcal{T}$ = 25 trials for the coagulation problem.
+}\label{fig-exp-DOPS-b4}
+\end{figure}
+
+  \begin{figure}[h!]
   \caption{\csentence{Sample figure title.}
@@
 
-%\section*{Additional Files}
+\section*{Additional Files}
+\setcounter{equation}{0}
+\setcounter{table}{0}
+\setcounter{figure}{0}
 %  \subsection*{Additional file 1 --- Sample additional file title}
@@
 %    Additional file descriptions text.
+\subsection*{Additional file S\ref{fig-sims-b1}-(Data fits for CHO metabolism problem)}
+\begin{figure}[ht]
+\centering
+%\includegraphics[width=1.00\textwidth]{./figs/Figure_S1_B1_sims.pdf}
+\includegraphics[width=1.00\textwidth]{./rachelfigs/RecreatedFigureS1usingMS2_save.pdf}
+\caption{\textbf {(Data fits for CHO metabolism problem)} Pseudo-experimental data (red x) vs.
+optimal solution obtained using DOPS (solid blue lines) for the 44 observed states.
+X axis: time [s]; Y axis: metabolite concentrations [mM].
+}\label{fig-sims-b1}
+\end{figure}
+
+\subsection*{Additional file S\ref{fig-sims-b4}-(Data fits for S.cerevisiae metabolism problem)}
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth]{./rachelfigs/RecreatedFigureS2}
+\caption{\textbf {(Data fits for S.cerevisiae metabolism problem)} Pseudo-experimental data (red x) vs.
+optimal solution obtained using DOPS (solid blue lines) for the 13 observed states.
+X axis: time [s]; Y axis: metabolite concentrations [mM].
+}\label{fig-sims-b4}
+\end{figure}
+
+\subsection*{Additional file S3-(Comparison of states and parameters)}
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.00\textwidth]{./rachelfigs/fig_s3-crop.pdf}
+\caption{ \textbf {(A)} Difference between nominal and optimal parameters for genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
+\textbf {(B)} Difference between experimental (measured) data and data simulated with optimal parameters for genome wide kinetic model of \textit{S.cerevisiae} with 1759 unknown parameters.
+\textbf {(C)} Difference between nominal and optimal parameters for metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
+\textbf {(D)} Difference between experimental (measured) data and data simulated with optimal parameters for metabolic model of Chinese  Hamster Ovary Cells (CHO) cells with 117 parameters.
+}\label{fig-benchmark}
+\end{figure}
+
+\subsection*{Additional file S4-(Time Comparison)}
+\begin{figure}[ht]
+\centering
+\includegraphics[trim = {0 5cm 0 5cm }, clip, width=1.00\textwidth]{./rachelfigs/TimeComparsion.pdf}
+\caption{ Comparison of the runtime of the different optimization methods used for comparison with  $\mathcal{T}$ = 25 trials per method.
+All methods used take about the same amount of time to perform 4,000 function evaluations on the coagulation problem, as this problem is very stiff, so the majority of the time is spent solving the system of differential equations.
+}\label{fig-time-comparsion}
+\end{figure}
+
+\subsection*{Additional file S5-(Convergence Curves)}
+\begin{figure}
+\includegraphics[width=1.00\textwidth]{./rachelfigs/fig_s4-crop}
+    \caption{Mean convergence curves for different metaheuristics for (a) Ackley 300 dimensional and (b) Rastrigin 300 dimensional with $\mathcal{T}$ = 25 trials per method.
+DOPS not only finds a better solution than any other technique, it finds it with fewer function evaluations}%
+    \label{fig-more-methods}%
+\end{figure}
+
+\subsection*{Additional file S6-(Comparison of DOPS to ESS)}
+\begin{figure}
+\includegraphics[width=1.00\textwidth]{./rachelfigs/fig_s5-crop} %fix me
+  \caption{Mean convergence curves for DOPS and ESS for (a) CHO model  (b) coagulation with $\mathcal{T}$ = 25 trials per method.}%
+    \label{fig-compare-DOPS-ESS}%
+\end{figure}
 

### 2017-12-05 Paper_v7.tex
adding last figures

@@
 \begin{figure}
-\includegraphics[width=1.00\textwidth]{./rachelfigs/fig_s5-crop} %fix me
+\includegraphics[width=1.00\textwidth]{./rachelfigs/fig_s5-crop}
   \caption{Mean convergence curves for DOPS and ESS for (a) CHO model  (b) coagulation with $\mathcal{T}$ = 25 trials per method.}%
@@
 
+\subsection*{Additional file S7-(Dispersion Curves)}
+\begin{figure}
+\includegraphics[width=1.00\textwidth]{./rachelfigs/dispersion_curves-crop}
+    \caption{Dispersion curves for DOPS on (a) CHO model (b) coagulation with $\mathcal{T}$ = 25 trials per problem .}%
+    \label{fig-DOPS-Dispersion-Curves}%
+\end{figure}
+
+\subsection*{Additional file S8-(Comparison of functional values)}
+\begin{figure}
+\includegraphics[width=1.00\textwidth]{./rachelfigs/compare_functional_vals-crop}
+    \caption{Variability  analysis in best objective value for $\mathcal{T}$ = 25 trials. (a) Ackley  300 dimensional (b) Rastrigin 300 dimensional (c) CHO model (d) coagulation. }%
+    \label{fig-DOPS-Variability}%
+\end{figure}
+
 \end{backmatter}

### 2017-12-05 Paper_v7.tex
updating authors

@@
 \author[
-   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
-   corref={aff1},                       % id of corresponding address, if any
+   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}                       % id of corresponding address, if any
    noteref={n1},                        % id's of article notes, if any
@@
    addressref={aff1},
-   email={ril34@cornell.edu}
+   email={ril34@cornell.edu},
+   noteref={n1}
 ]{\fnm{Rachel} \snm{LeCover}}
@@
    addressref={aff1},
+   corref={aff1},
    email={jdv27@cornell.edu}
@@
 \begin{figure}
-\includegraphics[width=1.00\textwidth]{./rachelfigs/compare_functional_vals-crop}
+\includegraphics[width=1.0\textwidth]{./rachelfigs/compare_functional_vals-crop}
     \caption{Variability  analysis in best objective value for $\mathcal{T}$ = 25 trials. (a) Ackley  300 dimensional (b) Rastrigin 300 dimensional (c) CHO model (d) coagulation. }%

### 2017-12-14 DOPS/utilities/DOPS_PSO_only.m
more clean up

@@ -1,0 +1,209 @@
+%  ----------------------------------------------------------------------------------- 
+%  Copyright (c) 2016 Varnerlab
+%  School of Chemical and Biomolecular Engineering
+%  Cornell University, Ithaca NY 14853 USA
+%
+%----------------*****  Code Author Information *****----------------------
+%   Code Author (Implementation Questions, Bug Reports, etc.): 
+%       Adithya Sagar: asg242@cornell.edu
+% 
+%  Permission is hereby granted, free of charge, to any person obtaining a copy
+%  of this software and associated documentation files (the "Software"), to deal
+%  in the Software without restriction, including without limitation the rights
+%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+%  copies of the Software, and to permit persons to whom the Software is
+%  furnished to do so, subject to the following conditions:
+
+%  The above copyright notice and this permission notice shall be included
+%  in
+%  all copies or substantial portions of the Software.
+%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+%  THE SOFTWARE.
+%  ----------------------------------------------------------------------------------- 
+
+%% SWARM SEARCH
+
+function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=DOPS_PSO_only(optFunction,MAXJ,MINJ,NP,NI,NS,G,r)
+
+% Parameters for swarm search
+Max_Inertia_weight=0.9;
+Min_Inertia_weight=0.4;
+C1=1.5;C2=1.5;
+N=length(MINJ);
+SUB_SWARM_SIZE=round(NP/NS);
+failure_counter=0;
[…]

### 2017-12-14 DOPS/utilities/RunEssOnSelectedFunction.m

@@ -1,0 +1,17 @@
+function [ output_args ] = RunEssOnSelectedFunction( lowerIter,upperIter )
+%run ESS on Coag and problem b4
+    addpath(genpath('CoagulationFiles')) %add coagulation functions to path
+    addpath(genpath('eSS')); % add ess files to path
+    functions_to_test = {@b4_obj, @fitCoag};
+    bounds_files = { 'b4_bounds.mat', 'coagBounds.mat'};
+    num_repeats = 25;
+    num_f_eval = 4000;
+    for j = 2:size(functions_to_test,2)
+        load(bounds_files{j});
+       for k =lowerIter:upperIter
+           fprintf('On function %d of %d, on ESS repeat %d of %d', j,size(functions_to_test,2),k,num_repeats);
+           runEss(func2str(functions_to_test{j}), lb, ub, num_f_eval,k,1,k);
+       end
+    end
+end
+

### 2017-12-14 DOPS/utilities/compareAllTechniques.m

@@ -1,0 +1,75 @@
+function[]= compareAllTechniques()
+    %this function compares serveral techniques to DOPS on the coagulation
+    %problem to replication figure 4 from the paper
+    [allDOPS,adjustedAllDOPS,allExperimental,allDOPSTimes]=compareExperimentalDOPS();
+    [allSAResults,allSATimes] = processSAResults();
+    [allDDSResults,allDDSTimes] = processDDSResults();
+    [allDEResults,allDETimes] = processDEResults();
+    close('all');
+    f=figure();
+    hold('on')
+    plot(mean(adjustedAllDOPS,1), 'LineWidth', 2.0, 'Color', 'k');
+    plot(mean(allSAResults,1), 'LineWidth', 2.0, 'Color','g');
+    plot(mean(allDDSResults,1), 'LineWidth',2.0, 'Color', 'b');
+    plot(1:40:4000,mean(allDEResults,1), 'LineWidth', 2.0,'Color','r');
+    set(gca, 'YScale', 'log');
+    axis([0,4000,7E4,2E7])
+    legend('DOPS', 'Simulated Annealing', 'DDS', 'DE');
+    xlabel('Objective Function Evaluations')
+    ylabel({'Mean Objective Function Error', '(Number of Trials = 25)'});
+    saveas(f,'../DOPS_Results/figures/RecreateFigure4.pdf', 'pdf')
+
+    %to create DOPS dispersion curve
+    f=figure();
+    hold('on')
+    for j=1:25
+       semilogy(adjustedAllDOPS(j,:), 'k') 
+       alpha(.5); %make lines somewhat transparent
+    end
+    xlabel('Number of Function Evaluations')
+    ylabel('Functional Value');
+    set(gca, 'YScale', 'log');
+     saveas(f,'../DOPS_Results/figures/DOPSDispersionCurvesOnCoag.pdf', 'pdf')
+end
+
+function [allSAResults,allTimes] =processSAResults()
+    numTrials = 25;
+    numIters = 4000;
+    allSAResults = zeros(numTrials,numIters);
+    allTimes = zeros(numTrials,1);
+    for j = 1:numTrials
+       currdata =load(strcat('/home/rachel/Documents/DOPS/SA_Results/fitCoag/FuntionalValuesTrial',num2str(j),'.txt'));  
+       allSAResults(j,:) =currdata(1:4000);
[…]

### 2017-12-14 DOPS/utilities/compareDOPSVersionsOnSelectedFunction.m

@@ -1,0 +1,76 @@
+function [allDOPS,adjustedAllDOPS,allExperimental,allDOPSTimes]=compareDOPSVersionsOnSelectedFunction(functionName)
+    %This function plots the mean convergence curves for DOPS and
+    %experimental DOPS assuming that the results are stored at where the
+    %paths point.
+    if(contains(functionName, 'fitCoag'))
+        addpath('/home/rachel/Documents/DOPS/DOPS_Results/fitCoag/')
+        NUM_ITERS = 25;
+    else
+       NUM_ITERS = 250; 
+        addpath(strcat('/home/rachel/Documents/DOPS/DOPS_Results/',functionName));
+    %else
+     %   NUM_ITERS = 500;
+     %   addpath(strcat('/home/rachel/Documents/DOPS/DOPS_Results/',functionName));
+    end
+   close('all');
+    
+    
+    NUM_EVALS = 4000;
+    NUM_PARTCILES = 40;
+    allDOPS = zeros(NUM_ITERS,NUM_EVALS);
+    adjustedAllDOPS = zeros(NUM_ITERS,NUM_EVALS);
+    allExperimental = zeros(NUM_ITERS,NUM_EVALS);
+    allDOPSTimes = zeros(NUM_ITERS,1);
+    %load DOPS data
+    for j=1:NUM_ITERS
+        load(strcat('DOPS_error_iter',num2str(j),'.mat'));
+        PSO_error = g_best_solution{j};
+        corrected_PSO_error = fillInPSO(NUM_PARTCILES,PSO_error);
+        load(strcat('DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter', num2str(j),'.mat'));
+        DDS_error = bestval_dds_swarm{j};
+        alldata= cat(2,PSO_error,DDS_error);
+        correctedAllData = cat(2,corrected_PSO_error,DDS_error);
+        correctedAllData = correctedAllData(1:NUM_EVALS);
+        %allDOPS(j,:) =alldata;
+        adjustedAllDOPS(j,:) = correctedAllData;
+        
+        currTime = load(strcat('DOPS_time_iter', num2str(j), '.txt'));
+        currTime = currTime(j);
+        allDOPSTimes(j) = currTime;
+    end
+   
+    %load experimental DOPS data
+    for k=1:NUM_ITERS
[…]

### 2017-12-14 DOPS/utilities/compareESSToDOPS.m

@@ -1,0 +1,137 @@
+function [] = compareESSToDOPS()
+    %this function plots the performance of ESS and DOPS on the coagulation
+    %problem and problem b4
+    close('all');
+    NUM_ITERS = 25;
+    NUM_EVALS = 4000;
+    NUM_PARTCILES = 40;
+    allDOPS = zeros(NUM_ITERS,NUM_EVALS);
+    adjustedAllDOPS = zeros(NUM_ITERS,NUM_EVALS);
+    allESS = zeros(NUM_ITERS,NUM_EVALS+1);
+    addpath('/home/rachel/Documents/DOPS/DOPS_Results/fitCoag');
+    for j=1:NUM_ITERS
+        load(strcat('DOPS_error_iter',num2str(j),'.mat'));
+        PSO_error = g_best_solution{j};
+        corrected_PSO_error = fillInPSO(NUM_PARTCILES,PSO_error);
+        load(strcat('DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter', num2str(j),'.mat'));
+        DDS_error = bestval_dds_swarm{j};
+        alldata= cat(2,PSO_error,DDS_error);
+        correctedAllData = cat(2,corrected_PSO_error,DDS_error);
+        correctedAllData = correctedAllData(1:NUM_EVALS);
+        adjustedAllDOPS(j,:) = correctedAllData;
+    end
+   
+    for k=1:NUM_ITERS
+        load(strcat('/home/rachel/Documents/DOPS/ESS_Results/fitCoag/functionalInterpIter', num2str(k),'.mat'));
+        exp_err= f_interp(1:NUM_EVALS+1);
+        allESS(k,:) = exp_err;
+    end
+    f=figure();
+    hold('on')
+    plot(1:NUM_EVALS,nanmean(adjustedAllDOPS,1), 'LineWidth', 2, 'Color', 'r')
+    plot(0:NUM_EVALS,nanmean(allESS,1), 'LineWidth', 2, 'Color','k')
+    set(gca, 'YScale', 'log')
+    set(gca, 'fontsize', 20);
+    xlabel('Number of Function Evaluations', 'FontSize', 32)
+    ylabel('Functional Value', 'FontSize',32)
+    legend('DOPS', 'ESS')
+     f.PaperPositionMode = 'auto';
+    f.PaperUnits = 'inches';
+    f.PaperPosition = [0 0 12 12];
[…]

### 2017-12-14 DOPS/utilities/compareExperimentalDOPS.m

@@ -1,0 +1,103 @@
+function [allDOPS,adjustedAllDOPS,allExperimental,allDOPSTimes]=compareExperimentalDOPS()
+    addpath('/home/rachel/Documents/DOPS/DOPS_Results/fitCoag');
+    close('all');
+    NUM_ITERS = 25;
+    NUM_EVALS = 4000;
+    NUM_PARTCILES = 40;
+    allDOPS = zeros(NUM_ITERS,NUM_EVALS);
+    adjustedAllDOPS = zeros(NUM_ITERS,NUM_EVALS);
+    allExperimental = zeros(NUM_ITERS,NUM_EVALS);
+    allExperimentalbeta = zeros(NUM_ITERS,NUM_EVALS);
+    allExperimentalgamma = zeros(NUM_ITERS,NUM_EVALS);
+    allDOPSTimes = zeros(NUM_ITERS,1);
+    for j=1:NUM_ITERS
+       % if(j ==5) %need to skip 5 since rerunning at the moment
+       %    adjustedAllDOPS(j,:) = NaN;
+       %    continue;
+       % end
+        load(strcat('DOPS_error_iter',num2str(j),'.mat'));
+        PSO_error = g_best_solution{j};
+        corrected_PSO_error = fillInPSO(NUM_PARTCILES,PSO_error);
+        load(strcat('DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter', num2str(j),'.mat'));
+        DDS_error = bestval_dds_swarm{j};
+        alldata= cat(2,PSO_error,DDS_error);
+        correctedAllData = cat(2,corrected_PSO_error,DDS_error);
+        correctedAllData = correctedAllData(1:NUM_EVALS);
+        %allDOPS(j,:) =alldata;
+        adjustedAllDOPS(j,:) = correctedAllData;
+        
+        currTime = load(strcat('DOPS_time_iter', num2str(j), '.txt'));
+        currTime = currTime(j);
+        allDOPSTimes(j) = currTime;
+    end
+   
+    for k=1:NUM_ITERS
+       if(k ==1||k==2|| k ==20)
+           allExperimental(k,:) = NaN;
+           continue;
+        end
+        
+        load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/ExperimentalV2/fitCoag/DOPS_error_iter', num2str(k),'.mat'));
+        exp_err= g_best_solution{k};
[…]

### 2017-12-14 DOPS/utilities/compareSwarmNumbers.m

@@ -1,0 +1,60 @@
+function []=compareSwarmNumbers()
+    %this function plots the results of changing the number of particles
+    %in a swarm during PSO
+    [allDOPS,adjustedAllDOPS,allExperimental]=compareExperimentalDOPS();
+    swarmNum = [1,2,4,5,8];
+    close('all');
+    f = figure();
+    hold('on')
+    numEvals = 4000;
+    numFEvals = 1:1:numEvals;
+    %plot DOPS results
+    set(gca, 'YScale', 'log')
+    semilogy(numFEvals, mean(adjustedAllDOPS,1), 'LineWidth', 2.0,'Color', 'k');
+    numRepeats = 25;
+    NUM_PARTCILES = 40;
+    colors = {repmat(75/255,1,3), repmat(125/255,1,3), repmat(175/255,1,3),[0,0,0], repmat(225/255,1,3)};
+    
+    for k = 1:size(swarmNum,2)
+       adjustedAllDOPS = [];
+       currSwarmNum = swarmNum(k);
+       for j = 1:numRepeats
+           
+           load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/PSO_Only/fitCoag/',num2str(currSwarmNum), 'Swarms/DOPS_error_iter',num2str(j),'.mat'));
+           PSO_error = g_best_solution{j};
+           corrected_PSO_error = fillInPSO(NUM_PARTCILES,PSO_error);
+           %load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/PSO_Only/fitCoag/',num2str(currSwarmNum), 'Swarms/DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter', num2str(j),'.mat'));
+           %DDS_error = bestval_dds_swarm{j};
+           %alldata= cat(2,PSO_error,DDS_error);
+           correctedAllData = corrected_PSO_error;
+           correctedAllData = correctedAllData(1:numEvals);
+           adjustedAllDOPS(j,:) = correctedAllData;
+       end
+       if (currSwarmNum ==5)
+           mDOPS=mean(adjustedAllDOPS,1);
+         semilogy(numFEvals(1:40:end), mDOPS(1:40:end), 'LineWidth', 1, 'Color', colors{k}, 'Marker', 'o','MarkerFaceColor', 'w');  
+       else
[…]

### 2017-12-14 DOPS/utilities/compareVariationInFunctionalVals.m

@@ -1,0 +1,76 @@
+function []=compareVariationInFunctionalVals(functionName)
+    %for the supplimental figures-this function compares the variation in
+    %functional values between the different trials of the various methods
+    numRepeats = 25;
+    close('all');
+    CMAES_f = zeros(1,numRepeats);
+    DDS_f = zeros(1,numRepeats);
+    DOPS_f = zeros(1,numRepeats);
+    ESS_f = zeros(1,numRepeats);
+    SA_f = zeros(1,numRepeats);
+    DE_f = zeros(1, numRepeats);
+    
+    %read in functional values or regenerate if cheap to evaluate
+    if(strcmp('ackley',functionName) || strcmp('rast',functionName))
+       [allCMAES,allSA,allDE,allDDS,allDOPS,allESS]= replicateFigureS4(functionName,numRepeats);
+       for j = 1:numRepeats
+          CMAES_f(j) = allCMAES(end,j);
+          SA_f(j) = allSA(end,j);
+          DE_f(j)=allDE(end,j);
+          DDS_f(j) = allDDS(end,j);
+          DOPS_f(j) = allDOPS(end,j);
+          ESS_f(j) = allESS(end,j);
+       end
+    end
+    
+    if(strcmp('b4_obj', functionName))
+        for j =1:numRepeats
+            load(strcat('../ESS_Results/b4_obj/functionalInterpIter', num2str(j), '.mat'));
+            ESS_f(j) = f_interp(end);
+            load(strcat('../DOPS_Results/b4_obj/DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter' ,num2str(j), '.mat'));
+            DDS_error = bestval_dds_swarm{j};
+            DOPS_f(j) = DDS_error(end);
+        end
+    end
+    
+   if(strcmp('fitCoag', functionName))
+        for j =1:numRepeats
+          %  load(strcat('../ESS_Results/fitCoag/functionalInterpIter', num2str(j), '.mat'));
+          %  ESS_f(j) = f_interp(end);
+            try
+                load(strcat('../DOPS_Results/fitCoag/DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter' ,num2str(j), '.mat'));
+                DDS_error = bestval_dds_swarm{j};
+            catch    
+                DDS_error = 0;
+            end
[…]

### 2017-12-14 DOPS/utilities/makeFig2Figs.jl

@@ -1,0 +1,58 @@
+using DataFrames
+using PyPlot
+using PyCall
+@pyimport matplotlib.patches as mpatches
+
+function makeFig2Figs()
+	close("all")
+	df = readtable("../DOPS_Results/performanceComparsion.csv")
+	colors = ["#000000","#708090" , "#696969", "#A9A9A9", "#DCDCDC"]
+	optmethod = ["SA","PSO", "DE", "DOPS", "DDS"]
+	legend_patches = []
+	cmap=ColorMap("gray")
+	df=hcat(df, ["" for r in 1:size(df,1)]) #adding column for color
+	rename!(df, [:x1], [:barcolor])
+	#fill in the colors
+	for k in collect(1:size(df,1))
+		if(df[:opt_method][k]=="SA")
+			df[:barcolor][k]=colors[1] #cmap(1)
+		elseif(df[:opt_method][k]=="PSO")
+			df[:barcolor][k]=colors[2] #cmap(1)
+		elseif(df[:opt_method][k]=="DE")
+			df[:barcolor][k]=colors[3] #cmap(1)
+		elseif(df[:opt_method][k]=="Z_DOPS") #to force at the end when sorting
+			df[:barcolor][k]=colors[4]
+		elseif(df[:opt_method][k]=="DDS")
+			df[:barcolor][k]=colors[5]
+		end
+	end
+
+	for k in collect(1:size(colors,1))
+		push!(legend_patches,mpatches.Patch(color=colors[k], label=optmethod[k]))
+	end
+
+	
+	@show df
+	@show size(legend_patches)
+	alllimits = [[.5,16,0,1.3],[.5,16,0,1.3], [.5,16,0,1.3]]
+	functions = ["ackley_10d", "rast_10d", "rast_300d"]
+	xlabels = ["40 Evaluations", "2000 Evaluations", "4000 Evaluations"]
+	count = 1
+	for f in functions
+		reldata = df[df[:_function].==f,:]
+		reldata = sort!(reldata, cols = [order(:num_evals), order(:opt_method)])
+		figure(figsize=[15,7.5])
+		bar(collect(1:size(reldata,1)), reldata[:avg_scaled_err], yerr=reldata[:std_scaled_err], color=reldata[:barcolor],ecolor="k")
+		#title(f, fontsize=20)
+		ylabel("Scaled Error (mean+std)", fontsize=30)
+		ax=gca()
+		axis(alllimits[count])
[…]

### 2017-12-14 DOPS/utilities/makeFig6Plot.m

@@ -1,0 +1,42 @@
+function makeFig6Plot(allcurve1, allcurve2,allcurve3,DFIN)
+    %plot coagulation curves for figure 6
+    close('all');
+    f = figure();
+    plot(mean(allcurve1,1), 'k');
+    hold('on')
+    plot(mean(allcurve2,1), 'k');
+    plot(mean(allcurve3,1), 'k');
+    numRepeats = 25;
+    top1 = mean(allcurve1,1)+ 2.58*std(allcurve1,1)/sqrt(numRepeats);
+    bottom1 = mean(allcurve1,1)- 2.58*std(allcurve1,1)/sqrt(numRepeats);
+    x = 1:1:250;
+    idx = find(bottom1<0);
+    bottom1(idx) = 0; %remove things less than 0
+    fill([x,fliplr(x)],[top1, fliplr(bottom1)], [.25, .25, .25],'facealpha', .5)
+    top2 = mean(allcurve2,1)+ 2.58*std(allcurve2,1)/sqrt(numRepeats);
+    bottom2 = mean(allcurve2,1)- 2.58*std(allcurve2,1)/sqrt(numRepeats);
+    x = 1:1:250;
+    idx = find(bottom2<0);
+    bottom2(idx) = 0; %remove things less than 0
+    fill([x,fliplr(x)],[top2, fliplr(bottom2)], [.25, .25, .25], 'facealpha', .5)
+    top3 = mean(allcurve3,1)+ 2.58*std(allcurve3,1)/sqrt(numRepeats);
+    bottom3 = mean(allcurve3,1)- 2.58*std(allcurve3,1)/sqrt(numRepeats);
+    x = 1:1:250;
+    idx = find(bottom3<0);
+    bottom3(idx) = 0; %remove things less than 0
+    fill([x,fliplr(x)],[top3, fliplr(bottom3)], [.25, .25, .25],'facealpha', .5)
+    DATA1 = DFIN.EXPT_DATA_FIG2E2;
+    DATA2 = DFIN.EXPT_DATA_FIG2E3;
+    DATA3 = DFIN.EXPT_DATA_FIG2E4;
+    plot(DATA1(:,1), DATA1(:,2), 'b.', 'MarkerSize', 12);
+    plot(DATA2(:,1), DATA2(:,2), 'm.', 'MarkerSize',12);
+    plot(DATA3(:,1), DATA3(:,2), '.', 'MarkerSize',12, 'MarkerFaceColor', [238/256,118/256,0]);
[…]

### 2017-12-14 DOPS/utilities/recreateFigureS1.m

@@ -1,0 +1,93 @@
+function [x,bestparams,initialvals,functionalvals] = recreateFigureS1()
+    %also makes figures S3a and S3b
+    close('all')
+    NUM_REPEATS = 76;
+    functionalvals = [];
+    calculatedfunctionalvals = [];
+    initialvals = [];
+    num_states = 276;
+    for j =1:NUM_REPEATS
+        if(j==20 || j ==26 ||j==28 || j ==29|| j==33 || j==35||j==39||j==41||j==42||j==44|| j==45||j==46||j==47||j==48||j==52||j==53||j==65||j==67||j==70||j==73||j==74||j==75)
+           params(j,:) = NaN; 
+           functionalvals(j)=10^20;
+           continue;
+        end
+        load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/b1_obj/DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter',num2str(j),'.mat'))
+        functionalvals(j) = bestval_dds_swarm{j}(end);
+        load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/b1_obj/DOPS_error_iter',num2str(j),'.mat'))
+        initialvals(j) = g_best_solution{j}(1);
+        load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/b1_obj/DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_particledds_iter',num2str(j),'.mat'))
+        params(j,:) =best_particle_dds_swarm{j}(:,end);
+    end
+    %[smallestf, minIdx] = nanmin(functionalvals); %ignore NAN
+    minIdx = 2; %this is just hard coding which parameter set performs the best
+    bestparams = params(minIdx,:);
+%      for j = 1:NUM_REPEATS
+%          fprintf('Calculating functional value for param set %d \n', j);
+%         if(isnan(params(j,:)))
+%            calculatedfunctionalvals(j) = 1E20;
+%            continue;
+%         else
+%              [objective,constraints,residuals,x_sel,x,t,scaled_data] = b1_obj_forS1(params(j,:),j);
+%              calculatedfunctionalvals(j) = objective;
+%         end
+%         fprintf('Calculated functional value for param set %d is %f\n', j, objective);
+%      end
+     
[…]

### 2017-12-14 DOPS/utilities/repeatDDS.m

@@ -1,0 +1,15 @@
+function [ mean_perf,std_perf ] = repeatDDS( input_args )
+    %Repeat DDS for figure2 
+    num_repeats = 25;
+    allfvals = [];
+    allinitial_vals = [];
+    for j=1:num_repeats
+        fprintf('On repeat %d\n', j);
+        [fvals, inital_vals]=runDDSforTable2(j);
+        allfvals(j,:,:) = fvals;
+        allinitial_vals(j,:,:) = inital_vals;
+    end
+    mean_perf = mean(allfvals./allinitial_vals,1);
+    std_perf = std(allfvals./allinitial_vals,1);
+end
+

### 2017-12-14 DOPS/utilities/repeatDE.m

@@ -1,0 +1,15 @@
+function [ mean_perf,std_perf ] = repeatDE( input_args )
+    %Repeat DE for figure2 
+    num_repeats = 25;
+    allfvals = [];
+    allinitial_vals = [];
+    for j=1:num_repeats
+        fprintf('On repeat %d\n', j);
+        [fvals, inital_vals]=runDE(j);
+        allfvals(j,:,:) = fvals;
+        allinitial_vals(j,:,:) = inital_vals;
+    end
+    mean_perf = mean(allfvals./allinitial_vals,1);
+    std_perf = std(allfvals./allinitial_vals,1);
+end
+

### 2017-12-14 DOPS/utilities/repeatPSO.m

@@ -1,0 +1,15 @@
+function [ mean_perf,std_perf ] = repeatPSO( input_args )
+    %Repeat PSO for figure2 
+    num_repeats = 25;
+    allfvals = [];
+    allinitial_vals = [];
+    for j=1:num_repeats
+        fprintf('On repeat %d\n', j);
+        [best_params, fvals, outputs, inital_vals]=runPSO(j);
+        allfvals(j,:,:) = fvals;
+        allinitial_vals(j,:,:) = inital_vals;
+    end
+    mean_perf = mean(allfvals./allinitial_vals,1);
+    std_perf = std(allfvals./allinitial_vals,1);
+end
+

### 2017-12-14 DOPS/utilities/repeatSimulatedAnealing.m

@@ -1,0 +1,15 @@
+function [ mean_perf,std_perf ] = repeatSimulatedAnealing( input_args )
+    %Repeat Simulated Annealing for figure2 
+    num_repeats = 25;
+    allfvals = [];
+    allinitial_vals = [];
+    for j=1:num_repeats
+        fprintf('On repeat %d\n', j);
+        [best_params, fvals, outputs, inital_vals]=runSimulatedAnnealing(j);
+        allfvals(j,:,:) = fvals;
+        allinitial_vals(j,:,:) = inital_vals;
+    end
+    mean_perf = mean(allfvals./allinitial_vals,1);
+    std_perf = std(allfvals./allinitial_vals,1);
+end
+

### 2017-12-14 DOPS/utilities/replicateFig5.m

@@ -1,0 +1,87 @@
+function [all_curve1,all_curve2] = replicateFig5() %for DE, need 2 inputs
+    if(nargin<2)
+       y = []; 
+    end
+	DF=DataFile(0,0,0,[]); %Loads the coagulation data structure
+	DFIN=DF;
+    num_iters = 25;
+    all_params = readInParams(num_iters);
+    all_curve1 = [];
+    all_curve2 = [];
+    thrombinIdx = 48;
+    for j = 1:num_iters
+        fprintf('On parameter set %d of %d\n', j, num_iters);
+        DFIN.RATE_CONSTANT_VECTOR=all_params(j,:); %Assigns the rate constant vector as the current solution
+        [t1,curve1]=SimFunctionFig2E1(DFIN);
+        thrombin1 = curve1(:,thrombinIdx);
+        [t2,curve2]=SimFunctionFig2E5(DFIN); %Calculates the fitness values by passing the rate constant vector to the objectives E1 and E5
+        thrombin2 = curve2(:,thrombinIdx);
+        all_curve1(j,:) = thrombin1;
+        all_curve2(j,:) = thrombin2;
+    end
+   makePlot(all_curve1,all_curve2,DFIN);
+end
+function[all_params]= readInParams(num_iters)
+    %num_iters = 25;
+    addpath('/home/rachel/Documents/DOPS/DOPS_Results/fitCoag');
+    num_params = 148;
+    all_params = zeros(num_iters, num_params);
+    numParticles =40;
+    for j = 1:num_iters
+        load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/fitCoag/DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_particledds_iter', num2str(j),'.mat'));
+        load(strcat('DOPS_error_iter',num2str(j),'.mat'));
+        PSO_error = g_best_solution{j};
+        numFEvalsInPSO = size(PSO_error,2)*numParticles;
+        adjIdx = 4000-numFEvalsInPSO;
+        best_params = best_particle_dds_swarm{j}(:,adjIdx); %correct for interations in PSO
+        all_params(j,:) = best_params;
+    end
+end
+
[…]

### 2017-12-14 DOPS/utilities/replicateFig6.m

@@ -1,0 +1,95 @@
+function [all_curve1,all_curve2,all_curve3,DFIN] = replicateFig6() %for DE, need 2 inputs
+    if(nargin<2)
+       y = []; 
+    end
+	DF=DataFile(0,0,0,[]); %Loads the coagulation data structure
+	DFIN=DF;
+    num_iters = 25;
+    all_params = readInParams(num_iters);
+    all_curve1 = [];
+    all_curve2 = [];
+    all_curve3 = [];
+    thrombinIdx = 48;
+    for j = 1:num_iters
+        fprintf('On parameter set %d of %d\n', j, num_iters);
+        DFIN.RATE_CONSTANT_VECTOR=all_params(j,:); %Assigns the rate constant vector as the current solution
+        [t1,curve1]=SimFunctionFig2E2(DFIN);
+        thrombin1 = curve1(:,thrombinIdx);
+        [t2,curve2]=SimFunctionFig2E3(DFIN); %Calculates the fitness values by passing the rate constant vector to the objectives E1 and E5
+        thrombin2 = curve2(:,thrombinIdx);
+        [t3,curve3] = SimFunctionFig2E4(DFIN);
+        thrombin3 = curve3(:,thrombinIdx);
+        dyn1 = checkForDynamics(thrombin1);
+        dyn2 = checkForDynamics(thrombin2);
+        dyn3 = checkForDynamics(thrombin3);
+        allDyn = [dyn1, dyn2, dyn3];
+        for m = 1:max(size(allDyn))
+           if(allDyn(m)==0)
+              fprintf('On param set %d, curve %d did not have dynamics', j, m); 
+           end
+               
+        end
+        
+        all_curve1(j,:) = thrombin1;
+        all_curve2(j,:) = thrombin2;
+        all_curve3(j,:) = thrombin3;
+    end
+   makeFig6Plot(all_curve1,all_curve2,all_curve3,DFIN);
+    DATA1 = DFIN.EXPT_DATA_FIG2E2;
+    DATA2 = DFIN.EXPT_DATA_FIG2E3;
+    DATA3 = DFIN.EXPT_DATA_FIG2E4;
+    x = [1:1:250];
+    NSE10pm = calculateNSE(DATA1,mean(all_curve1,1), x);
+    NSE50pm = calculateNSE(DATA2,mean(all_curve2,1),x);
+    NSE500pm = calculateNSE(DATA3,mean(all_curve3,1),x);
+    
[…]

### 2017-12-14 DOPS/utilities/replicateFigureS4.m

@@ -1,0 +1,73 @@
+function[allCMAES,allSA,allDE,allDDS,allDOPS,allEss]= replicateFigureS4(Infunction, numRepeats)
+    %compares optimization methods for figure S4
+    numDims = 300;
+    functions = {Infunction};
+    ack_bounds = [-15,30];
+    rast_bounds = [-5.12,5.12];
+    numFEvals = 4000;
+    seed = 14850;
+    
+    for j=1:size(functions,2)
+        close('all');
+       if(j ==1)
+           lb = repmat(ack_bounds(1),numDims,1);
+           ub = repmat(ack_bounds(2), numDims,1);
+       elseif(j ==2)
+               lb = repmat(rast_bounds(1),numDims,1);
+               ub = repmat(rast_bounds(2), numDims,1);
+       end
+       f = figure();
+       hold('on');
+       allCMAES = [];
+       allSA = [];
+       allDE = [];
+       allDDS = [];
+       allDOPS = [];
+       allEss = [];
+       for k=1:numRepeats
+           seed = k;
+           cmasesavestr = strcat('../../DOPS/CMAES_Results/', functions{j}, '/Iter', num2str(k));
+           [evalCounter,functionalVal]= runCmaes(cmasesavestr,lb,ub,functions{j});
+           [best_params, fvalsSA, outputs, inital_vals]=runBasicSimulatedAnnealing(numFEvals,functions{j},lb,ub,seed);
+           [best_params, fvalsDE, outputs, inital_vals]=runBasicAdisDE(numFEvals,functions{j},lb,ub,seed);
+           [fvalsDDS]=runDDSBasic(numFEvals,functions{j},lb,ub,seed);
+           [g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls,correctedAllData]=run_DOPSBasic(numFEvals,functions{j},lb,ub,seed);
+           [res,f_iterp] = runEss(functions{j},lb,ub, numFEvals,seed,0,k);
+       
+        %iterpolate for CMAES
+        functionalVal_i = interp1(evalCounter,functionalVal,1:1:numFEvals);
+        allCMAES(:,k) = functionalVal_i;
+        allSA(:,k) = fvalsSA;
+        allDE(:,k) = fvalsDE;
+        allDDS(:,k) = fvalsDDS;
[…]

### 2017-12-14 DOPS/utilities/runAdisDE.m

@@ -1,0 +1,46 @@
+function[best_params, fvals, outputs, inital_vals]=runAdisDE(seed)
+    %runs DE for comparsion with DOPS
+    function [stop,options,optchanged] = myoutput(options,optimvalues,flag)
+        stop = false;
+        if isequal(flag,'iter')
+              history = [history; optimvalues.fval];
+        end
+        optchanged = false;
+    end
+    
+
+    numParticles = 40;
+    ctl.constr= 0;
+    ctl.const = [];
+    ctl.NP    = numParticles;         % NP will be set later
+    ctl.F     = 0.8;
+    ctl.CR    = 0.9;
+    ctl.strategy = 1;
+    ctl.refresh  = 0;
+    ctl.VTR   = -Inf;
+    ctl.tol   = 0;
+    ctl.maxnfe  = 1e6;
+    poss_iter_nums =[40,2000,4000];
+    functions_to_test = {@ackley, @rast, @rast};
+    bounds = {[30;-15], [5.12;-5.12], [5.12;-5.12]};
+    dimensions =[10,10,300];
+    best_params = {};
+    fvals = zeros(size(dimensions,2),size(functions_to_test,2));
+    inital_vals = zeros(size(dimensions,2), size(functions_to_test,2));
+    outputs = {};
+    for k =1:size(poss_iter_nums,2)
+        NUM_ITERATIONS=poss_iter_nums(k);
+         options = optimoptions(@simulannealbnd,'MaxIterations',NUM_ITERATIONS, 'OutputFcn', @myoutput);
+        for j=1:size(functions_to_test,2)
+            %create bounds
+            rng(seed)
+            currlimits = bounds{j};
+            MAXJ = repmat(currlimits(1),dimensions(j),1);
+            MINJ = repmat(currlimits(2), dimensions(j),1);
+            x0 = MAXJ-MINJ.*rand(size(MAXJ,1),numParticles)+MINJ; %create an initial guess within bounds
+            inital_vals(j,k) = min(feval(functions_to_test{j},x0));
+            [bestparticle, g_best_solution,nfeval] = de(functions_to_test{j},x0,MINJ,MAXJ,poss_iter_nums(k)/numParticles);
+            fvals(j,k) = g_best_solution(end);
+        end
+    end
+end

### 2017-12-14 DOPS/utilities/runDDSforTable2.m

@@ -1,0 +1,20 @@
+function[fvals, initial_vals]=runDDSforTable2(seed)
+    functions_to_test = {@ackley, @rast, @rast};
+    bounds_files = {'ackley10d_bounds.mat','rast10d_bounds.mat', 'rast300d_bounds.mat'};
+    poss_num_evals = [40,2000,4000];
+    r=0.2; 
+    initial_vals = zeros(size(functions_to_test,2), size(poss_num_evals,2));
+    fvals=zeros(size(functions_to_test,2), size(poss_num_evals,2));
+    for j=1:size(poss_num_evals,2)
+        for k=1:size(functions_to_test,2)
+            rng(seed)
+            load(bounds_files{k});
+            MAXJ = ub;
+            MINJ  = lb;
+            x0 = MINJ+rand()*(MAXJ-MINJ);
+            initial_vals(j,k) = feval(functions_to_test{k}, x0);
+            [best,x_best,dds_swarm_flag]=DOPS_DDS(functions_to_test{k},x0,MAXJ,MINJ,r,poss_num_evals(j));
+            fvals(j,k) = min(best);
+        end
+    end
+end

### 2017-12-14 DOPS/utilities/runDOPSforTable2.m

@@ -1,0 +1,26 @@
+function[mean_data, std_data]=runDOPSforTable2()
+    %calculates statistics about DOPS performance
+    functions_to_test = {@ackley, @rast, @rast, @rosenbrock, @BukinF6, @Beale};
+    bounds_files = {'ackley10d_bounds.mat','rast10d_bounds.mat', 'rast300d_bounds.mat', 'rosenbrock10d_bounds.mat', 'BukinF6Bounds.mat', 'BealeBounds.mat'};
+    NUM_TRIALS=25;
+    fvals = zeros(size(functions_to_test,2)-3,NUM_TRIALS);
+    initial_vals = zeros(size(functions_to_test,2)-3,NUM_TRIALS);
+    for j=1:size(functions_to_test,2)-3
+        load(bounds_files{j});
+         MAXJ = ub;
+         MINJ = lb;
+         [g_best_solution1,bestparticle1,particle1,fitness1,bestval_dds_swarm1,best_particle_dds_swarm1,best_particles_ls1]=run_DOPS(functions_to_test{j},MAXJ,MINJ,NUM_TRIALS);
+        [g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=run_DOPS(functions_to_test{j},MAXJ,MINJ,NUM_TRIALS);
+         %[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=run_DOPS_onlyPSO(functions_to_test{j},MAXJ,MINJ);
+        for k=1:NUM_TRIALS
+            if(size(bestval_dds_swarm{k},1) ==0)
+                fvals(j,k) = g_best_solution{k}(1); %min(fitness{k}(:,end));
+            else
+                fvals(j,k) = bestval_dds_swarm{k}(end);
+            end
+            initial_vals(j,k)=min(fitness1{k}(:,1));
+        end
+        mean_data(j) = mean(fvals(j,:)./initial_vals(j,:));
+        std_data(j) = std(fvals(j,:)./initial_vals(j,:));
+    end
+end

### 2017-12-14 DOPS/utilities/runForTable1DOPS_PSO_Only.m

@@ -1,0 +1,14 @@
+function[]=runForTable1DOPS_PSO_Only(lowerIter,upperIter,numSwarms)
+    addpath(genpath('CoagulationFiles')) %add coagulation functions to path
+   % functions_to_test = {@rast, @ackley,@b4_obj, @b1_obj_v2, @fitCoag};
+    functions_to_test = {@rast, @ackley,@fitCoag};
+   % bounds_files = {'rast_bounds.mat', 'ackley_bounds.mat', 'b4_bounds.mat','b1_bounds.mat', 'coagBounds.mat'};
+    bounds_files = {'rast_bounds.mat', 'ackley_bounds.mat', 'coagBounds.mat'}; 
+   for j=3:size(functions_to_test,2)
+        load(bounds_files{j});
+        %deal with transposed files
+            MAXJ = ub;
+            MINJ = lb;
+        run_DOPS_PSO_only(functions_to_test{j},MAXJ,MINJ,lowerIter,upperIter,numSwarms);
+    end
+end

### 2017-12-14 DOPS/utilities/runPSO.m

@@ -1,0 +1,32 @@
+function[best_params, fvals, outputs, inital_vals]=runPSO(seed)
+    poss_iter_nums =[40,2000,4000];
+    functions_to_test = {@ackley, @rast, @rast};
+    bounds = {[30;-15], [5.12;-5.12], [5.12;-5.12]};
+    dimensions =[10,10,300];
+    best_params = {};
+    fvals = zeros(size(dimensions,2),size(functions_to_test,2));
+    inital_vals = zeros(size(dimensions,2), size(functions_to_test,2));
+    outputs = {};
+    NUM_PARTCLES = 40;
+    for k =1:size(poss_iter_nums,2)
+        NUM_ITERATIONS=poss_iter_nums(k);
+        for j=1:size(functions_to_test,2)
+            %create bounds
+            rng(seed)
+            options = optimoptions('particleswarm','MaxIterations',1, 'SwarmSize',40);
+            currlimits = bounds{j};
+            MAXJ = repmat(currlimits(1),dimensions(j),1);
+            MINJ = repmat(currlimits(2), dimensions(j),1);
+            x0 = MAXJ-MINJ.*rand(size(MAXJ,1),1)+MINJ; %create an initial guess within bounds
+            [x,fval,exitflag,output]=particleswarm(functions_to_test{j},size(MINJ,1),MINJ,MAXJ,options);
+            inital_vals(j,k) = fval; %call our inital val the result after 1 iteration of PSO
+             options = optimoptions('particleswarm','MaxIterations',NUM_ITERATIONS/NUM_PARTCLES, 'SwarmSize',40);% 'MinNeighborsFraction', 0);
+             %now run PSO for actual number of interations
+             rng(seed);
+            [x,fval,exitflag,output]=particleswarm(functions_to_test{j},size(MINJ,1),MINJ,MAXJ,options);
+            best_params{j,k} = x;
+            fvals(j,k) = fval;
+            outputs{j,k} = output;
+        end
+    end
+end

### 2017-12-14 DOPS/utilities/runSimulatedAnnealing.m

@@ -1,0 +1,37 @@
+function[best_params, fvals, outputs, inital_vals]=runSimulatedAnnealing(seed)
+    function [stop,options,optchanged] = myoutput(options,optimvalues,flag)
+        stop = false;
+        if isequal(flag,'iter')
+              history = [history; optimvalues.fval];
+        end
+        optchanged = false;
+    end
+    
+
+    poss_iter_nums =[40,2000,4000];
+    functions_to_test = {@ackley, @rast, @rast};
+    bounds = {[30;-15], [5.12;-5.12], [5.12;-5.12]};
+    dimensions =[10,10,300];
+    best_params = {};
+    fvals = zeros(size(dimensions,2),size(functions_to_test,2));
+    inital_vals = zeros(size(dimensions,2), size(functions_to_test,2));
+    outputs = {};
+    for k =1:size(poss_iter_nums,2)
+        NUM_ITERATIONS=poss_iter_nums(k);
+         options = optimoptions(@simulannealbnd,'MaxIterations',NUM_ITERATIONS, 'OutputFcn', @myoutput);
+        for j=1:size(functions_to_test,2)
+            %create bounds
+            rng(seed)
+            currlimits = bounds{j};
+            MAXJ = repmat(currlimits(1),dimensions(j),1);
+            MINJ = repmat(currlimits(2), dimensions(j),1);
+            x0 = MAXJ-MINJ.*rand(size(MAXJ,1),1)+MINJ; %create an initial guess within bounds
+            inital_vals(j,k) = feval(functions_to_test{j},x0);
+            history = [];
+            [x,fval,exitflag,output]=simulannealbnd(functions_to_test{j},x0,MINJ,MAXJ,options);
+            best_params{j,k} = x;
+            fvals(j,k) = fval;
+            outputs{j,k} = output;
+        end
+    end
+end

### 2017-12-14 DOPS/utilities/runSimulatedAnnealingOnCoag.m

@@ -1,0 +1,35 @@
+function [allhistory]= runSimulatedAnnealingOnCoag()
+	seed = 1;
+	b=load('coagBounds.mat');
+	lb = b.lb;
+	ub = b.ub;
+   function [stop,options,optchanged] = myoutput(options,optimvalues,flag)
+        stop = false;
+        if isequal(flag,'iter')
+              history = [history; optimvalues.fval];
+            fprintf('Iter %d and functional value %f\n', optimvalues.iteration, optimvalues.fval);  
+            optimvalues.iteration; %so I can see what's going on'
+        end
+        optchanged = false;
+    end
+	NUM_ITERATIONS = 2000;
+	options = optimoptions(@simulannealbnd,'MaxIterations',NUM_ITERATIONS, 'OutputFcn', @myoutput);
+	NUM_REPEATS = 25;
+	allhistory = zeros(NUM_REPEATS, NUM_ITERATIONS+1);
+	for j=10:NUM_REPEATS
+        rng(j);
+		fprintf("On trial %d\n", j)
+		history = [];
+		%create initial guess
+		x0 = lb-ub.*rand(size(lb,1),1)+lb;
+		tstart = tic();
+		[x,fval,exitflag,output] = simulannealbnd(@fitCoag,x0,lb,ub,options);
+		timeSA(j) = toc(tstart)
+		%size(history) %history is one longer than number of iterations 
+		cmd1= ['save -ascii ../SA_Results/fitCoag/FuntionalValuesTrial', num2str(j), '.txt history'];
+		eval(cmd1)
+		cmd2= ['save -ascii ../SA_Results/fitCoag/TimesTrial', num2str(j), '.txt timeSA'];
+		eval(cmd2)
+		allhistory(j,:) = history;
+	end
+end

### 2017-12-14 DOPS/utilities/run_DOPS_PSO_only.m

@@ -1,0 +1,77 @@
+
+%  ----------------------------------------------------------------------------------- 
+%  Copyright (c) 2016 Varnerlab
+%  School of Chemical and Biomolecular Engineering
+%  Cornell University, Ithaca NY 14853 USA
+%
+%----------------*****  Code Author Information *****----------------------
+%   Code Author (Implementation Questions, Bug Reports, etc.): 
+%       Adithya Sagar: asg242@cornell.edu
+% 
+%  Permission is hereby granted, free of charge, to any person obtaining a copy
+%  of this software and associated documentation files (the "Software"), to deal
+%  in the Software without restriction, including without limitation the rights
+%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+%  copies of the Software, and to permit persons to whom the Software is
+%  furnished to do so, subject to the following conditions:
+
+%  The above copyright notice and this permission notice shall be included
+%  in
+%  all copies or substantial portions of the Software.
+%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+%  THE SOFTWARE.
+%  ----------------------------------------------------------------------------------- 
+function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls,DDS_iters]=run_DOPS_PSO_only(optFunction, MAXJ,MINJ,lowerIter,upperIter,numSwarms)
+    %only run PSO, no DDS
+    %clear all;
+    more off;
+    warning('off','all');
+
+    %% PARAMETERS FOR SWARM SEARCH AND DYNAMICALLY DIMENSIONED SEARCH
+
+    NP=40;    %default 40                 %Number of particles in the swarm
+    NI=100;                    %Number of iterations
[…]

### 2017-12-14 DOPS/utilities/run_DOPS_with_Initial_Guess.m

@@ -1,0 +1,88 @@
+
+%  ----------------------------------------------------------------------------------- 
+%  Copyright (c) 2016 Varnerlab
+%  School of Chemical and Biomolecular Engineering
+%  Cornell University, Ithaca NY 14853 USA
+%
+%----------------*****  Code Author Information *****----------------------
+%   Code Author (Implementation Questions, Bug Reports, etc.): 
+%       Adithya Sagar: asg242@cornell.edu
+% 
+%  Permission is hereby granted, free of charge, to any person obtaining a copy
+%  of this software and associated documentation files (the "Software"), to deal
+%  in the Software without restriction, including without limitation the rights
+%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+%  copies of the Software, and to permit persons to whom the Software is
+%  furnished to do so, subject to the following conditions:
+
+%  The above copyright notice and this permission notice shall be included
+%  in
+%  all copies or substantial portions of the Software.
+%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+%  THE SOFTWARE.
+%  ----------------------------------------------------------------------------------- 
+function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=run_DOPS_with_Initial_Guess(optFunction, MAXJ,MINJ,lowerIter,upperIter,initial_guess)
+    %for running DOPS with an initial parameter guess
+    %clear all;
+    more off;
+    warning('off','all');
+
+    %% PARAMETERS FOR SWARM SEARCH AND DYNAMICALLY DIMENSIONED SEARCH
+
+    NP=40;    %default 40                 %Number of particles in the swarm
+    NI=4000;                    %Number of iterations
[…]

### 2017-12-14 DOPS/BukinF6.m
Delete BukinF6.m

@@ -1,11 +1,0 @@
-function [ f ] = BukinF6( x,y)
-%UNTITLED Summary of this function goes here
-%   Detailed explanation goes here
-if(nargin<2)
-    y = [];
-end
-
-f = 100*sqrt(abs(x(2)-.01*x(1)^2))+.01*abs(x(1)+10);
-
-end
-

### 2017-12-14 DOPS/ExpDOPS_core/DOPS_DDS_ExperimentalV2gamma.m
more reorg

@@ -1,0 +1,154 @@
+%  ----------------------------------------------------------------------------------- 
+%  Copyright (c) 2016 Varnerlab
+%  School of Chemical and Biomolecular Engineering
+%  Cornell University, Ithaca NY 14853 USA
+%
+%----------------*****  Code Author Information *****----------------------
+%   Code Author (Implementation Questions, Bug Reports, etc.): 
+%       Adithya Sagar: asg242@cornell.edu
+% 
+%  Permission is hereby granted, free of charge, to any person obtaining a copy
+%  of this software and associated documentation files (the "Software"), to deal
+%  in the Software without restriction, including without limitation the rights
+%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+%  copies of the Software, and to permit persons to whom the Software is
+%  furnished to do so, subject to the following conditions:
+
+%  The above copyright notice and this permission notice shall be included
+%  in
+%  all copies or substantial portions of the Software.
+%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+%  THE SOFTWARE.
+%  ------------------------------------------------------------------------
+%  ----------- 
+
+%% Dynamically Dimensioned Search
+function[best,x_best,num_iters_remaining]=DOPS_DDS_ExperimentalV2gamma(optFunction,IC,MAXJ,MINJ,r,NI)
+fprintf('In DDS. NI= %d\n', NI);
+global num_method_switches;
+global best_PSO_val;        %keep track of best functional value found by PSO
[…]

### 2017-12-14 DOPS/ExpDOPS_core/DOPS_Main_ExperimentalV2gamma.m

@@ -1,0 +1,84 @@
+function [ g_best_solution,bestparticle,particle,fitness, bestval_dds_swarm, best_particle_dds_swarm,best_particles_ls,DDS_iters] =DOPS_Main_ExperimentalV2gamma( optFunction,MAXJ,MINJ,NP,NI,NS,G,r)
+%to switch back between PSO and DDS in experimental version of DOPS
+    global best_PSO_val;        %keep track of best functional value found by PSO
+    global best_DDS_val;        %keep track of best functional value found by DDS
+    global best_PSO_x;          % keep track of best parameters found by PSO
+    global best_DDS_x;          %keep track of best parameters found by DDS
+    global num_method_switches; %keep track of how many times we've switched methods
+    global solve_tol;
+num_iters_remaining = NI;
+g_best_solution=[];
+bestparticle=[];
+particle=[];
+fitness=[]; 
+bestval_dds_swarm=[]; best_particle_dds_swarm=[];best_particles_ls = [];
+DDS_iters = [];
+count = 0;
+while((num_iters_remaining > 0) && size(g_best_solution,2)<NI)
+    %call PSO, first time without feeding back guess
+    if(count ==0)
+        [g_best_solution_r,bestparticle_r,particle_r,fitness_r,num_iters_remaining] = DOPS_PSO_ExperimentalV2gamma(optFunction,MAXJ,MINJ,NP,num_iters_remaining,NS,G,r);
+    else
+        if(num_iters_remaining == 1)
+            %so that PSO will actually run
+            num_iters_remaining = 2;
+        end
+        [g_best_solution_r,bestparticle_r,particle_r,fitness_r,num_iters_remaining] = DOPS_PSO_ExperimentalV2gamma(optFunction,MAXJ,MINJ,NP,num_iters_remaining,NS,G,r,IC);
+    end
+    if(size(g_best_solution,2)>NI)
+        break;
+    end
+    %concat results
+    filled_in_g_best_solution=fillInPSO(NP,g_best_solution_r);
+    g_best_solution = cat(2,g_best_solution,filled_in_g_best_solution);
+    
+    bestparticle =cat(2,bestparticle,bestparticle_r);
+    particle = cat(2,particle, particle_r);
+    fitness = cat(2,fitness,fitness_r);
+    if(count~=0)
+        if(best_DDS_val<best_PSO_val)
[…]

### 2017-12-14 DOPS/ExpDOPS_core/DOPS_PSO_ExperimentalV2gamma.m

@@ -1,0 +1,254 @@
+%  ----------------------------------------------------------------------------------- 
+%  Copyright (c) 2016 Varnerlab
+%  School of Chemical and Biomolecular Engineering
+%  Cornell University, Ithaca NY 14853 USA
+%
+%----------------*****  Code Author Information *****----------------------
+%   Code Author (Implementation Questions, Bug Reports, etc.): 
+%       Adithya Sagar: asg242@cornell.edu
+% 
+%  Permission is hereby granted, free of charge, to any person obtaining a copy
+%  of this software and associated documentation files (the "Software"), to deal
+%  in the Software without restriction, including without limitation the rights
+%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+%  copies of the Software, and to permit persons to whom the Software is
+%  furnished to do so, subject to the following conditions:
+
+%  The above copyright notice and this permission notice shall be included
+%  in
+%  all copies or substantial portions of the Software.
+%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+%  THE SOFTWARE.
+%  ----------------------------------------------------------------------------------- 
+
+%% SWARM SEARCH WITH ADAPTIVE SWITCHING STRATEGY
+
+function[g_best_solution,bestparticle,particle,fitness,num_iters_remaining]=DOPS_PSO_ExperimentalV2gamma(optFunction,MAXJ,MINJ,NP,NI,NS,G,r, initial_guess)
+if(nargin ==8)
+   initial_guess = []; 
+end
+fprintf('In PSO. NI = %d\n', NI);
+global num_method_switches;
+global best_PSO_val;        %keep track of best functional value found by PSO
[…]

### 2017-12-14 DOPS/ExpDOPS_core/fit.m

@@ -1,0 +1,38 @@
+%  ----------------------------------------------------------------------------------- 
+%  Copyright (c) 2016 Varnerlab
+%  School of Chemical and Biomolecular Engineering
+%  Cornell University, Ithaca NY 14853 USA
+%
+%----------------*****  Code Author Information *****----------------------
+%   Code Author (Implementation Questions, Bug Reports, etc.): 
+%       Adithya Sagar: asg242@cornell.edu
+% 
+%  Permission is hereby granted, free of charge, to any person obtaining a copy
+%  of this software and associated documentation files (the "Software"), to deal
+%  in the Software without restriction, including without limitation the rights
+%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+%  copies of the Software, and to permit persons to whom the Software is
+%  furnished to do so, subject to the following conditions:
+
+%  The above copyright notice and this permission notice shall be included
+%  in
+%  all copies or substantial portions of the Software.
+%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+%  THE SOFTWARE.
+%  ----------------------------------------------------------------------------------- 
+
+%% Returns a numerical error value to DOPS 
+function fitness = fit(x, optFunction)
+        %optFunction = function to be minimized, x parameters
+        optFuncStr = func2str(optFunction);
+       % if(contains(optFuncStr, 'b4'))
+        %    fitness=optFunction(x);      %An example here is ackley function. This function is user defined an basically determines the error between simulation and experiment. 
+        %else
+            fitness=optFunction(x');
+        %end
+end

### 2017-12-14 DOPS/ExpDOPS_core/runForTable1ExperimentalV2gamma.m

@@ -1,0 +1,21 @@
+function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls,DDS_iters]=runForTable1ExperimentalV2gamma(lowerIter,upperIter)
+    addpath(genpath('CoagulationFiles')) %add coagulation functions to path
+    functions_to_test = {@rast, @ackley,@b4_obj, @b1_obj_v2, @fitCoag, @rast300, @BukinF6, @rosenbrock, @Eggholder, @mccorm, @hart6, @stybtang};
+   % functions_to_test = {@rast, @ackley,@rosenbrock, @ButkinF6, @fitCoag};
+    bounds_files = {'rast_bounds.mat', 'ackley_bounds.mat', 'b4_bounds.mat','b1_bounds.mat', 'coagBounds.mat', 'rast300d_bounds.mat', 'BukinF6Bounds.mat','rosenbrock10d_bounds.mat', 'EggholderBounds.mat', 'mccorm_bounds.mat'...
+        'hart6_bounds' ,'stybtang100d_bounds'};
+    %bounds_files = {'rast_bounds.mat', 'ackley_bounds.mat','rosenbrock10d_bounds.mat', 'BukinF6Bounds.mat','coagBounds.mat'}; 
+   for j=5:5%size(functions_to_test,2):size(functions_to_test,2)%-4
+        load(bounds_files{j});
+        %deal with transposed files
+        if(j ==3 || j ==4)
+           MAXJ = ub';
+           MINJ = lb';
+        else
+        
+            MAXJ = ub;
+            MINJ = lb;
+        end
+        [g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls,DDS_iters]=run_DOPS_ExperimentalV2gamma(functions_to_test{j},MAXJ,MINJ,lowerIter,upperIter);
+    end
+end

### 2017-12-14 DOPS/ExpDOPS_core/run_DOPS_ExperimentalV2gamma.m

@@ -1,0 +1,93 @@
+
+%  ----------------------------------------------------------------------------------- 
+%  Copyright (c) 2016 Varnerlab
+%  School of Chemical and Biomolecular Engineering
+%  Cornell University, Ithaca NY 14853 USA
+%
+%----------------*****  Code Author Information *****----------------------
+%   Code Author (Implementation Questions, Bug Reports, etc.): 
+%       Adithya Sagar: asg242@cornell.edu
+% 
+%  Permission is hereby granted, free of charge, to any person obtaining a copy
+%  of this software and associated documentation files (the "Software"), to deal
+%  in the Software without restriction, including without limitation the rights
+%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+%  copies of the Software, and to permit persons to whom the Software is
+%  furnished to do so, subject to the following conditions:
+
+%  The above copyright notice and this permission notice shall be included
+%  in
+%  all copies or substantial portions of the Software.
+%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+%  THE SOFTWARE.
+%  ----------------------------------------------------------------------------------- 
+function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls,DDS_iters]=run_DOPS_ExperimentalV2gamma(optFunction, MAXJ,MINJ,lowerIter,upperIter)
+
+    %clear all;
+    more off;
+    warning('off','all');
+
+    %% PARAMETERS FOR SWARM SEARCH AND DYNAMICALLY DIMENSIONED SEARCH
+
+    NP=40;    %default 40                 %Number of particles in the swarm
+    NI=4000;                    %Number of iterations
+    NS=5;              %default 5        %Number of sub swarms
[…]

### 2017-12-14 DOPS/DOPS_DDS.m
Delete DOPS_DDS.m

@@ -1,131 +1,0 @@
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ------------------------------------------------------------------------
-%  ----------- 
-
-%% Dynamically Dimensioned Search
-function[best,x_best,dds_swarm_flag]=DOPS_DDS(optFunction,IC,MAXJ,MINJ,r,NI)
-
-% DDS search parameters
-N=length(MINJ);
-x_best(:,1)=IC;          % Initialize the solution
-x=IC;
-J=1:N;                   % Specify initial dimensions being perturbed
[…]

### 2017-12-14 DOPS/DOPS_PSO.m
Delete DOPS_PSO.m

@@ -1,225 +1,0 @@
- %  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ----------------------------------------------------------------------------------- 
-
-%% SWARM SEARCH WITH ADAPTIVE SWITCHING STRATEGY
-
-function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=DOPS_PSO(optFunction,MAXJ,MINJ,NP,NI,NS,G,r)
-
-% Parameters for swarm search
-Max_Inertia_weight=0.9;
-Min_Inertia_weight=0.4;
-C1=1.5;C2=1.5;
-N=length(MINJ);
[…]

### 2017-12-14 DOPS/DOPS_Main_ExperimentalV2gamma.m
Delete DOPS_Main_ExperimentalV2gamma.m

@@ -1,88 +1,0 @@
-function [ g_best_solution,bestparticle,particle,fitness, bestval_dds_swarm, best_particle_dds_swarm,best_particles_ls,DDS_iters] =DOPS_Main_ExperimentalV2gamma( optFunction,MAXJ,MINJ,NP,NI,NS,G,r)
-%to switch back between PSO and DDS in experimental version of DOPS
-    global best_PSO_val;        %keep track of best functional value found by PSO
-    global best_DDS_val;        %keep track of best functional value found by DDS
-    global best_PSO_x;          % keep track of best parameters found by PSO
-    global best_DDS_x;          %keep track of best parameters found by DDS
-    global num_method_switches; %keep track of how many times we've switched methods
-    global solve_tol;
-num_iters_remaining = NI;
-g_best_solution=[];
-bestparticle=[];
-particle=[];
-fitness=[]; 
-bestval_dds_swarm=[]; best_particle_dds_swarm=[];best_particles_ls = [];
-DDS_iters = [];
-count = 0;
-while((num_iters_remaining > 0) && size(g_best_solution,2)<NI)
-    %call PSO, first time without feeding back guess
-    if(count ==0)
-        [g_best_solution_r,bestparticle_r,particle_r,fitness_r,num_iters_remaining] = DOPS_PSO_ExperimentalV2gamma(optFunction,MAXJ,MINJ,NP,num_iters_remaining,NS,G,r);
-    else
-        if(num_iters_remaining == 1)
-            %so that PSO will actually run
-            num_iters_remaining = 2;
-        end
-        [g_best_solution_r,bestparticle_r,particle_r,fitness_r,num_iters_remaining] = DOPS_PSO_ExperimentalV2gamma(optFunction,MAXJ,MINJ,NP,num_iters_remaining,NS,G,r,IC);
-    end
-    if(size(g_best_solution,2)>NI)
-        break;
-    end
-    %concat results
-    filled_in_g_best_solution=fillInPSO(NP,g_best_solution_r);
-    g_best_solution = cat(2,g_best_solution,filled_in_g_best_solution);
-    
-    bestparticle =cat(2,bestparticle,bestparticle_r);
-    particle = cat(2,particle, particle_r);
-    fitness = cat(2,fitness,fitness_r);
-    if(count~=0)
-        if(best_DDS_val<best_PSO_val)
[…]

### 2017-12-14 DOPS/DOPS_Main_ExperimentalV2beta.m
Delete DOPS_Main_ExperimentalV2beta.m

@@ -1,88 +1,0 @@
-function [ g_best_solution,bestparticle,particle,fitness, bestval_dds_swarm, best_particle_dds_swarm,best_particles_ls,DDS_iters] =DOPS_Main_ExperimentalV2beta( optFunction,MAXJ,MINJ,NP,NI,NS,G,r)
-%to switch back between PSO and DDS in experimental version of DOPS
-    global best_PSO_val;        %keep track of best functional value found by PSO
-    global best_DDS_val;        %keep track of best functional value found by DDS
-    global best_PSO_x;          % keep track of best parameters found by PSO
-    global best_DDS_x;          %keep track of best parameters found by DDS
-    global num_method_switches; %keep track of how many times we've switched methods
-    global solve_tol;
-num_iters_remaining = NI;
-g_best_solution=[];
-bestparticle=[];
-particle=[];
-fitness=[]; 
-bestval_dds_swarm=[]; best_particle_dds_swarm=[];best_particles_ls = [];
-DDS_iters = [];
-count = 0;
-while((num_iters_remaining > 0) && size(g_best_solution,2)<NI)
-    %call PSO, first time without feeding back guess
-    if(count ==0)
-        [g_best_solution_r,bestparticle_r,particle_r,fitness_r,num_iters_remaining] = DOPS_PSO_ExperimentalV2(optFunction,MAXJ,MINJ,NP,num_iters_remaining,NS,G,r);
-    else
-        if(num_iters_remaining == 1)
-            %so that PSO will actually run
-            num_iters_remaining = 2;
-        end
-        [g_best_solution_r,bestparticle_r,particle_r,fitness_r,num_iters_remaining] = DOPS_PSO_ExperimentalV2(optFunction,MAXJ,MINJ,NP,num_iters_remaining,NS,G,r,IC);
-    end
-    if(size(g_best_solution,2)>NI)
-        break;
-    end
-    %concat results
-    filled_in_g_best_solution=fillInPSO(NP,g_best_solution_r);
-    g_best_solution = cat(2,g_best_solution,filled_in_g_best_solution);
-    
-    bestparticle =cat(2,bestparticle,bestparticle_r);
-    particle = cat(2,particle, particle_r);
-    fitness = cat(2,fitness,fitness_r);
-    if(count~=0)
-        if(best_DDS_val<best_PSO_val)
[…]

### 2017-12-14 DOPS/DOPS_Main_Experimental.m
Delete DOPS_Main_Experimental.m

@@ -1,88 +1,0 @@
-function [ g_best_solution,bestparticle,particle,fitness, bestval_dds_swarm, best_particle_dds_swarm,best_particles_ls,DDS_iters] =DOPS_Main_Experimental( optFunction,MAXJ,MINJ,NP,NI,NS,G,r)
-%to switch back between PSO and DDS in experimental version of DOPS
-    global best_PSO_val;        %keep track of best functional value found by PSO
-    global best_DDS_val;        %keep track of best functional value found by DDS
-    global best_PSO_x;          % keep track of best parameters found by PSO
-    global best_DDS_x;          %keep track of best parameters found by DDS
-    global num_method_switches; %keep track of how many times we've switched methods
-    global solve_tol;
-num_iters_remaining = NI;
-g_best_solution=[];
-bestparticle=[];
-particle=[];
-fitness=[]; 
-bestval_dds_swarm=[]; best_particle_dds_swarm=[];best_particles_ls = [];
-DDS_iters = [];
-count = 0;
-while((num_iters_remaining > 0) && size(g_best_solution,2)<NI)
-    %call PSO, first time without feeding back guess
-    if(count ==0)
-        [g_best_solution_r,bestparticle_r,particle_r,fitness_r,num_iters_remaining] = DOPS_PSO_ExperimentalV2(optFunction,MAXJ,MINJ,NP,num_iters_remaining,NS,G,r);
-    else
-        if(num_iters_remaining == 1)
-            %so that PSO will actually run
-            num_iters_remaining = 2;
-        end
-        [g_best_solution_r,bestparticle_r,particle_r,fitness_r,num_iters_remaining] = DOPS_PSO_ExperimentalV2(optFunction,MAXJ,MINJ,NP,num_iters_remaining,NS,G,r,IC);
-    end
-    if(size(g_best_solution,2)>NI)
-        break;
-    end
-    %concat results
-    filled_in_g_best_solution=fillInPSO(NP,g_best_solution_r);
-    g_best_solution = cat(2,g_best_solution,filled_in_g_best_solution);
-    
-    bestparticle =cat(2,bestparticle,bestparticle_r);
-    particle = cat(2,particle, particle_r);
-    fitness = cat(2,fitness,fitness_r);
-    if(count~=0)
-        if(best_DDS_val<best_PSO_val)
-           IC = best_DDS_x;
-        else
[…]

### 2017-12-14 DOPS/DOPS_DDS_ExperimentalV2gamma.m
Delete DOPS_DDS_ExperimentalV2gamma.m

@@ -1,160 +1,0 @@
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ------------------------------------------------------------------------
-%  ----------- 
-
-%% Dynamically Dimensioned Search
-function[best,x_best,num_iters_remaining]=DOPS_DDS_ExperimentalV2gamma(optFunction,IC,MAXJ,MINJ,r,NI)
-fprintf('In DDS. NI= %d\n', NI);
-global num_method_switches;
-global best_PSO_val;        %keep track of best functional value found by PSO
[…]

### 2017-12-14 DOPS/DOPS_DDS_ExperimentalV2beta.m
Delete DOPS_DDS_ExperimentalV2beta.m

@@ -1,163 +1,0 @@
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ------------------------------------------------------------------------
-%  ----------- 
-
-%% Dynamically Dimensioned Search
-function[best,x_best,num_iters_remaining]=DOPS_DDS_ExperimentalV2beta(optFunction,IC,MAXJ,MINJ,r,NI)
-fprintf('In DDS. NI= %d\n', NI);
-global num_method_switches;
-global best_PSO_val;        %keep track of best functional value found by PSO
[…]

### 2017-12-14 DOPS/DOPS_DDS_ExperimentalV2.m
Delete DOPS_DDS_ExperimentalV2.m

@@ -1,167 +1,0 @@
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ------------------------------------------------------------------------
-%  ----------- 
-
-%% Dynamically Dimensioned Search
-function[best,x_best,num_iters_remaining]=DOPS_DDS_ExperimentalV2(optFunction,IC,MAXJ,MINJ,r,NI)
-fprintf('In DDS. NI= %d\n', NI);
-global num_method_switches;
-global best_PSO_val;        %keep track of best functional value found by PSO
-global best_DDS_val;        %keep track of best functional value found by DDS
[…]

### 2017-12-14 DOPS/DOPS_DDS_Experimental.m
Delete DOPS_DDS_Experimental.m

@@ -1,140 +1,0 @@
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ------------------------------------------------------------------------
-%  ----------- 
-
-%% Dynamically Dimensioned Search
-function[best,x_best,dds_swarm_flag,num_iters_remaining]=DOPS_DDS(optFunction,IC,MAXJ,MINJ,r,NI,NS,G,NP)
-fprintf("In DDS. NI= %d\n", NI);
-global num_method_switches;
-global best_PSO_val;        %keep track of best functional value found by PSO
[…]

### 2017-12-14 DOPS/DOPS_PSO_ExperimentalV2.m
Delete DOPS_PSO_ExperimentalV2.m

@@ -1,249 +1,0 @@
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ----------------------------------------------------------------------------------- 
-
-%% SWARM SEARCH WITH ADAPTIVE SWITCHING STRATEGY
-
-function[g_best_solution,bestparticle,particle,fitness,num_iters_remaining]=DOPS_PSO_ExperimentalV2(optFunction,MAXJ,MINJ,NP,NI,NS,G,r, initial_guess)
-if(nargin ==8)
-   initial_guess = []; 
-end
-fprintf('In PSO. NI = %d\n', NI);
-global num_method_switches;
-global best_PSO_val;        %keep track of best functional value found by PSO
[…]

### 2017-12-14 DOPS/DOPS_PSO_parallel.m
Delete DOPS_PSO_parallel.m

@@ -1,220 +1,0 @@
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ----------------------------------------------------------------------------------- 
-
-%% SWARM SEARCH WITH ADAPTIVE SWITCHING STRATEGY
-
-function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=DOPS_PSO_parallel(optFunction,MAXJ,MINJ,NP,NI,NS,G,r)
-
-% Parameters for swarm search
-Max_Inertia_weight=0.9;
-Min_Inertia_weight=0.4;
-C1=1.5;C2=1.5;
-N=length(MINJ);
[…]

### 2017-12-14 DOPS/b1_obj.m
Delete b1_obj.m

@@ -1,545 +1,0 @@
-
-function [objective,constraints,residuals] = b1_obj(par)
-
-global p
-
-% 'b1.m' yields a 'par' vector with 1770 elements (of which 1759 are unknown)
-% If the 'par' vector is taken from an optimization result, it will in principle have
-% only 1759 elements. In this case we add the 11 known extra pars:
-if numel(par)==1759
-    knownpars = [
-    0,... %s_0364
-    1,... %s_0420
-    0,...%s_0458
-    0,... %s_0681
-    0,... %s_0766
-    1,... %s_1277
-    1,... %s_1324
-    1,... %s_1468	
-    0,... %s_1521
-    1,... %cell
-    1 %extracellular
-    ];
-    p = [par,knownpars];
-end
-    
-p = par;
-
-
-%% EXPERIMENTAL DATA
-
-exp_data{1}			= [ ...
-									0.000840417,	0.169184,	0.0163175,	0.00239284,	0.127935,	0.203522,	56.8005,	1.11884,	0.0210244,	0.375397,	4.43861,	1.03812,	0.0704771,	1.32223,	0.152018,	0.226579,	0.156606,	0.328797,	45.55,	0.125229,	0.0733177,	0.0570895,	0.923224,	0.141514,	0.00989585,	0.280125,	0.232816,	0.219758,	0.194598,	1.36405,	0.102268,	0.131983,	0.473955,	0.00449796,	0.254449,	0.403419,	0.579927,	0.0421182,	3.76888,	5.7798,	5.03726,	0.0821183,	0.027064,	0.00906198;
-									0.000802963,	0.158521,	0.0161075,	0.00307146,	0.109955,	0.160931,	49.574,	1.10448,	0.016191,	0.450099,	4.82148,	1.37501,	0.0692448,	1.47821,	0.11133,	0.182669,	0.145041,	0.243723,	60.181,	0.150438,	0.0722532,	0.0726141,	1.10651,	0.132761,	0.00878002,	0.35718,	0.181043,	0.185836,	0.19663,	1.66104,	0.0745045,	0.165711,	0.423285,	0.00450396,	0.327811,	0.457869,	0.635343,	0.0500988,	2.99164,	7.15244,	4.85506,	0.0718256,	0.0281856,	0.00809582;
[…]

### 2017-12-14 DOPS/b1_obj_forS1.m
Delete b1_obj_forS1.m

@@ -1,548 +1,0 @@
-
-function [objective,constraints,residuals,x_sel,x,t_int,exp_data] = b1_obj_forS1(par)
-
-global p
-
-% 'b1.m' yields a 'par' vector with 1770 elements (of which 1759 are unknown)
-% If the 'par' vector is taken from an optimization result, it will in principle have
-% only 1759 elements. In this case we add the 11 known extra pars:
-if numel(par)==1759
-    knownpars = [
-    0,... %s_0364
-    1,... %s_0420
-    0,...%s_0458
-    0,... %s_0681
-    0,... %s_0766
-    1,... %s_1277
-    1,... %s_1324
-    1,... %s_1468	
-    0,... %s_1521
-    1,... %cell
-    1 %extracellular
-    ];
-    p = [par,knownpars];
-end
-    
-p = par;
-
-
-%% EXPERIMENTAL DATA
-
-exp_data{1}			= [ ...
-									0.000840417,	0.169184,	0.0163175,	0.00239284,	0.127935,	0.203522,	56.8005,	1.11884,	0.0210244,	0.375397,	4.43861,	1.03812,	0.0704771,	1.32223,	0.152018,	0.226579,	0.156606,	0.328797,	45.55,	0.125229,	0.0733177,	0.0570895,	0.923224,	0.141514,	0.00989585,	0.280125,	0.232816,	0.219758,	0.194598,	1.36405,	0.102268,	0.131983,	0.473955,	0.00449796,	0.254449,	0.403419,	0.579927,	0.0421182,	3.76888,	5.7798,	5.03726,	0.0821183,	0.027064,	0.00906198;
[…]

### 2017-12-14 DOPS/b1_obj_v2.m
Delete b1_obj_v2.m

@@ -1,545 +1,0 @@
-
-function [objective,constraints,residuals,t_int,ms,expdata] = b1_obj_v2(par)
-
-global p
-
-% 'b1.m' yields a 'par' vector with 1770 elements (of which 1759 are unknown)
-% If the 'par' vector is taken from an optimization result, it will in principle have
-% only 1759 elements. In this case we add the 11 known extra pars:
-if numel(par)==1759
-    knownpars = [
-    0,... %s_0364
-    1,... %s_0420
-    0,...%s_0458
-    0,... %s_0681
-    0,... %s_0766
-    1,... %s_1277
-    1,... %s_1324
-    1,... %s_1468	
-    0,... %s_1521
-    1,... %cell
-    1 %extracellular
-    ];
-    p = [par,knownpars];
-end
-    
-p = par;
-
-
-%% EXPERIMENTAL DATA
-
-exp_data{1}			= [ ...
-									0.000840417,	0.169184,	0.0163175,	0.00239284,	0.127935,	0.203522,	56.8005,	1.11884,	0.0210244,	0.375397,	4.43861,	1.03812,	0.0704771,	1.32223,	0.152018,	0.226579,	0.156606,	0.328797,	45.55,	0.125229,	0.0733177,	0.0570895,	0.923224,	0.141514,	0.00989585,	0.280125,	0.232816,	0.219758,	0.194598,	1.36405,	0.102268,	0.131983,	0.473955,	0.00449796,	0.254449,	0.403419,	0.579927,	0.0421182,	3.76888,	5.7798,	5.03726,	0.0821183,	0.027064,	0.00906198;
[…]

### 2017-12-14 DOPS/b4_obj.m
Delete b4_obj.m

@@ -1,311 +1,0 @@
-
-function [objective,constraints,residuals,x] = b4_obj(par)
-
-global p
-p = par;
-
-%% EXPERIMENTAL DATA
-
-% steady-state reference concentrations:
-      css(1)=100000;
-      css(2)=1;
-      css(3)=1000;
-      css(4)=1000;
-      css(5)=30;
-      css(6)=1000;
-      css(7)=1000;
-      css(8)=1000;
-      css(9)=1000;
-      css(10)=1000;
-      css(11)=1000;
-      css(12)=1000;
-      css(13)=3000;
-      css(14)=1000;
-      css(15)=1000;
-      css(16)=1000;
-      css(17)=100;
-      css(18)=100;
-      css(19)=100;
-      css(20)=100;
-      css(21)=1000;
-      css(22)=1000;
-      css(23)=1000;
-      css(24)=1000;
-      css(25)=1000;
-      css(26)=1000;
-      css(27)=1000;
-      css(28)=1000;
-      css(29)=1000;
-      css(30)=3000;
-      css(31)=100000;
-      css(32)=1000;
-      css(33)=1000;
-      css(34)=1000;
-                              
-t_s      = [23.9759 48.4337 72.8916 90.3614 115.984 140.442 164.9 210.321 233.614 258.072 282.53 300];
-timeseries_artificalData_noised = [
-% Productprotein	L-Methionine	L-Leucine	L-Lactate	beta-D-Glucose	L-Aspartate	L-Malate	Pyruvate	Oxaloacetate	ATP	ATP	ADP	ADP_c
-% Time	f	f	f	f	f	c	c	c	m	c	m	m	c
-% min	mumol/l_fermenter	mumol/l_fermenter	mumol/l_fermenter	mumol/l_fermenter	mumol/l_fermenter	mumol/l_cell	mumol/l_cell	mumol/l_cell	mumol/l_cell	mumol/l_cell	mumol/l_cell	mumol/l_cell	mumol/l_cell
-0	    29.54174187	5110.119691	4999.956639	1.056682831	449958.4343	1009.945764	988.3517969	996.1405443	839.7703035	3314.558385	2968.296835	1016.447754	1029.663149
-23.9759	29.06135825	4427.537773	4990.471667	19033.05068	499628.9405	514.889607	1883.981851	4115.730274	443.5222307	3298.082924	3588.716676	701.1769714	841.6656476
[…]

### 2017-12-14 DOPS/b4_obj_forS2.m
Delete b4_obj_forS2.m

@@ -1,314 +1,0 @@
-
-function [objective,constraints,residuals,x_all,x_useful,t,timeseries_artificalData_noised,scaled_data] = b4_obj_forS2(par)
-
-global p
-p = par;
-
-%% EXPERIMENTAL DATA
-
-% steady-state reference concentrations:
-      css(1)=100000;
-      css(2)=1;
-      css(3)=1000;
-      css(4)=1000;
-      css(5)=30;
-      css(6)=1000;
-      css(7)=1000;
-      css(8)=1000;
-      css(9)=1000;
-      css(10)=1000;
-      css(11)=1000;
-      css(12)=1000;
-      css(13)=3000;
-      css(14)=1000;
-      css(15)=1000;
-      css(16)=1000;
-      css(17)=100;
-      css(18)=100;
-      css(19)=100;
-      css(20)=100;
-      css(21)=1000;
-      css(22)=1000;
-      css(23)=1000;
-      css(24)=1000;
-      css(25)=1000;
-      css(26)=1000;
-      css(27)=1000;
-      css(28)=1000;
-      css(29)=1000;
-      css(30)=3000;
-      css(31)=100000;
-      css(32)=1000;
-      css(33)=1000;
-      css(34)=1000;
-                              
-t_s      = [23.9759 48.4337 72.8916 90.3614 115.984 140.442 164.9 210.321 233.614 258.072 282.53 300];
-timeseries_artificalData_noised = [
-% Productprotein	L-Methionine	L-Leucine	L-Lactate	beta-D-Glucose	L-Aspartate	L-Malate	Pyruvate	Oxaloacetate	ATP	ATP	ADP	ADP_c
-% Time	f	f	f	f	f	c	c	c	m	c	m	m	c
-% min	mumol/l_fermenter	mumol/l_fermenter	mumol/l_fermenter	mumol/l_fermenter	mumol/l_fermenter	mumol/l_cell	mumol/l_cell	mumol/l_cell	mumol/l_cell	mumol/l_cell	mumol/l_cell	mumol/l_cell	mumol/l_cell
-0	    29.54174187	5110.119691	4999.956639	1.056682831	449958.4343	1009.945764	988.3517969	996.1405443	839.7703035	3314.558385	2968.296835	1016.447754	1029.663149
[…]

### 2017-12-14 DOPS/b4_test.m
Delete b4_test.m

@@ -1,7 +1,0 @@
-% b4_test
-% performs a test integration of the b4 model with nominal parameter values
-% and displays the objective function value
-
-[ydot, xinit, par] = b4(0);
-objective = b4_obj(par);
-disp(objective)

### 2017-12-14 DOPS/run_DOPS.m
Delete run_DOPS.m

@@ -1,95 +1,0 @@
-
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ----------------------------------------------------------------------------------- 
-function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=run_DOPS(optFunction, MAXJ,MINJ,lowerIter,upperIter)
-
-    %clear all;
-    more off;
-    warning('off','all');
-
-    %bounds=load('bounds.txt');
-    %MAXJ=bounds(:,2);
-    %MINJ=bounds(:,1);
-    %for ackley, bounds -5 to 5
-    %MAXJ=[5.12;5.12];
[…]

### 2017-12-14 DOPS/run_DOPS_ExperimentalV2.m
Delete run_DOPS_ExperimentalV2.m

@@ -1,103 +1,0 @@
-
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ----------------------------------------------------------------------------------- 
-function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls,DDS_iters]=run_DOPS_ExperimentalV2(optFunction, MAXJ,MINJ,lowerIter,upperIter)
-
-    %clear all;
-    more off;
-    warning('off','all');
-
-
-    %bounds=load('bounds.txt');
-    %MAXJ=bounds(:,2);
-    %MINJ=bounds(:,1);
[…]

### 2017-12-14 DOPS/run_DOPS_ExperimentalV2gamma.m
Delete run_DOPS_ExperimentalV2gamma.m

@@ -1,103 +1,0 @@
-
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ----------------------------------------------------------------------------------- 
-function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls,DDS_iters]=run_DOPS_ExperimentalV2gamma(optFunction, MAXJ,MINJ,lowerIter,upperIter)
-
-    %clear all;
-    more off;
-    warning('off','all');
-
-
-    %bounds=load('bounds.txt');
-    %MAXJ=bounds(:,2);
-    %MINJ=bounds(:,1);
[…]

### 2017-12-14 DOPS/run_DOPS_PSO_only.m
Delete run_DOPS_PSO_only.m

@@ -1,87 +1,0 @@
-
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ----------------------------------------------------------------------------------- 
-function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls,DDS_iters]=run_DOPS_PSO_only(optFunction, MAXJ,MINJ,lowerIter,upperIter,numSwarms)
-
-    %clear all;
-    more off;
-    warning('off','all');
-
-
-    %bounds=load('bounds.txt');
-    %MAXJ=bounds(:,2);
-    %MINJ=bounds(:,1);
-    %for ackley, bounds -5 to 5
[…]

### 2017-12-14 DOPS/run_DOPS_selected_function.m
Delete run_DOPS_selected_function.m

@@ -1,35 +1,0 @@
-function[]= run_DOPS_selected_function(currfun,MAXJ,MINJ)
-
-    [g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=run_DOPS(currfunc,MAXJ,MINJ);
-
-    NUMBER_TRIALS=25;
-    %fig = figure();
-    close('all');
-    maxlen = 0;
-    for j=1:NUMBER_TRIALS
-        currlen=size([g_best_solution{j},bestval_dds_swarm{j}],2);
-        if(currlen>maxlen)
-           maxlen = currlen;
-        end
-       % alldata(j,:)=([g_best_solution{j},bestval_dds_swarm{j}]);
-    end
-    alldata = zeros(maxlen,NUMBER_TRIALS);
-    for k=1:NUMBER_TRIALS
-       currdata = ([g_best_solution{k}, bestval_dds_swarm{k}]);
-       if(size(currdata,2)<maxlen)
-          currdata(size(currdata):maxlen)=1E-16; %should find a better way of making these numbers not count 
-       end
-       size(currdata)
-       alldata(:,k)=currdata;
-    end
-
-    size(alldata)
-    close('all');
-    iters = 1:maxlen;
-    semilogy(iters,mean(alldata,2),'kx');
-    xlabel('Iteration')
-    ylabel('Error')
-    axis([0,maxlen,1E-16,.8])
-    set(gca,'yscale','log')
-    saveas(strcat('testFigures/MeanErrorFor', func2str(currfunc)), 'pdf')
-end

### 2017-12-14 DOPS/run_DOPS_with_Initial_Guess.m
Delete run_DOPS_with_Initial_Guess.m

@@ -1,95 +1,0 @@
-
-%  ----------------------------------------------------------------------------------- 
-%  Copyright (c) 2016 Varnerlab
-%  School of Chemical and Biomolecular Engineering
-%  Cornell University, Ithaca NY 14853 USA
-%
-%----------------*****  Code Author Information *****----------------------
-%   Code Author (Implementation Questions, Bug Reports, etc.): 
-%       Adithya Sagar: asg242@cornell.edu
-% 
-%  Permission is hereby granted, free of charge, to any person obtaining a copy
-%  of this software and associated documentation files (the "Software"), to deal
-%  in the Software without restriction, including without limitation the rights
-%  to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-%  copies of the Software, and to permit persons to whom the Software is
-%  furnished to do so, subject to the following conditions:
-
-%  The above copyright notice and this permission notice shall be included
-%  in
-%  all copies or substantial portions of the Software.
-%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-%  IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-%  AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-%  OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
-%  THE SOFTWARE.
-%  ----------------------------------------------------------------------------------- 
-function[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=run_DOPS_with_Initial_Guess(optFunction, MAXJ,MINJ,lowerIter,upperIter,initial_guess)
-
-    %clear all;
-    more off;
-    warning('off','all');
-
-    %bounds=load('bounds.txt');
-    %MAXJ=bounds(:,2);
-    %MINJ=bounds(:,1);
[…]

### 2017-12-14 DOPS/stybtang.m
Delete stybtang.m

@@ -1,45 +1,0 @@
-function [y] = stybtang(xx)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% STYBLINSKI-TANG FUNCTION
-%
-% Authors: Sonja Surjanovic, Simon Fraser University
-%          Derek Bingham, Simon Fraser University
-% Questions/Comments: Please email Derek Bingham at dbingham@stat.sfu.ca.
-%
-% Copyright 2013. Derek Bingham, Simon Fraser University.
-%
-% THERE IS NO WARRANTY, EXPRESS OR IMPLIED. WE DO NOT ASSUME ANY LIABILITY
-% FOR THE USE OF THIS SOFTWARE.  If software is modified to produce
-% derivative works, such modified software should be clearly marked.
-% Additionally, this program is free software; you can redistribute it 
-% and/or modify it under the terms of the GNU General Public License as 
-% published by the Free Software Foundation; version 2.0 of the License. 
-% Accordingly, this program is distributed in the hope that it will be 
-% useful, but WITHOUT ANY WARRANTY; without even the implied warranty 
-% of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU 
-% General Public License for more details.
-%
-% For function details and reference information, see:
-% http://www.sfu.ca/~ssurjano/
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% INPUT:
-%
-% xx = [x1, x2, ..., xd]
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-d = length(xx);
-sum = 0;
-for ii = 1:d
-	xi = xx(ii);
-	new = xi^4 - 16*xi^2 + 5*xi;
-	sum = sum + new;
-end
-
-y = sum/2;
-
-end

### 2017-12-14 DOPS/compareDOPSVersionsOnB4.m
Delete compareDOPSVersionsOnB4.m

@@ -1,121 +1,0 @@
-function [allDOPS,adjustedAllDOPS,allExperimental,allDOPSTimes]=compareDOPSVersionsOnB4()
-    %this function was developed with testing different versions of
-    %experimental DOPS. It plots the performance of different versions on
-    %the b4 problem
-    addpath('/home/rachel/Documents/DOPS/DOPS_Results/b4_obj');
-    close('all');
-    NUM_ITERS = 25;
-    NUM_EVALS = 4000;
-    NUM_PARTCILES = 40;
-    allDOPS = zeros(NUM_ITERS,NUM_EVALS);
-    adjustedAllDOPS = zeros(NUM_ITERS,NUM_EVALS);
-    allExperimental = zeros(NUM_ITERS,NUM_EVALS);
-    allExperimentalbeta = zeros(NUM_ITERS,NUM_EVALS);
-    allExperimentalgamma = zeros(NUM_ITERS,NUM_EVALS);
-    
-    allExperimentalgamma_shrinkfactor1E5 = zeros(NUM_ITERS,NUM_EVALS);
-    allExperimentalgamma_shrinkfactor1E6 = zeros(NUM_ITERS,NUM_EVALS);
-    allDOPSTimes = zeros(NUM_ITERS,1);
-    for j=1:NUM_ITERS
-        load(strcat('DOPS_error_iter',num2str(j),'.mat'));
-        PSO_error = g_best_solution{j};
-        corrected_PSO_error = fillInPSO(NUM_PARTCILES,PSO_error);
-        load(strcat('DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter', num2str(j),'.mat'));
-        DDS_error = bestval_dds_swarm{j};
-        alldata= cat(2,PSO_error,DDS_error);
-        correctedAllData = cat(2,corrected_PSO_error,DDS_error);
-        correctedAllData = correctedAllData(1:NUM_EVALS);
-        adjustedAllDOPS(j,:) = correctedAllData;
-        
-        currTime = load(strcat('DOPS_time_iter', num2str(j), '.txt'));
-        currTime = currTime(j);
-        allDOPSTimes(j) = currTime;
-    end
-   
-    for k=1:NUM_ITERS
-        load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/ExperimentalV2/b4_obj/DOPS_error_iter', num2str(k),'.mat'));
[…]

### 2017-12-14 DOPS/compareDOPSVersionsOnSelectedFunction.m
Delete compareDOPSVersionsOnSelectedFunction.m

@@ -1,76 +1,0 @@
-function [allDOPS,adjustedAllDOPS,allExperimental,allDOPSTimes]=compareDOPSVersionsOnSelectedFunction(functionName)
-    %This function plots the mean convergence curves for DOPS and
-    %experimental DOPS assuming that the results are stored at where the
-    %paths point.
-    if(contains(functionName, 'fitCoag'))
-        addpath('/home/rachel/Documents/DOPS/DOPS_Results/fitCoag/')
-        NUM_ITERS = 25;
-    else
-       NUM_ITERS = 250; 
-        addpath(strcat('/home/rachel/Documents/DOPS/DOPS_Results/',functionName));
-    %else
-     %   NUM_ITERS = 500;
-     %   addpath(strcat('/home/rachel/Documents/DOPS/DOPS_Results/',functionName));
-    end
-   close('all');
-    
-    
-    NUM_EVALS = 4000;
-    NUM_PARTCILES = 40;
-    allDOPS = zeros(NUM_ITERS,NUM_EVALS);
-    adjustedAllDOPS = zeros(NUM_ITERS,NUM_EVALS);
-    allExperimental = zeros(NUM_ITERS,NUM_EVALS);
-    allDOPSTimes = zeros(NUM_ITERS,1);
-    %load DOPS data
-    for j=1:NUM_ITERS
-        load(strcat('DOPS_error_iter',num2str(j),'.mat'));
-        PSO_error = g_best_solution{j};
-        corrected_PSO_error = fillInPSO(NUM_PARTCILES,PSO_error);
-        load(strcat('DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter', num2str(j),'.mat'));
-        DDS_error = bestval_dds_swarm{j};
-        alldata= cat(2,PSO_error,DDS_error);
-        correctedAllData = cat(2,corrected_PSO_error,DDS_error);
-        correctedAllData = correctedAllData(1:NUM_EVALS);
-        %allDOPS(j,:) =alldata;
-        adjustedAllDOPS(j,:) = correctedAllData;
-        
-        currTime = load(strcat('DOPS_time_iter', num2str(j), '.txt'));
-        currTime = currTime(j);
-        allDOPSTimes(j) = currTime;
-    end
-   
-    %load experimental DOPS data
-    for k=1:NUM_ITERS
[…]

### 2017-12-14 DOPS/compareESSToDOPSOnCoag.m
Delete compareESSToDOPSOnCoag.m

@@ -1,65 +1,0 @@
-function [] = compareESSToDOPS()
-    close('all');
-    NUM_ITERS = 25;
-    NUM_EVALS = 4000;
-    NUM_PARTCILES = 40;
-    allDOPS = zeros(NUM_ITERS,NUM_EVALS);
-    adjustedAllDOPS = zeros(NUM_ITERS,NUM_EVALS);
-    allESS = zeros(NUM_ITERS,NUM_EVALS+1);
-    addpath('/home/rachel/Documents/DOPS/DOPS_Results/fitCoag');
-    for j=1:25
-        load(strcat('DOPS_error_iter',num2str(j),'.mat'));
-        PSO_error = g_best_solution{j};
-        corrected_PSO_error = fillInPSO(NUM_PARTCILES,PSO_error);
-        load(strcat('DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter', num2str(j),'.mat'));
-        DDS_error = bestval_dds_swarm{j};
-        alldata= cat(2,PSO_error,DDS_error);
-        correctedAllData = cat(2,corrected_PSO_error,DDS_error);
-        correctedAllData = correctedAllData(1:NUM_EVALS);
-       % allDOPS(j,:) =alldata;
-        adjustedAllDOPS(j,:) = correctedAllData;
-    end
-   
-    for k=1:NUM_ITERS
-%         if(k ==18 || k ==19 || k ==20)
-%            allESS(k,:) = NaN;
-%            continue;
-%         end
-        load(strcat('/home/rachel/Documents/DOPS/ESS_Results/fitCoag/functionalInterpIter', num2str(k),'.mat'));
-        exp_err= f_interp(1:NUM_EVALS+1);
-        allESS(k,:) = exp_err;
-    end
-    f=figure();
-    hold('on')
-    plot(1:NUM_EVALS,nanmean(adjustedAllDOPS,1), 'LineWidth', 2, 'Color', 'r')
-    plot(0:NUM_EVALS,nanmean(allESS,1), 'LineWidth', 2, 'Color','k')
-    %axis([0,4000,1E5,1E8])
-    set(gca, 'YScale', 'log')
-    set(gca, 'fontsize', 20);
-    xlabel('Number of Function Evaluations', 'FontSize', 32)
-    ylabel('Functional Value', 'FontSize',32)
-    legend('DOPS', 'ESS')
-     f.PaperPositionMode = 'auto';
[…]

### 2017-12-14 DOPS/compareSwarmNumbers.m
Delete compareSwarmNumbers.m

@@ -1,62 +1,0 @@
-function []=compareSwarmNumbers()
-    [allDOPS,adjustedAllDOPS,allExperimental]=compareExperimentalDOPS();
-    swarmNum = [1,2,4,5,8];
-    close('all');
-    f = figure();
-    hold('on')
-    numEvals = 4000;
-    numFEvals = 1:1:numEvals;
-    %plot DOPS results
-    set(gca, 'YScale', 'log')
-    semilogy(numFEvals, mean(adjustedAllDOPS,1), 'LineWidth', 2.0,'Color', 'k');
-    numRepeats = 25;
-    NUM_PARTCILES = 40;
-    colors = {repmat(75/255,1,3), repmat(125/255,1,3), repmat(175/255,1,3),[0,0,0], repmat(225/255,1,3)};
-    
-    for k = 1:size(swarmNum,2)
-       adjustedAllDOPS = [];
-       currSwarmNum = swarmNum(k);
-       for j = 1:numRepeats
-           %deal with runs that haven't finished
-         %  if(currSwarmNum ==2 && j ==15)
-          %    continue; 
-           %end
-           
-           load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/PSO_Only/fitCoag/',num2str(currSwarmNum), 'Swarms/DOPS_error_iter',num2str(j),'.mat'));
-           PSO_error = g_best_solution{j};
-           corrected_PSO_error = fillInPSO(NUM_PARTCILES,PSO_error);
-           %load(strcat('/home/rachel/Documents/DOPS/DOPS_Results/PSO_Only/fitCoag/',num2str(currSwarmNum), 'Swarms/DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter', num2str(j),'.mat'));
-           %DDS_error = bestval_dds_swarm{j};
-           %alldata= cat(2,PSO_error,DDS_error);
-           correctedAllData = corrected_PSO_error;
-           correctedAllData = correctedAllData(1:numEvals);
-           adjustedAllDOPS(j,:) = correctedAllData;
-       end
-       if (currSwarmNum ==5)
-           mDOPS=mean(adjustedAllDOPS,1);
-         semilogy(numFEvals(1:40:end), mDOPS(1:40:end), 'LineWidth', 1, 'Color', colors{k}, 'Marker', 'o','MarkerFaceColor', 'w');  
-       else
[…]

### 2017-12-14 DOPS/compareVariationInFunctionalVals.m
Delete compareVariationInFunctionalVals.m

@@ -1,78 +1,0 @@
-function []=compareVariationInFunctionalVals(functionName)
-    numRepeats = 25;
-    close('all');
-    CMAES_f = zeros(1,numRepeats);
-    DDS_f = zeros(1,numRepeats);
-    DOPS_f = zeros(1,numRepeats);
-    ESS_f = zeros(1,numRepeats);
-    SA_f = zeros(1,numRepeats);
-    DE_f = zeros(1, numRepeats);
-    
-    %read in functional values or regenerate if cheap to evaluate
-    if(strcmp('ackley',functionName) || strcmp('rast',functionName))
-       [allCMAES,allSA,allDE,allDDS,allDOPS,allESS]= replicateFigureS4(functionName,numRepeats);
-       for j = 1:numRepeats
-          CMAES_f(j) = allCMAES(end,j);
-          SA_f(j) = allSA(end,j);
-          DE_f(j)=allDE(end,j);
-          DDS_f(j) = allDDS(end,j);
-          DOPS_f(j) = allDOPS(end,j);
-          ESS_f(j) = allESS(end,j);
-       end
-    end
-    
-    if(strcmp('b4_obj', functionName))
-        for j =1:numRepeats
-            load(strcat('../ESS_Results/b4_obj/functionalInterpIter', num2str(j), '.mat'));
-            ESS_f(j) = f_interp(end);
-            load(strcat('../DOPS_Results/b4_obj/DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter' ,num2str(j), '.mat'));
-            DDS_error = bestval_dds_swarm{j};
-            DOPS_f(j) = DDS_error(end);
-        end
-    end
-    
-   if(strcmp('fitCoag', functionName)) %need to wait for ESS to finish running
-        for j =1:numRepeats
-          %  load(strcat('../ESS_Results/fitCoag/functionalInterpIter', num2str(j), '.mat'));
-          %  ESS_f(j) = f_interp(end);
-            try
-                load(strcat('../DOPS_Results/fitCoag/DDSPSO_strategy1_5_swarms_results_DDSPSO_strategy1_errordds_iter' ,num2str(j), '.mat'));
-                DDS_error = bestval_dds_swarm{j};
-            catch    
-                DDS_error = 0;
-            end
-            DOPS_f(j) = DDS_error(end);
-            
[…]

### 2017-12-14 DOPS/de.m
Delete de.m

@@ -1,437 +1,0 @@
-%% Copyright (C) 1996, 1997 R. Storn
-%% Copyright (C) 2009-2010 Christian Fischer <cfischer@itm.uni-stuttgart.de>
-%%
-%% This program is free software; you can redistribute it and/or modify it under
-%% the terms of the GNU General Public License as published by the Free Software
-%% Foundation; either version 3 of the License, or (at your option) any later
-%% version.
-%%
-%% This program is distributed in the hope that it will be useful, but WITHOUT
-%% ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License for more
-%% details.
-%%
-%% You should have received a copy of the GNU General Public License along with
-%% this program; if not, see <http://www.gnu.org/licenses/>.
-
-%% de_min: global optimisation using differential evolution
-%%
-%% Usage: [x, obj_value, nfeval, convergence] = de_min(fcn, control)
-%%
-%% minimization of a user-supplied function with respect to x(1:D),
-%% using the differential evolution (DE) method based on an algorithm
-%% by  Rainer Storn (http://www.icsi.berkeley.edu/~storn/code.html)
-%% See: http://www.softcomputing.net/tevc2009_1.pdf
-%%
-%%
-%% Arguments:  
-%% ---------------
-%% fcn        string : Name of function. Must return a real value
-%% control    vector : (Optional) Control variables, described below
-%%         or struct
-%%
-%% Returned values:
-%% ----------------
-%% x          vector : parameter vector of best solution
-%% obj_value  scalar : objective function value of best solution
-%% nfeval     scalar : number of function evaluations
-%% convergence       : 1 = best below value to reach (VTR)
-%%                     0 = population has reached defined quality (tol)
-%%                    -1 = some values are close to constraints/boundaries
-%%                    -2 = max number of iterations reached (maxiter)
-%%                    -3 = max number of functions evaluations reached (maxnfe)
-%%
-%% Control variable:   (optional) may be named arguments (i.e. "name",value
-%% ----------------    pairs), a struct, or a vector, where
-%%                     NaN's are ignored.
-%%
[…]

### 2017-12-14 DOPS/runSimulatedAnnealingOnCoag.m
Delete runSimulatedAnnealingOnCoag.m

@@ -1,35 +1,0 @@
-function [allhistory]= runSimulatedAnnealingOnCoag()
-	seed = 1;
-	b=load('coagBounds.mat');
-	lb = b.lb;
-	ub = b.ub;
-   function [stop,options,optchanged] = myoutput(options,optimvalues,flag)
-        stop = false;
-        if isequal(flag,'iter')
-              history = [history; optimvalues.fval];
-            fprintf('Iter %d and functional value %f\n', optimvalues.iteration, optimvalues.fval);  
-            optimvalues.iteration; %so I can see what's going on'
-        end
-        optchanged = false;
-    end
-	NUM_ITERATIONS = 2000;
-	options = optimoptions(@simulannealbnd,'MaxIterations',NUM_ITERATIONS, 'OutputFcn', @myoutput);
-	NUM_REPEATS = 25;
-	allhistory = zeros(NUM_REPEATS, NUM_ITERATIONS+1);
-	for j=10:NUM_REPEATS
-        rng(j);
-		fprintf("On trial %d\n", j)
-		history = [];
-		%create initial guess
-		x0 = lb-ub.*rand(size(lb,1),1)+lb;
-		tstart = tic();
-		[x,fval,exitflag,output] = simulannealbnd(@fitCoag,x0,lb,ub,options);
-		timeSA(j) = toc(tstart)
-		%size(history) %history is one longer than number of iterations 
-		cmd1= ['save -ascii ../SA_Results/fitCoag/FuntionalValuesTrial', num2str(j), '.txt history'];
-		eval(cmd1)
-		cmd2= ['save -ascii ../SA_Results/fitCoag/TimesTrial', num2str(j), '.txt timeSA'];
-		eval(cmd2)
-		allhistory(j,:) = history;
-	end
-end

### 2017-12-14 DOPS/runSimulatedAnnealing.m
Delete runSimulatedAnnealing.m

@@ -1,37 +1,0 @@
-function[best_params, fvals, outputs, inital_vals]=runSimulatedAnnealing(seed)
-    function [stop,options,optchanged] = myoutput(options,optimvalues,flag)
-        stop = false;
-        if isequal(flag,'iter')
-              history = [history; optimvalues.fval];
-        end
-        optchanged = false;
-    end
-    
-
-    poss_iter_nums =[40,2000,4000];
-    functions_to_test = {@ackley, @rast, @rast};
-    bounds = {[30;-15], [5.12;-5.12], [5.12;-5.12]};
-    dimensions =[10,10,300];
-    best_params = {};
-    fvals = zeros(size(dimensions,2),size(functions_to_test,2));
-    inital_vals = zeros(size(dimensions,2), size(functions_to_test,2));
-    outputs = {};
-    for k =1:size(poss_iter_nums,2)
-        NUM_ITERATIONS=poss_iter_nums(k);
-         options = optimoptions(@simulannealbnd,'MaxIterations',NUM_ITERATIONS, 'OutputFcn', @myoutput);
-        for j=1:size(functions_to_test,2)
-            %create bounds
-            rng(seed)
-            currlimits = bounds{j};
-            MAXJ = repmat(currlimits(1),dimensions(j),1);
-            MINJ = repmat(currlimits(2), dimensions(j),1);
-            x0 = MAXJ-MINJ.*rand(size(MAXJ,1),1)+MINJ; %create an initial guess within bounds
-            inital_vals(j,k) = feval(functions_to_test{j},x0);
-            history = [];
-            [x,fval,exitflag,output]=simulannealbnd(functions_to_test{j},x0,MINJ,MAXJ,options);
-            best_params{j,k} = x;
-            fvals(j,k) = fval;
-            outputs{j,k} = output;
-        end
-    end
-end

### 2017-12-14 DOPS/runPSOOnCoag.m
Delete runPSOOnCoag.m

@@ -1,38 +1,0 @@
-function [allhistory]= runPSOOnCoag()
-	seed = 1;
-	b=load('coagBounds.mat');
-	lb = b.lb;
-	ub = b.ub;
-   function [stop] = myoutput(optimvalues,flag)
-        stop = false;
-        if isequal(flag,'iter')
-              history = [history; optimvalues.meanfval];
-              besthistory =[besthistory;optimvalues.bestfval];
-		optimvalues.iteration %so I can see what's going on'
-        end
-        optchanged = false;
-    end
-	NUM_ITERATIONS = 4000;
-	options = optimoptions('particleswarm','MaxIterations',NUM_ITERATIONS, 'OutputFcn', @myoutput, 'SwarmSize',40, 'UseParallel', true);
-	NUM_REPEATS = 25;
-	allhistory = [];
-	for j=1:NUM_REPEATS
-		fprintf("On trial %d\n", j)
-		history = [];
-        besthistory=[];
-		%create initial guess
-        rng(j,'twister');
-		x0 = lb-ub.*rand(size(lb,1),1)+lb;
-		tstart = tic();
-		[x,fval,exitflag,output] = particleswarm(@fitCoag,size(lb,1),lb,ub,options);
-		timeSA(j) = toc(tstart)
-		size(history) %history is one longer than number of iterations 
-		cmd1= ['save -ascii ../PSO_Results/fitCoag/MeanFuntionalValuesTrial', num2str(j), '.txt history'];
-		eval(cmd1)
-		cmd2= ['save -ascii ../PSO_Results/fitCoag/TimesTrial', num2str(j), '.txt timeSA'];
-		eval(cmd2)
-        cmd3= ['save -ascii ../PSO_Results/fitCoag/BestFuntionalValuesTrial', num2str(j), '.txt besthistory'];
-		eval(cmd3)
-		%allhistory(j,:) = history;
-	end
-end

### 2017-12-14 DOPS/runDDSBasic.m
Delete runDDSBasic.m

@@ -1,10 +1,0 @@
-function[best]=runDDSBasic(numIters,functionToTest,MINJ,MAXJ,seed)
-    r=0.2; 
-    NUM_ITERATIONS = numIters;
-    load('coagBounds.mat')
-        rng(seed,'twister');
-        tstart = tic();
-        x0 = MINJ+rand()*(MAXJ-MINJ);
-        [best,x_best,dds_swarm_flag]=DOPS_DDS(str2func(functionToTest),x0,MAXJ,MINJ,r,NUM_ITERATIONS);
-        times = toc(tstart);
-end

### 2017-12-14 DOPS/runDOPSforTable2.m
Delete runDOPSforTable2.m

@@ -1,25 +1,0 @@
-function[mean_data, std_data]=runDOPSforTable2()
-    functions_to_test = {@ackley, @rast, @rast, @rosenbrock, @BukinF6, @Beale};
-    bounds_files = {'ackley10d_bounds.mat','rast10d_bounds.mat', 'rast300d_bounds.mat', 'rosenbrock10d_bounds.mat', 'BukinF6Bounds.mat', 'BealeBounds.mat'};
-    NUM_TRIALS=25;
-    fvals = zeros(size(functions_to_test,2)-3,NUM_TRIALS);
-    initial_vals = zeros(size(functions_to_test,2)-3,NUM_TRIALS);
-    for j=1:size(functions_to_test,2)-3
-        load(bounds_files{j});
-         MAXJ = ub;
-         MINJ = lb;
-         [g_best_solution1,bestparticle1,particle1,fitness1,bestval_dds_swarm1,best_particle_dds_swarm1,best_particles_ls1]=run_DOPS(functions_to_test{j},MAXJ,MINJ,NUM_TRIALS);
-        [g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=run_DOPS(functions_to_test{j},MAXJ,MINJ,NUM_TRIALS);
-         %[g_best_solution,bestparticle,particle,fitness,bestval_dds_swarm,best_particle_dds_swarm,best_particles_ls]=run_DOPS_onlyPSO(functions_to_test{j},MAXJ,MINJ);
-        for k=1:NUM_TRIALS
-            if(size(bestval_dds_swarm{k},1) ==0)
-                fvals(j,k) = g_best_solution{k}(1); %min(fitness{k}(:,end));
-            else
-                fvals(j,k) = bestval_dds_swarm{k}(end);
-            end
-            initial_vals(j,k)=min(fitness1{k}(:,1));
-        end
-        mean_data(j) = mean(fvals(j,:)./initial_vals(j,:));
-        std_data(j) = std(fvals(j,:)./initial_vals(j,:));
-    end
-end

### 2017-12-14 DOPS/runDE.m
Delete runDE.m

@@ -1,62 +1,0 @@
-function[fvals,inital_vals]=runDE(seed)
-    % VTR		"Value To Reach" (stop when ofunc < VTR)
-            VTR = -Inf; 
-
-    % D		number of parameters of the objective function 
-            D = 2; 
-
-    % XVmin,XVmax   vector of lower and bounds of initial population
-    %    		the algorithm seems to work well only if [XVmin,XVmax] 
-    %    		covers the region where the global minimum is expected
-    %               *** note: these are no bound constraints!! ***
-            XVmin = [-2 -2]; 
-            XVmax = [2 2];
-
-    % y		problem data vector (remains fixed during optimization)
-            y=[]; 
-
-    % NP            number of population members
-            NP = 40; %to match with DOPS 
-
-    % itermax       maximum number of iterations (generations)
-            itermax = 200; 
-
-    % F             DE-stepsize F ex [0, 2]
-            F = 0.8; 
-
-    % CR            crossover probabililty constant ex [0, 1]
-            CR = 0.9; 
-
-    % strategy       1 --> DE/best/1/exp           6 --> DE/best/1/bin
-    %                2 --> DE/rand/1/exp           7 --> DE/rand/1/bin
-    %                3 --> DE/rand-to-best/1/exp   8 --> DE/rand-to-best/1/bin
-    %                4 --> DE/best/2/exp           9 --> DE/best/2/bin
-    %                5 --> DE/rand/2/exp           else  DE/rand/2/bin
-
-            strategy = 1;
-
-    % refresh       intermediate output will be produced after "refresh"
-    %               iterations. No intermediate output will be produced
-    %               if refresh is < 1
-            refresh = 0; 
-    poss_iter_nums =[40,2000,4000];
-    functions_to_test = {'ackley', 'rast', 'rast'};
-    bounds = {[30;-15], [5.12;-5.12], [5.12;-5.12]};
-    dimensions =[10,10,300];
-    best_params = {};
-    fvals = zeros(size(dimensions,2),size(functions_to_test,2));
-    inital_vals = zeros(size(dimensions,2), size(functions_to_test,2));
-    for k=1:size(poss_iter_nums,2)
-        for j =1:size(functions_to_test,2)            
-            currlimits = bounds{j};
-            MAXJ = repmat(currlimits(1),dimensions(j),1);
[…]

### 2017-12-14 DOPS/runDDSforTable2.m
Delete runDDSforTable2.m

@@ -1,20 +1,0 @@
-function[fvals, initial_vals]=runDDSforTable2(seed)
-    functions_to_test = {@ackley, @rast, @rast};
-    bounds_files = {'ackley10d_bounds.mat','rast10d_bounds.mat', 'rast300d_bounds.mat'};
-    poss_num_evals = [40,2000,4000];
-    r=0.2; 
-    initial_vals = zeros(size(functions_to_test,2), size(poss_num_evals,2));
-    fvals=zeros(size(functions_to_test,2), size(poss_num_evals,2));
-    for j=1:size(poss_num_evals,2)
-        for k=1:size(functions_to_test,2)
-            rng(seed)
-            load(bounds_files{k});
-            MAXJ = ub;
-            MINJ  = lb;
-            x0 = MINJ+rand()*(MAXJ-MINJ);
-            initial_vals(j,k) = feval(functions_to_test{k}, x0);
-            [best,x_best,dds_swarm_flag]=DOPS_DDS(functions_to_test{k},x0,MAXJ,MINJ,r,poss_num_evals(j));
-            fvals(j,k) = min(best);
-        end
-    end
-end

### 2017-12-14 DOPS/runDDSOnCoag.m
Delete runDDSOnCoag.m

@@ -1,24 +1,0 @@
-function[fvals, initial_vals]=runDDSOnCoag()
-    r=0.2; 
-    NUM_TRIALS = 25;
-    NUM_ITERATIONS = 4000;
-    fvals = zeros(NUM_TRIALS,NUM_ITERATIONS);
-    load('coagBounds.mat')
-    MAXJ = ub;
-    MINJ  = lb;
-    times = [];
-    for j = 1:NUM_TRIALS
-        rng(j,'twister');
-        tstart = tic();
-        x0 = MINJ+rand()*(MAXJ-MINJ);
-        [best,x_best,dds_swarm_flag]=DOPS_DDS(@fitCoag,x0,MAXJ,MINJ,r,NUM_ITERATIONS);
-        times(j) = toc(tstart)
-        cmd1= ['save -ascii ../DDS_Results/fitCoag/FuntionalValuesTrial', num2str(j), '.txt best'];
-		eval(cmd1)
-        cmd2= ['save -ascii ../DDS_Results/fitCoag/ParameterValuesTrial', num2str(j), '.txt x_best'];
-		eval(cmd2)
-        cmd3= ['save -ascii ../DDS_Results/fitCoag/TimesTrial', num2str(j), '.txt times'];
-		eval(cmd3)
-        fvals(j,:) = (best);
-    end
-end

### 2017-12-14 Paper_v7.tex
Working on LATeX issues

@@
 %%% Load packages
-%\usepackage{amsthm,amsmath}
-%\RequirePackage{natbib}
+\usepackage{amsthm,amsmath}
+\usepackage{amssymb}
+\usepackage{graphicx}
+\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
+
+% Imports required for p-code of JuPOETs
+\usepackage[linesnumbered]{algorithm2e}
+\usepackage{courier}
+%\usepackage{algorithm}
+\RequirePackage{natbib}
 %\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
-%\RequirePackage{hyperref}
+\RequirePackage{hyperref}
 \usepackage[utf8]{inputenc} %unicode support
@@
 %\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
-% Load packages
-\usepackage{url}  % Formatting web addresses
-\usepackage{ifthen}  % Conditional
-\usepackage{multicol}   %Columns
-\usepackage[utf8]{inputenc} %unicode support
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{epsfig}
-\usepackage{epstopdf}
-\usepackage{graphicx}
-\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
-\usepackage{setspace}
-%\usepackage{longtable}
-\usepackage{colortbl}
-%\usepackage{palatino,lettrine}
-%\usepackage{times}
-%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
-%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
-\usepackage[wide]{sidecap}
-%\usepackage[authoryear,round,comma,sort&compress]{natbib}
-\usepackage[square,sort,comma,numbers,sort&compress]{natbib}
-%\usepackage[authoryear,round]{natbib}
-\usepackage{supertabular}
-\usepackage{simplemargins}
-\usepackage{fullpage}
-\usepackage{comment}
-\usepackage{lineno}
-%\usepackage{chicago}
-\usepackage{textcomp}
-\usepackage{multirow}
-\usepackage{amsmath}
-\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
-\DeclareMathOperator*{\argmin}{\arg\!\min}
-
-\usepackage{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{subfig}
-%\usepackage[space]{cite}
-\urlstyle{rm}
 
@@
 
-%\def\includegraphic{}
-%\def\includegraphics{}
+\def\includegraphic{}
+\def\includegraphics{}
 
@@
 
-\title{Parameter estimation in biochemical models}
+\title{Dynamic Optimization with Particle Swarms (DOPS): A meta-heuristic for parameter estimation in biochemical models}
 
@@
 \author[
+   addressref={aff2},
+   email={jdv27@cornell.edu}
+]{\fnm{Christine} \snm{Shoemaker}}
+\author[
    addressref={aff1},
@@
 ]{\inits{D}\fnm{Jeffrey} \snm{Varner}}
-\author[
-   addressref={aff2},
-   email={jdv27@cornell.edu}
-]{\fnm{Christine} \snm{Shoemaker}}
 
@@
 The DDS phase refined the globally best particle to produce significantly lower error on average when compared to multi-swarm particle swarm optimization alone.
-Thus, the automated switching strategy was critical to the overall performance of DOPS.
-
-\subsection*{Continued Switching May Be Beneficial}
-We explored the performance of DOPS if it was permitted to switch between PSO and DDS modes.
-This mode ('Experimental DOPS') performed comparably to DOPS on 10-d Ackley and Rastrigin as well as on 300-dimensional Rastrigin.
-It performed better than DOPS on the CHO metabolism problem (Fig.
-\ref{fig-exp-DOPS-b4}a), with the average functional value for Experimental DOPS being nearly half of that of DOPS.
-To further distinguish DOPS from Experimental DOPS, we compared the performance of each algorithm on the Eggholder function, which is considered to be a difficult function to optimize, owing to its multiple minima.
-On this function, Experimental DOPS outperformed DOPS however, neither version of DOPS reached the true minimum at -959.6407 on any trial with a budget of 4,000 function evaluations (Fig.
+Thus, the automated switching strategy was critical to the overall performance of DOPS. However, it was unclear if multiple strategy switches could further improve perfomance.
+
+We explored the performance of DOPS if it was permitted to switch between PSO and DDS modes multiple times.
+This mode (\texttt{Experimental~DOPS}) performed comparably to DOPS on 10-d Ackley and Rastrigin as well as on 300-dimensional Rastrigin function.
+\texttt{Experimental~DOPS} performed better than DOPS on the CHO metabolism problem (Fig. \ref{fig-exp-DOPS-b4}a),
+with the average functional value being nearly half of that of DOPS.
+To further distinguish DOPS from \texttt{Experimental~DOPS}, we compared the performance of each algorithm on the Eggholder function,
+a difficult function to optimize given its its multiple minima.
+On the Eggholder function, \texttt{Experimental~DOPS} outperformed DOPS however, neither version of DOPS reached the true minimum at -959.6407 on any trial with a budget of 4,000 function evaluations (Fig.
 \ref{fig-exp-DOPS-b4}b).
-We also explored the performance of both versions of DOPS on the 100 dimensional Styblinksi-Tang function \cite{jamil2013literature} (Fig.
-\ref{fig-exp-DOPS-b4}c).
-In this comparison, experimental DOPS far outperforms DOPS, succeeding in finding the true minimum before exhausting its budget of functional evaluations, whereas DOPS does not reach the minimum.
-Since the performance of Experimental DOPS was promising on these problems, we decided to measure its performance on the coagulation problem.
-At 4,000 function evaluations, DOPS performed better than Experimental DOPS, however, the difference between the two method's final average functional values was small (Fig.
-\ref{fig-exp-DOPS-b4}d), with the final average objective value for DOPS reaching .9413\% of the initial functional value, compared to .9428\% for Experimental DOPS.
+We also explored the performance of both versions of DOPS on the 100 dimensional Styblinksi-Tang function \cite{jamil2013literature} (Fig. \ref{fig-exp-DOPS-b4}c).
+In this comparison, \texttt{Experimental~DOPS} significantly outperformed DOPS, finding the true minimum before exhausting its functional evaluation budget, while DOPS does not reach the minimum.
+Since the performance of \texttt{Experimental~DOPS} was promising on these problems, we measured its performance on the coagulation problem.
+Surprisingly, DOPS performed slightly better than \texttt{Experimental~DOPS} on the coagulation problem (Fig. \ref{fig-exp-DOPS-b4}d);
+the final average objective value for DOPS reached 0.9413\% of the initial functional value, compared to 0.9428\% for \texttt{Experimental~DOPS}.
 
@@
 \subsection*{Experimental DOPS}
-We sought to investigate if switching methods more than once would result in better performance.
-Experimental DOPS begins with the PSO phase and uses the same criteria as DOPS to switch to the DDS phase.
+We investigated where switching methods more than once would result in better performance; this DOPS variant is referred to as \texttt{Experimental~DOPS}.
+\texttt{Experimental~DOPS} begins with the PSO phase and uses the same criteria as DOPS to switch to the DDS phase.
 The technique switches back to PSO phase when the DDS phase has reduced the functional value to 90\% of its initial value.
@@
 When the switch from DDS to PSO occurs, we use the best solution from DDS to seed the particle swarm.
-
 DOPS and Experimental DOPS source code is available for download under a MIT license at http://www.varnerlab.org.
@@
 
+\section*{Ethics approval and consent to participate}
+Not applicable
+
+\section*{Consent to publish}
+Not applicable
+
 \section*{Competing interests}
@@
 \section*{Author's contributions}
-    A.S. originally developed algorithm. R.L developed experimental DOPS and generated all of the figures. JDV assisted in manuscript preparation.
-
-\section*{Acknowledgements}
- This study was supported by an award from the Army Research Office (ARO \#59155-LS).
+    A.S. developed original DOPS algorithm. R.L developed experimental DOPS and generated all of the figures.
+    CS developed the dynamic dimensional search approach, and assisted in editing the manuscript.
+    JDV directed the study, prepared and edited the manuscript.
+
+\section*{Availability of data and materials}
+DOPS is open source, available under an MIT software license.
+The DOPS source code is freely available from the DOPS GitHub repository at https://github.com/varnerlab/POETs.jl.
+All samples used in this study are included in the \texttt{sample/biochemical} and \texttt{sample/test\_functions} subdirectories of the JuPOETs GitHub repository.
+
+\section*{Funding}
+This study was supported by an award from the US Army and Systems Biology of Trauma Induced Coagulopathy (W911NF-10-1-0376).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\clearpage
+
 \section*{Figures}
@@
 \centering
-\includegraphics[width=.85\textwidth]{./figs/Figure_1_Algorithm.pdf}
+\includegraphics[width=0.85\textwidth]{./figs/Figure_1_Algorithm.pdf}
 \caption{\csentence{Schematic of the dynamic optimization with particle swarms (DOPS) approach.} \textbf{A}: Each particle represents an N dimensional parameter vector.
@@
 
+\clearpage
+
 \subsection*{Additional file S8-(Comparison of functional values)}

### 2017-12-14 Paper_v7.tex
Working on LATeX issues
adding syblinkski tang ref

@@
 
-\def\includegraphic{}
-\def\includegraphics{}
+% \def\includegraphic{}
+% \def\includegraphics{}
 
@@
 \ref{fig-exp-DOPS-b4}b).
-We also explored the performance of both versions of DOPS on the 100 dimensional Styblinksi-Tang function \cite{jamil2013literature} (Fig. \ref{fig-exp-DOPS-b4}c).
+We also explored the performance of both versions of DOPS on the 100 dimensional Styblinksi-Tang function \cite{styblinski1990experiments} (Fig. \ref{fig-exp-DOPS-b4}c).
 In this comparison, \texttt{Experimental~DOPS} significantly outperformed DOPS, finding the true minimum before exhausting its functional evaluation budget, while DOPS does not reach the minimum.
@@
 
+\clearpage
+
 \begin{figure}[h!]
@@
 \end{figure}
+
+\clearpage
 
@@
 
-  \begin{figure}[h!]
-  \caption{\csentence{Sample figure title.}
-      A short description of the figure content
-      should go here.}
-      \end{figure}
+\clearpage
 
@@
 \end{table}
+
+\clearpage
 
@@
 
+\clearpage
+
 \section*{Additional Files}
